### 2017-11-30 paper/paper.tex
added script and template for paper

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4-1}
%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-1}
\documentclass[aps,prl,reprint,groupedaddress]{revtex4-1}

% You should use BibTeX and apsrev.bst for references
% Choosing a journal automatically selects the correct APS
% BibTeX style file (bst file), so only uncomment the line
% below if necessary.
%\bibliographystyle{apsrev4-1}

\begin{document}

% Use the \preprint command to place your local institutional report
% number in the upper righthand corner of the title page in preprint mode.
% Multiple \preprint commands are allowed.
% Use the 'preprintnumbers' class option to override journal defaults
% to display numbers if necessary
%\preprint{}

%Title of paper
\title{The Magnetorotational Instability Prefers Three Dimensions}

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current
% author. Explanatory text should go in the []'s, actual e-mail
% address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro foreach each type of information

% \affiliation command applies to all authors since the last
% \affiliation command. The \affiliation command should follow the
% other information
% \affiliation can be followed by \email, \homepage, \thanks as well.
\author{Dedalus}
%\email[]{Your e-mail address}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\affiliation{}

%Collaboration name if desired (requires use of superscriptaddress
%option in \documentclass). \noaffiliation is required (may also be
%used with the \author command).
%\collaboration can be followed by \email, \homepage, \thanks as well.
%\collaboration{}
%\noaffiliation

\date{\today}

\begin{abstract}
The magnetorotational instability shows growth rates in three dimensions where the two dimensional problem shows nothing. NOTHING!
\end{abstract}

% insert suggested PACS numbers in braces on next line
\pacs{}
% insert suggested keywords - APS authors don't need to do this
%\keywords{}

%\maketitle must follow title, authors, abstract, \pacs, and \keywords
\maketitle

% body of paper here - Use proper section commands
% References should be done using the \cite, \ref, and \label commands

\section{Introduction}
\label{sec:intro}

\section{Method}
\label{sec:method}

% If in two-column mode, this environment will change to single-column
% format so that long equations can be displayed. Use
% sparingly.
%\begin{widetext}
% put long equation here
%\end{widetext}

% figures should be put into the text as floats.
% Use the graphics or graphicx packages (distributed with LaTeX2e)
% and the \includegraphics macro defined in those packages.
% See the LaTeX Graphics Companion by Michel Goosens, Sebastian Rahtz,
% and Frank Mittelbach for instance.
%
% Here is an example of the general form of a figure:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Use the figure* environment if the figure should span across the
% entire page. There is no need to do explicit centering.

% \begin{figure}
% \includegraphics{}%
% \caption{\label{}}
% \end{figure}

% Surround figure environment with turnpage environment for landscape
% figure
% \begin{turnpage}
% \begin{figure}
% \includegraphics{}%
% \caption{\label{}}
% \end{figure}
% \end{turnpage}

% tables should appear as floats within the text
%
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
% The ruledtabular enviroment adds doubled rules to table and sets a
% reasonable default table settings.
% Use the table* environment to get a full-width table in two-column
% Add \usepackage{longtable} and the longtable (or longtable*}
% environment for nicely formatted long tables. Or use the the [H]
% placement option to break a long table (with less control than 
% in longtable).
% \begin{table}%[H] add [H] placement to break table across pages
% \caption{\label{}}
% \begin{ruledtabular}
% \begin{tabular}{}
% Lines of table here ending with \\
% \end{tabular}
% \end{ruledtabular}
% \end{table}

% Surround table environment with turnpage environment for landscape
% table
% \begin{turnpage}
% \begin{table}
% \caption{\label{}}
% \begin{ruledtabular}
% \begin{tabular}{}
% \end{tabular}
% \end{ruledtabular}
% \end{table}
% \end{turnpage}

% Specify following sections are appendices. Use \appendix* if there
% only one appendix.
%\appendix
%\section{}

% If you have acknowledgments, this puts in the proper section head.
%\begin{acknowledgments}
% put your acknowledgments here.
%\end{acknowledgments}

% Create the reference section using BibTeX:
\bibliography{basename of .bib file}

\end{document}
%
% ****** End of file apstemplate.tex ******

### 2017-11-30 python/mri.py

@@ -1,0 +1,180 @@
+"""
+The magnetorotational instability prefers three dimensions.
+    
+Dedalus script for calculating the maximum growth rates for the
+3D MRI over a range of wavenumbers.
+
+This script can be ran serially or in parallel, and produces a plot of the
+highest growth rate found for each wavenumber.
+
+To run using 4 processes, for instance, you could use:
+    $ mpiexec -n 4 python3 mri.py
+
+"""
+
+import time
+import numpy as np
+import matplotlib.pyplot as plt
+import dedalus.public as de
+from mpi4py import MPI
+CW = MPI.COMM_WORLD
+import logging
+logger = logging.getLogger(__name__)
+
+
+# Global parameters
+Nx, Lx, B = 128, np.pi, 1
+
+R      =  1.01 # 1   = critical
+q      =  3/4. # 3/4 = Keplerian
+
+kx     =  np.pi/Lx
+S      = -R*B*kx*np.sqrt(q)
+f      =  R*B*kx/np.sqrt(q)
+cutoff =  kx*np.sqrt(R**2 - 1)
+
+kymin, kymax, Nky = 0.00, 0.8, 101 # 2*pi*(Lx/Ly)
+kzmin, kzmax, Nkz = 0.01, 1.0, 100 # 2*pi*(Lx/Ly)
+
+ν, η = 1e-5, 1e-5
+
+# Create bases and domain
+# Use COMM_SELF so keep calculations independent between processes
+x_basis = de.Chebyshev('x', Nx, interval=(-Lx/2, Lx/2))
+domain = de.Domain([x_basis], grid_dtype=np.complex128, comm=MPI.COMM_SELF)
+
+# 3D MRI
+problem = de.EVP(domain, variables=['p','vx','vy','vz','ωy','ωz','bx','by','bz','jxx'], eigenvalue='gamma')
+problem.meta[:]['x']['dirichlet'] = True
+
+# Local parameters
+
+problem.parameters['S'] = S
+problem.parameters['f'] = f
+problem.parameters['B'] = B
+
+problem.parameters['ky'] = 0
+problem.parameters['kz'] = 0.5*cutoff
+
+problem.parameters['ν'] = ν
+problem.parameters['η'] = η
+
+# Operator substitutions for y,z, and t derivatives
+
[…]

### 2018-03-02 paper/paper.tex
added Andrew Swan to authors
some equations jazz

@@
 \documentclass[aps,prl,reprint,groupedaddress]{revtex4-1}
-
+\usepackage{graphicx}
 % You should use BibTeX and apsrev.bst for references
@@
 \author{Dedalus}
-%\email[]{Your e-mail address}
+\author{Andrew Swan}
+% \email[]{Your e-mail address}
 %\homepage[]{Your web page}
@@
 
+We solve the linearized magnetohydrodynamic equations in a rotating, shearing background. The Navier-Stokes equation is in a fairly standard form,
+\begin{equation}
+  \label{eq:ns}
+  \frac{D \mathbf{v}}{Dt} - f \hat{z} \times \mathbf{v} + S v_x \hat{z} + \mathbf{\nabla}{p} + \nu \mathbf{\nabla} \times \mathbf{\omega} = 0,
+\end{equation}
+where $\omega = \mathbf{\nabla} \times \mathbf{v}$ is the vorticity and $S$ is the background shear rate. We write the induction equation in terms of the $x$ component of the magnetic field,
+\begin{equation}
+  \label{eq:Bx}
+  \frac{D B_x}{Dt} - B_0 \partial_z v_x + \eta (\partial_y J_z - \partial_z J_y) = 0
+\end{equation}
+and the $x$ component of the current density,
+\begin{equation}
+  \label{eq:Jx}
+  \frac{D J_x}{Dt} - B_0 \partial_z \omega_x + S \partial_z B_x - \eta \nabla^2 J_x = 0
+\end{equation}
+We explicitly enforce the divergence constraint on the incompressible velocity field $v$ and the magnetic field,
+\begin{equation}
+  \label{eq:divu}
+  \mathbf{\nabla} \cdot \mathbf{v} = \mathbf{\nabla} \cdot \mathbf{B} = 0.
+\end{equation}
+We consider the behavior of the MRI in a vertically ($z$) and horizontally ($y$) periodic channel of width $L_x = \pi$. The walls at $x = \pm L_x/2$ are perfectly conducting and stress-free.
+
+Using the \emph{Dedalus} pseudospectral framework, we solve equations~(\ref{eq:ns})(\ref{eq:Bx})(\ref{eq:divu})
+\section{Results}
+\label{sec:results}
+
+\begin{figure}[h]
+  \includegraphics[width=0.95\columnwidth]{growth_rates.png}
+  \caption{Growth rate as a function of $k_z$ and $k_y$. The maximum growth rate is far from $k_y = 0$, indicating the MRI is most unstable in a fully three-dimensional configuration.}
+  \label{fig:growth_rate}
+\end{figure}
 % If in two-column mode, this environment will change to single-column

### 2018-03-02 python/mri.py
search for more modes...makes things *more* stable AND faster??

@@ -118,12 +118,11 @@ def ideal_2D(kz):
     return np.sqrt( ( - b + np.sqrt(b*b - 4*a*c + 0j) ) / (2*a) )
 
 # Create function to compute max growth rate for given ky, kz
-def growth_rate(ky,kz,target,N=2):
+def growth_rate(ky,kz,target,N=15):
     
     # Change ky, kz parameters
     problem.namespace['ky'].value = ky
     problem.namespace['kz'].value = kz
-    
     # Solve for eigenvalues with sparse search near target, rebuilding NCCs
     try:
         solver.solve_sparse(solver.pencils[0], N=N, target=target, rebuild_coeffs=True)

### 2018-03-02 paper/paper.tex
more equations

@@
 \end{equation}
-We consider the behavior of the MRI in a vertically ($z$) and horizontally ($y$) periodic channel of width $L_x = \pi$. The walls at $x = \pm L_x/2$ are perfectly conducting and stress-free.
+We consider the behavior of the MRI in a vertically ($z$) and horizontally ($y$) periodic channel of width $L_x = \pi$.
+The walls at $x = \pm L_x/2$ are perfectly conducting and stress-free.
+The MRI is a weak-field instability; in the invisicd case
 
-Using the \emph{Dedalus} pseudospectral framework, we solve equations~(\ref{eq:ns})(\ref{eq:Bx})(\ref{eq:divu})
+We solve equations~(\ref{eq:ns})--~(\ref{eq:divu}) by assuming harmonic perturbations in $y$ and $z$, $f = \hat{f}(x) e^{i(k_y y + k_z) + \gamma t}$.
+This reduces to an eigenvalue problem in $x$, which we solve using the \emph{Dedalus} pseudospectral framework.
+
 \section{Results}

### 2019-06-17 python/mri.py
Add dense solver option; calc metric for optimization
* optimization metric: (guess - opt)/guess where guess and opt are
  growth rates from the max on the grid and the output of the
  minimizer, respectively

@@ -43,6 +43,16 @@ config.read(str(filename))
 logger.info('Running mri.py with the following parameters:')
 logger.info(config.items('parameters'))
 
+try:
+    dense = config.getboolean('solver','dense')
+except:
+    dense = False
+if dense:
+    logger.info("Using dense solver.")
+    dense_threshold = config.getfloat('solver','dense_threshold')
+else:
+    logger.info("Using sparse solver.")
+
 Nx = config.getint('parameters','Nx')
 Lx = eval(config.get('parameters','Lx'))
 B = config.getfloat('parameters','B')
@@ -150,7 +160,7 @@ def ideal_2D(kz):
     return np.sqrt( ( - b + np.sqrt(b*b - 4*a*c + 0j) ) / (2*a) )
 
 # Create function to compute max growth rate for given ky, kz
-def growth_rate(ky,kz,target,N=15):
+def growth_rate(ky,kz,target,N=15, dense=False):
     eigvec = np.zeros((10,Nx),dtype=np.complex128)
     # Change ky, kz parameters
     problem.namespace['ky'].value = ky
@@ -158,7 +168,10 @@ def growth_rate(ky,kz,target,N=15):
     # Solve for eigenvalues with sparse search near target, rebuilding NCCs
     solver_failed = False
     try:
-        solver.solve_sparse(solver.pencils[0], N=N, target=target, rebuild_coeffs=True)
+        if dense:
+            solver.solve_dense(solver.pencils[0], rebuild_coeffs=True)
+        else:
+            solver.solve_sparse(solver.pencils[0], N=N, target=target, rebuild_coeffs=True)
     except:
         logger.info("Solver failed for (ky, kz) = (%f, %f)"%(ky, kz))
         solver_failed = True
@@ -170,6 +183,8 @@ def growth_rate(ky,kz,target,N=15):
         gamma.append(gamma_r + 1j*gamma_i)
     else:
         gamma = solver.eigenvalues
+        if dense:
+            gamma = gamma[np.abs(gamma) < dense_threshold]
         index = np.argsort(-gamma.real)
         gamma = gamma[index]
     
@@ -203,13 +218,13 @@ eigvec_local = eigvec_global[:,CW.rank::CW.size]
 
 t1 = time.time()
 for k, kz in enumerate(kz_local):
-    soln = growth_rate(0., kz, ideal_2D(kz), N=Nmodes)
[…]

### 2019-06-18 python/plot_alpha.py
Updated alpha plot to show growth rates
* also, uses 200x200 grid

@@ -9,17 +9,21 @@ import matplotlib.pyplot as plt
 from pathlib import Path
 
 basedir = Path("data/")
-runs = [1, 2, 3]
-ssc = np.array([1.01, 1.2, 2.0])**2
+runs = [21, 15, 11, 12, 16, 13, 17, 14, 18]
+R = [np.sqrt(0.5), 0.8, 1.01, 1.2, np.sqrt(1.75), np.sqrt(2.0), 1.5, np.sqrt(2.5), 2.0]
+ssc = np.array(R)**2
 alpha = []
 
+growth = []
 for run in runs:
     dfname = Path("run_{:d}_output.h5".format(run))
     df = h5py.File(basedir/dfname, "r")
-    alpha.append(np.arctan2(df['gamma'].attrs['max ky'], df['gamma'].attrs['max kz']))
+    #alpha.append(np.arctan2(df['gamma'].attrs['max ky'], df['gamma'].attrs['max kz']))
+    kk = np.sqrt(df['gamma'].attrs['max ky']**2 + df['gamma'].attrs['max kz']**2)
+    alpha.append(np.arcsin(df['gamma'].attrs['max ky']/kk))
+    growth.append(df['gamma'].attrs['max growth rate'])
     df.close()
-
-plt.plot(ssc, alpha,'ko')
+plt.scatter(ssc, alpha, marker='o', c=growth)
 plt.xlabel(r"$S/S_c$")
 plt.ylabel(r"$\alpha$")
 plt.savefig("plots/alpha_vs_ssc.png")

### 2019-06-18 paper/paper.tex
Updates to paper draft.

@@
 \documentclass[aps,prl,reprint,groupedaddress]{revtex4-1}
+\newcommand{\SSC}{S/S_{c}}
 \usepackage{graphicx}
@@
 \author{Andrew Swan}
+\author{Morgan Baxter}
+
 % \email[]{Your e-mail address}
@@
 \begin{abstract}
-The magnetorotational instability shows growth rates in three dimensions where the two dimensional problem shows nothing. NOTHING!
+Write abstract last.
 \end{abstract}
@@
 
-We solve the linearized magnetohydrodynamic equations in a rotating, shearing background. The Navier-Stokes equation is in a fairly standard form,
+We solve the linearized magnetohydrodynamic equations in a frame rotating with angular frequency $\Omega$. The Navier-Stokes equations is in a fairly standard form,
 \begin{equation}
@@
 \end{equation}
-where $\omega = \mathbf{\nabla} \times \mathbf{v}$ is the vorticity and $S$ is the background shear rate. We write the induction equation in terms of the $x$ component of the magnetic field,
+where $\omega = \mathbf{\nabla} \times \mathbf{v}$ is the vorticity, $f = 2 \Omega$ is the Coriolis parameter, and $S$ is the background shear rate.
+$Df/Dt = \partial f/\partial t + S x \partial f/\partial y$ is the advective derivative including the background linear shear velocity.
+We write the induction equation in terms of the $x$ component of the magnetic field,
 \begin{equation}
@@
 \end{equation}
+The magnetic field is expressed in Alfven units, $\mathbf{B} = \mathbf{B_*}/\sqrt{\mu_0 \rho}$.
 We consider the behavior of the MRI in a vertically ($z$) and horizontally ($y$) periodic channel of width $L_x = \pi$.
 The walls at $x = \pm L_x/2$ are perfectly conducting and stress-free.
-The MRI is a weak-field instability; in the invisicd case
+The MRI is a weak-field instability; in the invisicd, ideal case the critical shear rate for instability in two dimensions is given by
+\begin{equation}
+  \label{eq:Sc}
+  S_c = \frac{-\pi^2 B^2}{f d^2}.
+\end{equation}
+Here, we use $\SSC$ as our control parameter.
 
@@
 This reduces to an eigenvalue problem in $x$, which we solve using the \emph{Dedalus} pseudospectral framework.
+The advantage of this particular form is that it can be expressed as a set of ten first-order equations with Dirichlet boundary conditions on all ten variables.
 
@@
 \label{sec:results}
+Our most surprising result is that at very small values of supercriticality, the fastest growing mode is not axisymmetric.
+Figure~\ref{fig:growth_rate} shows the growth rates for three values of $\SSC$.
+At $\SSC = 1.01$, the maximum growth rate is $(k_y, k_z) \simeq (0.263, 0.447)$.
 
 \begin{figure}[h]
-  \includegraphics[width=0.95\columnwidth]{growth_rates.png}
+  \includegraphics[width=0.33\textwidth]{run_11_output_growthrates.png}
+  \includegraphics[width=0.33\textwidth]{run_12_output_growthrates.png}
+  \includegraphics[width=0.33\textwidth]{run_13_output_growthrates.png}
   \caption{Growth rate as a function of $k_z$ and $k_y$. The maximum growth rate is far from $k_y = 0$, indicating the MRI is most unstable in a fully three-dimensional configuration.}
@@
 \end{figure}
-% If in two-column mode, this environment will change to single-column
+
+Numerical simulations have consistently axisymmetric modes dominating early evolution of the MRI (cite Stone 95, etc) before parasitic modes cause a breakdown into MHD turbulence.
+However, our results recover this limit, as $\SSC$ becomes large.
+In this case
+
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{alpha_vs_ssc.png}
+  \caption{Angle $\alpha$ vs $\SSC$.}
+  \label{fig:alpha}
+\end{figure}
+\section{Non-Normality}
+\label{sec:non-normality}
+
+The linear operator for the MRI, like other shearing systems, is non-normal. 
+%If in two-column mode, this environment will change to single-column
 % format so that long equations can be displayed. Use

### 2019-06-24 paper/paper.tex
Updated figures; added to text a bit.

@@
 
+\begin{itemize}
+\item 3D $\to$ dynamo
+\item stars, NSSL
+\item experiments
+\end{itemize}
+
 \section{Method}
@@
 
-We solve equations~(\ref{eq:ns})--~(\ref{eq:divu}) by assuming harmonic perturbations in $y$ and $z$, $f = \hat{f}(x) e^{i(k_y y + k_z) + \gamma t}$.
+We solve equations~(\ref{eq:ns})--~(\ref{eq:divu}) by assuming harmonic perturbations in $y$ and $z$, $f = \hat{f}(x) e^{i(k_y y + k_z z) + \gamma t}$.
 This reduces to an eigenvalue problem in $x$, which we solve using the \emph{Dedalus} pseudospectral framework.
-The advantage of this particular form is that it can be expressed as a set of ten first-order equations with Dirichlet boundary conditions on all ten variables.
+The advantage of this particular form is that it can be expressed as a set of ten first-order equations with Dirichlet boundary conditions for either no-slip or stress-free boundary conditions.
+
+Our interest is in nearly ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions; we thus set $\eta=\nu=10^{-5}$, though we have confirmed our main results are insensitive to this particular value.
 
@@
 \begin{figure}[h]
-  \includegraphics[width=0.33\textwidth]{run_11_output_growthrates.png}
-  \includegraphics[width=0.33\textwidth]{run_12_output_growthrates.png}
-  \includegraphics[width=0.33\textwidth]{run_13_output_growthrates.png}
+  \includegraphics[width=\columnwidth]{run_11_output_growthrates.png}
+  \includegraphics[width=\columnwidth]{run_12_output_growthrates.png}
+  \includegraphics[width=\columnwidth]{run_13_output_growthrates.png}
   \caption{Growth rate as a function of $k_z$ and $k_y$. The maximum growth rate is far from $k_y = 0$, indicating the MRI is most unstable in a fully three-dimensional configuration.}
@@
 However, our results recover this limit, as $\SSC$ becomes large.
-In this case
+Figure~\ref{fig:alpha} plots $\alpha = \arctan k_y/k_z$ as  function of $\SSC$.
+Above $\SSC \gtrsim 2$, $\alpha$ is zero, indicating that axisymmetric modes have the fastest growth rates.
 
 \begin{figure}[h]
-  \includegraphics[width=\columnwidth]{alpha_vs_ssc.png}
-  \caption{Angle $\alpha$ vs $\SSC$.}
+  \includegraphics[width=\columnwidth]{alpha_vs_ssc_grid.pdf}
+  \caption{Angle $\alpha$ vs $\SSC$. Stability limit is at $\SSC \simeq 0.102$}
   \label{fig:alpha}
@@
 
+\section{Speculations}
+\label{sec:speculations}
+
+\begin{itemize}
+\item non-normal and $k_y$ even when $\alpha = 0$ implies important role for linear dynamics at early times. 
+\item Following Squire + Bhattacharjee, this suggests that important stellar layers may in fact be driven by linear dynamics
+\item why is MRI non-oscillatory when hydro shear is oscil or stable?
+\end{itemize}
 The linear operator for the MRI, like other shearing systems, is non-normal. 

### 2019-06-24 python/make_eigenvalue_plot.py
aesthetic considerations for plots

@@ -12,6 +12,8 @@ from pathlib import Path
 import numpy as np
 import matplotlib.pyplot as plt
 
+plt.style.use('prl')
+
 # Parses filename passed to script
 parser = argparse.ArgumentParser(description='Passes filename')
 parser.add_argument('filename', metavar='Rc', type=str, help='.h5 file to plot eigenvectors for maximum eigenvalue')
@@ -31,12 +33,13 @@ max_kz = datafile['gamma'].attrs['max kz']
 ky_global    = datafile['ky']
 kz_global    = datafile['kz']
 
-PCM = plt.pcolormesh(kz_global,ky_global,gamma_r)
+PCM = plt.pcolormesh(kz_global,ky_global,gamma_r, vmin=-0.01,vmax=0.071)
 plt.contour(kz_global,ky_global,gamma_r,10,colors='k')
 plt.plot(max_kz, max_ky, 'ro')
 plt.colorbar(PCM)
 plt.xlabel(r'$k_z$')
 plt.ylabel(r'$k_y$')
+plt.tight_layout()
 #plt.title(r'3D Keplerian MRI growth rates/f  $\left( S/S_{\mathrm{crit.}} = %.2f\right)$' %(R[i]))
 plot_file_name = Path(filename.stem + '_growthrates.png')
 plt.savefig(outbase/plot_file_name, dpi=300)

### 2019-06-24 python/plot_alpha.py

@@ -4,26 +4,128 @@ alpha = arctan(ky/kz)
 """
 
 import numpy as np
+import numpy.polynomial.polynomial as poly
 import h5py
 import matplotlib.pyplot as plt
 from pathlib import Path
+from basic_units import radians
+
+plt.style.use('prl')
+
+use_opt = False
 
 basedir = Path("data/")
-runs = [21, 15, 11, 12, 16, 13, 17, 14, 18]
-R = [np.sqrt(0.5), 0.8, 1.01, 1.2, np.sqrt(1.75), np.sqrt(2.0), 1.5, np.sqrt(2.5), 2.0]
+runs = [34, 33, 31, 21, 15, 11, 12, 16, 30, 13, 27, 28, 29, 26, 24, 17, 14, 18]
+R = [np.sqrt(0.2), np.sqrt(0.3), np.sqrt(0.4), np.sqrt(0.5), 0.8, 1.01, 1.2, np.sqrt(1.75), 1.375, np.sqrt(2.0), np.sqrt(2.01), np.sqrt(2.015), 1.425, np.sqrt(2.05), np.sqrt(2.1), 1.5, np.sqrt(2.5), 2.0]
 ssc = np.array(R)**2
 alpha = []
 
 growth = []
+dk = []
 for run in runs:
     dfname = Path("run_{:d}_output.h5".format(run))
     df = h5py.File(basedir/dfname, "r")
-    #alpha.append(np.arctan2(df['gamma'].attrs['max ky'], df['gamma'].attrs['max kz']))
-    kk = np.sqrt(df['gamma'].attrs['max ky']**2 + df['gamma'].attrs['max kz']**2)
-    alpha.append(np.arcsin(df['gamma'].attrs['max ky']/kk))
-    growth.append(df['gamma'].attrs['max growth rate'])
+    if use_opt:
+        #alpha.append(np.arctan2(df['gamma'].attrs['max ky'], df['gamma'].attrs['max kz']))
+        ky = df['gamma'].attrs['max ky']
+        kx = df['gamma'].attrs['max kz']
+        gamma_max = df['gamma'].attrs['max growth rate']
+    else:
+        gamma_global = df['gamma'][:,:]
+        max_where = np.unravel_index(gamma_global.argmax(), gamma_global.shape)
[…]

### 2019-07-18 paper/paper.tex
Full author list

@@
 %\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-1}
-\documentclass[aps,prl,reprint,groupedaddress]{revtex4-1}
+\documentclass[aps,prl, reprint,superscriptaddress]{revtex4-1}
 \newcommand{\SSC}{S/S_{c}}
@@
 % \affiliation can be followed by \email, \homepage, \thanks as well.
-\author{Dedalus}
+\author{Jeffrey~S.~Oishi}
+\email[]{joishi@bates.edu}
+\affiliation{Bates College Department of Physics and Astronomy, Lewiston, ME 04240}
+
+\author{Geoffrey~M.~Vasil}
+\affiliation{University of Sydney School of Mathematics and Statistics, Sydney, NSW, Australia}
+\author{Morgan Baxter}
+\affiliation{Bates College Department of Physics and Astronomy, Lewiston, ME 04240}
 \author{Andrew Swan}
-\author{Morgan Baxter}
+\affiliation{Faculty of Mathematics, Cambridge University, Cambridge, United Kingdom}
+\author{Keaton~J.~Burns}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010}
+\affiliation{Massachusetts Institute of Technology Department of Physics, Cambridge, MA 02139}
+\author{Daniel~Lecoanet}
+\affiliation{Princeton Center for Theoretical Science and Princeton University Department of Astrophysical Sciences, Princeton, NJ 08544}
+\author{Benjamin~P.~Brown}
+\affiliation{University of Colorado Laboratory for Atmospheric and Space Physics and Department of Astrophysical and Planetary Sciences, Boulder, CO 80309}
 
-% \email[]{Your e-mail address}
+% 
 %\homepage[]{Your web page}
@@
 %\altaffiliation{}
-\affiliation{}
 

### 2019-07-19 python/plot_fig1.py
Added a bunch of runs; fixed fig1 so asymptotics and numerics align
* just needed more resolution!

@@ -11,7 +11,7 @@ plt.style.use('prl')
 filebase = 'data/run_{:d}_output.h5'
 runs = [15, 39, 12, 13]
 SSC = [0.64, 1.002001, 1.0201, 2.]
-index = 50
+index = 0
 
 fig = plt.figure(figsize=(16,8))
 # grid = AxesGrid(fig, (0.125, 0.5,0.8,0.35),
@@ -42,7 +42,7 @@ for i,r in enumerate(runs):
     contour_levels = np.linspace(0,gamma_r.max(),5)
     vmin = 0.
     vmax = 0.071
-    c = ax.pcolormesh(kz_global,ky_global,gamma_r, vmin=vmin,vmax=vmax)
+    c = ax.imshow(gamma_r, extent=[kz_global[:].min(),kz_global[:].max(),ky_global[:].min(),ky_global[:].max()], vmin=vmin,vmax=vmax, origin='lower')
 
     ax.contour(kz_global,ky_global,gamma_r,levels=contour_levels,colors='k')
     ax.contour(kz_global,ky_global,gamma_r,levels=[0.],colors='w')
@@ -68,7 +68,7 @@ xloc = grid[0].get_position().x0
 w_lower = 2*w + pad
 asymp_ax = fig.add_axes([0.1,0.1,w_lower,0.35])
 
-filename = 'data/run_39_output.h5'
+filename = 'data/run_41_output.h5'
 data = h5py.File(filename, "r")
 gamma_global = data['gamma'][:]
 gamma_r = gamma_global.real
@@ -83,7 +83,7 @@ R = 1.001
 B = 1
 d = np.pi
 q = 0.75
-ky_an = np.linspace(0.055,0.2,200)
+ky_an = np.linspace(0.05,0.2,200)
 omega = calc_asymptotic_growth(R, q, B, d, ky_an, kz)
 
 asymp_ax.plot(ky,gamma_r[:,index], label='numerical')

### 2019-07-22 paper/paper.tex
Lots of updates to text.

@@
 %\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-1}
-\documentclass[aps,prl, reprint,superscriptaddress]{revtex4-1}
+\documentclass[aps,prl,reprint,superscriptaddress]{revtex4-1}
 \newcommand{\SSC}{S/S_{c}}
+\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
+\newcommand{\apjs}{The Astrophysical Journal Supplement Series}
 \usepackage{graphicx}
@@
 \begin{abstract}
-Write abstract last.
+The magnetorotational instability (MRI) is a powerful instability in which conducting fluids with weak magnetic fields and decreasing angular velocity are linearly unstable.
+The MRI is thought to be essential to the operation of astrophysical accretion disks, where the angular velocity graident is supplied by the gravitational field of the central object.
+However, the MRI may also be important in the internal shear layers of stars, where the MRI may operate in a poorly studied regime where the shear is near critical but the dissipative forces are negligable.
+Here, we consider the linear stability of three-dimensional perturbations to a nearly inviscid and ideal magnetofluid and show that the fastest growing modes are three dimensional when the instability is near onset.
+These three dimensional modes are unstable at shear rates down to 0.10 $S_c$, the critical value in two dimensions, and they have growth rates exceeding the fastest two dimensional modes until the shear parameter is about $2.05 S_c$.
+These modes are significant, as they suggest the possibility of direct dynamo action from the linear instability itself.
+Furthermore, even when two dimensional modes have the fastest growth rates, significant numbers of three dimensional modes are unstable.
+In such a system, the non-normal character of the MRI cannot be ignored anytime the shear is near critical.
+Because evidence from weakly non-linear analysis suggests that the MRI saturates by pushing the shear rate to its critical value, non-normality is never unimportant for systems, like stars, that can rearrange their angular velocity profiles.
 \end{abstract}
@@
 \label{sec:intro}
+The magnetorotational instability (MRI) is an extremely important instability in astrophysical fluid dynamics.
+By changing the stabiliity criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient, a weak magnetic field allows the Keplerian velocity profile due to a central point mass to drive turbulence \citep[e.g.][]{1998RvMP...70....1B}.
+This discovery provided a pathway to explain the ubiquitous accretion onto compact objects at rates compatible with observations and may play a significant role in the formation of planets \citep[e.g.][]{2007Natur.448.1022J}.
+In these objects, the gravitational field dominates the local plasma dynamics, and thus the MRI cannot significantly affect the background shear.
+It must saturate by other means in these situations \citep{2018MNRAS.474.3451X}.
+However, stars have differential rotation profiles driven by much weaker stresses.
+In these regions, the MRI could be active and because the driving is much weaker, it can sature by pushing the background shear close to critical \citep{2017ApJ...841....1C,2017ApJ...841....2C}.
 
-\begin{itemize}
-\item 3D $\to$ dynamo
-\item stars, NSSL
-\item experiments
-\end{itemize}
+Here, we investigate the linear stability of three dimensional perturbations near the critical shear rate $S_c$ for a nearly invisicd, ideal MHD flow.
+We find that the first modes to become destabilized are three-dimensional and thus act as a dynamo even in the absence of secondary instability.
+These results strongly suggest that the non-normality of the linear operator leading to the MRI are always important, even near onset.
+% \begin{itemize}
+% \item 3D $\to$ dynamo
+% \item stars, NSSL
+% \item experiments
+% \item Squire's theorem violation? Check MHD Squire, check S instead of Re...
+% \item Do eigenvalue solution for B growth? i.e.\ check if it is a dynamo?
+% \end{itemize}
 
@@
 \label{sec:method}
-
-We solve the linearized magnetohydrodynamic equations in a frame rotating with angular frequency $\Omega$. The Navier-Stokes equations is in a fairly standard form,
+We solve the linearized magnetohydrodynamic equations in a frame rotating with angular frequency $\Omega$. The Navier-Stokes equation is in a fairly standard form,
 \begin{equation}
@@
 The walls at $x = \pm L_x/2$ are perfectly conducting and stress-free.
-The MRI is a weak-field instability; in the invisicd, ideal case the critical shear rate for instability in two dimensions is given by
+The MRI is a weak-field instability; in the inviscid, ideal case the critical shear rate for instability in two dimensions is given by
 \begin{equation}
@@
 We solve equations~(\ref{eq:ns})--~(\ref{eq:divu}) by assuming harmonic perturbations in $y$ and $z$, $f = \hat{f}(x) e^{i(k_y y + k_z z) + \gamma t}$.
-This reduces to an eigenvalue problem in $x$, which we solve using the \emph{Dedalus} pseudospectral framework.
-The advantage of this particular form is that it can be expressed as a set of ten first-order equations with Dirichlet boundary conditions for either no-slip or stress-free boundary conditions.
+This reduces to an eigenvalue problem in $x$, which we solve using the \emph{Dedalus} framework.
+The advantage of this particular form is that it can be expressed as a set of ten first-order equations with Dirichlet boundary conditions for the stress-free, perfect conductor boundary conditions we use.
 
-Our interest is in nearly ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions; we thus set $\eta=\nu=10^{-5}$, though we have confirmed our main results are insensitive to this particular value.
+Our interest is in nearly ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions.
+However, in order to avoid critical layers in the stable solutions, we set $\eta=\nu=10^{-5}$.
+We have confirmed our main results are insensitive to this particular value.
 
@@
 Our most surprising result is that at very small values of supercriticality, the fastest growing mode is not axisymmetric. 
-Figure~\ref{fig:growth_rate} shows the growth rates for three values of $\SSC$.
-At $\SSC = 1.01$, the maximum growth rate is $(k_y, k_z) \simeq (0.263, 0.447)$.
+Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
+At $\SSC = 1.02$, the maximum growth rate is $(k_y, k_z) \simeq (0.263, 0.447)$.
+Figure~\ref{fig:growth_rate}b compares the numerically growth rates for $\SSC=1.002$ between $0 <= k_y < 0.2$ to an asymptotic calculation.
+By expanding equations~(\ref{eq:ns}-~\ref{eq:divu}) in terms of $S = \SSC + \epsilon R$, we compute the leading order correction to the growth rate from three-dimensional effects as
+\begin{equation}
+  \label{eq:asymp}
+  \left(\pi ^2 B^2+d^2 f^2 \right) \omega^{2} - B^4 k^2 \left(d^2 k^2-\pi ^2 R\right)   \ + \ \frac{\pi ^2 B^4 \ell ^2}{12}  \left(\left(6+\pi ^2\right) q+\pi ^2-6\right) \ = \ 0.
+\end{equation}
 
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{run_11_output_growthrates.png}
-  \includegraphics[width=\columnwidth]{run_12_output_growthrates.png}
-  \includegraphics[width=\columnwidth]{run_13_output_growthrates.png}
-  \caption{Growth rate as a function of $k_z$ and $k_y$. The maximum growth rate is far from $k_y = 0$, indicating the MRI is most unstable in a fully three-dimensional configuration.}
+%
+\onecolumngrid
+
+\begin{figure*}[b!]
+  \includegraphics[width=\textwidth]{fig_1.pdf}
+  \caption{Stuff}
   \label{fig:growth_rate}
-\end{figure}
+\end{figure*}
+\twocolumngrid
+%
+Numerical simulations have consistently shown axisymmetric modes dominating early evolution of the MRI before parasitic modes cause a breakdown into MHD turbulence \citep[e.g.][]{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
+In order to reconcile our results with these simulations, in figure~\ref{fig:alpha} we plot the angle $\alpha = \arctan k_y/k_z$ of the fastest growing mode as function of $\SSC$.
+Above $\SSC \gtrsim 2.05$, $\alpha$ is zero, indicating that axisymmetric modes have the fastest growth rates.
+For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC \simeq 4.84$ (similar values are found in most works); thus we recover these results.
 
-Numerical simulations have consistently axisymmetric modes dominating early evolution of the MRI (cite Stone 95, etc) before parasitic modes cause a breakdown into MHD turbulence.
-However, our results recover this limit, as $\SSC$ becomes large.
-Figure~\ref{fig:alpha} plots $\alpha = \arctan k_y/k_z$ as  function of $\SSC$.
-Above $\SSC \gtrsim 2$, $\alpha$ is zero, indicating that axisymmetric modes have the fastest growth rates.
-
-\begin{figure}[h]
+\begin{figure}[h!]
   \includegraphics[width=\columnwidth]{alpha_vs_ssc_grid.pdf}
@@
 \end{figure}
+
 \section{Non-Normality}
@@
 % Create the reference section using BibTeX:
-\bibliography{basename of .bib file}
+\bibliography{mri}
 

### 2019-07-22 paper/paper.tex
Fix long equation.
Right asymptotic equation this time...
More words.

@@
 
-\section{Introduction}
-\label{sec:intro}
 The magnetorotational instability (MRI) is an extremely important instability in astrophysical fluid dynamics.
@@
 
-\section{Method}
-\label{sec:method}
 We solve the linearized magnetohydrodynamic equations in a frame rotating with angular frequency $\Omega$. The Navier-Stokes equation is in a fairly standard form,
@@
 where $\omega = \mathbf{\nabla} \times \mathbf{v}$ is the vorticity, $f = 2 \Omega$ is the Coriolis parameter, and $S$ is the background shear rate.
+It is common in the accretion disk literature to relate $S$ and $f$ via the $q = S/f$ parameter; for all runs here, $q = 0.75$.
 $Df/Dt = \partial f/\partial t + S x \partial f/\partial y$ is the advective derivative including the background linear shear velocity.
@@
 
-We solve equations~(\ref{eq:ns})--~(\ref{eq:divu}) by assuming harmonic perturbations in $y$ and $z$, $f = \hat{f}(x) e^{i(k_y y + k_z z) + \gamma t}$.
+We solve equations~(\ref{eq:ns})--~(\ref{eq:divu}) by assuming harmonic perturbations in $y$ and $z$, $f = \hat{f}(x) e^{i(k_y y + k_z z) + \sigma t}$, where $\sigma = \gamma + i\omega$ with $\gamma$ and $\omega$ real.
 This reduces to an eigenvalue problem in $x$, which we solve using the \emph{Dedalus} framework.
 The advantage of this particular form is that it can be expressed as a set of ten first-order equations with Dirichlet boundary conditions for the stress-free, perfect conductor boundary conditions we use.
-
 Our interest is in nearly ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions.
@@
 
-\section{Results}
-\label{sec:results}
-Our most surprising result is that at very small values of supercriticality, the fastest growing mode is not axisymmetric. 
+Our most surprising result is that at small values of supercriticality, the fastest growing mode is not axisymmetric. 
 Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
-At $\SSC = 1.02$, the maximum growth rate is $(k_y, k_z) \simeq (0.263, 0.447)$.
-Figure~\ref{fig:growth_rate}b compares the numerically growth rates for $\SSC=1.002$ between $0 <= k_y < 0.2$ to an asymptotic calculation.
+At $\SSC = 1.02$, the maximum growth rate is found at $(k_y, k_z) \simeq (0.263, 0.447)$.
 By expanding equations~(\ref{eq:ns}-~\ref{eq:divu}) in terms of $S = \SSC + \epsilon R$, we compute the leading order correction to the growth rate from three-dimensional effects as
+\begin{widetext}
 \begin{equation}
   \label{eq:asymp}
-  \left(\pi ^2 B^2+d^2 f^2 \right) \omega^{2} - B^4 k^2 \left(d^2 k^2-\pi ^2 R\right)   \ + \ \frac{\pi ^2 B^4 \ell ^2}{12}  \left(\left(6+\pi ^2\right) q+\pi ^2-6\right) \ = \ 0.
+\omega^{2} \ = \ \frac{q\,B^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
+   ^2-6}{12 } \, k_{y}^{2}  \right] \ + \ ...
 \end{equation}
-
+\end{widetext}
+Figure~\ref{fig:growth_rate}b compares the numerically growth rates for $\SSC=1.002$ between $0 <= k_y < 0.2$ to this asymptotic approximation, showing that the numerical calculations accurately reproduce the analytic ones in the region where the latter is valid.
+All unstable modes are purely growing (figure~\ref{fig:growth_rate}c); the stable modes are rotationally modified Alfven waves found in left- and right- going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
 %
@@
 %
-Numerical simulations have consistently shown axisymmetric modes dominating early evolution of the MRI before parasitic modes cause a breakdown into MHD turbulence \citep[e.g.][]{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
+Numerical simulations have consistently shown axisymmetric modes dominating early evolution of the MRI before parasitic modes cause a breakdown into MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
 In order to reconcile our results with these simulations, in figure~\ref{fig:alpha} we plot the angle $\alpha = \arctan k_y/k_z$ of the fastest growing mode as function of $\SSC$.
@@
 For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC \simeq 4.84$ (similar values are found in most works); thus we recover these results.
-
+%
 \begin{figure}[h!]
@@
 \end{figure}
+%
+However, even for $\SSC> 2.05$ there are significant swaths of unstable three-dimensional modes with growth rates comparable to the two-dimensional mode (figure~\ref{fig:growth_rate}a).
 
-\section{Non-Normality}
-\label{sec:non-normality}
+The linear operator for the MRI, like other shearing systems, is non-normal.
+As such, significant non-modal transient amplification occurs even for stable systems.
 
-\section{Speculations}
-\label{sec:speculations}
+% \begin{itemize}
+% \item non-normal and $k_y$ even when $\alpha = 0$ implies important role for linear dynamics at early times. 
+% \item Following Squire + Bhattacharjee, this suggests that important stellar layers may in fact be driven by linear dynamics
+% \item why is MRI non-oscillatory when hydro shear is oscil or stable?
+% \end{itemize}
 
-\begin{itemize}
-\item non-normal and $k_y$ even when $\alpha = 0$ implies important role for linear dynamics at early times. 
-\item Following Squire + Bhattacharjee, this suggests that important stellar layers may in fact be driven by linear dynamics
-\item why is MRI non-oscillatory when hydro shear is oscil or stable?
-\end{itemize}
-The linear operator for the MRI, like other shearing systems, is non-normal. 
 %If in two-column mode, this environment will change to single-column

### 2019-07-23 paper/paper.tex
Discussion of results

@@
 The magnetic field is expressed in Alfven units, $\mathbf{B} = \mathbf{B_*}/\sqrt{\mu_0 \rho}$.
-We consider the behavior of the MRI in a vertically ($z$) and horizontally ($y$) periodic channel of width $L_x = \pi$.
-The walls at $x = \pm L_x/2$ are perfectly conducting and stress-free.
+We consider the behavior of the MRI in a vertically ($z$) and horizontally ($y$) periodic channel of width $d = \pi$.
+The walls at $x = \pm d/2$ are perfectly conducting and stress-free.
 The MRI is a weak-field instability; in the inviscid, ideal case the critical shear rate for instability in two dimensions is given by
@@
 
-Our most surprising result is that at small values of supercriticality, the fastest growing mode is not axisymmetric. 
+Our main result is that at small values of supercriticality, the fastest growing mode is not axisymmetric. 
 Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
 At $\SSC = 1.02$, the maximum growth rate is found at $(k_y, k_z) \simeq (0.263, 0.447)$.
-By expanding equations~(\ref{eq:ns}-~\ref{eq:divu}) in terms of $S = \SSC + \epsilon R$, we compute the leading order correction to the growth rate from three-dimensional effects as
+In fact, at $\SSC = 0.640$, there are no unstable two-dimensional modes, but three-dimensional instability is present.
+
+While surprising, this result can be predicted analytically.
+By expanding equations~(\ref{eq:ns}-\ref{eq:divu}) in terms of $S = S_c + \epsilon R$, we compute the leading order correction to the growth rate from three-dimensional effects as
 \begin{widetext}
@@
   \label{eq:asymp}
-\omega^{2} \ = \ \frac{q\,B^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
+\sigma^{2} \ = -\frac{q\,B^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
    ^2-6}{12 } \, k_{y}^{2}  \right] \ + \ ...
@@
 \end{widetext}
-Figure~\ref{fig:growth_rate}b compares the numerically growth rates for $\SSC=1.002$ between $0 <= k_y < 0.2$ to this asymptotic approximation, showing that the numerical calculations accurately reproduce the analytic ones in the region where the latter is valid.
-All unstable modes are purely growing (figure~\ref{fig:growth_rate}c); the stable modes are rotationally modified Alfven waves found in left- and right- going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
+The first term in equation~(\ref{eq:asymp}) is the result from a two-dimensional calculation.
+The second term is negative definite and proportional to $k_y^2$: for sufficiently large $k_y$, it leads to instability even when the two-dimensional modes are stable.
+
+Figure~\ref{fig:growth_rate}b compares the numerically growth rates for $\SSC=1.002$ between $0 \le k_y \le 0.2$ to this asymptotic approximation, showing that the numerical calculations accurately reproduce the analytic ones in the region where the latter is valid.
+We provide the full spectrum for the most unstable parameter values for $\SSC = 1.02$ in figure~\ref{fig:growth_rate}c.
+The single unstable mode is purely growing (orange dot), and its complex conjugate is also present (blue dot at $\omega = 0$), consistent with equation~(\ref{eq:asymp}).
+The other stable modes (blue dots) are rotationally modified Alfven waves found in left- and right- going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
 %

### 2019-07-23 paper/paper.tex
Added bibliography and words words words

@@
-%\documentclass[aps,prl,preprint,groupedaddress]{revtex4-1}
+% \documentclass[aps,prl,preprint,groupedaddress]{revtex4-1}
 %\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-1}
@@
 It must saturate by other means in these situations \citep{2018MNRAS.474.3451X}.
-However, stars have differential rotation profiles driven by much weaker stresses.
-In these regions, the MRI could be active and because the driving is much weaker, it can sature by pushing the background shear close to critical \citep{2017ApJ...841....1C,2017ApJ...841....2C}.
+However, stars and liquid metal Taylor-Couette expereiments have differential rotation profiles driven by much weaker stresses.
+Where the MRI is active in these flows, it can sature by pushing the background shear close to critical \citep{2017ApJ...841....1C,2017ApJ...841....2C}, analogous to the saturation of convection by relaxing to a nearly constant entropy profile.
+Stellar interiors are at extremely high fluid and magnetic Reynolds numbers, but may potentially operate at or near the critical shear rate for the MRI\@.
+This critical shear is set by the background magnetic field strength for axisymmetric perturbations in a channel of a given size.
+Despite the extensive literature on the accretion disk case of strong shear and high Reynolds numbers and the liquid metal literature on weak shear and lower Reynolds numbers, even the linear dynamics of the MRI are not well studied in the weak shear, high Reynolds number regime.
 
-Here, we investigate the linear stability of three dimensional perturbations near the critical shear rate $S_c$ for a nearly invisicd, ideal MHD flow.
+Here, we investigate the linear stability of three dimensional perturbations near the two dimensional critical shear rate $S_c$ for a nearly invisicd, ideal MHD flow.
 We find that the first modes to become destabilized are three-dimensional and thus act as a dynamo even in the absence of secondary instability.
-These results strongly suggest that the non-normality of the linear operator leading to the MRI are always important, even near onset.
+These results strongly suggest that the non-normality of the linear operator driving the MRI are always important, even when axisymmetric modes dominate.
 % \begin{itemize}
@@
 
-We solve the linearized magnetohydrodynamic equations in a frame rotating with angular frequency $\Omega$. The Navier-Stokes equation is in a fairly standard form,
+We solve the linearized magnetohydrodynamic equations in rotating plane Couette geometry.
+This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $v_{y0} = Sx$. 
+We cast the Navier-Stokes equation in a fairly standard form,
 \begin{equation}
@@
 The advantage of this particular form is that it can be expressed as a set of ten first-order equations with Dirichlet boundary conditions for the stress-free, perfect conductor boundary conditions we use.
-Our interest is in nearly ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions.
+Our interest is in ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions.
 However, in order to avoid critical layers in the stable solutions, we set $\eta=\nu=10^{-5}$.
 We have confirmed our main results are insensitive to this particular value. 
+1For each $(k_y, k_z)$ pair, we solve an eigenvalue problem with $n_x = 128$ modes; our results are identical at double the resolution\footnote{All code used in this paper can be found at \protect\url{https://github.com/jsoishi/mri\_prefers\_3d}}.
 
@@
 At $\SSC = 1.02$, the maximum growth rate is found at $(k_y, k_z) \simeq (0.263, 0.447)$.
-In fact, at $\SSC = 0.640$, there are no unstable two-dimensional modes, but three-dimensional instability is present.
+In fact, the first panel of figure~\ref{fig:growth_rate}a shows $\SSC = 0.640$, where there are no unstable two-dimensional modes, but three-dimensional instability is present.
+The critical shear for three dimensional perturbations is at $S_{c,3D} \simeq 0.
+102 S_c$.
 
-While surprising, this result can be predicted analytically.
-By expanding equations~(\ref{eq:ns}-\ref{eq:divu}) in terms of $S = S_c + \epsilon R$, we compute the leading order correction to the growth rate from three-dimensional effects as
+While surprising, that three-dimensional perturbations grow faster than two-dimensional ones can be predicted analytically.
+By expanding equations~(\ref{eq:ns}-\ref{eq:divu}) in terms of $S = S_c + \epsilon R$, such that the shear is asymptotically close to the critical rate for two dimensional perturbations, we compute the leading order correction to the growth rate from three-dimensional effects as
 \begin{widetext}
@@
 
-\begin{figure*}[b!]
+\begin{figure*}[h!]
   \includegraphics[width=\textwidth]{fig_1.pdf}
@@
 %
-Numerical simulations have consistently shown axisymmetric modes dominating early evolution of the MRI before parasitic modes cause a breakdown into MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
-In order to reconcile our results with these simulations, in figure~\ref{fig:alpha} we plot the angle $\alpha = \arctan k_y/k_z$ of the fastest growing mode as function of $\SSC$.
+Numerical simulations have consistently shown axisymmetric modes dominating early evolution of the MRI before breaking down into MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
+These axisymmetric modes, called ``channel modes,'' are exact non-linear solutions of the governing equations with shearing-periodic boundary conditions in $x$, and thus must saturate via parasitic modes from physics beyond non-linear mode coupling \citep{1994ApJ...432..213G}.
+Though we use impenetrable, stress-free boundary conditions that do not allow the unbounded growth of channel modes, the basic principle that axisymmetric disturbences grow fastest is not clearly caused by boundary conditions.
+thus, we must reconcile our results with those earlier simulations.
+In figure~\ref{fig:alpha}, we plot the angle $\alpha = \arctan k_y/k_z$ of the fastest growing mode with the $z$ axis as function of $\SSC$.
 Above $\SSC \gtrsim 2.05$, $\alpha$ is zero, indicating that axisymmetric modes have the fastest growth rates.
-For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC \simeq 4.84$ (similar values are found in most works); thus we recover these results.
+For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC \simeq 4.84$ (similar values are found in most works).
+Thus, our theory predicts that the linear dynamics should be dominated by axisymmetric modes for the parameters studied in prior numerical simulations.
 %
@@
 %
+
 However, even for $\SSC> 2.05$ there are significant swaths of unstable three-dimensional modes with growth rates comparable to the two-dimensional mode (figure~\ref{fig:growth_rate}a).
+There are two significant implications for this.
+First, the linear operator for the MRI, like other shearing systems, is non-normal; this effect appears whenever non-axisymmetric perturbations are allowed \citep[see][for a discussion relevant to the MRI]{1992MNRAS.255P..25K}.
+Because of this, non-modal transient amplification occurs even for stable modes.
+These effects can be strong enough to cause non-linear interactions and turbulence even when the entire flow is linearly stable.
+While we are here concerned with systems linearly unstable to the MRI, the presence of three dimensional modes implies that linear, non-normal behavior could play a very strong role in the MRI near onset.
 
-The linear operator for the MRI, like other shearing systems, is non-normal.
-As such, significant non-modal transient amplification occurs even for stable systems.
+The second implication of the three dimensional modes is of considerably more importance.
+These modes produce dynamo action in the linear regime, as evidenced by the presence of non-trivial eigenmodes for all three components of the magnetic field.
+This is of considerable significance, as it leads to the possibility of a laminar MRI dynamo operating directly in regions of weak shear.
+MRI dynamos have been studied in a number of different contexts \citep{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one have focused on the turbulent dynamo.
+The lone exception, \citet{2016MNRAS.462..818B}, is a numerical study that discovered the exponential growth of mean magnetic fields during the linear growth phase of the MRI far from stability.
+Our work explains this result in terms of purely linear dynamics:
+non-axisymmetric MRI unstable modes drive the exponential growth of magnetic fields.
 

### 2019-07-23 paper/paper.tex
More words. Acknowledgements. Removed onecolumngrid for fig 1
* this fixes the weirdness after the long equation. Can be put back
  later for more accurate length estimation.

@@
 
-Here, we investigate the linear stability of three dimensional perturbations near the two dimensional critical shear rate $S_c$ for a nearly invisicd, ideal MHD flow.
+Here, we investigate the stability of three dimensional perturbations near the two dimensional critical shear rate $S_c$ for a nearly invisicd, ideal MHD flow.
 We find that the first modes to become destabilized are three-dimensional and thus act as a dynamo even in the absence of secondary instability.
@@
 %
-\onecolumngrid
+%\onecolumngrid
 
@@
   \includegraphics[width=\textwidth]{fig_1.pdf}
-  \caption{Stuff}
+  \caption{Growth rates for three-dimensional MRI modes.
+a) growth rates $\gamma$ over a grid of $k_y$ and $k_z$ for four values of supercriticality $\SSC$.
+The white contour highlights $\gamma = 0$; the red dots indicate the fastest growing mode.
+At $\SSC = 0.640$, there are no two dimensional modes.
+b) Growth rate vs $k_y$ for $k_z = ???$ at $\SSC = 1.002$ (highlighted by the blue line in a).
+The orange line gives the asymptotic result from equation~(\ref{eq:asymp}) and the blue line are the numerical results.
+The asymptotic form is valid as $\SSC \to 1$ and $k_y \ll 1$.
+c) The full discrete spectrum of the MRI for $(k_y, k_z) \simeq (0.263, 0.447)$.
+The unstable mode is shown in orange; all other modes are stable ($\gamma \le 0$).}
   \label{fig:growth_rate}
 \end{figure*}
-\twocolumngrid
+%\twocolumngrid
 %
@@
   \includegraphics[width=\columnwidth]{alpha_vs_ssc_grid.pdf}
-  \caption{Angle $\alpha$ vs $\SSC$. Stability limit is at $\SSC \simeq 0.102$}
+  \caption{Angle $\alpha$ of mode with respect to $z$ axis vs $\SSC$.
+The inset shows the transition from axisymmetric modes to three dimensional modes as $\SSC$ decreases below $2.05$.
+The shaded region indicates stability to all perturbations; the stability limit for three dimensional modes is $\SSC \simeq 0.102$.}
   \label{fig:alpha}
@@
 % If you have acknowledgments, this puts in the proper section head.
-%\begin{acknowledgments}
-% put your acknowledgments here.
-%\end{acknowledgments}
+\begin{acknowledgments}
+We would like to thank Steve Tobias for helpful conversations on this work.
+JSO acknowledges funding from NASA LWS grant No.
+NNX16AC92G and Research Corporation Scialog Collaborative Award (TDA) ID\#24231.
+Computations were performed on the \emph{Leavitt} cluster at the Bates College High Performance Computing Center.
+Morgan Baxter acknowledges support from NASA LWS grant No. NNX16AC92G.
+\end{acknowledgments}
 

### 2019-08-16 python/calc_asymptotic_approx.py
updated fig 1 to include sigma^2

@@ -9,4 +9,4 @@ def calc_asymptotic_growth(R, q, B, d, ky, kz):
     omega2 = q*B**2/(q+1)*(kz**2*(d**2/np.pi**2 * kz**2 - Rprime) - ((6 + np.pi**2)*q + np.pi**2 - 6)*ky**2/12.)
     omega = np.sqrt(-omega2)
 
-    return omega
+    return omega, omega2

### 2019-08-16 python/plot_fig1.py

@@ -67,35 +67,43 @@ cbar.set_label(r'$\gamma$')
 xloc = grid[0].get_position().x0
 w_lower = 2*w + pad
 asymp_ax = fig.add_axes([0.1,0.1,w_lower,0.35])
+asymp_sub = asymp_ax.inset_axes([0.6,0.25,0.3,0.3*w_lower/0.35])
 
 filename = 'data/run_41_output.h5'
 data = h5py.File(filename, "r")
 gamma_global = data['gamma'][:]
 gamma_r = gamma_global.real
-
+gamma_i = gamma_global.imag
 ky_global    = data['ky']
 kz_global    = data['kz']
 
 ky = ky_global[:]
 kz = kz_global[index]
-grid[1].axvline(kz,alpha=0.4)
+grid[1].axvline(kz, ymin=0, ymax=0.2, alpha=0.7)
 R = 1.001
 B = 1
 d = np.pi
 q = 0.75
-ky_an = np.linspace(0.05,0.2,200)
-omega = calc_asymptotic_growth(R, q, B, d, ky_an, kz)
+ky_an = np.linspace(0.0,0.2,400)
+omega, omega2 = calc_asymptotic_growth(R, q, B, d, ky_an, kz)
 
 asymp_ax.plot(ky,gamma_r[:,index], label='numerical')
 asymp_ax.plot(ky_an, omega, label='asymptotic')
+asymp_sub.plot(ky,-(gamma_r[:,index] + 1j*gamma_i[:,index])**2, label='numerical')
+asymp_sub.plot(ky_an, omega2, label='asymptotic')
+
 #asymp_ax.legend(loc='upper right')
 asymp_ax.text(0.125, 0.048, 'numerical', fontsize=18)
 asymp_ax.text(0.035, 0.03, 'asymptotic', fontsize=18)
 asymp_ax.set_ylim(0, 0.06)
 asymp_ax.set_xlim(0,0.2)
 asymp_ax.set_xlabel(r'$k_y$')
-asymp_ax.set_ylabel(r'$\gamma$')
-asymp_ax.xaxis.set_major_locator(plt.MultipleLocator(0.1))
+asymp_ax.set_ylabel(r'$\gamma = Re(\sigma)$')
+asymp_ax.xaxis.set_major_locator(plt.MultipleLocator(0.05))
+asymp_sub.set_ylim(-0.005, 0.005)
+asymp_sub.set_xlim(0.03,0.09)
+asymp_sub.set_xlabel(r'$k_y$')
[…]

### 2019-08-16 python/plot_3D_eigenvalues.py
Added script to plot mean field alpha

@@ -44,8 +44,9 @@ kz = datafile['gamma'].attrs['max kz']
 Ly = 2*np.pi/ky
 Lz = 2*np.pi/kz
 
-y = np.linspace(0, Ly, 768, endpoint=False)
-z = np.linspace(0, Lz, 768, endpoint=False)
+n_yz = 128
+y = np.linspace(0, Ly, n_yz, endpoint=False)
+z = np.linspace(0, Lz, n_yz, endpoint=False)
 
 zz,yy,xx = np.meshgrid(z,y,x,indexing='ij')
 fig = plt.figure(figsize=(6,6))
@@ -59,6 +60,28 @@ w = (eigvec[y_max, z_max,3,:]*yz_dep).real
 Bx =(eigvec[y_max, z_max,6,:]*yz_dep).real
 By =(eigvec[y_max, z_max,7,:]*yz_dep).real
 Bz =(eigvec[y_max, z_max,8,:]*yz_dep).real
+
+save_vtk = True
+save_h5 = True
+if save_vtk:
+    vtkfile = filename.stem + '.vtk'
+    from pyevtk.hl import gridToVTK 
+    pointData = {'p':p.copy(), 'u':u.copy(), 'v':v.copy(), 'w':w.copy(),
+                 'Bx':Bx.copy(), 'By':By.copy(), 'Bz':Bz.copy()}
+    print("p flags", pointData['p'].flags)
+
+    gridToVTK(vtkfile, xx, yy, zz, pointData=pointData)
+if save_h5:
+    h5file = filename.stem + '.h5'
+    datafile = h5py.File(h5file, 'w')
+    datafile.create_dataset('u', data=u)
+    datafile.create_dataset('v', data=v)
+    datafile.create_dataset('w', data=w)
+    datafile.create_dataset('Bx', data=Bx)
+    datafile.create_dataset('By', data=By)
+    datafile.create_dataset('Bz', data=Bz)
+    datafile.create_dataset('x_grid', data=xx)
+    datafile.close()
 
 
 grid = ImageGrid(fig, 111, nrows_ncols=(4,1),axes_pad=0.5, cbar_mode='each', cbar_location='top',cbar_size="10%", cbar_pad=0)
@@ -84,7 +107,7 @@ plot_file_name = Path(filename.stem + '_3D_v_eigenvector.png')
 fig.savefig(outbase/plot_file_name, dpi=300)
 
 fig.clear()
[…]

### 2019-08-16 python/plot_mean_field_alpha.py

@@ -1,0 +1,29 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import h5py
+
+plt.style.use('prl')
+
+dfname = "run_11_output.h5"
+df = h5py.File(dfname,"r")
+x = df['x_grid'][:,:,:]
+u = df['u'][:,:,:]
+v = df['v'][:,:,:]
+w = df['w'][:,:,:]
+Bx = df['Bx'][:,:,:]
+By = df['By'][:,:,:]
+Bz = df['Bz'][:,:,:]
+cheby_factor = np.sqrt(np.pi**2 - x**2)
+
+u_mean = np.sqrt(np.mean((u**2 + v**2 + w**2)*cheby_factor)/np.mean(cheby_factor))
+B_mean = np.sqrt(np.mean((Bx**2 + By**2 + Bz**2)*cheby_factor)/np.mean(cheby_factor))
+norm = u_mean*B_mean
+alpha = (u*By - v*Bx)
+alpha_yz = alpha.mean(axis=0).mean(axis=0)/norm
+
+plt.plot(x[0,0,:], alpha_yz)
+plt.xlabel(r"$x$")
+plt.ylabel(r"$\alpha$")
+plt.tight_layout()
+plt.savefig('plots/mean_field_alpha_{}.png'.format(dfname))
+

### 2019-08-16 paper/paper.tex
paper edits

@@
 
-\begin{abstract}
-The magnetorotational instability (MRI) is a powerful instability in which conducting fluids with weak magnetic fields and decreasing angular velocity are linearly unstable.
-The MRI is thought to be essential to the operation of astrophysical accretion disks, where the angular velocity graident is supplied by the gravitational field of the central object.
-However, the MRI may also be important in the internal shear layers of stars, where the MRI may operate in a poorly studied regime where the shear is near critical but the dissipative forces are negligable.
+\begin{abstract}Ú
+The magnetorotational instability (MRI) is a powerful instability where a weak magnetic can destabilize an electrically conducting fluid with a decreasing angular velocity shear.
+The MRI is thought to be essential to the operation of astrophysical accretion disks, where the central object supplies a Keplerian angular velocity gradient.
+The MRI may also be important in the internal shear layers of stars.
+In stellar interiors, the shear can take a wide range to shear profiles, including near-critical, but with negligible dissipative the MRI may operate in a poorly studied regime where the shear is near critical but the dissipative forces are negligible.
 Here, we consider the linear stability of three-dimensional perturbations to a nearly inviscid and ideal magnetofluid and show that the fastest growing modes are three dimensional when the instability is near onset.
@@
 We solve the linearized magnetohydrodynamic equations in rotating plane Couette geometry.
-This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $v_{y0} = Sx$. 
+This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y} = Sx$. 
 We cast the Navier-Stokes equation in a fairly standard form,
 \begin{equation}
-  \label{eq:ns}
-  \frac{D \mathbf{v}}{Dt} - f \hat{z} \times \mathbf{v} + S v_x \hat{z} + \mathbf{\nabla}{p} + \nu \mathbf{\nabla} \times \mathbf{\omega} = 0,
+  \label{eq:mhd}
+  \frac{D \mathbf{v}}{Dt} + f \hat{z} \times \mathbf{v} + S v_x \hat{z} + \mathbf{\nabla}{p} + \nu \mathbf{\nabla} \times \mathbf{\omega} = B_{0} \partial_{z} \mathbf{b},
 \end{equation}
-where $\omega = \mathbf{\nabla} \times \mathbf{v}$ is the vorticity, $f = 2 \Omega$ is the Coriolis parameter, and $S$ is the background shear rate.
-It is common in the accretion disk literature to relate $S$ and $f$ via the $q = S/f$ parameter; for all runs here, $q = 0.75$.
-$Df/Dt = \partial f/\partial t + S x \partial f/\partial y$ is the advective derivative including the background linear shear velocity.
+where
+\begin{equation}
+\omega = \mathbf{\nabla} \times \mathbf{v}, \quad \text{and} \quad \frac{D}{Dt} \ = \ \partial_{t} + S x \partial_{y}
+\end{equation}
+The main input parameters are the Coriolis parameter, $f = 2 \Omega$; the background shear rate, $S = dV_{y}/dx < 0$;  the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0} \rho_{0} = 1$); and the horizontal domain width $-d/2 \le x \le d/2$.
+The accretion disk literature usually relates $S$ to $f$ via the $q = |S|/f$ parameter; for all calculations here, $q = 3/4$ (Keplerian).
 We write the induction equation in terms of the $x$ component of the magnetic field,
@@
   \label{eq:Bx}
-  \frac{D B_x}{Dt} - B_0 \partial_z v_x + \eta (\partial_y J_z - \partial_z J_y) = 0
+  \frac{D b_x}{Dt} - B_0 \partial_z v_x + \eta (\partial_y j_z - \partial_z j_y) = 0
 \end{equation}
-and the $x$ component of the current density,
+and the $x$ component of the current density ($j_{x} = \partial_{y}b_{z} - \partial_{z} b_{y}$),
 \begin{equation}
   \label{eq:Jx}
-  \frac{D J_x}{Dt} - B_0 \partial_z \omega_x + S \partial_z B_x - \eta \nabla^2 J_x = 0
+  \frac{D j_x}{Dt} - B_0 \partial_z \omega_x + S \partial_z B_x - \eta \nabla^2 j_x = 0
 \end{equation}
@@
   \label{eq:divu}
-  \mathbf{\nabla} \cdot \mathbf{v} = \mathbf{\nabla} \cdot \mathbf{B} = 0.
+  \mathbf{\nabla} \cdot \mathbf{v} = \mathbf{\nabla} \cdot \mathbf{b} = 0.
 \end{equation}
-The magnetic field is expressed in Alfven units, $\mathbf{B} = \mathbf{B_*}/\sqrt{\mu_0 \rho}$.
 We consider the behavior of the MRI in a vertically ($z$) and horizontally ($y$) periodic channel of width $d = \pi$.
-The walls at $x = \pm d/2$ are perfectly conducting and stress-free.
-The MRI is a weak-field instability; in the inviscid, ideal case the critical shear rate for instability in two dimensions is given by
+The boundary conditions are impenetrable stress-free and perfectly conducting; $v_{x} = \omega_{y} = \omega_{z} = b_{x} = \partial_{x}j_{x} = 0$.
+The MRI is a weak-field instability; in the inviscid, ideal case the critical shear rate for instability in \textit{two dimensions} is
 \begin{equation}
   \label{eq:Sc}
-  S_c = \frac{-\pi^2 B^2}{f d^2}.
+  S_c = -\frac{\pi^2 B_{0}^2}{f d^2}.
 \end{equation}
@@
 
-We solve equations~(\ref{eq:ns})--~(\ref{eq:divu}) by assuming harmonic perturbations in $y$ and $z$, $f = \hat{f}(x) e^{i(k_y y + k_z z) + \sigma t}$, where $\sigma = \gamma + i\omega$ with $\gamma$ and $\omega$ real.
-This reduces to an eigenvalue problem in $x$, which we solve using the \emph{Dedalus} framework.
-The advantage of this particular form is that it can be expressed as a set of ten first-order equations with Dirichlet boundary conditions for the stress-free, perfect conductor boundary conditions we use.
+We assume harmonic perturbations in $y$ and $z$, (e.g.
+pressure) $p = \hat{p}(x) e^{i(k_y y + k_z z) + \sigma t}$ .
+We use a complex-valued growth rate $\sigma = \gamma + i\omega$ with $\gamma$ and $\omega$ real.
+The system reduces to a $\sigma$-eigenvalue problem of ten first-order ODEs in $x$ with Dirichlet boundary conditions.
+We pose and solve the system in the same formulation as equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework.
+
 Our interest is in ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions.
-However, in order to avoid critical layers in the stable solutions, we set $\eta=\nu=10^{-5}$.
-We have confirmed our main results are insensitive to this particular value. 
-1For each $(k_y, k_z)$ pair, we solve an eigenvalue problem with $n_x = 128$ modes; our results are identical at double the resolution\footnote{All code used in this paper can be found at \protect\url{https://github.com/jsoishi/mri\_prefers\_3d}}.
+However, we set $\eta=\nu=10^{-5}$ to avoid critical layers in for the \textit{stable} solution branch.
+We confirmed our results are insensitive to the presence of small diffusion. 
+For each $(k_y, k_z)$ pair, we solve using $n_x = 128$ modes; all our results are identical at double the resolution\footnote{All code used in this paper can be found at \protect\url{https://github.com/jsoishi/mri\_prefers\_3d}}.
 
-Our main result is that at small values of supercriticality, the fastest growing mode is not axisymmetric. 
+Our two main results are: (I) When the first two-dimensional mode becomes unstable, there already exist three-dimensional modes with positive growth rate.
+(II) When the criticality becomes sufficiently large, the fastest-growing mode becomes purely two-dimensional.
+These results are pertinent in two ways.
+First, the MRI contains a ``Goldilocks regime" with possible direct dynamo action, and this regime likely most applies to stellar interiors.
+Second, our results accord with well-established results for accretion disk that expect primary two-dimensional linear modes.
 Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
-At $\SSC = 1.02$, the maximum growth rate is found at $(k_y, k_z) \simeq (0.263, 0.447)$.
-In fact, the first panel of figure~\ref{fig:growth_rate}a shows $\SSC = 0.640$, where there are no unstable two-dimensional modes, but three-dimensional instability is present.
-The critical shear for three dimensional perturbations is at $S_{c,3D} \simeq 0.
+At $\SSC = 1.02$, the maximum growth rate happens at $(k_y, k_z) \simeq (0.263, 0.447)$.
+The first panel of figure~\ref{fig:growth_rate}a shows $\SSC = 0.640$, which contains only three-dimensional instability.
+The overall critical shear for three-dimensional modes is $S_{c,3D} \simeq 0.
 102 S_c$.
@@
 While surprising, that three-dimensional perturbations grow faster than two-dimensional ones can be predicted analytically.
-By expanding equations~(\ref{eq:ns}-\ref{eq:divu}) in terms of $S = S_c + \epsilon R$, such that the shear is asymptotically close to the critical rate for two dimensional perturbations, we compute the leading order correction to the growth rate from three-dimensional effects as
+We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) with the shear asymptotically close to the critical two-dimensional rate,  and compute the leading-order correction to the growth rate from three-dimensional effects as
 \begin{widetext}
@@
   \label{eq:asymp}
-\sigma^{2} \ = -\frac{q\,B^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
-   ^2-6}{12 } \, k_{y}^{2}  \right] \ + \ ...
+\sigma^{2} \ = -\frac{q\,B^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R^{2}\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
+   ^2-6}{12 } \, k_{y}^{2}  \right] \ + \  \ldots,
 \end{equation}
 \end{widetext}
+where $S = S_c - \epsilon \sqrt{R}$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$.
 The first term in equation~(\ref{eq:asymp}) is the result from a two-dimensional calculation.
-The second term is negative definite and proportional to $k_y^2$: for sufficiently large $k_y$, it leads to instability even when the two-dimensional modes are stable.
+The second term is negative definite and proportional to $k_{y}^2$, which leads to an ultraviolet catastrophe in the absence of higher-order effects.
 

### 2019-08-17 paper/paper.tex
first full pass through edits

@@
 \usepackage{graphicx}
+
 % You should use BibTeX and apsrev.bst for references
@@
 \begin{abstract}Ú
-The magnetorotational instability (MRI) is a powerful instability where a weak magnetic can destabilize an electrically conducting fluid with a decreasing angular velocity shear.
-The MRI is thought to be essential to the operation of astrophysical accretion disks, where the central object supplies a Keplerian angular velocity gradient.
-The MRI may also be important in the internal shear layers of stars.
-In stellar interiors, the shear can take a wide range to shear profiles, including near-critical, but with negligible dissipative the MRI may operate in a poorly studied regime where the shear is near critical but the dissipative forces are negligible.
-Here, we consider the linear stability of three-dimensional perturbations to a nearly inviscid and ideal magnetofluid and show that the fastest growing modes are three dimensional when the instability is near onset.
-These three dimensional modes are unstable at shear rates down to 0.10 $S_c$, the critical value in two dimensions, and they have growth rates exceeding the fastest two dimensional modes until the shear parameter is about $2.05 S_c$.
-These modes are significant, as they suggest the possibility of direct dynamo action from the linear instability itself.
-Furthermore, even when two dimensional modes have the fastest growth rates, significant numbers of three dimensional modes are unstable.
-In such a system, the non-normal character of the MRI cannot be ignored anytime the shear is near critical.
-Because evidence from weakly non-linear analysis suggests that the MRI saturates by pushing the shear rate to its critical value, non-normality is never unimportant for systems, like stars, that can rearrange their angular velocity profiles.
+The magnetorotational instability (MRI) occurs when a weak magnetic field destabilizes a rotating electrically conducting fluid with an inwardly increasing angular velocity shear.
+The MRI is considered essential to the operation of astrophysical accretion disks, with the central object supplying a Keplerian rotation profile.
+The MRI should also exist for the internal shear layers of stars.
+Stellar-interior shears can take a wide range of profiles, including near-critical, but with negligible dissipation. 
+Here, we show that the fastest growing unstable modes of an ideal magnetofluid are three dimensional provided the shear rate, $S$, is close to the two-dimensional onset value, $S_c$.
+Three-dimensional modes are unstable down to $S \approx 0.10 S_c$, and dominate the two-dimensional modes until $S \approx 2.05 S_c$.
+Significant numbers of fast three dimensional modes remain well past $2.05 S_{c}$. 
+These finding are significant in three ways. 
+First, evidence from weakly nonlinear theory suggests that the MRI saturates by pushing the shear rate to its critical value. 
+This can happen for systems, like stars, that can rearrange their angular velocity profiles.
+Second, the non-normal character and large transient growth of the MRI should be important anytime strong three-dimensionality exists.
+Finally, three-dimensional growth suggest the possibility of direct dynamo action driven from the linear instability itself.
 \end{abstract}
@@
 
-The magnetorotational instability (MRI) is an extremely important instability in astrophysical fluid dynamics.
+The magnetorotational instability (MRI) is an extremely important in astrophysical fluid dynamics.
 By changing the stabiliity criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient, a weak magnetic field allows the Keplerian velocity profile due to a central point mass to drive turbulence \citep[e.g.][]{1998RvMP...70....1B}.
-This discovery provided a pathway to explain the ubiquitous accretion onto compact objects at rates compatible with observations and may play a significant role in the formation of planets \citep[e.g.][]{2007Natur.448.1022J}.
-In these objects, the gravitational field dominates the local plasma dynamics, and thus the MRI cannot significantly affect the background shear.
-It must saturate by other means in these situations \citep{2018MNRAS.474.3451X}.
-However, stars and liquid metal Taylor-Couette expereiments have differential rotation profiles driven by much weaker stresses.
-Where the MRI is active in these flows, it can sature by pushing the background shear close to critical \citep{2017ApJ...841....1C,2017ApJ...841....2C}, analogous to the saturation of convection by relaxing to a nearly constant entropy profile.
-Stellar interiors are at extremely high fluid and magnetic Reynolds numbers, but may potentially operate at or near the critical shear rate for the MRI\@.
+This discovery explained the ubiquitous accretion onto compact objects at rates compatible with observations, and may also influence the formation of planets \citep[e.g.][]{2007Natur.448.1022J}.
+In disks, the gravitational field dominates the local plasma dynamics, and thus the MRI cannot significantly affect the background shear; it must saturate by other means \citep{2018MNRAS.474.3451X}.
+However, stars and liquid metal Taylor-Couette experiments have differential rotation profiles driven by much weaker stresses.
+Where the MRI is active in these flows, it can saturate by pushing the background shear close to critical \citep{2017ApJ...841....1C,2017ApJ...841....2C}, analogous to convection mixing entropy.
+Stellar interiors have extremely high fluid and magnetic Reynolds numbers, but may potentially operate at or near the critical shear rate for the MRI.
 This critical shear is set by the background magnetic field strength for axisymmetric perturbations in a channel of a given size.
@@
 Here, we investigate the stability of three dimensional perturbations near the two dimensional critical shear rate $S_c$ for a nearly invisicd, ideal MHD flow.
-We find that the first modes to become destabilized are three-dimensional and thus act as a dynamo even in the absence of secondary instability.
+We find that the first destabilized modes are three-dimensional, and thus can act as a dynamo even in the absence of secondary instability.
 These results strongly suggest that the non-normality of the linear operator driving the MRI are always important, even when axisymmetric modes dominate.
@@
 
-We solve the linearized magnetohydrodynamic equations in rotating plane Couette geometry.
+We solve the linearized magnetohydrodynamic equations in rotating plane-Couette geometry.
 This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y} = Sx$. 
@@
 The main input parameters are the Coriolis parameter, $f = 2 \Omega$; the background shear rate, $S = dV_{y}/dx < 0$;  the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0} \rho_{0} = 1$); and the horizontal domain width $-d/2 \le x \le d/2$.
-The accretion disk literature usually relates $S$ to $f$ via the $q = |S|/f$ parameter; for all calculations here, $q = 3/4$ (Keplerian).
+Accretion disk modelling usually considers the Rossby number $q = -S/f < 1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
+For all calculations here, we use $q = 3/4$ (Keplerian).
 We write the induction equation in terms of the $x$ component of the magnetic field,
@@
 \end{equation}
-Here, we use $\SSC$ as our control parameter.
+We use $\SSC$ as our instability control parameter.
 
-We assume harmonic perturbations in $y$ and $z$, (e.g.
-pressure) $p = \hat{p}(x) e^{i(k_y y + k_z z) + \sigma t}$ .
-We use a complex-valued growth rate $\sigma = \gamma + i\omega$ with $\gamma$ and $\omega$ real.
-The system reduces to a $\sigma$-eigenvalue problem of ten first-order ODEs in $x$ with Dirichlet boundary conditions.
+We assume harmonic perturbations in $y$ and $z$, (e.g. pressure) $p = \hat{p}(x) e^{i(k_y y + k_z z) + \sigma t}$ . 
+We use a complex-valued growth rate $\sigma = \gamma + i\omega$ with $\gamma, \omega$ both real. 
+The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $x$ with Dirichlet boundary conditions.
 We pose and solve the system in the same formulation as equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework.
@@
 These results are pertinent in two ways.
-First, the MRI contains a ``Goldilocks regime" with possible direct dynamo action, and this regime likely most applies to stellar interiors.
-Second, our results accord with well-established results for accretion disk that expect primary two-dimensional linear modes.
+First, the MRI contains a substantial ``Goldilocks regime" with possible direct dynamo action, and this regime likely most applies to stellar interiors.
+Second, our results accord with well-established results for accretion disks that expect primary two-dimensional linear modes.
 Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
-At $\SSC = 1.02$, the maximum growth rate happens at $(k_y, k_z) \simeq (0.263, 0.447)$.
+At $\SSC = 1.02$, the maximum growth rate ($\sigma = ??$) happens at $(k_y, k_z) \simeq (0.263, 0.447)$.
 The first panel of figure~\ref{fig:growth_rate}a shows $\SSC = 0.640$, which contains only three-dimensional instability.
-The overall critical shear for three-dimensional modes is $S_{c,3D} \simeq 0.
-102 S_c$.
+The overall critical shear for three-dimensional modes is $S_{c,3D} \simeq 0.102 S_c$.
 
-While surprising, that three-dimensional perturbations grow faster than two-dimensional ones can be predicted analytically.
-We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) with the shear asymptotically close to the critical two-dimensional rate,  and compute the leading-order correction to the growth rate from three-dimensional effects as
-\begin{widetext}
+While surprising, preferred three-dimensional modes can be predicted analytically.
+We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) with the shear asymptotically close to the critical two-dimensional rate, and compute the leading-order correction to the growth rate from three-dimensional effects assuming where $S = S_c - \epsilon \sqrt{R}$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$,
+%\begin{widetext}
 \begin{equation}
   \label{eq:asymp}
-\sigma^{2} \ = -\frac{q\,B^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R^{2}\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
-   ^2-6}{12 } \, k_{y}^{2}  \right] \ + \  \ldots,
+\sigma^{2} =  -\frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left( \frac{d^2}{\pi^{2}} k_z^2 - R^{2}\right)  -  \Upsilon \, k_{y}^{2}  \right] +  \ldots,
 \end{equation}
-\end{widetext}
-where $S = S_c - \epsilon \sqrt{R}$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$.
-The first term in equation~(\ref{eq:asymp}) is the result from a two-dimensional calculation.
-The second term is negative definite and proportional to $k_{y}^2$, which leads to an ultraviolet catastrophe in the absence of higher-order effects.
+where
+\begin{equation}
+\Upsilon \ = \ \frac{\left(6+\pi ^2\right) q+\pi^2-6}{12 }\  \approx \ 1.31 \ \  \text{for}  \ \  q = \frac{3}{4}.
+\end{equation}
+%\end{widetext}
+The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
+The second term is negative definite and leads ultraviolet divergence in the absence of higher-order effects.
 
-Figure~\ref{fig:growth_rate}b compares the numerically growth rates for $\SSC=1.002$ between $0 \le k_y \le 0.2$ to this asymptotic approximation, showing that the numerical calculations accurately reproduce the analytic ones in the region where the latter is valid.
-We provide the full spectrum for the most unstable parameter values for $\SSC = 1.02$ in figure~\ref{fig:growth_rate}c.
-The single unstable mode is purely growing (orange dot), and its complex conjugate is also present (blue dot at $\omega = 0$), consistent with equation~(\ref{eq:asymp}).
-The other stable modes (blue dots) are rotationally modified Alfven waves found in left- and right- going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
+Figure~\ref{fig:growth_rate}b compares the numerical growth rates for $\SSC=1.002$ between $0 \le k_y \le 0.2$ to the asymptotic approximation, showing good agreement where the latter is valid.
+Figure~\ref{fig:growth_rate}c shows the full spectrum for $\SSC = 1.02$.
+The plot shows a purely growing/decaying complex-conjugate pair (orange/blue dots on the real axis) consistent with equation~(\ref{eq:asymp}).
+The other stable modes (blue dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
+%\onecolumngrid
 %
-%\onecolumngrid
-
 \begin{figure*}[h!]
@@
 %
-Numerical simulations have consistently shown axisymmetric modes dominating early evolution of the MRI before breaking down into MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
-These axisymmetric modes, called ``channel modes,'' are exact non-linear solutions of the governing equations with shearing-periodic boundary conditions in $x$, and thus must saturate via parasitic modes from physics beyond non-linear mode coupling \citep{1994ApJ...432..213G}.
-Though we use impenetrable, stress-free boundary conditions that do not allow the unbounded growth of channel modes, the basic principle that axisymmetric disturbences grow fastest is not clearly caused by boundary conditions.
-thus, we must reconcile our results with those earlier simulations.
-In figure~\ref{fig:alpha}, we plot the angle $\alpha = \arctan k_y/k_z$ of the fastest growing mode with the $z$ axis as function of $\SSC$.
-Above $\SSC \gtrsim 2.05$, $\alpha$ is zero, indicating that axisymmetric modes have the fastest growth rates.
-For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC \simeq 4.84$ (similar values are found in most works).
-Thus, our theory predicts that the linear dynamics should be dominated by axisymmetric modes for the parameters studied in prior numerical simulations.
+Numerical simulations consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
+These ``channel modes'' are exact non-linear solutions for $x$-shearing-periodic domains, and can only saturate via parasitic shear instabilities \citep{1994ApJ...432..213G}.
+We use impenetrable, stress-free boundary conditions that are more applicable to stellar interiors. 
+But even in disks, any kind of finite radial extent will cutoff the unbounded growth of channel modes. 
+We reconcile our results with earlier simulations by showing that the MRI indeed prefers two dimensions for the larger values of  criticality found in disk simulations. 
+Figure~\ref{fig:alpha} shows the phase angle $\phi = \arctan k_y/k_z$ of the fastest growing mode as function of $\SSC$.
+Above $\SSC \gtrsim 2.05$ $\phi=0$, indicating that axisymmetric modes have the fastest growth rates.
+For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC \simeq 4.84$ (most works use similar values).
+Thus, we predict axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
 %
@@
 
-However, even for $\SSC> 2.05$ there are significant swaths of unstable three-dimensional modes with growth rates comparable to the two-dimensional mode (figure~\ref{fig:growth_rate}a).
-There are two significant implications for this.
-First, the linear operator for the MRI, like other shearing systems, is non-normal; this effect appears whenever non-axisymmetric perturbations are allowed \citep[see][for a discussion relevant to the MRI]{1992MNRAS.255P..25K}.
-Because of this, non-modal transient amplification occurs even for stable modes.
-These effects can be strong enough to cause non-linear interactions and turbulence even when the entire flow is linearly stable.
-While we are here concerned with systems linearly unstable to the MRI, the presence of three dimensional modes implies that linear, non-normal behavior could play a very strong role in the MRI near onset.
+Even for $\SSC> 2.05$ there are significant swaths of unstable three-dimensional modes with growth rates comparable to the maximum (figure~\ref{fig:growth_rate}a).
+Also, non-normality generically accompanies non-axisymmetry in shearing systems \citep[see][for a discussion relevant to the MRI]{1992MNRAS.255P..25K}.
+In non-normal systems, transient amplification can occur for stable modes, and can even cause turbulence.
+Therefore significant three dimensionality likely implies important non-normal behavior near onset.
 

### 2019-08-17 paper/supplemental.tex
Added supplemental information document

\documentclass[aps,prl,preprint
,superscriptaddress]{revtex4-1}\newcommand{\SSC}{S/S_{c}}
\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
\newcommand{\apjs}{The Astrophysical Journal Supplement Series}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}

% You should use BibTeX and apsrev.bst for references
% Choosing a journal automatically selects the correct APS
% BibTeX style file (bst file), so only uncomment the line
% below if necessary.
%\bibliographystyle{apsrev4-1}

\newcommand\Beq{\begin{eqnarray}} 
\newcommand\Eeq{\end{eqnarray}}

\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Eqs}[2]{Eqs.~(\ref{#1})--(\ref{#2})}
\newcommand{\eq}[1]{eq.~(\ref{#1})}
\newcommand{\eqs}[2]{eqs.~(\ref{#1})--(\ref{#2})}

\newcommand{\Prob}[2]{\mathbb{P}[\, #1 \, | \, #2 \, ] }
\newcommand{\Pro}[1]{\mathbb{P}[\, #1 \, ] }

\newcommand{\inv}[1]{\frac{1}{#1}}
\newcommand{\pd}[1]{\partial_{#1}}

\newcommand{\Ra}{R}
\newcommand{\Ve}{\mbox{\textit{Ve}}}
\newcommand{\eps}{\varepsilon}
\newcommand{\del}{\delta}
\newcommand{\e}[1]{\eps^{#1}}
\newcommand{\f}{2\Omega}

\newcommand{\n}{\hat{n}}

\newcommand{\dat}{\cdot}
\newcommand{\dd}[1]{\,\mathrm{d}{#1}}
\newcommand{\cc}{\mathrm{c.c.}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\grad}{\nabla}
\renewcommand{\div}{\nabla \dat}
\newcommand{\Tr}{\text{Tr}}

\newcommand{\pren}[1]{ \left(#1\right) }
\newcommand{\brac}[1]{\left[\,#1\,\right]}

\newcommand{\ave}[1]{\big<#1\big>}

\renewcommand{\l}{\ell}

\newcommand{\txt}[1]{\quad \text{#1} \quad }

\begin{document}

% Use the \preprint command to place your local institutional report
% number in the upper righthand corner of the title page in preprint mode.
% Multiple \preprint commands are allowed.
% Use the 'preprintnumbers' class option to override journal defaults
% to display numbers if necessary
%\preprint{}

%Title of paper
\title{Supplemental Information for: The Magnetorotational Instability Prefers Three Dimensions}

\section{Table of Runs}
\label{sec:runs}

\section{Resolution and Reynolds number study}
\label{sec:resolution}

\section{Asymptotic Calculation}
\label{sec:asymp}
Define the domain as $ - d/2 < x < d/2$. Make it symmetric so to make the shear simpler. The shear has no net. Otherwise it means we are not exactly in the rotating frame we thought we were. 
\Beq
\int_{-d/2}^{d/2} S x \text{d} x  \ = \ 0
\Eeq
This is very important later.

The full linear equations:
\Beq
\pd{t} u + S x \pd{y} u - f v + \pd{x} p - B \pd{z} b_{x} &=& 0\\
\pd{t} v + S x \pd{y} v + (S+f) u + \pd{y} p - B \pd{z} b_{y} &=& 0\\
\pd{t} w + S x \pd{y} w + \pd{z} p - B \pd{z} b_{z} &=& 0 \\
\pd{x} u + \pd{y} v + \pd{z} w  &=& 0\\
\pd{t} b_{x} + S x \pd{y} b_{x} - B \pd{z} u &=& 0\\
\pd{t} b_{y} + S x \pd{y} b_{y} - B \pd{z} v - S b_{x}  &=& 0\\
\pd{t} b_{z} + S x \pd{y} b_{z} - B \pd{z} w &=& 0
\Eeq

Assume all variable take the form
\Beq
g = \hat{G}(x) e^{ i ( \omega t + k z + \ell y ) } 
\Eeq

Define the ``parameter''
\Beq
\omega_{S} \ \equiv \ \omega + S x \ell 
\Eeq

We can use equations (2)--(7) to find all amplitudes in terms of $\hat{U}(x)$ and $\hat{U}'(x)$. 
\Beq
\hat{V} &=& \frac{i \left(k^2 \hat{U}(x) \left(B^2 k^2 S-(f+S) \omega
   _S^2\right)+\ell  \omega _S \hat{U}'(x) \left(B^2 k^2-\omega
   _S^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S
   \left(B^2 k^2-\omega _S^2\right)}\\
\hat{W} &=& -\frac{i k \left(\ell  \hat{U}(x) \left(B^2 k^2 S-(f+S) \omega
   _S^2\right)+\omega _S U'(x) \left(\omega _S^2-B^2
   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S
   \left(B^2 k^2-\omega _S^2\right)}\\
   \hat{P} &=& -\frac{i \left(\ell  \hat{U}(x) \left(B^2 k^2 S-(f+S) \omega
   _S^2\right)+\omega _S U'(x) \left(\omega _S^2-B^2
   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S^2}\\
   \hat{B}_{x} &=& \frac{B k \hat{U}(x)}{\omega _S}\\
   \hat{B}_{y} &=& -\frac{i B k \left(\hat{U}(x) \left(B^2 k^2 S \ell ^2+\omega _S^2
   \left(f k^2-S \ell ^2\right)\right)+\ell  \omega _S
   U'(x) \left(\omega _S^2-B^2
   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S^2
   \left(B^2 k^2-\omega _S^2\right)}\\
   \hat{B}_{z} &=& -\frac{i B k^2 \left(\ell  \hat{U}(x) \left(B^2 k^2 S-(f+S)
   \omega _S^2\right)+\omega _S U'(x) \left(\omega _S^2-B^2
   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S^2
   \left(B^2 k^2-\omega _S^2\right)}
\Eeq

We can then substitute everything into equation (1) to get a second-order equation for $\hat{U}$.

When we do this, we get something of the form 
\Beq
\alpha(x) \hat{U}''(x) + \beta(x) \hat{U}'(x) + \gamma(x) \hat(U)(x) \ = \ 0
\Eeq
where
\Beq
\alpha(x) &=& \frac{i \left(B^2 k^2-\omega _s^2\right)}{\left(k^2+\ell
   ^2\right) \omega _s} \\ 
   \beta(x) &=& -\frac{2 i B^2 k^2 S \ell }{\left(k^2+\ell ^2\right) \omega
   _s^2} \\
   \gamma(x) &=& \frac{i \left(\omega _s^4 \left(\frac{f^2
   k^2}{\left(k^2+\ell ^2\right) \left(B^2 k^2-\omega
   _s^2\right)}+1\right)-k^2 \omega _s^2 \left(B^2+\frac{f
   S}{k^2+\ell ^2}\right)+\frac{2 B^2 k^2 S^2 \ell
   ^2}{k^2+\ell ^2}\right)}{\omega _s^3}
\Eeq

We can eliminate the first-order term via
\Beq
\hat{U}(x) \ = \ \chi(x) \psi(x) 
\Eeq
if 
\Beq
\chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
\Eeq
\Beq
\psi''(x) + \Phi(x) \psi(x) \ = \ 0
\Eeq
\Beq
\Phi(x)  \ = \ -\frac{-2 \beta (x) \alpha '(x)+2 \alpha (x) \left(\beta'(x)-2 \gamma (x)\right)+\beta (x)^2}{4 \alpha (x)^2}
\Eeq
Putting everything together 
\Beq
\Phi(x) \ = \ \tfrac{k^2 \omega _S^2 \left(2 B^2 \left(k^2+\ell
   ^2\right)+f (f+S)\right)-B^2 k^2 \left(B^2 k^2
   \left(k^2+\ell ^2\right)+S \left(f k^2-S \ell
   ^2\right)\right)-\left(k^2+\ell ^2\right) \omega
   _S^4}{\left(\omega _S^2-B^2 k^2\right){}^2}
\Eeq

Now we want to rescale everything in terms of a bookkeeping parameter, $\eps$.
\Beq
S  \ \to \  - \frac{\pi ^2 B^2}{d^2 f} ( 1 + \eps^{2} R) 
\Eeq 
\Beq
\omega \ \to \ \eps^{2} \omega , \quad k \ \to \ \eps k, \quad \ell \ \to \ \eps^{2} \ell 
\Eeq
Also 
\Beq
q  \ \equiv  \ \frac{\pi ^2 B^2}{d^2 f^{2}}
\Eeq

Expanding to $\mathcal{O}(\eps^{2})$, 
\Beq
\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \frac{\epsilon ^2 \left(-B^4 k^4+B^2 f^2 q \left(k^2 R+q
   \ell ^2\right)+f^2 (q+1) (\omega -f q x \ell
   )^2\right)}{B^4 k^2}
\Eeq

Now do the perturbations

\Beq
\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0 
\Eeq
\Beq
\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) + \Phi_{2}(x)\psi_{0}(x)\ = \ 0 
\Eeq

Remember we are on $-d/2 < x < d/2$ Therefore 
\Beq
\psi_{0}(x) \ = \ \cos(\pi x / d)
\Eeq
This satisfies the equation and boundary condition. The solvability condition for $\psi_{2}(x)$ is

\Beq
\int_{-d/2}^{d/2} \Phi_{2}(x)\psi_{0}(x)^{2} \text{d} x \ = \ 0.
\Eeq

This works out to 
\Beq
\left(\pi ^2 B^2+d^2 f^2 \right) \omega^{2} - B^4 k^2 \left(d^2 k^2-\pi ^2 R\right)   \ + \ \frac{\pi ^2 B^4 \ell ^2}{12}  \left(\left(6+\pi ^2\right) q+\pi ^2-6\right) \ = \ 0
\Eeq

\Beq
\omega^{2} \ = \ \frac{q\,B^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
   ^2-6}{12 } \, k_{y}^{2}  \right] \ + \ ...
\Eeq

\end{document}

### 2019-08-17 paper/paper.tex
Change alpha --> phi in text. Added figure.

@@
 We reconcile our results with earlier simulations by showing that the MRI indeed prefers two dimensions for the larger values of  criticality found in disk simulations. 
-Figure~\ref{fig:alpha} shows the phase angle $\phi = \arctan k_y/k_z$ of the fastest growing mode as function of $\SSC$.
+Figure~\ref{fig:phi} shows the phase angle $\phi = \arctan k_y/k_z$ of the fastest growing mode as function of $\SSC$.
 Above $\SSC \gtrsim 2.05$ $\phi=0$, indicating that axisymmetric modes have the fastest growth rates.
@@
 \begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{alpha_vs_ssc_grid.pdf}
-  \caption{Angle $\alpha$ of mode with respect to $z$ axis vs $\SSC$.
+  \includegraphics[width=\columnwidth]{phi_vs_ssc_grid.pdf}
+  \caption{Angle $\phi$ of mode with respect to $z$ axis vs $\SSC$.
 The inset shows the transition from axisymmetric modes to three dimensional modes as $\SSC$ decreases below $2.05$.
 The shaded region indicates stability to all perturbations; the stability limit for three dimensional modes is $\SSC \simeq 0.102$.}
-  \label{fig:alpha}
+  \label{fig:phi}
 \end{figure}
@@
 % \begin{itemize}
-% \item non-normal and $k_y$ even when $\alpha = 0$ implies important role for linear dynamics at early times. 
+% \item non-normal and $k_y$ even when $\phi = 0$ implies important role for linear dynamics at early times. 
 % \item Following Squire + Bhattacharjee, this suggests that important stellar layers may in fact be driven by linear dynamics

### 2019-08-17 python/plot_alpha.py
Change alpha --> sigma in phase angle plot

@@ -1,5 +1,5 @@
 """
-alpha = arctan(ky/kz)
+phi = arctan(ky/kz)
 
 """
 
@@ -18,7 +18,7 @@ basedir = Path("data/")
 runs = [34, 33, 31, 21, 15, 11, 12, 16, 30, 13, 27, 28, 29, 26, 24, 17, 14, 18]
 R = [np.sqrt(0.2), np.sqrt(0.3), np.sqrt(0.4), np.sqrt(0.5), 0.8, 1.01, 1.2, np.sqrt(1.75), 1.375, np.sqrt(2.0), np.sqrt(2.01), np.sqrt(2.015), 1.425, np.sqrt(2.05), np.sqrt(2.1), 1.5, np.sqrt(2.5), 2.0]
 ssc = np.array(R)**2
-alpha = []
+phi = []
 
 growth = []
 dk = []
@@ -26,7 +26,7 @@ for run in runs:
     dfname = Path("run_{:d}_output.h5".format(run))
     df = h5py.File(basedir/dfname, "r")
     if use_opt:
-        #alpha.append(np.arctan2(df['gamma'].attrs['max ky'], df['gamma'].attrs['max kz']))
+        #phi.append(np.arctan2(df['gamma'].attrs['max ky'], df['gamma'].attrs['max kz']))
         ky = df['gamma'].attrs['max ky']
         kx = df['gamma'].attrs['max kz']
         gamma_max = df['gamma'].attrs['max growth rate']
@@ -38,51 +38,51 @@ for run in runs:
         gamma_max = gamma_global.real.max()
     dk.append(df['ky'][1]-df['ky'][0])
     kk = np.sqrt(ky**2 + kz**2)
-    alpha.append(np.arcsin(ky/kk))
+    phi.append(np.arcsin(ky/kk))
     growth.append(gamma_max)
     df.close()
-alpha = np.array(alpha)
+phi = np.array(phi)
 ssc = np.array(ssc)
 dk = np.array(dk)
 
 fig, ax = plt.subplots()
-ax.scatter(ssc, alpha, marker='o', c=growth, yunits=radians, zorder=2)#,edgecolor='k',linewidth=0.8)
-#ax.colorbar(label='$\gamma$')
[…]

### 2019-08-17 python/plot_fig1.py
Fixed colors for spectrum

@@ -123,8 +123,10 @@ spectrum = data['eigvals'][:]
 
 thresh = 2
 spec_thresh = spectrum[np.abs(spectrum) < thresh]
-spec_ax.scatter(spec_thresh[spec_thresh.real <= 0].real, spec_thresh[spec_thresh.real <= 0].imag)
+spec_ax.scatter(spec_thresh[(spec_thresh.real < 0) & (np.abs(spec_thresh.imag) < 1e-8)].real,spec_thresh[(spec_thresh.real < 0) & (np.abs(spec_thresh.imag) < 1e-8)].imag,zorder=10)
 spec_ax.scatter(spec_thresh[spec_thresh.real > 0].real, spec_thresh[spec_thresh.real > 0].imag)
+spec_ax.scatter(spec_thresh[spec_thresh.real <= 0].real, spec_thresh[spec_thresh.real <= 0].imag, color='lightgrey',zorder=3)
+
 spec_ax.set_xlim(-0.25,0.1)
 spec_ax.set_ylim(-0.8,0.8)
 spec_ax.set_xlabel(r"$\gamma$")

### 2019-08-17 paper/paper.tex
Updated paper to reflect new spectrum colors.

@@
 The plot shows a purely growing/decaying complex-conjugate pair (orange/blue dots on the real axis) consistent with equation~(\ref{eq:asymp}).
-The other stable modes (blue dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
+The other stable modes (gray dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
 %\onecolumngrid
@@
 c) The full discrete spectrum of the MRI for $(k_y, k_z) \simeq (0.263, 0.447)$.
-The unstable mode is shown in orange; all other modes are stable ($\gamma \le 0$).}
+The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray.}
   \label{fig:growth_rate}

### 2019-08-19 python/plot_spectrum.py
Bug fix

@@ -43,8 +43,9 @@ spec_thresh = spectrum[np.abs(spectrum) < thresh]
 plt.scatter(spec_thresh[spec_thresh.real <= 0].real, spec_thresh[spec_thresh.real <= 0].imag)
 plt.scatter(spec_thresh[spec_thresh.real > 0].real, spec_thresh[spec_thresh.real > 0].imag)
 plt.gca().set_prop_cycle(None)
-plt.plot(-omega, 0., marker='^')
-plt.plot(omega, 0., marker='^')
+if asymptotic:
+    plt.plot(-omega, 0., marker='^')
+    plt.plot(omega, 0., marker='^')
 plt.xlim(-0.25,0.25)
 plt.ylim(-0.8,0.8)
 plt.xlabel("real")

### 2019-08-19 paper/paper.tex

@@
 While surprising, preferred three-dimensional modes can be predicted analytically.
-We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) with the shear asymptotically close to the critical two-dimensional rate, and compute the leading-order correction to the growth rate from three-dimensional effects assuming where $S = S_c - \epsilon \sqrt{R}$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$,
+We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) with the shear asymptotically close to the critical two-dimensional rate, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $S = S_c - \epsilon \sqrt{R}$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$,
 %\begin{widetext}
@@
   \label{eq:asymp}
-\sigma^{2} =  -\frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left( \frac{d^2}{\pi^{2}} k_z^2 - R^{2}\right)  -  \Upsilon \, k_{y}^{2}  \right] +  \ldots,
+\sigma^{2} =  \frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left( R^{2} - \frac{d^2}{\pi^{2}} k_z^2 \right)  +  \Upsilon \, k_{y}^{2}  \right] +  \ldots,
 \end{equation}

### 2019-08-19 python/mri_single_yz_mode.py
Modified single mode code to allow ideal MHD case
* also added script to plot eigenvectors of most unstable mode

@@ -4,7 +4,10 @@ The magnetorotational instability prefers three dimensions.
 Single mode calculation. Returns spectrum.
 
 Usage:
-    mri_single_yz_mode.py <config_file> <ky> <kz>
+    mri_single_yz_mode.py [--ideal] <config_file> <ky> <kz>
+
+Options:
+    --ideal    Use Ideal MHD
 """
 
 from docopt import docopt
@@ -23,6 +26,7 @@ logger = logging.getLogger(__name__)
 args = docopt(__doc__)
 ky = float(args['<ky>'])
 kz = float(args['<kz>'])
+ideal = args['--ideal']
 filename = Path(args['<config_file>'])
 outbase = Path("data")
 
@@ -75,7 +79,9 @@ domain = de.Domain([x_basis], grid_dtype=np.complex128, comm=MPI.COMM_SELF)
 
 # 3D MRI
 
-problem_variables = ['p','vx','vy','vz','ωy','ωz','bx','by','bz','jxx']
+problem_variables = ['p','vx','vy','vz','bx','by','bz']
+if not ideal:
+    problem_variables += ['ωy','ωz','jxx']
 problem = de.EVP(domain, variables=problem_variables, eigenvalue='gamma')
 problem.meta[:]['x']['dirichlet'] = True
 
@@ -107,36 +113,46 @@ problem.substitutions['jz'] = "dx(by) - dy(bx)"
 # Hydro equations: p, vx, vy, vz, ωy, ωz
 
 problem.add_equation("dx(vx) + dy(vy) + dz(vz) = 0")
+if ideal:
+    problem.add_equation("Dt(vx)  -     f*vy + dx(p) - B*dz(bx) = 0")
+    problem.add_equation("Dt(vy)  + (f+S)*vx + dy(p) - B*dz(by) = 0")
+    problem.add_equation("Dt(vz)             + dz(p) - B*dz(bz) = 0")
+else:
+    problem.add_equation("Dt(vx)  -     f*vy + dx(p) - B*dz(bx) + ν*(dy(ωz) - dz(ωy)) = 0")
+    problem.add_equation("Dt(vy)  + (f+S)*vx + dy(p) - B*dz(by) + ν*(dz(ωx) - dx(ωz)) = 0")
+    problem.add_equation("Dt(vz)             + dz(p) - B*dz(bz) + ν*(dx(ωy) - dy(ωx)) = 0")
 
[…]

### 2019-08-19 python/plot_single_mode_evecs.py

@@ -1,0 +1,27 @@
+import matplotlib.pyplot as plt
+import h5py
+import sys
+from pathlib import Path
+
+plt.style.use('prl')
+filename = Path(sys.argv[-1])
+
+rows = 2
+cols = 3
+names = [r'$v_x$',r'$v_y$',r'$v_z$',r'$b_x$',r'$b_y$',r'$b_z$']
+plt.figure(figsize=(12,6))
+with h5py.File(filename, 'r') as df:
+    x = df['eigvecs'].attrs['x']
+    for i in range(1,7):
+        plt.subplot(rows,cols,i)
+        plt.plot(x, df['eigvecs'][i,:].real, label='real')
+        plt.plot(x, df['eigvecs'][i,:].imag, label='imag')
+        if i == 1:
+            plt.legend()
+        plt.xlabel(r'$x$')
+        plt.ylabel(names[i-1])
+
+plt.tight_layout()
+
+plt.savefig('plots/' + filename.stem + 'eigvecs.pdf')
+

### 2019-08-19 python/plot_mean_field_alpha.py
Use pdf in mean field alpha plot

@@ -25,5 +25,5 @@ plt.plot(x[0,0,:], alpha_yz)
 plt.xlabel(r"$x$")
 plt.ylabel(r"$\alpha$")
 plt.tight_layout()
-plt.savefig('plots/mean_field_alpha_{}.png'.format(dfname))
+plt.savefig('plots/mean_field_alpha_{}.pdf'.format(dfname))
 

### 2019-08-19 python/make_eigenvalue_plot.py
Allow this code to work even if root finder isn't run.

@@ -26,12 +26,18 @@ datafile = h5py.File(filename,'r')
 gamma_global = datafile['gamma'][:]
 gamma_r = gamma_global.real
 #gamma_r[np.where(gamma_r<0)] = 0.0
-max_gamma = datafile['gamma'].attrs['max growth rate']
-max_ky = datafile['gamma'].attrs['max ky']
-max_kz = datafile['gamma'].attrs['max kz']
-
+#max_gamma = datafile['gamma'].attrs['max growth rate']
 ky_global    = datafile['ky']
 kz_global    = datafile['kz']
+print(ky_global.shape)
+print(gamma_r.shape)
+try:
+    max_ky = datafile['gamma'].attrs['max ky']
+    max_kz = datafile['gamma'].attrs['max kz']
+except KeyError:
+    ky_index, kz_index = np.where(gamma_r == gamma_r.max())
+    max_ky = ky_global[ky_index[0]]
+    max_kz = kz_global[kz_index[0]]
 
 print("max growth rate on grid = {}".format(gamma_r.max()))
 
@@ -47,6 +53,7 @@ plt.colorbar(PCM, label=r'$\gamma$')
 plt.xlabel(r'$k_z$')
 plt.ylabel(r'$k_y$')
 plt.tight_layout()
+
 #plt.title(r'3D Keplerian MRI growth rates/f  $\left( S/S_{\mathrm{crit.}} = %.2f\right)$' %(R[i]))
 plot_file_name = Path(filename.stem + '_growthrates.png')
 plt.savefig(outbase/plot_file_name, dpi=300)

### 2019-08-20 python/mri_single_yz_mode.py
ideal is now 2nd order in x.

@@ -94,8 +94,9 @@ problem.parameters['B'] = B
 problem.parameters['ky'] = ky
 problem.parameters['kz'] = kz
 
-problem.parameters['ν'] = ν
-problem.parameters['η'] = η
+if not ideal:
+    problem.parameters['ν'] = ν
+    problem.parameters['η'] = η
 
 # Operator substitutions for y,z, and t derivatives
 
@@ -105,10 +106,12 @@ problem.substitutions['Dt(A)'] = "gamma*f*A + S*x*dy(A)"
 
 # Variable substitutions
 
-problem.substitutions['ωx'] = "dy(vz) - dz(vy)"
-problem.substitutions['jx'] = "dy(bz) - dz(by)"
-problem.substitutions['jy'] = "dz(bx) - dx(bz)"
-problem.substitutions['jz'] = "dx(by) - dy(bx)"
+if not ideal:
+    problem.substitutions['ωx'] = "dy(vz) - dz(vy)"
+    problem.substitutions['jx'] = "dy(bz) - dz(by)"
+    problem.substitutions['jy'] = "dz(bx) - dx(bz)"
+    problem.substitutions['jz'] = "dx(by) - dy(bx)"
+    probelm.substitutions['L(A)'] = "d(A,y=2) + d(A,z=2)"
 
 # Hydro equations: p, vx, vy, vz, ωy, ωz
 
@@ -117,6 +120,12 @@ if ideal:
     problem.add_equation("Dt(vx)  -     f*vy + dx(p) - B*dz(bx) = 0")
     problem.add_equation("Dt(vy)  + (f+S)*vx + dy(p) - B*dz(by) = 0")
     problem.add_equation("Dt(vz)             + dz(p) - B*dz(bz) = 0")
+
+    # Frozen-in field
+    problem.add_equation("Dt(bx) - B*dz(vx)        = 0")
+    problem.add_equation("Dt(by) - B*dz(vy) - S*bx = 0")
+    problem.add_equation("Dt(bz) - B*dz(vz)        = 0")
+
 else:
     problem.add_equation("Dt(vx)  -     f*vy + dx(p) - B*dz(bx) + ν*(dy(ωz) - dz(ωy)) = 0")
     problem.add_equation("Dt(vy)  + (f+S)*vx + dy(p) - B*dz(by) + ν*(dz(ωx) - dx(ωz)) = 0")
@@ -125,30 +134,22 @@ else:
[…]

### 2019-08-20 python/mri_single_yz_mode.py
fixed small bug

@@ -111,7 +111,7 @@ if not ideal:
     problem.substitutions['jx'] = "dy(bz) - dz(by)"
     problem.substitutions['jy'] = "dz(bx) - dx(bz)"
     problem.substitutions['jz'] = "dx(by) - dy(bx)"
-    probelm.substitutions['L(A)'] = "d(A,y=2) + d(A,z=2)"
+    probelm.substitutions['L(A)'] = "dy(dy(A)) + dz(dz(A))"
 
 # Hydro equations: p, vx, vy, vz, ωy, ωz
 

### 2019-08-20 python/mri_single_yz_mode.py
Save the right eigenvector...

@@ -111,7 +111,7 @@ if not ideal:
     problem.substitutions['jx'] = "dy(bz) - dz(by)"
     problem.substitutions['jy'] = "dz(bx) - dx(bz)"
     problem.substitutions['jz'] = "dx(by) - dy(bx)"
-    probelm.substitutions['L(A)'] = "dy(dy(A)) + dz(dz(A))"
+    problem.substitutions['L(A)'] = "dy(dy(A)) + dz(dz(A))"
 
 # Hydro equations: p, vx, vy, vz, ωy, ωz
 
@@ -173,15 +173,17 @@ logger.info("Solve time: {}".format(t2-t1))
 
 gamma = solver.eigenvalues
 dense_threshold=1
-gamma = gamma[np.abs(gamma) < dense_threshold]
+
+thresh = (np.abs(gamma) < dense_threshold)
+gamma = gamma[thresh]
+solver.eigenvectors = solver.eigenvectors[:,thresh]
 index = np.argsort(-gamma.real)
-gamma = gamma[index]
-print("saving eigenvector with gamma = {}".format(gamma[0]))
+
+logger.info("saving eigenvector with gamma = {}".format(gamma[index[0]]))
 nvars = len(problem_variables)
 eigvec = np.zeros((nvars,Nx),dtype=np.complex128)
-
+solver.set_state(index[0])
 for k in range(nvars):
-    solver.set_state(index[0])
     eigvec[k,:] = solver.state[problem_variables[k]]['g']
 
 # Save either or both eigenvalues and eigenvectors to a single .h5 file

### 2019-08-20 python/mri_single_yz_mode.py
Add option to save VTK files for 3D viz.
* it turns out 3D viz is pretty boring...

@@ -203,4 +203,32 @@ if CW.rank == 0:
     dset_eval.attrs.create("d", Lx)
     dset_evec = output_file.create_dataset('eigvecs',data=eigvec)
     dset_evec.attrs.create('x',x_basis.grid())
-    
+
+    n_yz = 128
+    Ly = 2*np.pi/ky
+    Lz = 2*np.pi/kz
+    x = x_basis.grid()
+    y = np.linspace(0, Ly, n_yz, endpoint=False)
+    z = np.linspace(0, Lz, n_yz, endpoint=False)
+
+    zz,yy,xx = np.meshgrid(z,y,x,indexing='ij')
+    yz_dep = np.exp(1j*(ky*yy +kz*zz))
+
+    p = (eigvec[0,:]*yz_dep).real
+    u = (eigvec[1,:]*yz_dep).real
+    v = (eigvec[2,:]*yz_dep).real
+    w = (eigvec[3,:]*yz_dep).real
+    Bx =(eigvec[6,:]*yz_dep).real
+    By =(eigvec[7,:]*yz_dep).real
+    Bz =(eigvec[8,:]*yz_dep).real
+
+    save_vtk = False
+    if save_vtk:
+        vtkfile = filename.stem + '.vtk'
+        from pyevtk.hl import gridToVTK 
+        pointData = {'p':p.copy(), 'u':u.copy(), 'v':v.copy(), 'w':w.copy(),
+                     'Bx':Bx.copy(), 'By':By.copy(), 'Bz':Bz.copy()}
+        print("p flags", pointData['p'].flags)
+
+        gridToVTK(vtkfile, xx, yy, zz, pointData=pointData)
+

### 2019-08-20 python/plot_single_mode_evecs.py
Plot 2D x-y plane visualizations of eigenmodes as well as 1D

@@ -1,6 +1,7 @@
 import matplotlib.pyplot as plt
 import h5py
 import sys
+import numpy as np
 from pathlib import Path
 
 plt.style.use('prl')
@@ -9,6 +10,7 @@ filename = Path(sys.argv[-1])
 rows = 2
 cols = 3
 names = [r'$v_x$',r'$v_y$',r'$v_z$',r'$b_x$',r'$b_y$',r'$b_z$']
+cmaps = ['RdBu','RdBu','RdBu','PuOr','PuOr','PuOr']
 plt.figure(figsize=(12,6))
 with h5py.File(filename, 'r') as df:
     x = df['eigvecs'].attrs['x']
@@ -21,7 +23,34 @@ with h5py.File(filename, 'r') as df:
         plt.xlabel(r'$x$')
         plt.ylabel(names[i-1])
 
-plt.tight_layout()
+    plt.tight_layout()
+    plt.savefig('plots/' + filename.stem + 'eigvecs.pdf')
 
-plt.savefig('plots/' + filename.stem + 'eigvecs.pdf')
+    # 2D plots
+    n_yz = 128
+    ky = df['eigvals'].attrs['ky']
+    kz = df['eigvals'].attrs['kz']
+    Ly = 2*np.pi/ky
+    Lz = 2*np.pi/kz
 
+    y = np.linspace(0, Ly, n_yz, endpoint=False)
+
+    yyy, xxx = np.meshgrid(y,x,indexing='ij')
+    y_dep = np.exp(1j*ky*yyy)
+    plt.figure(figsize=(10,10))
+    eigvec = df['eigvecs']
+    for i in range(1,7):
+        ax = plt.subplot(rows,cols,i)
+        pcol = plt.pcolormesh(xxx,yyy,(eigvec[i,:]*y_dep).real,linewidth=0,rasterized=True, cmap=cmaps[i-1])
+        #pcol.set_edgecolor('face')
+        plt.title(names[i-1])
+        if i !=4:
+            ax.axes.get_xaxis().set_visible(False)
+            ax.axes.get_yaxis().set_visible(False)
+        else:
+            plt.xlabel(r'$x$')
+            plt.ylabel(r'$y$')
+            ax.set_xticklabels(["$-\pi/2$", 0, "$\pi/2$"])
+            ax.set_xticks([-np.pi/2, 0, np.pi/2])
+
+    plt.savefig('plots/eigvecs_xy_' + filename.stem +'.pdf')

### 2019-08-21 paper/supplemental.tex
tidy up supp. material

@@
 \label{sec:asymp}
-Define the domain as $ - d/2 < x < d/2$. Make it symmetric so to make the shear simpler. The shear has no net. Otherwise it means we are not exactly in the rotating frame we thought we were. 
+Define the domain as $ - d/2 < x < d/2$. Being in the rotating frame means the shear has no net
 \Beq
-\int_{-d/2}^{d/2} S x \text{d} x  \ = \ 0
+\int_{-d/2}^{d/2} V_{y}(x) \text{d} x  \ = \ 0, \quad \implies  \quad V_{y}(x) \ = \ S\, x
 \Eeq
-This is very important later.
-
-The full linear equations:
+The full linear ideal equations are:
 \Beq
-\pd{t} u + S x \pd{y} u - f v + \pd{x} p - B \pd{z} b_{x} &=& 0\\
-\pd{t} v + S x \pd{y} v + (S+f) u + \pd{y} p - B \pd{z} b_{y} &=& 0\\
-\pd{t} w + S x \pd{y} w + \pd{z} p - B \pd{z} b_{z} &=& 0 \\
+\pd{t} u + S x \pd{y} u - f v + \pd{x} p - B_{0} \pd{z} b_{x} &=& 0\\
+\pd{t} v + S x \pd{y} v + (S+f) u + \pd{y} p - B_{0} \pd{z} b_{y} &=& 0\\
+\pd{t} w + S x \pd{y} w + \pd{z} p - B_{0} \pd{z} b_{z} &=& 0 \\
 \pd{x} u + \pd{y} v + \pd{z} w  &=& 0\\
-\pd{t} b_{x} + S x \pd{y} b_{x} - B \pd{z} u &=& 0\\
-\pd{t} b_{y} + S x \pd{y} b_{y} - B \pd{z} v - S b_{x}  &=& 0\\
-\pd{t} b_{z} + S x \pd{y} b_{z} - B \pd{z} w &=& 0
+\pd{t} b_{x} + S x \pd{y} b_{x} - B_{0} \pd{z} u &=& 0\\
+\pd{t} b_{y} + S x \pd{y} b_{y} - B_{0} \pd{z} v - S b_{x}  &=& 0\\
+\pd{t} b_{z} + S x \pd{y} b_{z} - B_{0} \pd{z} w &=& 0
 \Eeq
@@
 \Eeq
-
-Define the ``parameter''
+Define the frequency ``parameters'' 
 \Beq
-\omega_{S} \ \equiv \ \omega + S x \ell 
+\omega_{S}(x) \ \equiv \ \omega + S x \ell, \quad \omega_{A}  \ \equiv \ B_{0} k
 \Eeq
-
 We can use equations (2)--(7) to find all amplitudes in terms of $\hat{U}(x)$ and $\hat{U}'(x)$. 
 \Beq
-\hat{V} &=& \frac{i \left(k^2 \hat{U}(x) \left(B^2 k^2 S-(f+S) \omega
-   _S^2\right)+\ell  \omega _S \hat{U}'(x) \left(B^2 k^2-\omega
-   _S^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S
-   \left(B^2 k^2-\omega _S^2\right)}\\
-\hat{W} &=& -\frac{i k \left(\ell  \hat{U}(x) \left(B^2 k^2 S-(f+S) \omega
-   _S^2\right)+\omega _S U'(x) \left(\omega _S^2-B^2
-   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S
-   \left(B^2 k^2-\omega _S^2\right)}\\
-   \hat{P} &=& -\frac{i \left(\ell  \hat{U}(x) \left(B^2 k^2 S-(f+S) \omega
-   _S^2\right)+\omega _S U'(x) \left(\omega _S^2-B^2
-   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S^2}\\
-   \hat{B}_{x} &=& \frac{B k \hat{U}(x)}{\omega _S}\\
-   \hat{B}_{y} &=& -\frac{i B k \left(\hat{U}(x) \left(B^2 k^2 S \ell ^2+\omega _S^2
-   \left(f k^2-S \ell ^2\right)\right)+\ell  \omega _S
-   U'(x) \left(\omega _S^2-B^2
-   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S^2
-   \left(B^2 k^2-\omega _S^2\right)}\\
-   \hat{B}_{z} &=& -\frac{i B k^2 \left(\ell  \hat{U}(x) \left(B^2 k^2 S-(f+S)
-   \omega _S^2\right)+\omega _S U'(x) \left(\omega _S^2-B^2
-   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S^2
-   \left(B^2 k^2-\omega _S^2\right)}
+\hat{V}(x) &=& \frac{i \left(\ell 
+   \hat{U}'(x) + k^{2} \frac{ \left((f+S) \omega
+   _S^2-S \omega _A^2\right)}{\omega _S
+   \left(\omega _S^2-\omega _A^2\right)} \hat{U}(x)\right)}{k^2+\ell ^2} \\
+\hat{W}(x) &=& \frac{i k \left(\hat{U}'(x)- \frac{ \left((f+S) \omega
+   _S^2-S \omega _A^2\right)}{\omega _S
+   \left(\omega _S^2-\omega _A^2\right)}  \ell 
+   \hat{U}(x)\right)}{k^2+\ell ^2} \\
+   \hat{P}(x) &=& \frac{i \left(\omega _A^2-\omega _S^2\right)
+   \left(\hat{U}'(x) -  \ell \frac{ \left((f+S) \omega
+   _S^2-S \omega _A^2\right)}{\omega _S
+   \left(\omega _S^2-\omega _A^2\right)}
+   \hat{U}(x)\right)}{\left(k^2+\ell ^2\right)
+   \omega _S} \\
+   \hat{B}_{x}(x) &=& \frac{\omega _A \hat{U}(x)}{\omega _S} \\
+   \hat{B}_{y}(x) &=&\frac{\omega _A \hat{V}(x)}{\omega _S} -\frac{i S \omega _A \hat{U}(x)}{\omega _S^2} \\
+   \hat{B}_{z}(x) &=& \frac{\omega _A \hat{W}(x)}{\omega _S}
    \Eeq
-
-We can then substitute everything into equation (1) to get a second-order equation for $\hat{U}$.
-
-When we do this, we get something of the form 
+We can then substitute everything into equation (1) to get a second-order equation for $\hat{U}(x)$ of the general form 
 \Beq
-\alpha(x) \hat{U}''(x) + \beta(x) \hat{U}'(x) + \gamma(x) \hat(U)(x) \ = \ 0
+\alpha(x) \hat{U}''(x) + \beta(x) \hat{U}'(x) + \gamma(x) \hat{U}(x) \ = \ 0
 \Eeq
@@
 \Beq
-\alpha(x) &=& \frac{i \left(B^2 k^2-\omega _s^2\right)}{\left(k^2+\ell
-   ^2\right) \omega _s} \\ 
-   \beta(x) &=& -\frac{2 i B^2 k^2 S \ell }{\left(k^2+\ell ^2\right) \omega
-   _s^2} \\
-   \gamma(x) &=& \frac{i \left(\omega _s^4 \left(\frac{f^2
-   k^2}{\left(k^2+\ell ^2\right) \left(B^2 k^2-\omega
-   _s^2\right)}+1\right)-k^2 \omega _s^2 \left(B^2+\frac{f
-   S}{k^2+\ell ^2}\right)+\frac{2 B^2 k^2 S^2 \ell
-   ^2}{k^2+\ell ^2}\right)}{\omega _s^3}
+\alpha(x) &=& \omega _S^2-\omega _A^2 \\ 
+   \beta(x) &=& \frac{2 S \ell  \omega _A^2}{\omega _S} \\
+   \gamma(x) &=& \omega _A^2 \left(\frac{f^2 k^2}{\omega
+   _S^2-\omega _A^2}-\frac{2 S^2 \ell
+   ^2}{\omega _S^2}+k^2+\ell ^2\right)+f k^2
+   (f+S)-\left(k^2+\ell ^2\right) \omega _S^2
 \Eeq
@@
 \Beq
-\hat{U}(x) \ = \ \chi(x) \psi(x) 
-\Eeq
-if 
-\Beq
-\chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
+\hat{U}(x) \ = \ \chi(x) \psi(x) \quad \text{where} \quad \chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
 \Eeq
 \Beq
-\psi''(x) + \Phi(x) \psi(x) \ = \ 0
+-\psi''(x) + \Phi(x) \psi(x) \ = \ 0
 \Eeq
 \Beq
-\Phi(x)  \ = \ -\frac{-2 \beta (x) \alpha '(x)+2 \alpha (x) \left(\beta'(x)-2 \gamma (x)\right)+\beta (x)^2}{4 \alpha (x)^2}
+\Phi(x)  \ = \ \frac{-2 \beta (x) \alpha '(x)+2 \alpha (x)
+   \left(\beta '(x)-2 \gamma (x)\right)+\beta
+   (x)^2}{4 \alpha (x)^2}
 \Eeq
@@
 \Beq
-\Phi(x) \ = \ \tfrac{k^2 \omega _S^2 \left(2 B^2 \left(k^2+\ell
-   ^2\right)+f (f+S)\right)-B^2 k^2 \left(B^2 k^2
-   \left(k^2+\ell ^2\right)+S \left(f k^2-S \ell
-   ^2\right)\right)-\left(k^2+\ell ^2\right) \omega
-   _S^4}{\left(\omega _S^2-B^2 k^2\right){}^2}
+\Phi(x) \ = \ \frac{S  \left(f k^2-S \ell
+   ^2\right) \omega _A^2 -f k^2 (f+S) \omega
+   _S^2}{\left(\omega _A^2-\omega
+   _S^2\right){}^2} + k^{2} + \ell^{2}
 \Eeq
 
-Now we want to rescale everything in terms of a bookkeeping parameter, $\eps$.
+Rescale everything in terms of a bookkeeping parameter, $\eps$.
 \Beq
-S  \ \to \  - \frac{\pi ^2 B^2}{d^2 f} ( 1 + \eps^{2} R) 
+S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R) 
 \Eeq 
@@
 \Beq
-q  \ \equiv  \ \frac{\pi ^2 B^2}{d^2 f^{2}}
+q  \ \equiv  \ \frac{\pi ^2 B_{0}^2}{d^2 f^{2}}
 \Eeq
-
 Expanding to $\mathcal{O}(\eps^{2})$, 
 \Beq
-\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \frac{\epsilon ^2 \left(-B^4 k^4+B^2 f^2 q \left(k^2 R+q
+\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \frac{\epsilon ^2 \left(-B_{0}^4 k^4+B_{0}^2 f^2 q \left(k^2 R+q
    \ell ^2\right)+f^2 (q+1) (\omega -f q x \ell
-   )^2\right)}{B^4 k^2}
+   )^2\right)}{B_{0}^4 k^2}
 \Eeq
-
-Now do the perturbations
-
+Going order-by-order
 \Beq
@@
 \Eeq
-
-Remember we are on $-d/2 < x < d/2$ Therefore 
+Therefore 
 \Beq
@@
 This satisfies the equation and boundary condition. The solvability condition for $\psi_{2}(x)$ is 
-
 \Beq
@@
 \Eeq
-
-This works out to 
+This implies 
 \Beq
-\left(\pi ^2 B^2+d^2 f^2 \right) \omega^{2} - B^4 k^2 \left(d^2 k^2-\pi ^2 R\right)   \ + \ \frac{\pi ^2 B^4 \ell ^2}{12}  \left(\left(6+\pi ^2\right) q+\pi ^2-6\right) \ = \ 0
-\Eeq
-
-\Beq
-\omega^{2} \ = \ \frac{q\,B^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
+\omega^{2} \ = \ \frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
    ^2-6}{12 } \, k_{y}^{2}  \right] \ + \ ...

### 2019-08-21 python/plot_fig1.py
Change zero contour color on figure 1

@@ -45,7 +45,7 @@ for i,r in enumerate(runs):
     c = ax.imshow(gamma_r, extent=[kz_global[:].min(),kz_global[:].max(),ky_global[:].min(),ky_global[:].max()], vmin=vmin,vmax=vmax, origin='lower')
 
     ax.contour(kz_global,ky_global,gamma_r,levels=contour_levels,colors='k')
-    ax.contour(kz_global,ky_global,gamma_r,levels=[0.],colors='w')
+    ax.contour(kz_global,ky_global,gamma_r,levels=[0.],colors='grey')
     ax.text(0.05,0.9,r'$S/S_c = {:5.3f}$'.format(SSC[i]), color='w',fontsize=18)
     ax.plot(max_kz, max_ky, 'ro')
     #ax.colorbar(PCM, label=r'$\gamma$')

### 2019-08-22 python/make_eigenvalue_plot.py
Rescale growth rates by S

@@ -1,13 +1,20 @@
 """
 This script plots both the largest growth rates of the MRI data
-passed to the script. To run, pass it a file as follows:
+passed to the script. 
 
-python3 make_eigenvector_plots.py mri_dataset_845.h5
+Usage:
+    make_eigenvector_plots.py [--vmin=<vmin> --vmax=<vmax> --q=<q> --no-contours] <filename>
+
+Options:
+    --vmin=<vmin>    min for colorbar [default: 0]
+    --vmax=<vmax>    max for colorbar [default: 0.701]
+    --q=<q>          q = S/f [default: 0.75]
+    --no-contours    turn off contours
 
 """
 
 import h5py
-import argparse
+from docopt import docopt
 from pathlib import Path
 import numpy as np
 import matplotlib.pyplot as plt
@@ -15,16 +22,20 @@ import matplotlib.pyplot as plt
 plt.style.use('prl')
 
 # Parses filename passed to script
-parser = argparse.ArgumentParser(description='Passes filename')
-parser.add_argument('filename', metavar='Rc', type=str, help='.h5 file to plot eigenvectors for maximum eigenvalue')
-args = parser.parse_args()
-filename = Path(vars(args)['filename'])
+args = docopt(__doc__)
+
+filename = Path(args['<filename>'])
+vmin = float(args['--vmin'])
+vmax = float(args['--vmax'])
+q = float(args['--q'])
+contours = not args['--no-contours']
 outbase = Path("plots")
 
 # Plot growth rates
 datafile = h5py.File(filename,'r')
 gamma_global = datafile['gamma'][:]
-gamma_r = gamma_global.real
+gamma_r = gamma_global.real/q # scale by S...gamma is already multiplied by f in script
+
 #gamma_r[np.where(gamma_r<0)] = 0.0
 #max_gamma = datafile['gamma'].attrs['max growth rate']
 ky_global    = datafile['ky']
@@ -42,18 +53,16 @@ except KeyError:
 print("max growth rate on grid = {}".format(gamma_r.max()))
 
 contour_levels = np.linspace(0,gamma_r.max(),10)
-vmin = 0.
-vmax = 0.071
-PCM = plt.pcolormesh(kz_global,ky_global,gamma_r, vmin=vmin,vmax=vmax)
-
[…]

### 2019-08-22 python/plot_fig1.py

@@ -11,6 +11,7 @@ plt.style.use('prl')
 filebase = 'data/run_{:d}_output.h5'
 runs = [15, 39, 12, 13]
 SSC = [0.64, 1.002001, 1.0201, 2.]
+q = 0.75
 index = 0
 
 fig = plt.figure(figsize=(16,8))
@@ -34,14 +35,14 @@ for i,r in enumerate(runs):
     ky_global    = datafile['ky']
     kz_global    = datafile['kz']
     gamma_global = datafile['gamma'][:]
-    gamma_r = gamma_global.real
+    gamma_r = gamma_global.real/q # scale by S
     #gamma_r[np.where(gamma_r<0)] = 0.0
     max_gamma = datafile['gamma'].attrs['max growth rate']
     max_ky = datafile['gamma'].attrs['max ky']
     max_kz = datafile['gamma'].attrs['max kz']
     contour_levels = np.linspace(0,gamma_r.max(),5)
     vmin = 0.
-    vmax = 0.071
+    vmax = 0.1
     c = ax.imshow(gamma_r, extent=[kz_global[:].min(),kz_global[:].max(),ky_global[:].min(),ky_global[:].max()], vmin=vmin,vmax=vmax, origin='lower')
 
     ax.contour(kz_global,ky_global,gamma_r,levels=contour_levels,colors='k')
@@ -61,7 +62,7 @@ for i,r in enumerate(runs):
 
 cax = fig.add_axes([xgutter+4*(w+pad), 0.55,0.02,h])
 cbar = fig.colorbar(c, cax=cax)
-cbar.set_label(r'$\gamma$')
+cbar.set_label(r'$\gamma/S$')
 
 # asymptotics
 xloc = grid[0].get_position().x0
@@ -87,23 +88,23 @@ q = 0.75
 ky_an = np.linspace(0.0,0.2,400)
 omega, omega2 = calc_asymptotic_growth(R, q, B, d, ky_an, kz)
 
-asymp_ax.plot(ky,gamma_r[:,index], label='numerical')
-asymp_ax.plot(ky_an, omega, label='asymptotic')
-asymp_sub.plot(ky,-(gamma_r[:,index] + 1j*gamma_i[:,index])**2, label='numerical')
[…]

### 2019-08-22 paper/supplemental.tex

@@
 \documentclass[aps,prl,preprint
-,superscriptaddress]{revtex4-1}\newcommand{\SSC}{S/S_{c}}
+,superscriptaddress]{revtex4-1}
+\newcommand{\SSC}{S/S_{c}}
 \newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
@@
 
-\section{Table of Runs}
-\label{sec:runs}
+\section{Numerical Methods}
+\label{sec:methods}
+
+For both the ideal and non-ideal MHD equations, we solve the eigenvalue problem in the $x$ direction using the EigenValueProblem mode in \emph{Dedalus} for a grid of $N_y \times N_z$ modes in the $y$ and $z$ directions.
+For most of our runs, we use a targeted, sparse eigenvalue solver solving for the 15 modes closest to a guess for the maximum growth rate.
+We use the ideal 2D growth rate as input for the smallest $k_y > 0$ mode at each $k_z$ and then use the output from each previous $k_y$ as an input guess for the next mode.
+Our solver is parallelized over the $k_y, k_z$ modes.
+We have confirmed that dense solvers retrieve identical results.
+For the spectrum in figure 1 of the main text, we used a dense eigenvalue solver.
+
+In order to ensure our results are converged, we have repeated runs at $n_x=256$ Chebyshev modes as well as doubling the number of modes in $y$ and $z$.
+
+\begin{tabular}{cccccccccccccc}
+\textbf{Run} & \textbf{$\SSC$} & \textbf{$\nu$} & \textbf{$\eta$} & \textbf{$q$} & \textbf{$N_x$} & \textbf{$N_{k_y}$} & \textbf{$N_{k_z}$} & \textbf{Sparse/Dense}& \\
+1  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & resolution study\\
+2  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 256 & 200 & 200 & sparse & \\
+3  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & dense  & \\
+4  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 512 & 512 & sparse & \\
+5  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 100 & 100 & sparse & \\
+6  &   0.2  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & $\SSC$ variation\\
+7  &   0.3  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+8  &   0.4  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+9  &   0.5  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+10 &   0.64 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+11 & 1.020  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+12 & 1.44   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+13 & 1.75   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+14 & 1.891  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+15 & 2.     & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+16 & 2.01   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+17 & 2.015  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+18 & 2.031  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+19 & 2.05   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+20 & 2.1    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+21 & 2.25   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+22 & 2.5    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+23 & 4      & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+24 & 1.02   & $10^{-6}$ & $10^{-6}$ & 0.75 & 128 & 200 & 200 & sparse & Reynolds number study\\
+25 & 1.02   & $10^{-4}$ & $10^{-4}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+26 & 1.02   & $10^{-3}$ & $10^{-3}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+27 & 1.02   & $10^{-2}$ & $10^{-2}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+28 & 1.02   & $10^{-5}$ & $10^{-5}$ & 0.1  & 128 & 200 & 200 & sparse & Solar Rossby\\ 
+29 & 1.02   & $10^{-6}$ & $10^{-2}$ & 0.1  & 128 & 200 & 200 & sparse & Liquid metal case\\ 
+\end{tabular}
+%\label{simulations}
 

### 2019-08-22 paper/paper.tex
Many changes to text
* added discussion and figures for q=0.1 and Pm=1e-5 cases
* corrected "negative definite" --> "positive definite" to reflect
  sign change in equation
* added discussion and figure for eigenvectors

@@
+% PRL Length limits
+% Displayed Math 	The word equivalent for displayed math is 16 words per row for single-column equations. Two-column equations count as 32 words per row.
+% Figures 	To estimate the word equivalent for figures use the figure’s aspect ratio (width / height).
+The estimate is [(150 / aspect ratio) + 20 words] for single-column figures, and [300 / (0.5 * aspect ratio)] + 40 words for double-column figures.
+
+% Fig 1: A = w/h = 2. Double column, words = 340
+% Fig 2: A = 4/3. Single column, words = 133
+% Fig 3: A = 1. Single column, words = 170
+% Fig 4: A = 0.5 Single column, words = 320
+% equations: 8 * 16 = 128
+
+% current word count: 1780 + 128 + 340 + 133 + 170 + 320
+% limit: 3750
 % \documentclass[aps,prl,preprint,groupedaddress]{revtex4-1}
@@
 \newcommand{\SSC}{S/S_{c}}
+\newcommand{\Prm}{\mbox{\textit{Pm}}}
 \newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
@@
 
-\begin{abstract}Ú
+\begin{abstract}
 The magnetorotational instability (MRI) occurs when a weak magnetic field destabilizes a rotating electrically conducting fluid with an inwardly increasing angular velocity shear.
@@
 For each $(k_y, k_z)$ pair, we solve using $n_x = 128$ modes; all our results are identical at double the resolution\footnote{All code used in this paper can be found at \protect\url{https://github.com/jsoishi/mri\_prefers\_3d}}.
+\begin{figure*}[ht]
+  \includegraphics[width=\textwidth]{fig_1.pdf}
+  \caption{Growth rates for three-dimensional MRI modes.
+a) growth rates $\gamma$ over a grid of $k_y$ and $k_z$ for four values of supercriticality $\SSC$.
+The gray contour highlights $\gamma = 0$; the red dots indicate the fastest growing mode.
+At $\SSC = 0.640$, there are no two dimensional modes.
+b) Growth rate vs $k_y$ for $k_z = ???$ at $\SSC = 1.002$ (highlighted by the blue line in a).
+The orange line gives the asymptotic result from equation~(\ref{eq:asymp}) and the blue line are the numerical results.
+The asymptotic form is valid as $\SSC \to 1$ and $k_y \ll 1$.
+c) The full discrete spectrum of the MRI for $(k_y, k_z) \simeq (0.263, 0.447)$.
+The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray.}
+  \label{fig:growth_rate}
+\end{figure*}
 
@@
 The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
-The second term is negative definite and leads ultraviolet divergence in the absence of higher-order effects.
+The second term is positive definite: it \emph{always} leads to enhanced growth rates when $k_y \neq 0$.
+Of course, it ultimately leads to ultraviolet divergence in the absence of higher-order effects.
 
@@
 The other stable modes (gray dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
-%\onecolumngrid
+
+It is useful to examine the eigenvectors for these three dimensional modes.
+Because we are only interested in the \emph{unstable} modes, which have $\omega = 0$, critical layers do not form.
+We therefore do not need viscosity or resistivity to regularize the solutions.
+In order to test that our non-ideal solutions are sufficiently close to the ideal case we are interested in, we reformulated equations~(\ref{eq:mhd}--~(\ref{eq:divu})) without viscous and resistive terms and solved the corresponding eigenvalue problem.
+The ideal MHD system is second order in $x$ and thus only two boundary conditions can be set.
+We choose impenetrable ($v_x = 0$) conditions on both walls in this case.
+Figure~\ref{fig:eigvec} shows the eigenvectors for the most unstable mode at $\SSC = 1.02$ using ideal MHD. 
+The tilted structures of $v_x$ and $v_y$ as well as $b_x$ and $b_y$, show that the eigenmodes have non-trivial Reynolds and Maxwell stresses.
+The eigenfunctions are indistinguishable from those at finite resistivity and viscosity, lending additional confidence that our other results are sufficiently in the ideal regime.
+
 %
-\begin{figure*}[h!]
-  \includegraphics[width=\textwidth]{fig_1.pdf}
-  \caption{Growth rates for three-dimensional MRI modes.
-a) growth rates $\gamma$ over a grid of $k_y$ and $k_z$ for four values of supercriticality $\SSC$.
-The white contour highlights $\gamma = 0$; the red dots indicate the fastest growing mode.
-At $\SSC = 0.640$, there are no two dimensional modes.
-b) Growth rate vs $k_y$ for $k_z = ???$ at $\SSC = 1.002$ (highlighted by the blue line in a).
-The orange line gives the asymptotic result from equation~(\ref{eq:asymp}) and the blue line are the numerical results.
-The asymptotic form is valid as $\SSC \to 1$ and $k_y \ll 1$.
-c) The full discrete spectrum of the MRI for $(k_y, k_z) \simeq (0.263, 0.447)$.
-The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray.}
-  \label{fig:growth_rate}
-\end{figure*}
-%\twocolumngrid
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\columnwidth]{eigvecs_xy_run_11_ideal_single_mode.pdf}
+  \caption{Eigenvectors of the velocity and magnetic field perturbations for the most unstable mode when $\SSC = 1.02$.
+This calculation is \emph{ideal}: $\eta = \nu = 0$; calculations at $\eta = \nu = 10^{-5}$ are indistinguishable.
+The top row shows $v_x$, $v_y$, $v_z$ (red is negative, blue positive); the bottom row shows $b_x$, $b_y$, $b_z$ (purple is negative, orange positive).
+As these are linear calculations, the amplitudes are arbitrary.}
+  \label{fig:eigvec}
+\end{figure}
 %
-Numerical simulations consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
+
+Numerical simulations aimed at understanding the MRI in accretion disks consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
 These ``channel modes'' are exact non-linear solutions for $x$-shearing-periodic domains, and can only saturate via parasitic shear instabilities \citep{1994ApJ...432..213G}.
@@
 Therefore significant three dimensionality likely implies important non-normal behavior near onset.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\columnwidth]{run_47_output_growthrates.pdf}\\
+  \includegraphics[width=\columnwidth]{run_49_output_growthrates.pdf}
+  \caption{(top) Growth rates for q=0.1. (bottom) Growth rates for $\nu=10^{-6}$, $\eta = 0.1$.}
+  \label{fig:other_params}
+\end{figure}
+Finally, we vary two important parameters:
+$q$ and the magnetic Prandtl number $\Prm = \nu/\eta$.
+Figure~\ref{fig:other_params} shows the growth rates for $\SSC = 1.02$.
+The left panel shows $q = 0.1$ holding all other parameters equal to their fiducial values; the right panel shows $\Prm = 10^{-5}$ with all other parameters equal to their fiducial.
+The former is a Rossby number relevant to the Sun, while the latter is $\Prm$ close to those found in liquid metal experiments.
+For the liquid metal run, we choose $\nu = 10^{-6}$, implying a Reynolds number of $10^6$.
+At $\Prm = 10^{-5}$, the MRI is just supercritical; this is very similar to the experimental conditions of the Princeton MRI experiment \citep{2002JFM...462..365G}.
+Importantly, for both cases, the MRI remains dominated by three dimensional modes.
 

### 2019-08-22 paper/paper.tex
Changes to text
* added references
* fixed growth rates to be scaled by S
* removed commented out junk
* minor other changes

@@
 
-% current word count: 1780 + 128 + 340 + 133 + 170 + 320
+% current word count: 2057 + 128 + 340 + 133 + 170 + 320
 % limit: 3750
-% \documentclass[aps,prl,preprint,groupedaddress]{revtex4-1}
-%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-1}
 \documentclass[aps,prl,reprint,superscriptaddress]{revtex4-1}
@@
 
-% You should use BibTeX and apsrev.bst for references
-% Choosing a journal automatically selects the correct APS
-% BibTeX style file (bst file), so only uncomment the line
-% below if necessary.
-%\bibliographystyle{apsrev4-1}
-
 \begin{document}
 
-% Use the \preprint command to place your local institutional report
-% number in the upper righthand corner of the title page in preprint mode.
-% Multiple \preprint commands are allowed.
-% Use the 'preprintnumbers' class option to override journal defaults
-% to display numbers if necessary
-%\preprint{}
-
-%Title of paper
 \title{The Magnetorotational Instability Prefers Three Dimensions}
 
-% repeat the \author .. \affiliation  etc. as needed
-% \email, \thanks, \homepage, \altaffiliation all apply to the current
-% author. Explanatory text should go in the []'s, actual e-mail
-% address or url should go in the {}'s for \email and \homepage.
-% Please use the appropriate macro foreach each type of information
-
-% \affiliation command applies to all authors since the last
-% \affiliation command. The \affiliation command should follow the
-% other information
-% \affiliation can be followed by \email, \homepage, \thanks as well.
 \author{Jeffrey~S.~Oishi}
@@
 
-% 
-%\homepage[]{Your web page}
-%\thanks{}
-%\altaffiliation{}
-
-%Collaboration name if desired (requires use of superscriptaddress
-%option in \documentclass). \noaffiliation is required (may also be
-%used with the \author command).
-%\collaboration can be followed by \email, \homepage, \thanks as well.
-%\collaboration{}
-%\noaffiliation
-
 \date{\today}
@@
 Finally, three-dimensional growth suggest the possibility of direct dynamo action driven from the linear instability itself.
+We demonstrate that three dimensional modes dominate the dynamics for shear profiles relevant to the Solar interior and at the very low magnetic Prandtl numbers relevant to liquid metal laboratory experiments.
 \end{abstract}
 
-% insert suggested PACS numbers in braces on next line
 \pacs{}
-% insert suggested keywords - APS authors don't need to do this
-%\keywords{}
-
-%\maketitle must follow title, authors, abstract, \pacs, and \keywords
 \maketitle
 
-% body of paper here - Use proper section commands
-% References should be done using the \cite, \ref, and \label commands
-
-The magnetorotational instability (MRI) is an extremely important in astrophysical fluid dynamics.
-By changing the stabiliity criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient, a weak magnetic field allows the Keplerian velocity profile due to a central point mass to drive turbulence \citep[e.g.][]{1998RvMP...70....1B}.
+The magnetorotational instability (MRI) is extremely important in astrophysical fluid dynamics.
+By changing the stability criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient, a weak magnetic field allows the Keplerian velocity profile due to a central point mass to drive turbulence \citep[e.g.][]{1998RvMP...70....1B,2010RSPTA.368.1607J}.
 This discovery explained the ubiquitous accretion onto compact objects at rates compatible with observations, and may also influence the formation of planets \citep[e.g.][]{2007Natur.448.1022J}.
@@
 However, stars and liquid metal Taylor-Couette experiments have differential rotation profiles driven by much weaker stresses.
-Where the MRI is active in these flows, it can saturate by pushing the background shear close to critical \citep{2017ApJ...841....1C,2017ApJ...841....2C}, analogous to convection mixing entropy.
+Where the MRI is active in these flows, it can saturate by pushing the background shear close to critical \citep{2015RSPSA.47140699V,2017ApJ...841....1C,2017ApJ...841....2C}, analogous to convection mixing entropy.
 Stellar interiors have extremely high fluid and magnetic Reynolds numbers, but may potentially operate at or near the critical shear rate for the MRI.
@@
 These results strongly suggest that the non-normality of the linear operator driving the MRI are always important, even when axisymmetric modes dominate.
-% \begin{itemize}
-% \item 3D $\to$ dynamo
-% \item stars, NSSL
-% \item experiments
-% \item Squire's theorem violation? Check MHD Squire, check S instead of Re...
-% \item Do eigenvalue solution for B growth? i.e.\ check if it is a dynamo?
-% \end{itemize}
 
@@
 At $\SSC = 0.640$, there are no two dimensional modes.
-b) Growth rate vs $k_y$ for $k_z = ???$ at $\SSC = 1.002$ (highlighted by the blue line in a).
+b) Growth rate vs $k_y$ for $k_z = 0.259$ at $\SSC = 1.002$ (highlighted by the blue line in a).
 The orange line gives the asymptotic result from equation~(\ref{eq:asymp}) and the blue line are the numerical results.
@@
 Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
-At $\SSC = 1.02$, the maximum growth rate ($\sigma = ??$) happens at $(k_y, k_z) \simeq (0.263, 0.447)$.
+At $\SSC = 1.02$, the maximum growth rate ($\gamma/S = 0.
+093$) happens at $(k_y, k_z) \simeq (0.263, 0.447)$.
 The first panel of figure~\ref{fig:growth_rate}a shows $\SSC = 0.640$, which contains only three-dimensional instability.
@@
 
-While surprising, preferred three-dimensional modes can be predicted analytically.
+While surprising, the MRI's preference for three-dimensional modes can be predicted analytically.
 We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) with the shear asymptotically close to the critical two-dimensional rate, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $S = S_c - \epsilon \sqrt{R}$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$,
-%\begin{widetext}
 \begin{equation}
@@
 \end{equation}
-%\end{widetext}
 The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
@@
 
-%
 \begin{figure}[h!]
@@
 \end{figure}
-%
 
@@
 Thus, we predict axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
-%
 \begin{figure}[h!]
@@
 Therefore significant three dimensionality likely implies important non-normal behavior near onset.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=\columnwidth]{run_47_output_growthrates.pdf}\\
-  \includegraphics[width=\columnwidth]{run_49_output_growthrates.pdf}
-  \caption{(top) Growth rates for q=0.1. (bottom) Growth rates for $\nu=10^{-6}$, $\eta = 0.1$.}
-  \label{fig:other_params}
-\end{figure}
-Finally, we vary two important parameters:
-$q$ and the magnetic Prandtl number $\Prm = \nu/\eta$.
-Figure~\ref{fig:other_params} shows the growth rates for $\SSC = 1.02$.
-The left panel shows $q = 0.1$ holding all other parameters equal to their fiducial values; the right panel shows $\Prm = 10^{-5}$ with all other parameters equal to their fiducial.
-The former is a Rossby number relevant to the Sun, while the latter is $\Prm$ close to those found in liquid metal experiments.
-For the liquid metal run, we choose $\nu = 10^{-6}$, implying a Reynolds number of $10^6$.
-At $\Prm = 10^{-5}$, the MRI is just supercritical; this is very similar to the experimental conditions of the Princeton MRI experiment \citep{2002JFM...462..365G}.
-Importantly, for both cases, the MRI remains dominated by three dimensional modes.
 
@@
 
-% \begin{itemize}
-% \item non-normal and $k_y$ even when $\phi = 0$ implies important role for linear dynamics at early times. 
-% \item Following Squire + Bhattacharjee, this suggests that important stellar layers may in fact be driven by linear dynamics
-% \item why is MRI non-oscillatory when hydro shear is oscil or stable?
-% \end{itemize}
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\columnwidth]{run_47_output_growthrates.pdf}\\
+  \includegraphics[width=\columnwidth]{run_49_output_growthrates.pdf}
+  \caption{(top) Growth rates for q=0.1 relevant to the Solar interior. (bottom) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta = 0.1$.}
+  \label{fig:other_params}
+\end{figure}
+Finally, we vary two important parameters:
+$q$ and the magnetic Prandtl number $\Prm = \nu/\eta$.
+Figure~\ref{fig:other_params} shows the growth rates for $\SSC = 1.02$.
+The upper panel shows $q = 0.1$ holding all other parameters equal to their fiducial values.
+This corresponds a Rossby number relevant to the Sun, and it shows very similar growth rates to the Keplerian cases presented above, $\gamma_{max}/S \simeq 0.1$.
+Again, MRI remains dominated by three dimensional modes near onset.
+The lower panel shows $\nu = 10^{-6}$ and $\eta = 0.1$ with all other parameters equal to their fiducial. 
+This is $\Prm = 10^{-5}$ close to those found in liquid metal experiments.
+At this $\Prm$, the MRI is only slightly supercritical; this is very similar to the experimental conditions of the Princeton MRI experiment \citep{2002JFM...462..365G}.
+Our results are for the rotating Plane Couette geometry\footnote{This is equivalent to Taylor-Couette flow in the limit that the gap between cylinders $R_2 - R_1 \to 0$.} with highly idealized boundary conditions.
+Nevertheless, it is very suggestive that near onset in both shear and resistivity at very low $\Prm$, the MRI is \emph{only} unstable to three dimensional modes.
 
-%If in two-column mode, this environment will change to single-column
-% format so that long equations can be displayed. Use
-% sparingly.
-%\begin{widetext}
-% put long equation here
-%\end{widetext}
+Our results show that that whenever the MRI is near its critical shear values,  three-dimensional modes grow faster than their better-studied two dimensional counterparts.
+While we have focused on the Keplerian case ($q=0.75$), we have also demonstrated that the effect is robust and occurs at Rossby numbers $q=0.1$ relevant to the Sun.
+There are several important future directions this work suggests.
+First, the Sun possesses two internal shear layers, the tachocline and the near surface shear layer.
+The latter in particular, though unstable to convection, may also host slower MRI-driven dynamics.
+In order to further elucidate this processes, understanding the non-linear saturation of the linear instability presented here is crucial, as is its robustness to perturbations driven by the fast convective dynamics.
+Second, our results at low $\Prm$ suggest that three-dimensional MRI modes may be the easiest to excite in liquid metal experiments.
+Followup work including more complex boundary conditions and full, cylindrical Taylor-Couette geometries are required to determine if experiments should consider the signatures of these non-axisymmetric modes.
+Our prior work on the axisymmetric MRI \citep{2017ApJ...841....1C,2017ApJ...841....2C} shows that differences in linear dynamics between rotating Plane Couette and Taylor-Couette geometries are small in that case; whether or not such a result carries over to non-axisymmetric modes remains to be seen.
 
-% figures should be put into the text as floats.
-% Use the graphics or graphicx packages (distributed with LaTeX2e)
-% and the \includegraphics macro defined in those packages.
-% See the LaTeX Graphics Companion by Michel Goosens, Sebastian Rahtz,
-% and Frank Mittelbach for instance.
-%
-% Here is an example of the general form of a figure:
-% Fill in the caption in the braces of the \caption{} command. Put the label
-% that you will use with \ref{} command in the braces of the \label{} command.
-% Use the figure* environment if the figure should span across the
-% entire page. There is no need to do explicit centering.
-
-% \begin{figure}
-% \includegraphics{}%
-% \caption{\label{}}
-% \end{figure}
-
-% Surround figure environment with turnpage environment for landscape
-% figure
-% \begin{turnpage}
-% \begin{figure}
-% \includegraphics{}%
-% \caption{\label{}}
-% \end{figure}
-% \end{turnpage}
-
-% tables should appear as floats within the text
-%
-% Here is an example of the general form of a table:
-% Fill in the caption in the braces of the \caption{} command. Put the label
-% that you will use with \ref{} command in the braces of the \label{} command.
-% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
-% \begin{tabular}{} command.
-% The ruledtabular enviroment adds doubled rules to table and sets a
-% reasonable default table settings.
-% Use the table* environment to get a full-width table in two-column
-% Add \usepackage{longtable} and the longtable (or longtable*}
-% environment for nicely formatted long tables. Or use the the [H]
-% placement option to break a long table (with less control than 
-% in longtable).
-% \begin{table}%[H] add [H] placement to break table across pages
-% \caption{\label{}}
-% \begin{ruledtabular}
-% \begin{tabular}{}
-% Lines of table here ending with \\
-% \end{tabular}
-% \end{ruledtabular}
-% \end{table}
-
-% Surround table environment with turnpage environment for landscape
-% table
-% \begin{turnpage}
-% \begin{table}
-% \caption{\label{}}
-% \begin{ruledtabular}
-% \begin{tabular}{}
-% \end{tabular}
-% \end{ruledtabular}
-% \end{table}
-% \end{turnpage}
-
-% Specify following sections are appendices. Use \appendix* if there
-% only one appendix.
-%\appendix
-%\section{}
-
-% If you have acknowledgments, this puts in the proper section head.
 \begin{acknowledgments}
@@
 
-% Create the reference section using BibTeX:
 \bibliography{mri}
@@
 \end{document}
-%
-% ****** End of file apstemplate.tex ******

### 2019-08-22 python/mri.py
Added option to start with a dense solve at kz = 0.
* this is good for runs that are very far from ideal conditions
* it's very expensive at high resolution, though!

@@ -74,6 +74,12 @@ try:
 except configparser.NoOptionError:
     ky_start_guess = False
 
+try:
+    ky_start_dense = config.getboolean('parameters','ky_start_dense')
+    dense_threshold = config.getfloat('solver','dense_threshold')
+except configparser.NoOptionError:
+    ky_start_dense = False
+
 kzmin = config.getfloat('parameters','kzmin')
 kzmax = config.getfloat('parameters','kzmax')
 Nkz = config.getint('parameters','Nkz')
@@ -181,7 +187,7 @@ def growth_rate(ky,kz,target,N=15, dense=False):
         else:
             solver.solve_sparse(solver.pencils[0], N=N, target=target, rebuild_coeffs=True)
     except:
-        logger.info("Solver failed for (ky, kz) = (%f, %f)"%(ky, kz))
+        logger.warning("Solver failed for (ky, kz) = (%f, %f)"%(ky, kz))
         solver_failed = True
 
     if solver_failed:
@@ -228,6 +234,9 @@ t1 = time.time()
 for k, kz in enumerate(kz_local):
     if ky_start_guess:
         soln = growth_rate(ky_global[0], kz, ky_start_guess, N=Nmodes, dense=dense)
+    elif ky_start_dense:
+        logger.info("Using dense solve at ky=0 for initial guess.")
+        soln = growth_rate(0., kz, ideal_2D(kz), dense=True)
     else:
         soln = growth_rate(0., kz, ideal_2D(kz), N=Nmodes, dense=dense)
     gamma_local[0,k] = soln[0]
@@ -263,7 +272,7 @@ if CW.rank == 0:
         dset_evec = output_file.create_dataset('eigvec',data=eigvec_global)
         dset_x = output_file.create_dataset('x', data=x_basis.grid())
 
-find_max = True
+find_max = False
 if find_max:
     if CW.rank == 0:
         max_growth_guess = gamma_global.max()

### 2019-08-22 python/plot_fig1.py
Print useful info for text when making fig 1

@@ -43,6 +43,7 @@ for i,r in enumerate(runs):
     contour_levels = np.linspace(0,gamma_r.max(),5)
     vmin = 0.
     vmax = 0.1
+    print("max gamma_r/S = {}".format(gamma_r.max()))
     c = ax.imshow(gamma_r, extent=[kz_global[:].min(),kz_global[:].max(),ky_global[:].min(),ky_global[:].max()], vmin=vmin,vmax=vmax, origin='lower')
 
     ax.contour(kz_global,ky_global,gamma_r,levels=contour_levels,colors='k')
@@ -80,6 +81,7 @@ kz_global    = data['kz']
 
 ky = ky_global[:]
 kz = kz_global[index]
+print("kx for asymptotic comparison: {}".format(kz))
 grid[1].axvline(kz, ymin=0, ymax=0.2, alpha=0.7)
 R = 1.001
 B = 1

### 2019-08-22 python/plot_phi.py
Corrected legend in phi vs SSC figure

@@ -47,7 +47,7 @@ dk = np.array(dk)
 
 fig, ax = plt.subplots()
 c = ax.scatter(ssc, phi, marker='o', c=growth, yunits=radians, zorder=2)#,edgecolor='k',linewidth=0.8)
-fig.colorbar(c, ax = ax, label='$\gamma$')
+fig.colorbar(c, ax = ax, label='$\gamma/S$')
 ax.set_xlabel(r"$S/S_c$")
 ax.set_ylabel(r"$\phi$")
 ax.set_yticklabels([0, "$\pi/8$", "$\pi/4$"])

### 2019-08-22 python/plot_fig1.py
Move asymptotic figure labels to reflect scaling by q

@@ -81,7 +81,7 @@ kz_global    = data['kz']
 
 ky = ky_global[:]
 kz = kz_global[index]
-print("kx for asymptotic comparison: {}".format(kz))
+print("kx for asymptotic comparison: {}".format(kz));
 grid[1].axvline(kz, ymin=0, ymax=0.2, alpha=0.7)
 R = 1.001
 B = 1
@@ -96,8 +96,8 @@ asymp_sub.plot(ky,-(gamma_r[:,index]/q + 1j*gamma_i[:,index]/q)**2, label='numer
 asymp_sub.plot(ky_an, omega2/q**2, label='asymptotic')
 
 #asymp_ax.legend(loc='upper right')
-asymp_ax.text(0.125, 0.048, 'numerical', fontsize=18)
-asymp_ax.text(0.035, 0.03, 'asymptotic', fontsize=18)
+asymp_ax.text(0.125, 0.048/q, 'numerical', fontsize=18)
+asymp_ax.text(0.035, 0.03/q, 'asymptotic', fontsize=18)
 asymp_ax.set_ylim(0, 0.06/q)
 asymp_ax.set_xlim(0,0.2)
 asymp_ax.set_xlabel(r'$k_y$')

### 2019-08-22 python/plot_fig1.py
incorporate phi vs ssc into fig 1

@@ -4,7 +4,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 from mpl_toolkits.axes_grid1 import AxesGrid
 from calc_asymptotic_approx import calc_asymptotic_growth
-
+from basic_units import radians
 plt.style.use('prl')
 
 # params
@@ -61,15 +61,15 @@ for i,r in enumerate(runs):
         
     datafile.close()
 
-cax = fig.add_axes([xgutter+4*(w+pad), 0.55,0.02,h])
+cax = fig.add_axes([xgutter+4*(w+pad), 0.55,0.015,h])
 cbar = fig.colorbar(c, cax=cax)
 cbar.set_label(r'$\gamma/S$')
 
 # asymptotics
 xloc = grid[0].get_position().x0
-w_lower = 2*w + pad
+w_lower = 1.2*w + pad
 asymp_ax = fig.add_axes([0.1,0.1,w_lower,0.35])
-asymp_sub = asymp_ax.inset_axes([0.6,0.25,0.3,0.3*w_lower/0.35])
+asymp_sub = asymp_ax.inset_axes([0.7,0.25,0.25,0.3*w_lower/0.35])
 
 filename = 'data/run_41_output.h5'
 data = h5py.File(filename, "r")
@@ -96,14 +96,15 @@ asymp_sub.plot(ky,-(gamma_r[:,index]/q + 1j*gamma_i[:,index]/q)**2, label='numer
 asymp_sub.plot(ky_an, omega2/q**2, label='asymptotic')
 
 #asymp_ax.legend(loc='upper right')
-asymp_ax.text(0.125, 0.048/q, 'numerical', fontsize=18)
-asymp_ax.text(0.035, 0.03/q, 'asymptotic', fontsize=18)
+asymp_ax.text(0.11, 0.058, 'numerical', fontsize=18)
+asymp_ax.text(0.03, 0.07, 'asymptotic', fontsize=18)
 asymp_ax.set_ylim(0, 0.06/q)
 asymp_ax.set_xlim(0,0.2)
 asymp_ax.set_xlabel(r'$k_y$')
 asymp_ax.set_ylabel(r'$\gamma/S$')
 asymp_ax.xaxis.set_major_locator(plt.MultipleLocator(0.05))
 asymp_sub.set_ylim(-0.005/q**2, 0.005/q**2)
+asymp_sub.set_xticks([0.045,0.08])
 asymp_sub.set_xlim(0.03,0.09)
[…]

### 2019-08-22 python/plot_phi.py

@@ -14,118 +14,123 @@ plt.style.use('prl')
 
 use_opt = False
 
-basedir = Path("data/")
-runs = [34, 33, 31, 21, 15, 11, 12, 16, 30, 13, 27, 28, 29, 26, 24, 17, 14, 18]
-R = [np.sqrt(0.2), np.sqrt(0.3), np.sqrt(0.4), np.sqrt(0.5), 0.8, 1.01, 1.2, np.sqrt(1.75), 1.375, np.sqrt(2.0), np.sqrt(2.01), np.sqrt(2.015), 1.425, np.sqrt(2.05), np.sqrt(2.1), 1.5, np.sqrt(2.5), 2.0]
-ssc = np.array(R)**2
-phi = []
+def calc_phi():
+    basedir = Path("data/")
+    runs = [34, 33, 31, 21, 15, 11, 12, 16, 30, 13, 27, 28, 29, 26, 24, 17, 14, 18]
+    R = [np.sqrt(0.2), np.sqrt(0.3), np.sqrt(0.4), np.sqrt(0.5), 0.8, 1.01, 1.2, np.sqrt(1.75), 1.375, np.sqrt(2.0), np.sqrt(2.01), np.sqrt(2.015), 1.425, np.sqrt(2.05), np.sqrt(2.1), 1.5, np.sqrt(2.5), 2.0]
+    ssc = np.array(R)**2
+    phi = []
 
-growth = []
-dk = []
-for run in runs:
-    dfname = Path("run_{:d}_output.h5".format(run))
-    df = h5py.File(basedir/dfname, "r")
+    growth = []
+    dk = []
+    for run in runs:
+        dfname = Path("run_{:d}_output.h5".format(run))
+        df = h5py.File(basedir/dfname, "r")
+        if use_opt:
+            #phi.append(np.arctan2(df['gamma'].attrs['max ky'], df['gamma'].attrs['max kz']))
+            ky = df['gamma'].attrs['max ky']
+            kx = df['gamma'].attrs['max kz']
+            gamma_max = df['gamma'].attrs['max growth rate']
+        else:
+            gamma_global = df['gamma'][:,:]
+            max_where = np.unravel_index(gamma_global.argmax(), gamma_global.shape)
+            ky = df['ky'][max_where[0]] 
+            kz = df['kz'][max_where[1]]
[…]

### 2019-08-22 paper/paper.tex
Added phi vs SSC to figure 1, use mean field alpha plot instead

@@
 \begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{phi_vs_ssc_grid.pdf}
-  \caption{Angle $\phi$ of mode with respect to $z$ axis vs $\SSC$.
-The inset shows the transition from axisymmetric modes to three dimensional modes as $\SSC$ decreases below $2.05$.
-The shaded region indicates stability to all perturbations; the stability limit for three dimensional modes is $\SSC \simeq 0.102$.}
+  \includegraphics[width=\columnwidth]{{mean_field_alpha_run_11_output.h5}.pdf}
+  \caption{Mean field $\alpha$ coefficient.}
   \label{fig:phi}

### 2019-08-22 paper/paper.tex
more edits

@@
 Here, we show that the fastest growing unstable modes of an ideal magnetofluid are three dimensional provided the shear rate, $S$, is close to the two-dimensional onset value, $S_c$.
-Three-dimensional modes are unstable down to $S \approx 0.10 S_c$, and dominate the two-dimensional modes until $S \approx 2.05 S_c$.
-Significant numbers of fast three dimensional modes remain well past $2.05 S_{c}$. 
+For a Keplerian shear, three-dimensional modes are unstable down to $S \approx 0.10 S_c$, and dominate the two-dimensional modes until $S \approx 2.05 S_c$.
+Significant numbers of fast three-dimensional modes remain well past $2.05 S_{c}$. 
 These finding are significant in three ways. 
@@
 Finally, three-dimensional growth suggest the possibility of direct dynamo action driven from the linear instability itself.
-We demonstrate that three dimensional modes dominate the dynamics for shear profiles relevant to the Solar interior and at the very low magnetic Prandtl numbers relevant to liquid metal laboratory experiments.
+We also demonstrate that three-dimensional modes dominate for shear profiles relevant to stellar interiors and at the tiny magnetic Prandtl numbers relevant to liquid metal laboratory experiments.
 \end{abstract}
@@
 The magnetorotational instability (MRI) is extremely important in astrophysical fluid dynamics.
-By changing the stability criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient, a weak magnetic field allows the Keplerian velocity profile due to a central point mass to drive turbulence \citep[e.g.][]{1998RvMP...70....1B,2010RSPTA.368.1607J}.
+By changing the stability criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient, a weak magnetic field catalyzes turbulence in a Keplerian shear \citep[e.g.][]{1998RvMP...70....1B,2010RSPTA.368.1607J}.
 This discovery explained the ubiquitous accretion onto compact objects at rates compatible with observations, and may also influence the formation of planets \citep[e.g.][]{2007Natur.448.1022J}.
@@
 
-Here, we investigate the stability of three dimensional perturbations near the two dimensional critical shear rate $S_c$ for a nearly invisicd, ideal MHD flow.
-We find that the first destabilized modes are three-dimensional, and thus can act as a dynamo even in the absence of secondary instability.
+Here, we investigate the stability of three-dimensional perturbations near the two-dimensional critical shear rate $S_c$ for a nearly invisicd, ideal MHD flow.
+We find that the first destabilized modes are three-dimensional, and thus could act as a dynamo even in the absence of secondary instability.
 These results strongly suggest that the non-normality of the linear operator driving the MRI are always important, even when axisymmetric modes dominate.
@@
 We solve the linearized magnetohydrodynamic equations in rotating plane-Couette geometry.
-This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y} = Sx$. 
-We cast the Navier-Stokes equation in a fairly standard form,
+This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y} = Sx$; \citep[see][]{2015RSPSA.47140699V}.
+We cast the Navier-Stokes equation in the form,
 \begin{equation}
@@
 The main input parameters are the Coriolis parameter, $f = 2 \Omega$; the background shear rate, $S = dV_{y}/dx < 0$;  the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0} \rho_{0} = 1$); and the horizontal domain width $-d/2 \le x \le d/2$.
-Accretion disk modelling usually considers the Rossby number $q = -S/f < 1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
-For all calculations here, we use $q = 3/4$ (Keplerian).
+Accretion-disk modelling usually considers the Rossby number $q = -S/f < 1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
+Unless otherwise stated $q = 3/4$ (Keplerian).
 We write the induction equation in terms of the $x$ component of the magnetic field,
@@
 \end{equation}
-We explicitly enforce the divergence constraint on the incompressible velocity field $v$ and the magnetic field,
+We explicitly enforce divergence-free velocity and magnetic field,
 \begin{equation}
@@
 \end{equation}
-We consider the behavior of the MRI in a vertically ($z$) and horizontally ($y$) periodic channel of width $d = \pi$.
+The spatial domain is a vertically ($z$) and horizontally ($y$) periodic channel of width $d = \pi$. 
 The boundary conditions are impenetrable stress-free and perfectly conducting; $v_{x} = \omega_{y} = \omega_{z} = b_{x} = \partial_{x}j_{x} = 0$. 
@@
 The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $x$ with Dirichlet boundary conditions.
-We pose and solve the system in the same formulation as equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework.
+We pose and solve equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework.
 
-Our interest is in ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions.
+Our main interest is in ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions.
 However, we set $\eta=\nu=10^{-5}$ to avoid critical layers in for the \textit{stable} solution branch.
-We confirmed our results are insensitive to the presence of small diffusion. 
+We confirmed our results are totally insensitive to small diffusion. 
 For each $(k_y, k_z)$ pair, we solve using $n_x = 128$ modes; all our results are identical at double the resolution\footnote{All code used in this paper can be found at \protect\url{https://github.com/jsoishi/mri\_prefers\_3d}}.
@@
 
-Our two main results are: (I) When the first two-dimensional mode becomes unstable, there already exist three-dimensional modes with positive growth rate.
+Our two main results are: 
+(I) When the first two-dimensional mode becomes unstable, there already exist three-dimensional modes with positive growth rate. 
 (II) When the criticality becomes sufficiently large, the fastest-growing mode becomes purely two-dimensional. 
@@
 
-While surprising, the MRI's preference for three-dimensional modes can be predicted analytically.
-We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) with the shear asymptotically close to the critical two-dimensional rate, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $S = S_c - \epsilon \sqrt{R}$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$,
+The MRI's preference for three-dimensional modes can be predicted analytically.
+We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC = 1 + \epsilon^{2} R$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$,
 \begin{equation}
   \label{eq:asymp}
-\sigma^{2} =  \frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left( R^{2} - \frac{d^2}{\pi^{2}} k_z^2 \right)  +  \Upsilon \, k_{y}^{2}  \right] +  \ldots,
+\sigma^{2} =  \frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left( R - \frac{d^2}{\pi^{2}} k_z^2 \right)  +  \Upsilon \, k_{y}^{2}  \right] +  \ldots,
 \end{equation}
@@
 \begin{equation}
-\Upsilon \ = \ \frac{\left(6+\pi ^2\right) q+\pi^2-6}{12 }\  \approx \ 1.31 \ \  \text{for}  \ \  q = \frac{3}{4}.
+\Upsilon \ = \ \frac{\left(\pi ^2+6\right) q+\pi^2-6}{12 }\  \approx \ 1.31 \ \  \text{for}  \ \  q = \frac{3}{4}.
 \end{equation}
 The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
-The second term is positive definite: it \emph{always} leads to enhanced growth rates when $k_y \neq 0$.
-Of course, it ultimately leads to ultraviolet divergence in the absence of higher-order effects.
+The second term is positive definite: it \emph{always} leads to enhanced growth rates when $k_y \neq 0$ and ultraviolet divergence in the absence of higher-order effects.
 
@@
 
-It is useful to examine the eigenvectors for these three dimensional modes.
-Because we are only interested in the \emph{unstable} modes, which have $\omega = 0$, critical layers do not form.
-We therefore do not need viscosity or resistivity to regularize the solutions.
-In order to test that our non-ideal solutions are sufficiently close to the ideal case we are interested in, we reformulated equations~(\ref{eq:mhd}--~(\ref{eq:divu})) without viscous and resistive terms and solved the corresponding eigenvalue problem.
-The ideal MHD system is second order in $x$ and thus only two boundary conditions can be set.
-We choose impenetrable ($v_x = 0$) conditions on both walls in this case.
+It is useful to examine the eigenvectors for the three-dimensional unstable modes.
+No critical layers can form for $\gamma \ne 0$.
+We can therefore solve equations~(\ref{eq:mhd}--~(\ref{eq:divu})) without dissipation as a 2nd-order system in $\partial_{x}$;
+we impose only $v_x = 0$ on both walls.
 Figure~\ref{fig:eigvec} shows the eigenvectors for the most unstable mode at $\SSC = 1.02$ using ideal MHD. 
-The tilted structures of $v_x$ and $v_y$ as well as $b_x$ and $b_y$, show that the eigenmodes have non-trivial Reynolds and Maxwell stresses.
-The eigenfunctions are indistinguishable from those at finite resistivity and viscosity, lending additional confidence that our other results are sufficiently in the ideal regime.
+The tilted structures of $v_x$ and $v_y$ as well as $b_x$ and $b_y$, imply non-trivial Reynolds and Maxwell stresses.
+The eigenfunctions are indistinguishable from those at finite dissipation, lending additional confidence to our other results.
 
@@
 
-Numerical simulations aimed at understanding the MRI in accretion disks consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
+Numerical simulations of the MRI in accretion disks consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
 These ``channel modes'' are exact non-linear solutions for $x$-shearing-periodic domains, and can only saturate via parasitic shear instabilities \citep{1994ApJ...432..213G}.
-We use impenetrable, stress-free boundary conditions that are more applicable to stellar interiors. 
+Impenetrable and stress-free boundary conditions are more applicable to stellar interiors. 
 But even in disks, any kind of finite radial extent will cutoff the unbounded growth of channel modes. 
@@
   \includegraphics[width=\columnwidth]{{mean_field_alpha_run_11_output.h5}.pdf}
-  \caption{Mean field $\alpha$ coefficient.}
+  \caption{Mean-field $\alpha$ coefficient.}
   \label{fig:phi}
@@
 
-The second implication of the three dimensional modes is of considerably more importance.
-These modes produce dynamo action in the linear regime, as evidenced by the presence of non-trivial eigenmodes for all three components of the magnetic field.
-This is of considerable significance, as it leads to the possibility of a laminar MRI dynamo operating directly in regions of weak shear.
-MRI dynamos have been studied in a number of different contexts \citep{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one have focused on the turbulent dynamo.
+The second implication of the three-dimensional modes is considerably more important.
+These modes will generate a mean electromotive force in linear regime, as evidenced by the presence of non-trivial eigenmodes for all three components of the magnetic field.
+This is of considerable significance, as it leads to the possibility of a laminar MRI dynamo operating directly in regions of weak shear at large Reynolds number.
+MRI dynamos have been studied in a number of different contexts \citep{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one focused on turbulent dynamos. 
 The lone exception, \citet{2016MNRAS.462..818B}, is a numerical study that discovered the exponential growth of mean magnetic fields during the linear growth phase of the MRI far from stability.
 Our work explains this result in terms of purely linear dynamics:
-non-axisymmetric MRI unstable modes drive the exponential growth of magnetic fields.
+non-axisymmetric MRI unstable modes drive the dynamo growth of magnetic fields.
 
@@
 Figure~\ref{fig:other_params} shows the growth rates for $\SSC = 1.02$.
-The upper panel shows $q = 0.1$ holding all other parameters equal to their fiducial values.
-This corresponds a Rossby number relevant to the Sun, and it shows very similar growth rates to the Keplerian cases presented above, $\gamma_{max}/S \simeq 0.1$.
-Again, MRI remains dominated by three dimensional modes near onset.
+The upper panel shows $q = 1/10$ holding all other parameters equal to their fiducial values.
+This corresponds a Rossby number relevant to stellar interiors, and it shows very behavior to the $q=3/4$ cases, $\gamma_{max}/S \simeq 0.1$.
+Again, MRI remains dominated by three-dimensional modes near onset.
 The lower panel shows $\nu = 10^{-6}$ and $\eta = 0.1$ with all other parameters equal to their fiducial. 
@@
 At this $\Prm$, the MRI is only slightly supercritical; this is very similar to the experimental conditions of the Princeton MRI experiment \citep{2002JFM...462..365G}.
-Our results are for the rotating Plane Couette geometry\footnote{This is equivalent to Taylor-Couette flow in the limit that the gap between cylinders $R_2 - R_1 \to 0$.} with highly idealized boundary conditions.
-Nevertheless, it is very suggestive that near onset in both shear and resistivity at very low $\Prm$, the MRI is \emph{only} unstable to three dimensional modes.
+Our results are for the rotating plane-Couette geometry\footnote{This is equivalent to Taylor-Couette flow in the limit that the gap between cylinders $R_2 - R_1 \to 0$.} with highly idealized boundary conditions.
+Nevertheless, it is very suggestive that near onset in both shear and resistivity at very low $\Prm$, the MRI is \emph{only} unstable to three-dimensional modes.
 
-Our results show that that whenever the MRI is near its critical shear values,  three-dimensional modes grow faster than their better-studied two dimensional counterparts.
-While we have focused on the Keplerian case ($q=0.75$), we have also demonstrated that the effect is robust and occurs at Rossby numbers $q=0.1$ relevant to the Sun.
+Our results show that that whenever the MRI is near its critical shear values,  three-dimensional modes grow faster than their better-studied two-dimensional counterparts.
+While we have focused on the Keplerian case ($q=3/4$), we also demonstrate robustness for low Rossby numbers $q=1/10$.
 There are several important future directions this work suggests.
-First, the Sun possesses two internal shear layers, the tachocline and the near surface shear layer.
-The latter in particular, though unstable to convection, may also host slower MRI-driven dynamics.
+First, the Sun possesses two internal shear layers with inwardly increasing shear, the high-latitude tachocline, and the near-surface shear layer.
+The latter in particular, though unstable to small-scale convection, may also host slower MRI-driven dynamics.
 In order to further elucidate this processes, understanding the non-linear saturation of the linear instability presented here is crucial, as is its robustness to perturbations driven by the fast convective dynamics.

### 2019-08-23 paper/supplemental.tex
supplemental

@@
 27 & 1.02   & $10^{-2}$ & $10^{-2}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-28 & 1.02   & $10^{-5}$ & $10^{-5}$ & 0.1  & 128 & 200 & 200 & sparse & Solar Rossby\\ 
+28 & 1.02   & $10^{-5}$ & $10^{-5}$ & 0.1  & 128 & 200 & 200 & sparse & Low Rossby\\ 
 29 & 1.02   & $10^{-6}$ & $10^{-2}$ & 0.1  & 128 & 200 & 200 & sparse & Liquid metal case\\ 
@@
    \Eeq
-We can then substitute everything into equation (1) to get a second-order equation for $\hat{U}(x)$ of the general form 
+Substitute everything into equation (1) to get a second-order equation for $\hat{U}(x)$ of the general form 
 \Beq
@@
 \Beq
--\psi''(x) + \Phi(x) \psi(x) \ = \ 0
+-\psi''(x) + \Phi(x) \psi(x) \ = \ 0 \label{Schr-eq}
 \Eeq
@@
 \Eeq
-
-Rescale everything in terms of a bookkeeping parameter, $\eps$.
+We want to solve the Schr\:{o}dinger-type equation~\ref{Schr-eq} with the boundary conditions 
 \Beq
-S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R) 
+\psi(x=\pm d) = 0.
 \Eeq
+We solve perturbatively near the critical values for the 2D  instability. We rescale each parameters in terms of a bookkeeping parameter, $\eps$.
 \Beq
-\omega \ \to \ \eps^{2} \omega , \quad k \ \to \ \eps k, \quad \ell \ \to \ \eps^{2} \ell 
+S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k \ \to \ \eps k, \quad \ell \ \to \ \eps^{2} \ell
 \Eeq 
@@
 \Eeq
-Going order-by-order
+For leading order 
 \Beq
-\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0 
+\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0 , \quad \implies \quad \psi_{0}(x) \ = \ \cos(\pi x / d)
 \Eeq
+The next order is
 \Beq
@@
 \Eeq
-Therefore 
-\Beq
-\psi_{0}(x) \ = \ \cos(\pi x / d)
-\Eeq
-This satisfies the equation and boundary condition. The solvability condition for $\psi_{2}(x)$ is 
+The solvability condition for $\psi_{2}(x)$ is 
 \Beq
@@
 \Beq
-\omega^{2} \ = \ \frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
+-\omega^{2} \ = \ \frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left(R-\frac{d^2}{\pi^{2}} k_z^2\right) \ + \ \frac{\left(6+\pi ^2\right) q+\pi
    ^2-6}{12 } \, k_{y}^{2}  \right] \ + \ ...

### 2019-08-23 paper/paper.tex
a couple equation errors
lots of small changes.

@@
 
-% Fig 1: A = w/h = 2. Double column, words = 340
-% Fig 2: A = 4/3. Single column, words = 133
-% Fig 3: A = 1. Single column, words = 170
-% Fig 4: A = 0.5 Single column, words = 320
-% equations: 8 * 16 = 128
+% Fig 1: A=w/h=2. Double column, words=340
+% Fig 2: A=4/3. Single column, words=133
+% Fig 3: A=1. Single column, words=170
+% Fig 4: A=0.5 Single column, words=320
+% equations: 8 * 16=128
 
@@
 \usepackage{graphicx}
+\usepackage{amsmath}
 
@@
 Here, we show that the fastest growing unstable modes of an ideal magnetofluid are three dimensional provided the shear rate, $S$, is close to the two-dimensional onset value, $S_c$.
-For a Keplerian shear, three-dimensional modes are unstable down to $S \approx 0.10 S_c$, and dominate the two-dimensional modes until $S \approx 2.05 S_c$.
-Significant numbers of fast three-dimensional modes remain well past $2.05 S_{c}$. 
+For a Keplerian shear, three-dimensional modes are unstable down to $S\approx0.10S_c$, and dominate the two-dimensional modes until $S\approx2.05S_{c}$.
+Significant numbers of fast three-dimensional modes remain well past $2.05S_{c}$. 
 These finding are significant in three ways. 
@@
 Finally, three-dimensional growth suggest the possibility of direct dynamo action driven from the linear instability itself.
-We also demonstrate that three-dimensional modes dominate for shear profiles relevant to stellar interiors and at the tiny magnetic Prandtl numbers relevant to liquid metal laboratory experiments.
+We also demonstrate that three-dimensional modes dominate for shear profiles relevant to stars and at the tiny magnetic Prandtl numbers relevant to liquid-metal laboratory experiments.
 \end{abstract}
@@
 The magnetorotational instability (MRI) is extremely important in astrophysical fluid dynamics.
-By changing the stability criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient, a weak magnetic field catalyzes turbulence in a Keplerian shear \citep[e.g.][]{1998RvMP...70....1B,2010RSPTA.368.1607J}.
+A weak magnetic field catalyzes turbulence in a Keplerian shear by changing the stability criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient \citep[e.g.][]{1998RvMP...70....1B,2010RSPTA.368.1607J}.
 This discovery explained the ubiquitous accretion onto compact objects at rates compatible with observations, and may also influence the formation of planets \citep[e.g.][]{2007Natur.448.1022J}.
@@
 However, stars and liquid metal Taylor-Couette experiments have differential rotation profiles driven by much weaker stresses.
-Where the MRI is active in these flows, it can saturate by pushing the background shear close to critical \citep{2015RSPSA.47140699V,2017ApJ...841....1C,2017ApJ...841....2C}, analogous to convection mixing entropy.
-Stellar interiors have extremely high fluid and magnetic Reynolds numbers, but may potentially operate at or near the critical shear rate for the MRI.
-This critical shear is set by the background magnetic field strength for axisymmetric perturbations in a channel of a given size.
-Despite the extensive literature on the accretion disk case of strong shear and high Reynolds numbers and the liquid metal literature on weak shear and lower Reynolds numbers, even the linear dynamics of the MRI are not well studied in the weak shear, high Reynolds number regime.
+Where the MRI is active in these flows, it saturates by pushing the background shear close to critical \citep{2015RSPSA.47140699V,2017ApJ...841....1C,2017ApJ...841....2C}, analogous to convection mixing entropy.
+Stellar interiors have extremely high fluid and magnetic Reynolds numbers, but can operate at or near the critical shear rate for the MRI.
+This finite critical shear results from a finite-channel cutoff.
+Despite the extensive literature on accretion disks (strong shear, low dissipation) and liquid-metal experiments (weak shear, strong dissipation), even the linear MRI is not well understood in the weak-shear, low-dissipation regime.
 
-Here, we investigate the stability of three-dimensional perturbations near the two-dimensional critical shear rate $S_c$ for a nearly invisicd, ideal MHD flow.
+Here, we investigate the stability of three-dimensional perturbations near the two-dimensional critical shear rate $S_{c}$ for a nearly invisicd, ideal MHD flow.
 We find that the first destabilized modes are three-dimensional, and thus could act as a dynamo even in the absence of secondary instability.
@@
 We solve the linearized magnetohydrodynamic equations in rotating plane-Couette geometry.
-This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y} = Sx$; \citep[see][]{2015RSPSA.47140699V}.
+This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y}=Sx$; \citep[see][]{2015RSPSA.47140699V}.
 We cast the Navier-Stokes equation in the form,
-\begin{equation}
-  \label{eq:mhd}
-  \frac{D \mathbf{v}}{Dt} + f \hat{z} \times \mathbf{v} + S v_x \hat{z} + \mathbf{\nabla}{p} + \nu \mathbf{\nabla} \times \mathbf{\omega} = B_{0} \partial_{z} \mathbf{b},
+\begin{equation}\label{eq:mhd}
+\frac{D \boldsymbol{v}}{Dt}+f \boldsymbol{\hat{z}}\times\boldsymbol{v}+{S}v_{x}\,\boldsymbol{\hat{y}}+\boldsymbol{\nabla}{p}+\nu\,\boldsymbol{\nabla}\times\boldsymbol{\omega}=B_{0}\partial_{z}\boldsymbol{b},
 \end{equation}
@@
 \begin{equation}
-\omega = \mathbf{\nabla} \times \mathbf{v}, \quad \text{and} \quad \frac{D}{Dt} \ = \ \partial_{t} + S x \partial_{y}
+\boldsymbol{\omega}=\boldsymbol{\nabla}\times\boldsymbol{v},\quad\text{and}\quad\frac{D}{Dt}=\partial_{t}+{S}x\,\partial_{y}\end{equation}
+The main input parameters are the Coriolis parameter, $f=2 \Omega$; the background shear rate, $S=dV_{y}/dx<0$;  the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0}\rho_{0}=1$); and the horizontal domain width $-d/2\le{x}\le d/2$.
+Accretion-disk modelling usually considers the Rossby number $q=-S/f<1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
+Unless otherwise stated $q=3/4$ (Keplerian).
+We write the induction equation in terms of the $x$-component magnetic field,
+\begin{equation}\label{eq:Bx}
+\frac{Db_{x}}{Dt}+\eta(\partial_{y}j_{z}-\partial_{z}j_{y})=B_{0}\partial_{z}v_{x}
 \end{equation}
-The main input parameters are the Coriolis parameter, $f = 2 \Omega$; the background shear rate, $S = dV_{y}/dx < 0$;  the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0} \rho_{0} = 1$); and the horizontal domain width $-d/2 \le x \le d/2$.
-Accretion-disk modelling usually considers the Rossby number $q = -S/f < 1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
-Unless otherwise stated $q = 3/4$ (Keplerian).
-We write the induction equation in terms of the $x$ component of the magnetic field,
-\begin{equation}
-  \label{eq:Bx}
-  \frac{D b_x}{Dt} - B_0 \partial_z v_x + \eta (\partial_y j_z - \partial_z j_y) = 0
-\end{equation}
-and the $x$ component of the current density ($j_{x} = \partial_{y}b_{z} - \partial_{z} b_{y}$),
-\begin{equation}
-  \label{eq:Jx}
-  \frac{D j_x}{Dt} - B_0 \partial_z \omega_x + S \partial_z B_x - \eta \nabla^2 j_x = 0
+and the $x$-component current density ($j_{x}=\partial_{y}b_{z}-\partial_{z}b_{y}$),
+\begin{equation}\label{eq:Jx}
+\frac{Dj_{x}}{Dt}+{S}\partial_{z}b_{x}-\eta\nabla^{2}j_{x}=B_{0}\partial_{z}\omega_{x}
 \end{equation}
 We explicitly enforce divergence-free velocity and magnetic field,
-\begin{equation}
-  \label{eq:divu}
-  \mathbf{\nabla} \cdot \mathbf{v} = \mathbf{\nabla} \cdot \mathbf{b} = 0.
+\begin{equation}\label{eq:divu}
+ \boldsymbol{\nabla}\cdot\boldsymbol{v}=\boldsymbol{\nabla}\cdot\boldsymbol{b}=0.
 \end{equation}
-The spatial domain is a vertically ($z$) and horizontally ($y$) periodic channel of width $d = \pi$. 
-The boundary conditions are impenetrable stress-free and perfectly conducting; $v_{x} = \omega_{y} = \omega_{z} = b_{x} = \partial_{x}j_{x} = 0$. 
+The spatial domain is a vertically ($z$) and horizontally ($y$) periodic channel of width $d=\pi$. 
+The boundary conditions are impenetrable stress-free and perfectly conducting; $v_{x}=\omega_{y}=\omega_{z}=b_{x}=\partial_{x}j_{x}=0$ at $x=\pm{d}$. 
 The MRI is a weak-field instability; in the inviscid, ideal case the critical shear rate for instability in \textit{two dimensions} is
-\begin{equation}
-  \label{eq:Sc}
-  S_c = -\frac{\pi^2 B_{0}^2}{f d^2}.
+\begin{equation}\label{eq:Sc}
+  S_{c}=-\frac{\pi^{2}B_{0}^2}{fd^2}.
 \end{equation}
@@
 
-We assume harmonic perturbations in $y$ and $z$, (e.g. pressure) $p = \hat{p}(x) e^{i(k_y y + k_z z) + \sigma t}$ . 
-We use a complex-valued growth rate $\sigma = \gamma + i\omega$ with $\gamma, \omega$ both real. 
-The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $x$ with Dirichlet boundary conditions.
-We pose and solve equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework.
+We assume harmonic perturbations in $y$ and $z$, (e.g. pressure) $p=\hat{p}(x)e^{i(k_{y}y+k_{z}z)+\sigma{t}}$. 
+We use a complex-valued growth rate $\sigma=\gamma+i\omega$ with $\gamma,\omega$ both real. 
+The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $\partial_{x}$ with Dirichlet boundary conditions.
+We pose and solve equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework \citep{The Dedalus paper}.
 
-Our main interest is in ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions.
+Our main interest is in ideal ($\eta=0$), inviscid ($\nu=0$) conditions.
 However, we set $\eta=\nu=10^{-5}$ to avoid critical layers in for the \textit{stable} solution branch.
 We confirmed our results are totally insensitive to small diffusion. 
-For each $(k_y, k_z)$ pair, we solve using $n_x = 128$ modes; all our results are identical at double the resolution\footnote{All code used in this paper can be found at \protect\url{https://github.com/jsoishi/mri\_prefers\_3d}}.
+For each $(k_{y},k_{z})$ pair, we solve using $n_{x}=128$ modes; all our results are identical at double the resolution; see \protect\url{github.com/jsoishi/mri_prefers_3d} for all code used in this paper.
 \begin{figure*}[ht]
@@
   \caption{Growth rates for three-dimensional MRI modes.
-a) growth rates $\gamma$ over a grid of $k_y$ and $k_z$ for four values of supercriticality $\SSC$.
-The gray contour highlights $\gamma = 0$; the red dots indicate the fastest growing mode.
-At $\SSC = 0.640$, there are no two dimensional modes.
-b) Growth rate vs $k_y$ for $k_z = 0.259$ at $\SSC = 1.002$ (highlighted by the blue line in a).
+a) growth rates $\gamma$ over a grid of $k_{y}$ and $k_{z}$ for four values of supercriticality $\SSC$.
+The gray contour highlights $\gamma=0$; the red dots indicate the fastest growing mode.
+At $\SSC=0.640$, there are no two dimensional modes.
+b) Growth rate vs $k_{y}$ for $k_{z}=0.259$ at $\SSC=1.002$ (highlighted by the blue line in a).
 The orange line gives the asymptotic result from equation~(\ref{eq:asymp}) and the blue line are the numerical results.
-The asymptotic form is valid as $\SSC \to 1$ and $k_y \ll 1$.
-c) The full discrete spectrum of the MRI for $(k_y, k_z) \simeq (0.263, 0.447)$.
-The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray.}
+The asymptotic form is valid as $\SSC\to1$ and $k_{y}\ll1$.) c) The full discrete spectrum of the MRI for $(k_{y},k_{z})\simeq(0.263,0.447)$.
+The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray.
+d) The wavenumber angle $\arctan(k_{y}/k_{z})$ as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes.}
   \label{fig:growth_rate}
@@
 (I) When the first two-dimensional mode becomes unstable, there already exist three-dimensional modes with positive growth rate. 
-(II) When the criticality becomes sufficiently large, the fastest-growing mode becomes purely two-dimensional. 
+(II) At sufficiently large criticality, the fastest-growing mode becomes purely two-dimensional. 
 These results are pertinent in two ways. 
-First, the MRI contains a substantial ``Goldilocks regime" with possible direct dynamo action, and this regime likely most applies to stellar interiors. 
-Second, our results accord with well-established results for accretion disks that expect primary two-dimensional linear modes.
+First, the MRI contains a substantial ``Goldilocks regime" with possible direct dynamo action, and this regime likely most applies to stellar interiors and laboratory experiments. 
+Second, our results accord with well-established results for accretion disks that expect two-dimensional primary modes.
 Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
-At $\SSC = 1.02$, the maximum growth rate ($\gamma/S = 0.
-093$) happens at $(k_y, k_z) \simeq (0.263, 0.447)$.
-The first panel of figure~\ref{fig:growth_rate}a shows $\SSC = 0.640$, which contains only three-dimensional instability.
-The overall critical shear for three-dimensional modes is $S_{c,3D} \simeq 0.102 S_c$.
+At $\SSC=1.02$, the maximum growth rate ($\gamma/|S|=0.093$) happens at $(k_{y},k_{z})\simeq(0.263,0.447)$.
+The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability.
+The overall critical shear for three-dimensional modes is $S_{c,\text{3D}}\simeq0.102S_c$.
 
 The MRI's preference for three-dimensional modes can be predicted analytically.
-We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC = 1 + \epsilon^{2} R$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$,
-\begin{equation}
-  \label{eq:asymp}
-\sigma^{2} =  \frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left( R - \frac{d^2}{\pi^{2}} k_z^2 \right)  +  \Upsilon \, k_{y}^{2}  \right] +  \ldots,
+We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC=1+\epsilon^{2}R$, $k_{z}\sim\mathcal{O}(\epsilon)$, $\sigma{\sim}k_{y}\sim\mathcal{O}(\epsilon^{2})$,
+\begin{equation}\label{eq:asymp}
+\sigma^{2}=\frac{q\,B_{0}^{2}}{(q+1)}\left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
 \end{equation}
@@
 \begin{equation}
-\Upsilon \ = \ \frac{\left(\pi ^2+6\right) q+\pi^2-6}{12 }\  \approx \ 1.31 \ \  \text{for}  \ \  q = \frac{3}{4}.
+\Upsilon=\frac{\left(\pi^{2}+6\right) q+\pi^{2}-6}{12}\approx1.31\quad\text{for}\quad{q}=\frac{3}{4}.
 \end{equation}
 The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
-The second term is positive definite: it \emph{always} leads to enhanced growth rates when $k_y \neq 0$ and ultraviolet divergence in the absence of higher-order effects.
+The second term is positive definite: it \emph{always} leads to enhanced growth rates when $k_{y}\neq0$ and ultraviolet divergence in the absence of higher-order effects.
 
-Figure~\ref{fig:growth_rate}b compares the numerical growth rates for $\SSC=1.002$ between $0 \le k_y \le 0.2$ to the asymptotic approximation, showing good agreement where the latter is valid.
-Figure~\ref{fig:growth_rate}c shows the full spectrum for $\SSC = 1.02$.
+Figure~\ref{fig:growth_rate}b compares the numerical growth rates for $\SSC=1.002$ between $0\le{k_{y}}\le0.2$ to the asymptotic approximation, showing good agreement where the latter is valid.
+Figure~\ref{fig:growth_rate}c shows the full spectrum for $\SSC=1.02$.
 The plot shows a purely growing/decaying complex-conjugate pair (orange/blue dots on the real axis) consistent with equation~(\ref{eq:asymp}).
 The other stable modes (gray dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
-
-It is useful to examine the eigenvectors for the three-dimensional unstable modes.
-No critical layers can form for $\gamma \ne 0$.
-We can therefore solve equations~(\ref{eq:mhd}--~(\ref{eq:divu})) without dissipation as a 2nd-order system in $\partial_{x}$;
-we impose only $v_x = 0$ on both walls.
-Figure~\ref{fig:eigvec} shows the eigenvectors for the most unstable mode at $\SSC = 1.02$ using ideal MHD. 
-The tilted structures of $v_x$ and $v_y$ as well as $b_x$ and $b_y$, imply non-trivial Reynolds and Maxwell stresses.
-The eigenfunctions are indistinguishable from those at finite dissipation, lending additional confidence to our other results.
 
@@
   \includegraphics[width=\columnwidth]{eigvecs_xy_run_11_ideal_single_mode.pdf}
-  \caption{Eigenvectors of the velocity and magnetic field perturbations for the most unstable mode when $\SSC = 1.02$.
-This calculation is \emph{ideal}: $\eta = \nu = 0$; calculations at $\eta = \nu = 10^{-5}$ are indistinguishable.
-The top row shows $v_x$, $v_y$, $v_z$ (red is negative, blue positive); the bottom row shows $b_x$, $b_y$, $b_z$ (purple is negative, orange positive).
-As these are linear calculations, the amplitudes are arbitrary.}
+  \caption{\textit{Ideal} Eigenvectors of the velocity and magnetic field perturbations for the most unstable mode when $\SSC=1.02$, and $\eta=\nu=0$; calculations at $\eta=\nu=10^{-5}$ are indistinguishable.
+The top row shows $v_{x}$, $v_{y}$, $v_{z}$ (red is negative, blue positive); the bottom row shows $b_{x}$, $b_{y}$, $b_{z}$ (purple is negative, orange positive).
+The amplitudes are arbitrary because of linearity.}
   \label{fig:eigvec}
@@
 We reconcile our results with earlier simulations by showing that the MRI indeed prefers two dimensions for the larger values of criticality found in disk simulations. 
-Figure~\ref{fig:phi} shows the phase angle $\phi = \arctan k_y/k_z$ of the fastest growing mode as function of $\SSC$.
-Above $\SSC \gtrsim 2.05$ $\phi=0$, indicating that axisymmetric modes have the fastest growth rates.
-For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC \simeq 4.84$ (most works use similar values).
+Figure~\ref{fig:growth_rate}d shows the phase angle $\phi=\arctan(k_{y}/k_{z})$ of the fastest growing mode as function of $\SSC$.
+Above $\SSC\gtrsim2.05$ $\phi=0$, indicating that axisymmetric modes have the fastest growth rates.
+For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC\simeq4.84$ (most works use similar values).
 Thus, we predict axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
+
 \begin{figure}[h!]
@@
   \caption{Mean-field $\alpha$ coefficient.}
-  \label{fig:phi}
+  \label{fig:correlation}
 \end{figure}
@@
 Even for $\SSC> 2.05$ there are significant swaths of unstable three-dimensional modes with growth rates comparable to the maximum (figure~\ref{fig:growth_rate}a).
-Also, non-normality generically accompanies non-axisymmetry in shearing systems \citep[see][for a discussion relevant to the MRI]{1992MNRAS.255P..25K}.
+Also, non-normality generically accompanies non-axisymmetry in shearing systems \citep[see][]{1992MNRAS.255P..25K}.
 In non-normal systems, transient amplification can occur for stable modes, and can even cause turbulence.
@@
 
-The second implication of the three-dimensional modes is considerably more important.
-These modes will generate a mean electromotive force in linear regime, as evidenced by the presence of non-trivial eigenmodes for all three components of the magnetic field.
-This is of considerable significance, as it leads to the possibility of a laminar MRI dynamo operating directly in regions of weak shear at large Reynolds number.
+Figure~\ref{fig:eigvec} shows the eigenvectors for the most unstable mode at $\SSC=1.02$ using ideal MHD. 
+No critical layers can form for $\gamma\ne0$.
+We therefore solve equations~(\ref{eq:mhd})--~(\ref{eq:divu}) without dissipation as a 2nd-order system in $\partial_{x}$ and impose $v_{x}=0$ at $x=\pm{d}$.
+The eigenfunctions are indistinguishable from those at finite dissipation, lending additional confidence to our other results.
+The tilted structures of $v_{x}$ and $v_{y}$ as well as $b_{x}$ and $b_{y}$, imply non-trivial Reynolds and Maxwell stresses.
+
+We should expect the generation of a mean electromotive force in linear regime, \textit{i.e}
+$\left<\boldsymbol{v}\times\boldsymbol{b}\right>{\sim}\,\alpha\!\left<\boldsymbol{B}_{0}\right>$.
+This is significant because of the possibility of direct laminar dynamo action in regions of weak shear at large Reynolds number.
 MRI dynamos have been studied in a number of different contexts \citep{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one focused on turbulent dynamos. 
-The lone exception, \citet{2016MNRAS.462..818B}, is a numerical study that discovered the exponential growth of mean magnetic fields during the linear growth phase of the MRI far from stability.
+The lone exception, \citet{2016MNRAS.462..818B}, numerically found the exponential growth of mean magnetic fields during the linear growth phase of the MRI far from stability.
 Our work explains this result in terms of purely linear dynamics:
 non-axisymmetric MRI unstable modes drive the dynamo growth of magnetic fields.
+
+Figure~(\ref{fig:correlation}) shows... %I'm going to hold off on this until we know what happens with the non-mean field case.%
 
@@
   \includegraphics[width=\columnwidth]{run_49_output_growthrates.pdf}
-  \caption{(top) Growth rates for q=0.1 relevant to the Solar interior. (bottom) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta = 0.1$.}
+  \caption{(top) Growth rates for $q=1/10$ relevant to the stellar interiors. (bottom) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta=0.1$.}
   \label{fig:other_params}
@@
 Finally, we vary two important parameters:
-$q$ and the magnetic Prandtl number $\Prm = \nu/\eta$.
-Figure~\ref{fig:other_params} shows the growth rates for $\SSC = 1.02$.
-The upper panel shows $q = 1/10$ holding all other parameters equal to their fiducial values.
-This corresponds a Rossby number relevant to stellar interiors, and it shows very behavior to the $q=3/4$ cases, $\gamma_{max}/S \simeq 0.1$.
-Again, MRI remains dominated by three-dimensional modes near onset.
-The lower panel shows $\nu = 10^{-6}$ and $\eta = 0.1$ with all other parameters equal to their fiducial. 
-This is $\Prm = 10^{-5}$ close to those found in liquid metal experiments.
-At this $\Prm$, the MRI is only slightly supercritical; this is very similar to the experimental conditions of the Princeton MRI experiment \citep{2002JFM...462..365G}.
-Our results are for the rotating plane-Couette geometry\footnote{This is equivalent to Taylor-Couette flow in the limit that the gap between cylinders $R_2 - R_1 \to 0$.} with highly idealized boundary conditions.
-Nevertheless, it is very suggestive that near onset in both shear and resistivity at very low $\Prm$, the MRI is \emph{only} unstable to three-dimensional modes.
+$q$ and the magnetic Prandtl number $\Prm=\nu/\eta$.
+Figure~\ref{fig:other_params} shows the growth rates for $\SSC=1.02$.
+The upper panel shows $q=1/10$ (addressing a possible range within stellar interiors) holding all other parameters equal to their fiducial values.
+The growth rates show similar behavior to the $q=3/4$ cases, \textit{i.e.}, dominated by three-dimensional modes with $\gamma_{\text{max}}/|S|\simeq0.1$.
+The lower panel shows $\nu=10^{-6}$ and $\eta=0.1$ with all other parameters equal to their fiducial values. 
+This case is relevant to liquid metals e.g., the Princeton MRI experiment \citep{2002JFM...462..365G}
+at $\Prm=10^{-5}$. 
+Our results are for the rotating plane-Couette geometry (Taylor-Couette in the small-gap limit $R_{2}\approx{R}_{1}$) with highly idealized boundary conditions.
+It is nevertheless quite interesting that, near onset, the low-$\Prm$ MRI is \emph{only} unstable to three-dimensional modes.
 
-Our results show that that whenever the MRI is near its critical shear values,  three-dimensional modes grow faster than their better-studied two-dimensional counterparts.
-While we have focused on the Keplerian case ($q=3/4$), we also demonstrate robustness for low Rossby numbers $q=1/10$.
+Our results show that three-dimensional modes grow faster than two-dimensional modes whenever the MRI is near its critical shear values.
+While we mainly focus on the ideal Keplerian case ($q=3/4$), we also demonstrate robustness for low Rossby ($q=1/10$) and Prandtl numbers ($\Prm=10^{-5}$).
 There are several important future directions this work suggests.
-First, the Sun possesses two internal shear layers with inwardly increasing shear, the high-latitude tachocline, and the near-surface shear layer.
-The latter in particular, though unstable to small-scale convection, may also host slower MRI-driven dynamics.
-In order to further elucidate this processes, understanding the non-linear saturation of the linear instability presented here is crucial, as is its robustness to perturbations driven by the fast convective dynamics.
-Second, our results at low $\Prm$ suggest that three-dimensional MRI modes may be the easiest to excite in liquid metal experiments.
-Followup work including more complex boundary conditions and full, cylindrical Taylor-Couette geometries are required to determine if experiments should consider the signatures of these non-axisymmetric modes.
-Our prior work on the axisymmetric MRI \citep{2017ApJ...841....1C,2017ApJ...841....2C} shows that differences in linear dynamics between rotating Plane Couette and Taylor-Couette geometries are small in that case; whether or not such a result carries over to non-axisymmetric modes remains to be seen.
+First, the Sun possesses two internal shear layers with inwardly increasing shear, the high-latitude tachocline, and the near-surface shear layer (NSSL).
+The NSSL in particular, though containing small-scale convection, may also host slower MRI-driven dynamics.
+It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to noisy convection.
+Second, our low-$\Prm$ results suggest that three-dimensional MRI modes may be the easiest to excite in liquid metal experiments.
+Followup work using more realistic boundary conditions and geometry is required to determine possible non-axisymmetric signatures in Taylor-Couette experiments. 
+Our prior work on the axisymmetric MRI \citep{2017ApJ...841....1C,2017ApJ...841....2C} shows only small differences between rotating plane- and cylindrical- Taylor-Couette geometries. 
+We fully expect the general three-dimensional features of the MRI to persist in more complex applications.
 

### 2019-08-23 paper/paper.tex
couple small things

@@
 \begin{equation}\label{eq:Jx}
-\frac{Dj_{x}}{Dt}+{S}\partial_{z}b_{x}-\eta\nabla^{2}j_{x}=B_{0}\partial_{z}\omega_{x}
+\frac{Dj_{x}}{Dt}-\eta\nabla^{2}j_{x}=B_{0}\partial_{z}\omega_{x}-S\partial_{z}b_{x}
 \end{equation}
@@
 The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray.
-d) The wavenumber angle $\arctan(k_{y}/k_{z})$ as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes.}
+d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes.}
   \label{fig:growth_rate}
@@
 The MRI's preference for three-dimensional modes can be predicted analytically.
-We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC=1+\epsilon^{2}R$, $k_{z}\sim\mathcal{O}(\epsilon)$, $\sigma{\sim}k_{y}\sim\mathcal{O}(\epsilon^{2})$,
+We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC=1+\epsilon^{2}R$, $k_{z}\sim\mathcal{O}(\epsilon)$, $\sigma\sim{k_{y}}\sim\mathcal{O}(\epsilon^{2})$,
 \begin{equation}\label{eq:asymp}
@@
 We should expect the generation of a mean electromotive force in linear regime, \textit{i.e}
-$\left<\boldsymbol{v}\times\boldsymbol{b}\right>{\sim}\,\alpha\!\left<\boldsymbol{B}_{0}\right>$.
+$\left<\boldsymbol{v}\times\boldsymbol{b}\right>\sim\,\alpha\!\left<\boldsymbol{B}_{0}\right>$.
 This is significant because of the possibility of direct laminar dynamo action in regions of weak shear at large Reynolds number.
@@
 \begin{acknowledgments}
-We would like to thank Steve Tobias for helpful conversations on this work.
+We thank Steve Tobias for helpful conversations on this work.
 JSO acknowledges funding from NASA LWS grant No.

### 2019-08-24 paper/paper.tex
included doi's for a few solar refs + a few small tweaks.

@@
 This finite critical shear results from a finite-channel cutoff.
-Despite the extensive literature on accretion disks (strong shear, low dissipation) and liquid-metal experiments (weak shear, strong dissipation), even the linear MRI is not well understood in the weak-shear, low-dissipation regime.
+Despite the extensive literature on accretion disks (strong shear, low dissipation) and liquid-metal experiments (weak shear, large dissipation), even the linear MRI is not well understood in the weak-shear, low-dissipation regime.
 
@@
 We find that the first destabilized modes are three-dimensional, and thus could act as a dynamo even in the absence of secondary instability.
-These results strongly suggest that the non-normality of the linear operator driving the MRI are always important, even when axisymmetric modes dominate.
+These results also suggest that the non-normal growth the MRI is always important, even when axisymmetric modes dominate.
 
@@
 \begin{equation}\label{eq:Bx}
-\frac{Db_{x}}{Dt}+\eta(\partial_{y}j_{z}-\partial_{z}j_{y})=B_{0}\partial_{z}v_{x}
+\frac{Db_{x}}{Dt}+\eta(\partial_{y}j_{z}-\partial_{z}j_{y})=B_{0}\partial_{z}v_{x},
 \end{equation}
@@
 \begin{equation}\label{eq:Jx}
-\frac{Dj_{x}}{Dt}-\eta\nabla^{2}j_{x}=B_{0}\partial_{z}\omega_{x}-S\partial_{z}b_{x}
+\frac{Dj_{x}}{Dt}-\eta\nabla^{2}j_{x}=B_{0}\partial_{z}\omega_{x}-S\partial_{z}b_{x}.
 \end{equation}
@@
 \end{equation}
-The spatial domain is a vertically ($z$) and horizontally ($y$) periodic channel of width $d=\pi$. 
+The spatial domain is a doubly periodic channel in $y,\,z$ with width $d=\pi$. 
 The boundary conditions are impenetrable stress-free and perfectly conducting; $v_{x}=\omega_{y}=\omega_{z}=b_{x}=\partial_{x}j_{x}=0$ at $x=\pm{d}$. 
@@
 We assume harmonic perturbations in $y$ and $z$, (e.g. pressure) $p=\hat{p}(x)e^{i(k_{y}y+k_{z}z)+\sigma{t}}$. 
-We use a complex-valued growth rate $\sigma=\gamma+i\omega$ with $\gamma,\omega$ both real. 
+We use a complex-valued growth rate $\sigma=\gamma+i\omega$ with $\gamma,\,\omega$ both real. 
 The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $\partial_{x}$ with Dirichlet boundary conditions.
@@
 These results are pertinent in two ways. 
-First, the MRI contains a substantial ``Goldilocks regime" with possible direct dynamo action, and this regime likely most applies to stellar interiors and laboratory experiments. 
+First, the MRI contains a substantial ``Goldilocks regime" with possible direct dynamo action, and this regime likely applies to stellar interiors and laboratory experiments. 
 Second, our results accord with well-established results for accretion disks that expect two-dimensional primary modes.
@@
 No critical layers can form for $\gamma\ne0$.
-We therefore solve equations~(\ref{eq:mhd})--~(\ref{eq:divu}) without dissipation as a 2nd-order system in $\partial_{x}$ and impose $v_{x}=0$ at $x=\pm{d}$.
+We therefore solve equations~(\ref{eq:mhd}--\ref{eq:divu}) without dissipation as a 2nd-order system in $\partial_{x}$ and impose $v_{x}=0$ at $x=\pm{d}$.
 The eigenfunctions are indistinguishable from those at finite dissipation, lending additional confidence to our other results.
@@
 First, the Sun possesses two internal shear layers with inwardly increasing shear, the high-latitude tachocline, and the near-surface shear layer (NSSL).
-The NSSL in particular, though containing small-scale convection, may also host slower MRI-driven dynamics.
+Past work has already pointed out the possibility of the small-scale MRI in the Sun using local analysis \citep{doi:10.1086/522426,doi:10.1111/j.1745-3933.2010.00987.x,doi:10.1088/0004-637X/787/1/21}
+The NSSL, in particular, may also host slower MRI-driven dynamics; in spite of containing small-scale convection.
 It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to noisy convection.
 Second, our low-$\Prm$ results suggest that three-dimensional MRI modes may be the easiest to excite in liquid metal experiments.
-Followup work using more realistic boundary conditions and geometry is required to determine possible non-axisymmetric signatures in Taylor-Couette experiments. 
+Determining possible non-axisymmetric signatures in Taylor-Couette experiments requires followup work using more realistic boundary conditions and geometry. 
 Our prior work on the axisymmetric MRI \citep{2017ApJ...841....1C,2017ApJ...841....2C} shows only small differences between rotating plane- and cylindrical- Taylor-Couette geometries. 

### 2019-08-24 paper/paper.tex
small tweaks

@@
   \includegraphics[width=\columnwidth]{run_49_output_growthrates.pdf}
-  \caption{(top) Growth rates for $q=1/10$ relevant to the stellar interiors. (bottom) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta=0.1$.}
+  \caption{(top) Growth rates for $q=1/10$ relevant to the stellar interiors. 
+  (bottom) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta=0.1$.}
   \label{fig:other_params}
@@
 First, the Sun possesses two internal shear layers with inwardly increasing shear, the high-latitude tachocline, and the near-surface shear layer (NSSL). 
-Past work has already pointed out the possibility of the small-scale MRI in the Sun using local analysis \citep{doi:10.1086/522426,doi:10.1111/j.1745-3933.2010.00987.x,doi:10.1088/0004-637X/787/1/21}
+Past work has already pointed out the possibility of the small-scale MRI in the Sun using local analysis \citep{doi:10.1086/522426,doi:10.1111/j.1745-3933.2010.00987.x,doi:10.1088/0004-637X/787/1/21}.
 The NSSL, in particular, may also host slower MRI-driven dynamics; in spite of containing small-scale convection.

### 2019-08-24 paper/supplemental.tex
supp. critical layer comment

@@
 \Eeq
-
-Assume all variable take the form
+All variables take the form
 \Beq
@@
 \Beq
-\omega_{S}(x) \ \equiv \ \omega + S x \ell, \quad \omega_{A}  \ \equiv \ B_{0} k
+\omega_{S}(x) \ \equiv \ \omega + S x \ell, \quad \omega_{A}  \ \equiv \ B_{0} k.
 \Eeq
@@
 \Eeq
-We want to solve the Schr\:{o}dinger-type equation~\ref{Schr-eq} with the boundary conditions 
+If $\omega_{S}=-i \gamma+Sx\ell$, with $Re(\gamma)\neq0$ then the denominator of $\Phi(x)$ cannot vanish. 
+We want to solve the Schr\"{o}dinger-type equation~\ref{Schr-eq} with the boundary conditions 
 \Beq

### 2019-08-24 paper/supplemental.tex
supp. equation number fixes

@@
 The full linear ideal equations are:
-\Beq
+\Beq\label{u-eq}
 \pd{t} u + S x \pd{y} u - f v + \pd{x} p - B_{0} \pd{z} b_{x} &=& 0\\
-\pd{t} v + S x \pd{y} v + (S+f) u + \pd{y} p - B_{0} \pd{z} b_{y} &=& 0\\
+\pd{t} v + S x \pd{y} v + (f+S) u + \pd{y} p - B_{0} \pd{z} b_{y} &=& 0\label{v-eq} \\
 \pd{t} w + S x \pd{y} w + \pd{z} p - B_{0} \pd{z} b_{z} &=& 0 \\
@@
 \pd{t} b_{y} + S x \pd{y} b_{y} - B_{0} \pd{z} v - S b_{x}  &=& 0\\
-\pd{t} b_{z} + S x \pd{y} b_{z} - B_{0} \pd{z} w &=& 0
+\pd{t} b_{z} + S x \pd{y} b_{z} - B_{0} \pd{z} w &=& 0 \label{bz-eq}
 \Eeq
@@
 \Eeq
-Define the frequency ``parameters'' 
+We solve the ideal eigenvalue problem in \text{Dedalus} in the above formulation. To make analytical progress, define the frequency ``parameters'' 
 \Beq
@@
 \Eeq
-We can use equations (2)--(7) to find all amplitudes in terms of $\hat{U}(x)$ and $\hat{U}'(x)$. 
+We can use equations (\ref{v-eq}--\ref{bz-eq}) to find all amplitudes in terms of $\hat{U}(x)$ and $\hat{U}'(x)$. 
 \Beq
@@
    \Eeq
-Substitute everything into equation (1) to get a second-order equation for $\hat{U}(x)$ of the general form 
+Substitute everything into equation (\ref{u-eq}) to get a second-order equation for $\hat{U}(x)$ of the general form 
 \Beq
@@
 If $\omega_{S}=-i \gamma+Sx\ell$, with $Re(\gamma)\neq0$ then the denominator of $\Phi(x)$ cannot vanish. 
-We want to solve the Schr\"{o}dinger-type equation~\ref{Schr-eq} with the boundary conditions 
+We want to solve the Schr\"{o}dinger-type equation~(\ref{Schr-eq}) with the boundary conditions 
 \Beq
@@
 \Beq
-\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \frac{\epsilon ^2 \left(-B_{0}^4 k^4+B_{0}^2 f^2 q \left(k^2 R+q
-   \ell ^2\right)+f^2 (q+1) (\omega -f q x \ell
-   )^2\right)}{B_{0}^4 k^2}
+\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 q \left(k^2 R+q
+   \ell ^2\right) - B_{0}^4 k^4 +f^2 (q+1) (\omega -f q x \ell
+   )^2  }{B_{0}^4 k^2}
 \Eeq
@@
 \Beq
-\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0 , \quad \implies \quad \psi_{0}(x) \ = \ \cos(\pi x / d)
+\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0  \quad \implies \quad \psi_{0}(x) \ = \ \cos(\pi x / d).
 \Eeq
@@
 \Beq
-\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) + \Phi_{2}(x)\psi_{0}(x)\ = \ 0 
+\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) + \Phi_{2}(x)\psi_{0}(x)\ = \ 0.
 \Eeq

### 2019-08-24 paper/paper.tex
q convention note
Couple typos and words

@@
 We find that the first destabilized modes are three-dimensional, and thus could act as a dynamo even in the absence of secondary instability.
-These results also suggest that the non-normal growth the MRI is always important, even when axisymmetric modes dominate.
+These results also suggest that the non-normal growth of the MRI is always important, even when axisymmetric modes dominate.
 
-We solve the linearized magnetohydrodynamic equations in rotating plane-Couette geometry.
+We numerically solve the linearized magnetohydrodynamic equations in rotating plane-Couette geometry.
 This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y}=Sx$; \citep[see][]{2015RSPSA.47140699V}.
@@
 Accretion-disk modelling usually considers the Rossby number $q=-S/f<1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
+\textit{Note}: some sources use $q=-S/\Omega=3/2$ to represent the same quantity. 
 Unless otherwise stated $q=3/4$ (Keplerian).
@@
 We use a complex-valued growth rate $\sigma=\gamma+i\omega$ with $\gamma,\,\omega$ both real. 
-The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $\partial_{x}$ with Dirichlet boundary conditions.
+The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $x$ with Dirichlet boundary conditions.
 We pose and solve equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework \citep{The Dedalus paper}.
@@
 Our main interest is in ideal ($\eta=0$), inviscid ($\nu=0$) conditions.
-However, we set $\eta=\nu=10^{-5}$ to avoid critical layers in for the \textit{stable} solution branch.
-We confirmed our results are totally insensitive to small diffusion. 
-For each $(k_{y},k_{z})$ pair, we solve using $n_{x}=128$ modes; all our results are identical at double the resolution; see \protect\url{github.com/jsoishi/mri_prefers_3d} for all code used in this paper.
+However, we set $\eta=\nu=10^{-5}$ to avoid critical layers in the \textit{stable} solution branch.
+We confirmed our results for unstable solutions are insensitive to small diffusion. 
+For each $(k_{y},k_{z})$ pair, we solve the eigenvalue problem using $n_{x}=128$ modes; all our results are identical at double the resolution; see \protect\url{github.com/jsoishi/mri_prefers_3d} for all code used in this paper.
 \begin{figure*}[ht]
@@
   \includegraphics[width=\columnwidth]{eigvecs_xy_run_11_ideal_single_mode.pdf}
-  \caption{\textit{Ideal} Eigenvectors of the velocity and magnetic field perturbations for the most unstable mode when $\SSC=1.02$, and $\eta=\nu=0$; calculations at $\eta=\nu=10^{-5}$ are indistinguishable.
+  \caption{\textit{Ideal} eigenvectors of the velocity and magnetic field perturbations for the most unstable mode when $\SSC=1.02$, and $\eta=\nu=0$; calculations at $\eta=\nu=10^{-5}$ are indistinguishable.
   The top row shows $v_{x}$, $v_{y}$, $v_{z}$ (red is negative, blue positive); the bottom row shows $b_{x}$, $b_{y}$, $b_{z}$ (purple is negative, orange positive). 
@@
 
-We should expect the generation of a mean electromotive force in linear regime, \textit{i.e}
+We should expect the generation of a mean electromotive force in the linear regime, \textit{i.e}
 $\left<\boldsymbol{v}\times\boldsymbol{b}\right>\sim\,\alpha\!\left<\boldsymbol{B}_{0}\right>$.

### 2019-08-24 paper/supplemental.tex
typo
Word change in supplement

@@
 For both the ideal and non-ideal MHD equations, we solve the eigenvalue problem in the $x$ direction using the EigenValueProblem mode in \emph{Dedalus} for a grid of $N_y \times N_z$ modes in the $y$ and $z$ directions.
-For most of our runs, we use a targeted, sparse eigenvalue solver solving for the 15 modes closest to a guess for the maximum growth rate.
+For most of our runs, we use a targeted, sparse eigenvalue solver to find the 15 modes closest to a guess for the maximum growth rate.
 We use the ideal 2D growth rate as input for the smallest $k_y > 0$ mode at each $k_z$ and then use the output from each previous $k_y$ as an input guess for the next mode.
@@
 \Eeq
-We solve the ideal eigenvalue problem in \text{Dedalus} in the above formulation. To make analytical progress, define the frequency ``parameters'' 
+We solve the ideal eigenvalue problem in \textit{Dedalus} in the above formulation. To make analytical progress, define the frequency ``parameters'' 
 \Beq

### 2019-08-25 paper/paper.tex
better organise the equations a bit.

@@
 \boldsymbol{\omega}=\boldsymbol{\nabla}\times\boldsymbol{v},\quad\text{and}\quad\frac{D}{Dt}=\partial_{t}+{S}x\,\partial_{y}\end{equation}
-The main input parameters are the Coriolis parameter, $f=2 \Omega$; the background shear rate, $S=dV_{y}/dx<0$;  the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0}\rho_{0}=1$); and the horizontal domain width $-d/2\le{x}\le d/2$.
-Accretion-disk modelling usually considers the Rossby number $q=-S/f<1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
-\textit{Note}: some sources use $q=-S/\Omega=3/2$ to represent the same quantity. 
-Unless otherwise stated $q=3/4$ (Keplerian).
+
 We write the induction equation in terms of the $x$-component magnetic field,
@@
 \end{equation}
-The spatial domain is a doubly periodic channel in $y,\,z$ with width $d=\pi$. 
-The boundary conditions are impenetrable stress-free and perfectly conducting; $v_{x}=\omega_{y}=\omega_{z}=b_{x}=\partial_{x}j_{x}=0$ at $x=\pm{d}$. 
+The spatial domain is a doubly periodic channel in $y,\,z$ with width $-d/2\le{x}\le d/2$.
+The boundary conditions are impenetrable stress-free and perfectly conducting; $v_{x}=\omega_{y}=\omega_{z}=b_{x}=\partial_{x}j_{x}=0$ at $x=\pm{d/2}$.
+
+The main input parameters are the Coriolis parameter, $f=2 \Omega$; the background shear rate, $S=dV_{y}/dx<0$;  the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0}\rho_{0}=1$).
+Accretion-disk modelling usually considers the Rossby number $q=-S/f<1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
+\textit{Note}: some sources use $q=-S/\Omega$ to represent same effective quantity. 
+Unless otherwise stated $q=3/4$ (Keplerian).
+
 The MRI is a weak-field instability; in the inviscid, ideal case the critical shear rate for instability in \textit{two dimensions} is
@@
 Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
-At $\SSC=1.02$, the maximum growth rate ($\gamma/|S|=0.093$) happens at $(k_{y},k_{z})\simeq(0.263,0.447)$.
+At $\SSC=1.02$, the maximum growth rate ($\gamma_{\max}/|S|=0.093$) happens at $(k_{y},k_{z})\simeq(0.263,0.447)$.
 The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability.
@@
 No critical layers can form for $\gamma\ne0$.
-We therefore solve equations~(\ref{eq:mhd}--\ref{eq:divu}) without dissipation as a 2nd-order system in $\partial_{x}$ and impose $v_{x}=0$ at $x=\pm{d}$.
+We therefore solve equations~(\ref{eq:mhd}--\ref{eq:divu}) without dissipation as a 2nd-order system in $\partial_{x}$ and impose $v_{x}=0$ at $x=\pm{d/2}$.
 The eigenfunctions are indistinguishable from those at finite dissipation, lending additional confidence to our other results.
@@
 The upper panel shows $q=1/10$ (addressing a possible range within stellar interiors) holding all other parameters equal to their fiducial values.
-The growth rates show similar behavior to the $q=3/4$ cases, \textit{i.e.}, dominated by three-dimensional modes with $\gamma_{\text{max}}/|S|\simeq0.1$.
+The growth rates show similar behavior to the $q=3/4$ cases, \textit{i.e.}, dominated by three-dimensional modes with $\gamma_{\max}/|S|\simeq0.1$.
 The lower panel shows $\nu=10^{-6}$ and $\eta=0.1$ with all other parameters equal to their fiducial values. 

### 2019-08-25 python/make_eigenvalue_plot.py
Updated figures
* \gamma/S --> \gamma/|S|
* print \gamma_{max} on figure 1

@@ -58,7 +58,7 @@ if contours:
     plt.contour(kz_global,ky_global,gamma_r,levels=contour_levels,colors='k')
 plt.contour(kz_global,ky_global,gamma_r,levels=[0.],colors='w',alpha=0.5)
 plt.plot(max_kz, max_ky, 'ro')
-plt.colorbar(PCM, label=r'$\gamma/S$')
+plt.colorbar(PCM, label=r'$\gamma/|S|$')
 plt.xlabel(r'$k_z$')
 plt.ylabel(r'$k_y$')
 plt.tight_layout()

### 2019-08-25 python/plot_fig1.py

@@ -48,7 +48,8 @@ for i,r in enumerate(runs):
 
     ax.contour(kz_global,ky_global,gamma_r,levels=contour_levels,colors='k')
     ax.contour(kz_global,ky_global,gamma_r,levels=[0.],colors='grey')
-    ax.text(0.05,0.9,r'$S/S_c = {:5.3f}$'.format(SSC[i]), color='w',fontsize=18)
+    ax.text(0.05,0.9,r'$S/S_c = {:5.3f}$'.format(SSC[i]), color='w',fontsize=20)
+    ax.text(0.05,0.8,r'$\gamma_{{max}}/|S| = {:5.3f}$'.format(gamma_r.max()), color='w',fontsize=16)
     ax.plot(max_kz, max_ky, 'ro')
     #ax.colorbar(PCM, label=r'$\gamma$')
     ax.set_xlabel(r'$k_z$')
@@ -63,7 +64,7 @@ for i,r in enumerate(runs):
 
 cax = fig.add_axes([xgutter+4*(w+pad), 0.55,0.015,h])
 cbar = fig.colorbar(c, cax=cax)
-cbar.set_label(r'$\gamma/S$')
+cbar.set_label(r'$\gamma/|S|$')
 
 # asymptotics
 xloc = grid[0].get_position().x0
@@ -101,7 +102,7 @@ asymp_ax.text(0.03, 0.07, 'asymptotic', fontsize=18)
 asymp_ax.set_ylim(0, 0.06/q)
 asymp_ax.set_xlim(0,0.2)
 asymp_ax.set_xlabel(r'$k_y$')
-asymp_ax.set_ylabel(r'$\gamma/S$')
+asymp_ax.set_ylabel(r'$\gamma/|S|$')
 asymp_ax.xaxis.set_major_locator(plt.MultipleLocator(0.05))
 asymp_sub.set_ylim(-0.005/q**2, 0.005/q**2)
 asymp_sub.set_xticks([0.045,0.08])
@@ -133,8 +134,8 @@ spec_ax.scatter(spec_thresh[spec_thresh.real <= 0].real, spec_thresh[spec_thresh
 
 spec_ax.set_xlim(-0.25/q,0.1/q)
 spec_ax.set_ylim(-0.8/q,0.8/q)
-spec_ax.set_xlabel(r"$\gamma/S$")
-spec_ax.set_ylabel(r"$\omega/S$")
[…]

### 2019-08-25 python/make_eigenvalue_plot.py
Added labels to eigenvalue plots
* added merged Re plot for supplemental

@@ -3,14 +3,14 @@ This script plots both the largest growth rates of the MRI data
 passed to the script. 
 
 Usage:
-    make_eigenvector_plots.py [--vmin=<vmin> --vmax=<vmax> --q=<q> --no-contours] <filename>
+    make_eigenvector_plots.py [--vmin=<vmin> --vmax=<vmax> --q=<q> --no-contours --label=<label>] <filename>
 
 Options:
     --vmin=<vmin>    min for colorbar [default: 0]
     --vmax=<vmax>    max for colorbar [default: 0.701]
     --q=<q>          q = S/f [default: 0.75]
     --no-contours    turn off contours
-
+    --label=<label>  run label
 """
 
 import h5py
@@ -30,6 +30,7 @@ vmax = float(args['--vmax'])
 q = float(args['--q'])
 contours = not args['--no-contours']
 outbase = Path("plots")
+label = args['--label']
 
 # Plot growth rates
 datafile = h5py.File(filename,'r')
@@ -61,6 +62,9 @@ plt.plot(max_kz, max_ky, 'ro')
 plt.colorbar(PCM, label=r'$\gamma/|S|$')
 plt.xlabel(r'$k_z$')
 plt.ylabel(r'$k_y$')
+
+if label:
+    plt.text(0.05,0.9,r'${}$'.format(label), color='w',fontsize=20)
 plt.tight_layout()
 
 #plt.title(r'3D Keplerian MRI growth rates/f  $\left( S/S_{\mathrm{crit.}} = %.2f\right)$' %(R[i]))

### 2019-08-25 python/re_plots.sh

@@ -1,0 +1,5 @@
+python3 make_eigenvalue_plot.py --vmax=0.1 --label="Re = 10^{5}" data/run_11_output.h5
+python3 make_eigenvalue_plot.py --vmax=0.1 --label="Re = 10^{4}" data/run_43_output.h5
+python3 make_eigenvalue_plot.py --vmax=0.1 --label="Re = 10^{3}" data/run_44_output.h5
+python3 make_eigenvalue_plot.py --vmax=0.1 --label="Re = 10^{2}" data/run_45_output.h5
+python3 make_eigenvalue_plot.py --vmax=0.1 --label="Re = 10^{6}" data/run_46_output.h5

### 2019-08-25 python/dynamo_mri.py
include current terms for non-uniform background field.

@@ -1,0 +1,309 @@
+"""
+The magnetorotational instability prefers three dimensions.
+    
+Dedalus script for calculating the maximum growth rates for the
+3D MRI over a range of wavenumbers.
+
+This script can be ran serially or in parallel, and produces an h5py file
+that contains maximum eigenvalues and the corresponding eigenvectors.
+
+Pass it a .cfg file to specify parameters for the particular solve.
+
+To run using 4 processes, you would use:
+    $ mpiexec -n 4 python3 mri.py mri_params.cfg
+
+"""
+
+import time
+import configparser
+from configparser import ConfigParser
+import argparse
+from pathlib import Path
+import numpy as np
+import scipy.optimize as opt
+import os
+import h5py
+import matplotlib.pyplot as plt
+import dedalus.public as de
+from mpi4py import MPI
+CW = MPI.COMM_WORLD
+import logging
+logger = logging.getLogger(__name__)
+
+# Parses .cfg filename passed to script
+parser = argparse.ArgumentParser(description='Passes filename')
+parser.add_argument('filename', metavar='Rc', type=str, help='.h5 file to plot eigenvectors for maximum eigenvalue')
+args = parser.parse_args()
+filename = Path(vars(args)['filename'])
+outbase = Path("data")
+
+# Parse .cfg file to set global parameters for script
+config = ConfigParser()
+config.read(str(filename))
+
+logger.info('Running mri.py with the following parameters:')
+logger.info(config.items('parameters'))
+
+try:
+    dense = config.getboolean('solver','dense')
+except:
+    dense = False
+if dense:
+    logger.info("Using dense solver.")
+    dense_threshold = config.getfloat('solver','dense_threshold')
+else:
+    logger.info("Using sparse solver.")
+
+Nx = config.getint('parameters','Nx')
+Lx = eval(config.get('parameters','Lx'))
+B = config.getfloat('parameters','B')
+
+Nmodes = config.getint('parameters','Nmodes')
+
+R      =  config.getfloat('parameters','R')
+q      =  config.getfloat('parameters','q')
+
[…]

### 2019-08-25 paper/supplemental.tex
fixed small supp. error

@@
 
+\newcommand{\Reyn}{\mathrm{Re}}
+\newcommand{\Reym}{\mathrm{Rm}}
 \newcommand{\Ra}{R}
@@
 
-\section{Numerical Methods}
+\maketitle
+\section{Numerical Methods and Resolution Studies}
 \label{sec:methods}
 
-For both the ideal and non-ideal MHD equations, we solve the eigenvalue problem in the $x$ direction using the EigenValueProblem mode in \emph{Dedalus} for a grid of $N_y \times N_z$ modes in the $y$ and $z$ directions.
+For both the ideal and non-ideal MHD equations, we solve the eigenvalue problem in the $x$ direction using the \texttt{EigenValueProblem} solver in \emph{Dedalus} for a grid of $N_y \times N_z$ modes in the $y$ and $z$ directions.
 For most of our runs, we use a targeted, sparse eigenvalue solver to find the 15 modes closest to a guess for the maximum growth rate.
-We use the ideal 2D growth rate as input for the smallest $k_y > 0$ mode at each $k_z$ and then use the output from each previous $k_y$ as an input guess for the next mode.
-Our solver is parallelized over the $k_y, k_z$ modes.
-We have confirmed that dense solvers retrieve identical results.
-For the spectrum in figure 1 of the main text, we used a dense eigenvalue solver.
+In run 3, we have confirmed that dense solvers retrieve identical results.
+For nearly ideal runs, we use the ideal 2D growth rate as input for the smallest $k_y > 0$ mode at each $k_z$ and then use the output from each previous $k_y$ as an input guess for the next mode.
+For those runs that have significant $\eta$, we instead use a dense solve for the $k_y = 0$ modes and then step forward in $k_y$ at each $k_z$ as before.
+Our solver is embarrassingly parallelized over the $k_z$ modes.
+For the spectrum in figure 1 of the main text, we used a dense eigenvalue solver at $(k_y^{max},k_z^{max})$ for $\SSC= 1.02$.
 
@@
 
-\section{Resolution and Reynolds number study}
-\label{sec:resolution}
+\section{Reynolds Number and Departure from Ideal MHD}
+\label{sec:reyn}
+
+In order to ensure that our calculations are probing the ideal MHD regime we are interested in, we have performed as series of runs (24-27) with the $\eta = \nu$ ranging from $10^{-6}$ to $10^{-2}$.
+In our units, this is equivalent to varying the Reynolds and Magnetic Reynolds numbers $\Reyn = \Reym = 10^2 -- 10^6$.
+Figure~\pageref{fig:reynolds} shows the growth rates at $\SSC=1.05$ for five values of $\Reyn = \Reym$.
+For $\Reyn \gtrsim 10^4$, the maximum growth rate is located at the same position and the same growth rate; the unstable region expands, as expected.
+This suggests that our results at $\Reyn = 10^5$ are reasonably close to ideal MHD for the purposes of understanding the MRI's preference for 3D modes near shear onset.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\textwidth]{re_plots.pdf}
+  \caption{Growth rates at $\SSC = 1.02$ for varying diffusivities.
+Note that the top three panels are essentially identical, suggesting that above $\Reyn \gtrsim 10^4$ we are reasonably close to ideal conditions.}
+  \label{fig:reynolds}
+\end{figure}
 
@@
 \label{sec:asymp}
+Here, we outline the asymptotic calculation for the leading order correction to the 2D growth rates in when accouting for three dimensional effects via finite $k_y$.
 Define the domain as $ - d/2 < x < d/2$. Being in the rotating frame means the shear has no net
@@
 \Beq
-\psi(x=\pm d) = 0.
+\psi(x=\pm d/2) = 0.
 \Eeq

### 2019-08-25 python/plot_mean_field_alpha.py
Changed caption for mean field EMF figure. Cleaned up figure.

@@ -23,7 +23,12 @@ alpha_yz = alpha.mean(axis=0).mean(axis=0)/norm
 
 plt.plot(x[0,0,:], alpha_yz)
 plt.xlabel(r"$x$")
-plt.ylabel(r"$\alpha$")
+plt.ylabel(r"$(v_x B_y - v_y B_x)/|u||B|$")
+
+ax =plt.gca()
+ax.set_xlim(-np.pi/2,np.pi/2)
+ax.set_xticks([-np.pi/2, 0, np.pi/2])
+ax.set_xticklabels([r'$-\pi/2$', 0, r'$\pi/2$'])
 plt.tight_layout()
 plt.savefig('plots/mean_field_alpha_{}.pdf'.format(dfname))
 

### 2019-08-25 paper/paper.tex
small tweak
Minor changes; added first cut at mean field correlation text

@@
 \newcommand{\apjs}{The Astrophysical Journal Supplement Series}
+\newcommand{\apjl}{Astrophysical Journal Letters}
 \usepackage{graphicx}
@@
 For a Keplerian shear, three-dimensional modes are unstable down to $S\approx0.10S_c$, and dominate the two-dimensional modes until $S\approx2.05S_{c}$.
-Significant numbers of fast three-dimensional modes remain well past $2.05S_{c}$. 
+Significant numbers of rapidly growing three-dimensional modes remain well past $2.05S_{c}$. 
 These finding are significant in three ways. 
 First, evidence from weakly nonlinear theory suggests that the MRI saturates by pushing the shear rate to its critical value. 
-This can happen for systems, like stars, that can rearrange their angular velocity profiles.
-Second, the non-normal character and large transient growth of the MRI should be important anytime strong three-dimensionality exists.
+This can happen for systems, like stars and laboratory experiments, that can rearrange their angular velocity profiles.
+Second, the non-normal character and large transient growth of MRI modes should be important anytime strong three-dimensionality exists.
 Finally, three-dimensional growth suggest the possibility of direct dynamo action driven from the linear instability itself.
@@
 
-Here, we investigate the stability of three-dimensional perturbations near the two-dimensional critical shear rate $S_{c}$ for a nearly invisicd, ideal MHD flow.
+Here, we investigate the stability of three-dimensional perturbations near the two-dimensional critical shear rate $S_{c}$ for a nearly inviscid, ideal MHD flow.
 We find that the first destabilized modes are three-dimensional, and thus could act as a dynamo even in the absence of secondary instability.
@@
 
-We numerically solve the linearized magnetohydrodynamic equations in rotating plane-Couette geometry.
+We numerically solve the linearized magnetohydrodynamic equations in rotating plane Couette geometry.
 This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y}=Sx$; \citep[see][]{2015RSPSA.47140699V}.
@@
 The main input parameters are the Coriolis parameter, $f=2 \Omega$; the background shear rate, $S=dV_{y}/dx<0$;  the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0}\rho_{0}=1$).
-Accretion-disk modelling usually considers the Rossby number $q=-S/f<1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
-\textit{Note}: some sources use $q=-S/\Omega$ to represent same effective quantity. 
+Accretion-disk modeling usually considers the Rossby number $q=-S/f<1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear
+\footnote{The astrophysical disk community often uses $q' = -S/\Omega$ to represent the same effective quantity; Keplerian profiles have $q' = 3/2$ in this case.}.
 Unless otherwise stated $q=3/4$ (Keplerian).
@@
 The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $x$ with Dirichlet boundary conditions.
-We pose and solve equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework \citep{The Dedalus paper}.
+We pose and solve equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework \citep{2019arXiv190510388B}.
 
@@
 We confirmed our results for unstable solutions are insensitive to small diffusion. 
-For each $(k_{y},k_{z})$ pair, we solve the eigenvalue problem using $n_{x}=128$ modes; all our results are identical at double the resolution; see \protect\url{github.com/jsoishi/mri_prefers_3d} for all code used in this paper.
+For each $(k_{y},k_{z})$ pair, we solve the eigenvalue problem using $n_{x}=128$ modes; all our results are identical at double the resolution \footnote{see supplemental material for details}; see \protect\url{github.com/jsoishi/mri_prefers_3d} for all code used in this paper.
 \begin{figure*}[ht]
@@
   \includegraphics[width=\columnwidth]{{mean_field_alpha_run_11_output.h5}.pdf}
-  \caption{Mean-field $\alpha$ coefficient.}
+  \caption{Velocity-magnetic field correlations relevant to the mean electromotive force, $(v_x B_y - v_y B_x)/|u| |B|$ for the eigenvectors of the fastest growing mode at $\SSC = 1.02$.}
   \label{fig:correlation}
@@
 This is significant because of the possibility of direct laminar dynamo action in regions of weak shear at large Reynolds number.
+Figure~\ref{fig:correlation} shows the correlation between $u$ and $b$ as a function of $x$ over the domain for $\SSC = 1.02$, proportional to the mean electromotive force.
+The correlation is non-zero, demonstrating the tendency of the 3D MRI eigenmodes to act as a mean-field dynamo.
+
 MRI dynamos have been studied in a number of different contexts \citep{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one focused on turbulent dynamos. 
@@
 non-axisymmetric MRI unstable modes drive the dynamo growth of magnetic fields.
-
-Figure~(\ref{fig:correlation}) shows... %I'm going to hold off on this until we know what happens with the non-mean field case.%
 
@@
 at $\Prm=10^{-5}$. 
-Our results are for the rotating plane-Couette geometry (Taylor-Couette in the small-gap limit $R_{2}\approx{R}_{1}$) with highly idealized boundary conditions.
+Our results are for the rotating plane Couette geometry (Taylor-Couette in the small-gap limit $R_{2}\approx{R}_{1}$) with highly idealized boundary conditions.
 It is nevertheless quite interesting that, near onset, the low-$\Prm$ MRI is \emph{only} unstable to three-dimensional modes.
@@
 First, the Sun possesses two internal shear layers with inwardly increasing shear, the high-latitude tachocline, and the near-surface shear layer (NSSL). 
-Past work has already pointed out the possibility of the small-scale MRI in the Sun using local analysis \citep{doi:10.1086/522426,doi:10.1111/j.1745-3933.2010.00987.x,doi:10.1088/0004-637X/787/1/21}.
-The NSSL, in particular, may also host slower MRI-driven dynamics; in spite of containing small-scale convection.
+Past work has already pointed out the possibility of the small-scale MRI in the Sun using local analysis \citep{2007ApJ...667L.207P,2011MNRAS.411L..26M,2014ApJ...787...21K}.
+The NSSL, in particular, may also host slower MRI-driven dynamics; despite containing small-scale convection.
 It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to noisy convection.
@@
 Determining possible non-axisymmetric signatures in Taylor-Couette experiments requires followup work using more realistic boundary conditions and geometry. 
-Our prior work on the axisymmetric MRI \citep{2017ApJ...841....1C,2017ApJ...841....2C} shows only small differences between rotating plane- and cylindrical- Taylor-Couette geometries. 
+Our prior work on the axisymmetric MRI \citep{2017ApJ...841....1C,2017ApJ...841....2C} shows only small differences between rotating plane and cylindrical Taylor-Couette geometries. 
 We fully expect the general three-dimensional features of the MRI to persist in more complex applications.
@@
 \end{document}
+
+%  LocalWords:  Magnetorotational Oishi Lewiston Vasil NSW Lecoanet
+%  LocalWords:  magnetorotational Keplerian magnetofluid Prandtl MHD
+%  LocalWords:  dimensionality Couette axisymmetric linearized RSPSA
+%  LocalWords:  magnetohydrodynamic Navier Dt Dj dV dx Alfv Rossby fd
+%  LocalWords:  hydrodynamical ODEs Dedalus arXiv supercriticality nd
+%  LocalWords:  criticality eigenvectors fiducial axisymmetry NSSL
+%  LocalWords:  eigenfunctions eigenmodes diffusivities tachocline
+%  LocalWords:  Tobias JSO LWS NNX Scialog TDA Leavitt mri

### 2019-08-26 paper/paper.tex
correlation notaion changes in text and figure script

@@
   \includegraphics[width=\columnwidth]{{mean_field_alpha_run_11_output.h5}.pdf}
-  \caption{Velocity-magnetic field correlations relevant to the mean electromotive force, $(v_x B_y - v_y B_x)/|u| |B|$ for the eigenvectors of the fastest growing mode at $\SSC = 1.02$.}
+  \caption{Velocity-magnetic field correlations relevant to the mean electromotive force for the eigenvectors of the fastest growing mode at $\SSC = 1.02$.}
   \label{fig:correlation}
@@
 
-We should expect the generation of a mean electromotive force in the linear regime, \textit{i.e}
+We should expect the generation of a mean electromotive force (EMF) in the linear regime, \textit{i.e}
 $\left<\boldsymbol{v}\times\boldsymbol{b}\right>\sim\,\alpha\!\left<\boldsymbol{B}_{0}\right>$.
 This is significant because of the possibility of direct laminar dynamo action in regions of weak shear at large Reynolds number.
-Figure~\ref{fig:correlation} shows the correlation between $u$ and $b$ as a function of $x$ over the domain for $\SSC = 1.02$, proportional to the mean electromotive force.
-The correlation is non-zero, demonstrating the tendency of the 3D MRI eigenmodes to act as a mean-field dynamo.
+Figure~\ref{fig:correlation} shows the correlation (averaged over $y,\,z$) as a function of $x$ over the domain for $\SSC = 1.02$,
+\begin{equation}
+\text{EMF}\propto\frac{\left<\boldsymbol{\hat{z}}\cdot(\boldsymbol{v}\times\boldsymbol{b})\right>_{y,z}}{\boldsymbol{v}_{\text{rms}}\boldsymbol{b}_{\text{rms}}},
+\end{equation}
+where ``rms'' is denotes the average over the whole domain. 
+The correlation is independent of the arbitrary normalisation of the linear eigenvectors.
+The non-zero correlation shows the tendency for back reaction on the mean magnetic field.
 

### 2019-08-27 paper/paper.tex
Responding to Ben's comments.

--- 2019-08-25 paper/paper.tex
@@
   The gray contour highlights $\gamma=0$; the red dots indicate the fastest growing mode. 
+Two dimensional modes occur at $k_y = 0$ on the bottom of each figure.
   At $\SSC=0.640$, there are no two dimensional modes. 
@@
   The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray. 
-  d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes.}
+  d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes at $\phi = 0$. }
   \label{fig:growth_rate}
@@
 At $\SSC=1.02$, the maximum growth rate ($\gamma_{\max}/|S|=0.093$) happens at $(k_{y},k_{z})\simeq(0.263,0.447)$.
-The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability.
-The overall critical shear for three-dimensional modes is $S_{c,\text{3D}}\simeq0.102S_c$.
+Two-dimensional instability occurs for modes with finite growth rates at $k_y = 0$ along the bottom of the figure.
+In all cases in figure~\ref{fig:growth_rate}a, the maximum growth rate does not occur along this line, indicating the dominance of three-dimensional modes.
+The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability: the zero growth rate contour (highlighted in grey) does not intercept the x-axis.
 
@@
 Figure~\ref{fig:growth_rate}d shows the phase angle $\phi=\arctan(k_{y}/k_{z})$ of the fastest growing mode as function of $\SSC$.
+The overall critical shear for three-dimensional modes is $S_{c,\text{3D}}\simeq0.102S_c$.
 Above $\SSC\gtrsim2.05$ $\phi=0$, indicating that axisymmetric modes have the fastest growth rates.
 For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC\simeq4.84$ (most works use similar values).
-Thus, we predict axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
+Thus, our theory predicts axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
 
@@
   \caption{(top) Growth rates for $q=1/10$ relevant to the stellar interiors. 
-  (bottom) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta=0.1$.}
+  (bottom) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta=0.1$.
+The lack of two dimensional instability is evident by the fact that the zero contour does not intercept the x-axis.}
   \label{fig:other_params}
@@
 We thank Steve Tobias for helpful conversations on this work.
-JSO acknowledges funding from NASA LWS grant No.
-NNX16AC92G and Research Corporation Scialog Collaborative Award (TDA) ID\#24231.
+Oishi, Baxter, and Brown acknowledge support from NASA LWS grant No. NNX16AC92G.
+Oishi also acknowledges support from Research Corporation Scialog Collaborative Award (TDA) ID\#24231. 
 Computations were performed on the \emph{Leavitt} cluster at the Bates College High Performance Computing Center.
-Morgan Baxter acknowledges support from NASA LWS grant No. NNX16AC92G.
+
 \end{acknowledgments}

### 2019-08-27 python/mri_single_yz_mode.py
Check for spurious eigenmodes via eigentools

@@ -17,6 +17,7 @@ from pathlib import Path
 import numpy as np
 import os
 import h5py
+from eigentools import Eigenproblem
 import dedalus.public as de
 from mpi4py import MPI
 CW = MPI.COMM_WORLD
@@ -156,35 +157,22 @@ if not ideal:
     problem.add_bc("right(jxx) = 0")
 
 # GO
-
-solver = problem.build_solver()
-def ideal_2D(kz):
-    kk, BB = kz*kz, B*B
-    a = kx*kx + kk
-    b = kk*(2*BB*a + f*(f+S))
-    c = (BB*kk**2)*(BB*a + f*S)
-    return np.sqrt( ( - b + np.sqrt(b*b - 4*a*c + 0j) ) / (2*a) )
-
+EP = Eigenproblem(problem,sparse=False)
 t1 = time.time()
-solver.solve_dense(solver.pencils[0], rebuild_coeffs=True)
+gr, idx, freq = EP.growth_rate({})
 t2 = time.time()
-
 logger.info("Solve time: {}".format(t2-t1))
 
-gamma = solver.eigenvalues
-dense_threshold=1
-
-thresh = (np.abs(gamma) < dense_threshold)
-gamma = gamma[thresh]
-solver.eigenvectors = solver.eigenvectors[:,thresh]
+gamma = EP.evalues_good
+eigenvectors = EP.solver.eigenvectors[:,EP.evalues_good_index]
 index = np.argsort(-gamma.real)
 
 logger.info("saving eigenvector with gamma = {}".format(gamma[index[0]]))
 nvars = len(problem_variables)
 eigvec = np.zeros((nvars,Nx),dtype=np.complex128)
-solver.set_state(index[0])
+EP.solver.set_state(index[0])
 for k in range(nvars):
-    eigvec[k,:] = solver.state[problem_variables[k]]['g']
+    eigvec[k,:] = EP.solver.state[problem_variables[k]]['g']
 
 # Save either or both eigenvalues and eigenvectors to a single .h5 file
 # Output file will be the .cfg file name with _output.h5
@@ -194,7 +182,7 @@ if CW.rank == 0:
         tail = '_ideal' + tail
     output_file_name = Path(filename.stem + tail)
     output_file = h5py.File(outbase/output_file_name, 'w')
-    dset_eval = output_file.create_dataset('eigvals',data=solver.eigenvalues)
[…]

### 2019-08-27 python/plot_fig1.py
Updated figure 1
* added labels for subpanels
* fixed incorrect plot label
* only plot non-spurious spectra

@@ -10,7 +10,7 @@ plt.style.use('prl')
 # params
 filebase = 'data/run_{:d}_output.h5'
 runs = [15, 39, 12, 13]
-SSC = [0.64, 1.002001, 1.0201, 2.]
+SSC = [0.64, 1.002001, 1.2, 2.]
 q = 0.75
 index = 0
 
@@ -48,8 +48,10 @@ for i,r in enumerate(runs):
 
     ax.contour(kz_global,ky_global,gamma_r,levels=contour_levels,colors='k')
     ax.contour(kz_global,ky_global,gamma_r,levels=[0.],colors='grey')
-    ax.text(0.05,0.9,r'$S/S_c = {:5.3f}$'.format(SSC[i]), color='w',fontsize=20)
-    ax.text(0.05,0.8,r'$\gamma_{{max}}/|S| = {:5.3f}$'.format(gamma_r.max()), color='w',fontsize=16)
+    if i == 0:
+        ax.text(0.05,0.9,r'a)'.format(SSC[i]), color='w',fontsize=20)
+    ax.text(0.05,0.8,r'$S/S_c = {:5.3f}$'.format(SSC[i]), color='w',fontsize=20)
+    ax.text(0.05,0.73,r'$\gamma_{{max}}/|S| = {:5.3f}$'.format(gamma_r.max()), color='w',fontsize=16)
     ax.plot(max_kz, max_ky, 'ro')
     #ax.colorbar(PCM, label=r'$\gamma$')
     ax.set_xlabel(r'$k_z$')
@@ -95,7 +97,7 @@ asymp_ax.plot(ky,gamma_r[:,index]/q, label='numerical')
 asymp_ax.plot(ky_an, omega/q, label='asymptotic')
 asymp_sub.plot(ky,-(gamma_r[:,index]/q + 1j*gamma_i[:,index]/q)**2, label='numerical')
 asymp_sub.plot(ky_an, omega2/q**2, label='asymptotic')
-
+asymp_ax.text(0.005,0.07, 'b)',fontsize=20)
 #asymp_ax.legend(loc='upper right')
 asymp_ax.text(0.11, 0.058, 'numerical', fontsize=18)
 asymp_ax.text(0.03, 0.07, 'asymptotic', fontsize=18)
@@ -131,14 +133,16 @@ spec_thresh = spectrum[np.abs(spectrum) < thresh]/q
[…]

### 2019-08-27 paper/paper.tex
Minor updates
* added eigentools reference
* joined figure 4 into one pdf to comply with PRL
* minor text edits from Ben

@@
   The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray. 
-  d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes at $\phi = 0$. }
+  d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes ($\phi = 0$) at $\SSC = \gtrsim2.05$. }
   \label{fig:growth_rate}
@@
 The other stable modes (gray dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
+We reject spurious eigenvalues by using \texttt{eigentools} \footnote{\protect\url{https://bitbucket.org/jsoishi/eigentools}} to solve equations~(\ref{eq:mhd}-\ref{eq:divu}) at two resolutions and retain only pairs equal to within one part in $10^{-6}$.
 
@@
 The overall critical shear for three-dimensional modes is $S_{c,\text{3D}}\simeq0.102S_c$.
-Above $\SSC\gtrsim2.05$ $\phi=0$, indicating that axisymmetric modes have the fastest growth rates.
+Above $\SSC\gtrsim2.05$, $\phi$ becomes zero, indicating that axisymmetric modes have the fastest growth rates.
 For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC\simeq4.84$ (most works use similar values).
@@
   \centering
-  \includegraphics[width=\columnwidth]{run_47_output_growthrates.pdf}\\
-  \includegraphics[width=\columnwidth]{run_49_output_growthrates.pdf}
-  \caption{(top) Growth rates for $q=1/10$ relevant to the stellar interiors. 
-  (bottom) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta=0.1$.
+  \includegraphics[width=\columnwidth]{low_rossby_liquid_metal_growth_rates.pdf}
+  \caption{(a) Growth rates for $q=1/10$ relevant to the stellar interiors. 
+  (b) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta=0.1$.
 The lack of two dimensional instability is evident by the fact that the zero contour does not intercept the x-axis.}

### 2019-08-27 python/dynamo_mri.py
Removed script

@@ -1,309 +1,0 @@
-"""
-The magnetorotational instability prefers three dimensions.
-    
-Dedalus script for calculating the maximum growth rates for the
-3D MRI over a range of wavenumbers.
-
-This script can be ran serially or in parallel, and produces an h5py file
-that contains maximum eigenvalues and the corresponding eigenvectors.
-
-Pass it a .cfg file to specify parameters for the particular solve.
-
-To run using 4 processes, you would use:
-    $ mpiexec -n 4 python3 mri.py mri_params.cfg
-
-"""
-
-import time
-import configparser
-from configparser import ConfigParser
-import argparse
-from pathlib import Path
-import numpy as np
-import scipy.optimize as opt
-import os
-import h5py
-import matplotlib.pyplot as plt
-import dedalus.public as de
-from mpi4py import MPI
-CW = MPI.COMM_WORLD
-import logging
-logger = logging.getLogger(__name__)
-
-# Parses .cfg filename passed to script
-parser = argparse.ArgumentParser(description='Passes filename')
-parser.add_argument('filename', metavar='Rc', type=str, help='.h5 file to plot eigenvectors for maximum eigenvalue')
-args = parser.parse_args()
-filename = Path(vars(args)['filename'])
-outbase = Path("data")
-
-# Parse .cfg file to set global parameters for script
-config = ConfigParser()
-config.read(str(filename))
-
-logger.info('Running mri.py with the following parameters:')
-logger.info(config.items('parameters'))
-
-try:
-    dense = config.getboolean('solver','dense')
-except:
-    dense = False
-if dense:
-    logger.info("Using dense solver.")
-    dense_threshold = config.getfloat('solver','dense_threshold')
-else:
-    logger.info("Using sparse solver.")
-
-Nx = config.getint('parameters','Nx')
-Lx = eval(config.get('parameters','Lx'))
-B = config.getfloat('parameters','B')
-
-Nmodes = config.getint('parameters','Nmodes')
-
-R      =  config.getfloat('parameters','R')
-q      =  config.getfloat('parameters','q')
-
[…]

### 2019-08-27 paper/paper.tex
fixed a couple typos

@@
   The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray. 
-  d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes ($\phi = 0$) at $\SSC = \gtrsim2.05$. }
+  d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes ($\phi=0$) at $\SSC\gtrsim2.05$. }
   \label{fig:growth_rate}
@@
 In all cases in figure~\ref{fig:growth_rate}a, the maximum growth rate does not occur along this line, indicating the dominance of three-dimensional modes.
-The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability: the zero growth rate contour (highlighted in grey) does not intercept the x-axis.
+The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability: the zero growth rate contour (highlighted in grey) does not intercept the $x$-axis.
 

### 2019-08-27 paper/supplemental.tex
included 1st-order formulation in supp.

@@
 %\label{simulations}
+
+\section{1st-order formulation of the dissipative equations}
+\label{sec:formulation}
+
+We formulate the system in terms of the 10 problem \texttt{variables} 
+\Beq
+X \ = \ \left[\,p,v_{x},v_{y},v_{z},\omega_{y},\omega_{z},b_{x},b_{y},b_{z},j_{xx}\,\right] \nonumber
+\Eeq
+As sting-based \texttt{substitutions} 
+\Beq
+``\omega_{x}" &=& ``\pd{y} v_{z} - \pd{z} v_{y}" \nonumber \\
+``j_{x}" &=& ``\pd{y} b_{z} - \pd{z} b_{y}" \nonumber \\
+``j_{y}" &=& ``\pd{z} b_{x} - \pd{x} b_{z}" \nonumber \\
+``j_{z}" &=& ``\pd{x} b_{y} - \pd{y} b_{x}" \nonumber 
+\Eeq
+The expression for $\omega_{x},\,j_{x},\,j_{y},\,j_{z}$ are not legitimate variables that \textit{Dedalus} solves for.
+They are only short-hand aliases that mean literally the same thing as the expression on the right-hand side of the definition, which are defined strictly in terms of legitimate problem variables.
+
+We also define the following operator substitutions 
+\Beq
+``\pd{t} q" &=& ``\sigma q" \nonumber \\
+``\pd{y} q" &=& ``i k_{y} q" \nonumber \\
+``\pd{z} q" &=& ``i k_{z} q" \nonumber 
+\Eeq
+where $\sigma$ is the problem \texttt{eigenvalue}, and $k_{y},\,k_{z}$ are problem \texttt{parameters}, along with $f,\,S,\,B_{0},\, \nu,\, \eta$.
+
+The system of \texttt{equations} is:
+\Beq\label{u-eq}
+\pd{x} v_{x} + \pd{y} v_{y} + \pd{z} v_{z}  &=& 0\\
+\pd{t} v_{x} + S x \pd{y} v_{x} - f v_{y} + \pd{x} p - B_{0} \pd{z} b_{x}  + \nu ( \pd{y} \omega_{z} - \pd{z} \omega_{y}) &=& 0\\
+\pd{t} v_{y} + S x \pd{y} v_{y} + (f+S) v_{x} + \pd{y} p - B_{0} \pd{z} b_{y} + \nu ( \pd{z} \omega_{x} - \pd{x} \omega_{z})&=& 0\label{v-eq} \\
+\pd{t} v_{z} + S x \pd{y} v_{z} + \pd{z} p - B_{0} \pd{z} b_{z} + \nu ( \pd{x} \omega_{y} - \pd{y} \omega_{x})&=& 0 \\
+\omega_{y} - \pd{z} v_{x} + \pd{x} v_{z} &=& 0 \\
+\omega_{z} - \pd{x} v_{y} + \pd{y} v_{x} &=& 0 \\
+\pd{x} b_{x} + \pd{y} b_{y} + \pd{z} b_{z}   &=& 0\\
+\pd{t} b_{x} + S x \pd{y} b_{x} - B_{0} \pd{z} v_{x} + \eta ( \pd{y} j_{z} - \pd{z} j_{y})&=& 0\\
+\pd{t} j_{x}  + S x \pd{y} j_{x}- \eta( \pd{x}j_{xx} + \pd{y}^{2} j_{x} + \pd{z}^{2} j_{x} )  - B_{0} \pd{z} \omega_{x} + S \pd{z} b_{x} &=& 0\\
+j_{xx} - \pd{x} j_{x} &=& 0
+\Eeq
+We can see there are the same number of equations as variables, each equations has only 1st-order derivatives in $x$, and every variable has at least one $\pd{x}$.
+It appears (for example) that there are no $\pd{x}$ terms in the $b_{x}$ equation, but this due to the substitutions,
+\Beq
+\pd{y} j_{z} - \pd{z} j_{y} =  \pd{y}(\pd{x} b_{y} - \pd{y} b_{x}) - \pd{z}(\pd{z} b_{x} - \pd{x} b_{z})  =  \pd{x}( \pd{y}b_{y} + \pd{z}b_{z}) - (\pd{y}^{2} + \pd{z}^{2})b_{x}.
+\Eeq
+
+The entire system is formulated and solved as a generalized eigenvalue problem with non-constant coefficients 
+\Beq
+\sigma \, M \cdot X + L_{0}(x) \cdot X + L_{1}\cdot \pd{x} X \ = \ 0
+\Eeq
+The system is 10th order because $L_{1}$ has full rank. We therefore impose the 10 \texttt{boundary conditions} 
+\Beq
+v_{x} \ = \ \omega_{y} \ = \ \omega_{z} \ = \ b_{x} \ = \ j_{xx} \ = \ 0 \quad \text{at} \quad x \ = \ \pm d/2
+\Eeq
+We could choose any other set of boundary conditions constructed out of a linearly independent combinations of the legitimate problem variables.
+We choose impenetrable stress-free and perfectly conducting because it agrees with the purely impenetrable conditions for an ideal calculation.
 

### 2019-08-28 paper/supplemental.tex
tweaks to supp.

--- 2019-08-25 paper/supplemental.tex
@@
 %\label{simulations}
+
+\section{1st-order formulation of the dissipative equations}
+\label{sec:formulation}
+
+We formulate the system in terms of the 10 problem \texttt{variables} 
+\Beq
+X \ = \ \left[\,p,v_{x},v_{y},v_{z},\omega_{y},\omega_{z},b_{x},b_{y},b_{z},j_{xx}\,\right]^{\top} \nonumber
+\Eeq
+We use sting-based \texttt{substitutions} to expedite the input of the full the system of equations: 
+\Beq
+``\omega_{x}" &=& ``\pd{y} v_{z} - \pd{z} v_{y}" \nonumber \\
+``j_{x}" &=& ``\pd{y} b_{z} - \pd{z} b_{y}" \nonumber \\
+``j_{y}" &=& ``\pd{z} b_{x} - \pd{x} b_{z}" \nonumber \\
+``j_{z}" &=& ``\pd{x} b_{y} - \pd{y} b_{x}" \nonumber 
+\Eeq
+The expression for $\omega_{x},\,j_{x},\,j_{y},\,j_{z}$ are not legitimate variables that \textit{Dedalus} solves for.
+They are only short-hand aliases that mean literally the same thing as the expression on the right-hand side of the definition, which are defined strictly in terms of legitimate problem variables.
+
+We also define the following operator substitutions (with general argument $A$)
+\Beq
+``\pd{y} A" &=& ``i k_{y} A" \nonumber \\
+``\pd{z} A" &=& ``i k_{z} A" \nonumber \\
+``D_{t} A" &=& ``\sigma q +S x \pd{y} A " \nonumber \\
+``\Delta A" &=& ``\pd{y}^{2} A + \pd{z}^{2}A"
+\Eeq
+where $\sigma$ is the problem \texttt{eigenvalue}, and $k_{y},\,k_{z}$ are problem \texttt{parameters}, along with $f,\,S,\,B_{0},\, \nu,\, \eta$.
+
+The system of \texttt{equations} is:
+\Beq\label{u-eq}
+\pd{x} v_{x} + \pd{y} v_{y} + \pd{z} v_{z}  &=& 0\\
+D_{t} v_{x}  - f v_{y} + \pd{x} p - B_{0} \pd{z} b_{x}  + \nu ( \pd{y} \omega_{z} - \pd{z} \omega_{y}) &=& 0\\
+D_{t} v_{y}  + (f+S) v_{x} + \pd{y} p - B_{0} \pd{z} b_{y} + \nu ( \pd{z} \omega_{x} - \pd{x} \omega_{z})&=& 0\label{v-eq} \\
+D_{t} v_{z} + \pd{z} p - B_{0} \pd{z} b_{z} + \nu ( \pd{x} \omega_{y} - \pd{y} \omega_{x})&=& 0 \\
+\omega_{y} - \pd{z} v_{x} + \pd{x} v_{z} &=& 0 \\
+\omega_{z} - \pd{x} v_{y} + \pd{y} v_{x} &=& 0 \\
+\pd{x} b_{x} + \pd{y} b_{y} + \pd{z} b_{z}   &=& 0\\
+D_{t} b_{x} - B_{0} \pd{z} v_{x} + \eta ( \pd{y} j_{z} - \pd{z} j_{y})&=& 0\\
+D_{t} j_{x}  - \eta( \pd{x}j_{xx} +  \Delta j_{x} )  - B_{0} \pd{z} \omega_{x} + S \pd{z} b_{x} &=& 0\\
+j_{xx} - \pd{x} j_{x} &=& 0
+\Eeq
+We can see there are the same number of equations as variables, each equations has only 1st-order derivatives in $x$, and every variable has at least one $\pd{x}$.
+It appears (for example) that there are no $\pd{x}$ terms in the $b_{x}$ equation, but this due to the substitutions,
+\Beq
+\pd{y} j_{z} - \pd{z} j_{y} =  \pd{y}(\pd{x} b_{y} - \pd{y} b_{x}) - \pd{z}(\pd{z} b_{x} - \pd{x} b_{z})  =  \pd{y}\pd{x}b_{y} + \pd{z}\pd{x}b_{z} - (\pd{y}^{2} + \pd{z}^{2})b_{x}.
+\Eeq
+Also,
+\Beq
+\pd{x}j_{x} \ = \ \pd{y} \pd{x} b_{z} - \pd{z} \pd{x}b_{y}.
+\Eeq
+The entire system is formulated and solved as a generalized eigenvalue problem with non-constant coefficients 
+\Beq
+\sigma \, M \cdot X + L_{0}(x) \cdot X + L_{1}\cdot \pd{x} X \ = \ 0.
+\Eeq
+The problem is ``generalized" because $M$ is not full rank. The system is 10th order because $L_{1}$ is full rank. We therefore impose the 10 \texttt{boundary conditions} 
+\Beq
+v_{x} \ = \ \omega_{y} \ = \ \omega_{z} \ = \ b_{x} \ = \ j_{xx} \ = \ 0 \quad \text{at} \quad x \ = \ \pm d/2
+\Eeq
+We could choose any other set of boundary conditions constructed out of a linearly independent combinations of the legitimate problem variables.
+We choose impenetrable stress-free and perfectly conducting because it conforms with the purely impenetrable conditions for an ideal calculation, thus minimizing the arbitrary influence of boundary layers.
 

### 2019-08-28 python/mri_single_yz_mode.py
include no-slip + vacuum BC option in single ky-kz script

@@ -8,6 +8,7 @@ Usage:
 
 Options:
     --ideal    Use Ideal MHD
+    --hardwall Use experimental boundary conditions
 """
 
 from docopt import docopt
@@ -28,6 +29,7 @@ args = docopt(__doc__)
 ky = float(args['<ky>'])
 kz = float(args['<kz>'])
 ideal = args['--ideal']
+hardwall = args['--hardwall']
 filename = Path(args['<config_file>'])
 outbase = Path("data")
 
@@ -146,15 +148,30 @@ else:
 problem.add_bc("left(vx)   = 0")
 problem.add_bc("right(vx)  = 0")
 if not ideal:
-    problem.add_bc("left(bx)   = 0")
-    problem.add_bc("left(ωy)   = 0")
-    problem.add_bc("left(ωz)   = 0")
-    problem.add_bc("left(jxx)  = 0")
+    if hardwall:
+        problem.add_bc("left(vy)   = 0")
+        problem.add_bc("left(vz)   = 0")
+        problem.add_bc("left(jx)   = 0")
+        problem.add_bc("left(jx)   = 0")
+        problem.add_bc("left(dy(by)+dz(bz) + sqrt(ky**2 + kz**2)*bx) = 0")
+        
+        problem.add_bc("right(vy)   = 0")
+        problem.add_bc("right(vz)   = 0")
+        problem.add_bc("right(jx)   = 0")
+        problem.add_bc("right(dy(by)+dz(bz) - sqrt(ky**2 + kz**2)*bx) = 0")
+    
+    else:
+        problem.add_bc("left(ωy)   = 0")
+        problem.add_bc("left(ωz)   = 0")
+        problem.add_bc("left(bx)   = 0")
+        problem.add_bc("left(jxx)  = 0")
 
-    problem.add_bc("right(bx)  = 0")
-    problem.add_bc("right(ωy)  = 0")
-    problem.add_bc("right(ωz)  = 0")
-    problem.add_bc("right(jxx) = 0")
+        problem.add_bc("right(ωy)  = 0")
+        problem.add_bc("right(ωz)  = 0")
+        problem.add_bc("right(bx)  = 0")
+        problem.add_bc("right(jxx) = 0")
+
+
 
 # GO
 EP = Eigenproblem(problem,sparse=False)

### 2019-08-28 python/mri_single_yz_mode.py
fixed silly bug

@@ -152,7 +152,6 @@ if not ideal:
         problem.add_bc("left(vy)   = 0")
         problem.add_bc("left(vz)   = 0")
         problem.add_bc("left(jx)   = 0")
-        problem.add_bc("left(jx)   = 0")
         problem.add_bc("left(dy(by)+dz(bz) + sqrt(ky**2 + kz**2)*bx) = 0")
         
         problem.add_bc("right(vy)   = 0")

### 2019-08-28 paper/supplemental.tex
make the supp. consistent with notation between the sections

@@
 \Beq\label{u-eq}
-\pd{t} u + S x \pd{y} u - f v + \pd{x} p - B_{0} \pd{z} b_{x} &=& 0\\
-\pd{t} v + S x \pd{y} v + (f+S) u + \pd{y} p - B_{0} \pd{z} b_{y} &=& 0\label{v-eq} \\
-\pd{t} w + S x \pd{y} w + \pd{z} p - B_{0} \pd{z} b_{z} &=& 0 \\
-\pd{x} u + \pd{y} v + \pd{z} w  &=& 0\\
-\pd{t} b_{x} + S x \pd{y} b_{x} - B_{0} \pd{z} u &=& 0\\
-\pd{t} b_{y} + S x \pd{y} b_{y} - B_{0} \pd{z} v - S b_{x}  &=& 0\\
-\pd{t} b_{z} + S x \pd{y} b_{z} - B_{0} \pd{z} w &=& 0 \label{bz-eq}
+\pd{t} v_{x} + S x \pd{y} v_{x} - f v_{y} + \pd{x} p - B_{0} \pd{z} b_{x} &=& 0\\
+\pd{t} v_{y} + S x \pd{y} v_{y} + (f+S) v_{x} + \pd{y} p - B_{0} \pd{z} b_{y} &=& 0\label{v-eq} \\
+\pd{t} v_{z} + S x \pd{y} v_{z} + \pd{z} p - B_{0} \pd{z} b_{z} &=& 0 \\
+\pd{x} v_{x} + \pd{y} v_{y} + \pd{z} v_{z}  &=& 0\\
+\pd{t} b_{x} + S x \pd{y} b_{x} - B_{0} \pd{z} v_{x} &=& 0\\
+\pd{t} b_{y} + S x \pd{y} b_{y} - B_{0} \pd{z} v_{y} - S b_{x}  &=& 0\\
+\pd{t} b_{z} + S x \pd{y} b_{z} - B_{0} \pd{z} v_{z} &=& 0 \label{bz-eq}
 \Eeq
+The divergence-free condition for the velocity implies the same for the magnetic field.
+The system is 2nd order because $\pd{x}$ only appears in two places.
 All variables take the form
 \Beq
-g = \hat{G}(x) e^{ i ( \omega t + k z + \ell y ) } 
+p = \hat{P}(x) e^{ i ( \omega t + k z + \ell y ) } 
 \Eeq
@@
 \Eeq
-We can use equations (\ref{v-eq}--\ref{bz-eq}) to find all amplitudes in terms of $\hat{U}(x)$ and $\hat{U}'(x)$. 
+We can use equations (\ref{v-eq}--\ref{bz-eq}) to find all amplitudes in terms of $\hat{V}_{x}(x)$ and $\hat{V}_{x}'(x)$. 
 \Beq
-\hat{V}(x) &=& \frac{i \left(\ell 
-   \hat{U}'(x) + k^{2} \frac{ \left((f+S) \omega
+\hat{V}_{y}(x) &=& \frac{i \left(\ell 
+   \hat{V}_{x}'(x) + k^{2} \frac{ \left((f+S) \omega
    _S^2-S \omega _A^2\right)}{\omega _S
-   \left(\omega _S^2-\omega _A^2\right)} \hat{U}(x)\right)}{k^2+\ell ^2} \\
-\hat{W}(x) &=& \frac{i k \left(\hat{U}'(x)- \frac{ \left((f+S) \omega
+   \left(\omega _S^2-\omega _A^2\right)} \hat{V}_{x}(x)\right)}{k^2+\ell ^2} \\
+\hat{V}_{z}(x) &=& \frac{i k \left(\hat{V}_{x}'(x)- \frac{ \left((f+S) \omega
    _S^2-S \omega _A^2\right)}{\omega _S
    \left(\omega _S^2-\omega _A^2\right)}  \ell 
-   \hat{U}(x)\right)}{k^2+\ell ^2} \\
+   \hat{V}_{x}(x)\right)}{k^2+\ell ^2} \\
    \hat{P}(x) &=& \frac{i \left(\omega _A^2-\omega _S^2\right)
-   \left(\hat{U}'(x) -  \ell \frac{ \left((f+S) \omega
+   \left(\hat{V}_{x}'(x) -  \ell \frac{ \left((f+S) \omega
    _S^2-S \omega _A^2\right)}{\omega _S
    \left(\omega _S^2-\omega _A^2\right)}
-   \hat{U}(x)\right)}{\left(k^2+\ell ^2\right)
+   \hat{V}_{x}(x)\right)}{\left(k^2+\ell ^2\right)
    \omega _S} \\
-   \hat{B}_{x}(x) &=& \frac{\omega _A \hat{U}(x)}{\omega _S} \\
-   \hat{B}_{y}(x) &=&\frac{\omega _A \hat{V}(x)}{\omega _S} -\frac{i S \omega _A \hat{U}(x)}{\omega _S^2} \\
-   \hat{B}_{z}(x) &=& \frac{\omega _A \hat{W}(x)}{\omega _S}
+   \hat{B}_{x}(x) &=& \frac{\omega _A \hat{V}_{x}(x)}{\omega _S} \\
+   \hat{B}_{y}(x) &=&\frac{\omega _A \hat{V}_{y}(x)}{\omega _S} -\frac{i S \omega _A \hat{V}_{x}(x)}{\omega _S^2} \\
+   \hat{B}_{z}(x) &=& \frac{\omega _A \hat{V}_{z}(x)}{\omega _S}
    \Eeq
-Substitute everything into equation (\ref{u-eq}) to get a second-order equation for $\hat{U}(x)$ of the general form 
+Substitute everything into equation (\ref{u-eq}) to get a second-order equation for $\hat{V}_{x}(x)$ of the general form 
 \Beq
-\alpha(x) \hat{U}''(x) + \beta(x) \hat{U}'(x) + \gamma(x) \hat{U}(x) \ = \ 0
+\alpha(x) \hat{V}_{x}''(x) + \beta(x) \hat{V}_{x}'(x) + \gamma(x) \hat{V}_{x}(x) \ = \ 0
 \Eeq
@@
 \Beq
-\hat{U}(x) \ = \ \chi(x) \psi(x) \quad \text{where} \quad \chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
+\hat{V}_{x}(x) \ = \ \chi(x) \psi(x) \quad \text{where} \quad \chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
 \Eeq

### 2019-08-28 paper/paper.tex
tiny tweak
Dan's comments.

@@
 \author{Daniel~Lecoanet}
-\affiliation{Princeton Center for Theoretical Science and Princeton University Department of Astrophysical Sciences, Princeton, NJ 08544}
+\affiliation{Princeton University Department of Astrophysical Sciences, Princeton, NJ 08544}
 \author{Benjamin~P.~Brown}
@@
 The MRI is considered essential to the operation of astrophysical accretion disks, with the central object supplying a Keplerian rotation profile.
-The MRI should also exist for the internal shear layers of stars.
+Internal shear layers in stars may also be unstable to the MRI.
 Stellar-interior shears can take a wide range of profiles, including near-critical, but with negligible dissipation. 
@@
   a) growth rates $\gamma$ over a grid of $k_{y}$ and $k_{z}$ for four values of supercriticality $\SSC$. 
+  The five black contours are equally spaced between zero and the maximum growth rate.
   The gray contour highlights $\gamma=0$; the red dots indicate the fastest growing mode. 
@@
 \begin{equation}
-\text{EMF}\propto\frac{\left<\boldsymbol{\hat{z}}\cdot(\boldsymbol{v}\times\boldsymbol{b})\right>_{y,z}}{\boldsymbol{v}_{\text{rms}}\boldsymbol{b}_{\text{rms}}},
+\text{EMF}\propto\frac{\left<\boldsymbol{\hat{z}}\cdot(\boldsymbol{v}\times\boldsymbol{b})\right>_{\!y,z}}{\boldsymbol{v}_{\text{rms}}\boldsymbol{b}_{\text{rms}}},
 \end{equation}
@@
 The NSSL, in particular, may also host slower MRI-driven dynamics; despite containing small-scale convection.
-It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to noisy convection.
+It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to convection.
 Second, our low-$\Prm$ results suggest that three-dimensional MRI modes may be the easiest to excite in liquid metal experiments.

### 2019-08-29 paper/supplemental.tex
tweak to supp
*yet* another small tweak
fixed typo
full-rank L1
put in the actual matrices. this is a good example of how dedalus works
Minor changes; typo fixes; as submitted!

@@
 \Eeq
-We use sting-based \texttt{substitutions} to expedite the input of the full the system of equations: 
+We use string-based \texttt{substitutions} to expedite the input of the full the system of equations: 
 \Beq
@@
 \Eeq
-The expression for $\omega_{x},\,j_{x},\,j_{y},\,j_{z}$ are not legitimate variables that \textit{Dedalus} solves for.
-They are only short-hand aliases that mean literally the same thing as the expression on the right-hand side of the definition, which are defined strictly in terms of legitimate problem variables.
+The expression for $\omega_{x},\,j_{x},\,j_{y},\,j_{z}$ are not variables that \textit{Dedalus} solves for. 
+They are short-hand aliases that are directly replaced by the expressions on the right-hand side of the definition, which are themselves defined strictly in terms of problem variables.
 
@@
 ``\pd{z} A" &=& ``i k_{z} A" \nonumber \\
-``D_{t} A" &=& ``\sigma q +S x \pd{y} A " \nonumber \\
-``\Delta A" &=& ``\pd{y}^{2} A + \pd{z}^{2}A"
+``D_{t} A" &=& ``\sigma A +S x \pd{y} A " \nonumber \\
+``\Delta A" &=& ``\pd{y}^{2} A + \pd{z}^{2}A" \nonumber
 \Eeq
@@
 \Eeq
-The problem is ``generalized" because $M$ is not full rank. The system is 10th order because $L_{1}$ is full rank. We therefore impose the 10 \texttt{boundary conditions} 
+The problem is ``generalized" because $M$ is not full rank. The system is 10th order because $L_{1}$ is full rank,
+\Beq
+\text{det}(L_{1}) \ = \ -\eta ^2 \nu ^2 \left(k_y^2+k_z^2\right).
+\Eeq
+We therefore impose the 10 \texttt{boundary conditions} 
 \Beq
@@
 \Eeq
-We could choose any other set of boundary conditions constructed out of a linearly independent combinations of the legitimate problem variables.
+We could choose any other set of boundary conditions constructed out of a linearly independent combinations of the problem variables. 
 We choose impenetrable stress-free and perfectly conducting because it conforms with the purely impenetrable conditions for an ideal calculation, thus minimizing the arbitrary influence of boundary layers.
+
+Respectively $M,\,L_{0}(x),\,L_{1}$:
+\Beq
+&
+\tiny \nonumber
+\left[
+\begin{array}{cccccccccc}
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & -i k_z & i k_y & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+\end{array}
+\right]\\
+&
+\tiny \nonumber
+\left[
+\begin{array}{cccccccccc}
+ 0 & 0 & i k_y & i k_z & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & i S x k_y & -f & 0 & -i \nu  k_z & i \nu  k_y
+   & -i B_{0} k_z & 0 & 0 & 0 \\
+ i k_y & f+S & \nu  k_z^2+i S x k_y & -\nu  k_y
+   k_z & 0 & 0 & 0 & -i B_{0} k_z & 0 & 0 \\
+ i k_z & 0 & -\nu  k_y k_z & k_y \left(\nu  k_y+i
+   S x\right) & 0 & 0 & 0 & 0 & -i B_{0} k_z & 0 \\
+ 0 & -i k_z & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & i k_y & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & i k_y & i k_z & 0 \\
+ 0 & -i B_{0} k_z & 0 & 0 & 0 & 0 & \eta  |k|^2+i S x
+   k_y & 0 & 0 & 0 \\
+ 0 & 0 & -B_{0} k_z^2 & B_{0} k_y k_z & 0 & 0 & i S k_z &
+   k_z \left(S x k_y-i \eta  |k|^2\right) & k_y
+   \left(-S x k_y+i \eta  |k|^2\right) & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\
+\end{array}
+\right]\\
+&
+\tiny \nonumber
+\left[
+\begin{array}{cccccccccc}
+ 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 0 & 0 & -\nu  & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 0 & \nu  & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & -1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & i \eta  k_y & i \eta 
+   k_z & 0 \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & -\eta  \\
+ 0 & 0 & 0 & 0 & 0 & 0 & 0 & i k_z & -i k_y & 0 \\
+\end{array}
+\right]
+\Eeq
 
@@
 \label{sec:asymp}
-Here, we outline the asymptotic calculation for the leading order correction to the 2D growth rates in when accouting for three dimensional effects via finite $k_y$.
+Here, we outline the asymptotic calculation for the leading order correction to the 2D growth rates in when accounting for three dimensional effects via finite $k_y$.
 Define the domain as $ - d/2 < x < d/2$. Being in the rotating frame means the shear has no net
 \Beq
-\int_{-d/2}^{d/2} V_{y}(x) \text{d} x  \ = \ 0, \quad \implies  \quad V_{y}(x) \ = \ S\, x
+\int_{-d/2}^{d/2} V_{0,y}(x) \text{d} x  \ = \ 0 \quad \implies  \quad V_{0,y}(x) \ = \ S\, x
 \Eeq
@@
 \end{document}
+
+%  LocalWords:  Magnetorotational MHD EigenValueProblem Dedalus th nd
+%  LocalWords:  parallelized Chebyshev cccccccccccccc Rossby det Sx
+%  LocalWords:  cccccccccc diffusivities analytical Schr dinger
+%  LocalWords:  perturbatively rescale solvability

### 2019-09-04 paper/nature_phys_cover_letter.tex
Added cover letter template for nature physics.

\documentclass{letter}

\usepackage{graphicx}

\begin{document}
\begin{letter}{Nature Physics\\London, UK}

  \opening{Dear Editor:}

  We are submitting a manuscript titled ``The Magnetorotational Instability Prefers Three Dimensions'' for consideration in \emph{Nature Physics}.

\fromsig{\includegraphics[scale=0.5]{jso_sig.pdf}}\\
\fromname{Jeffrey S. Oishi,\\on behalf of the authors}}

\end{letter}
\end{document}

### 2019-09-05 paper/nature_phys_cover_letter.tex
Added letter content.

@@
 \documentclass{letter}
-
+\usepackage{geometry}
 \usepackage{graphicx}
@@
 \begin{document}
-\begin{letter}{Nature Physics\\London, UK}
-
+\begin{letter}{}
   \opening{Dear Editor:}
 
-  We are submitting a manuscript titled ``The Magnetorotational Instability Prefers Three Dimensions'' for consideration in \emph{Nature Physics}.
+  We are submitting a manuscript, ``The Magnetorotational Instability Prefers Three Dimensions,'' for consideration in \emph{Nature Physics}.
+The magnetorotational instability (MRI) is an extremely important part of astrophysical fluid dynamics, as it is believed to play a key role in accretion disks around compact objects and may also play an important role in stellar interiors.
+There are several active laboratory campaigns to make an experimental confirmation of the instability.
+Our work demonstrates for the first time that the MRI is three-dimensional at onset when the shear that drives it is close to its critical value.
+Far from being an unusual part of parameter space, driving the flow to near-critical shear is how the instability saturates in systems like stars and laboratory experiments, similar to convection driving the background entropy gradient to zero or Taylor-Couette instability forcing the angular momentum gradient toward a stable profile.
+This means that in the context of stellar interiors and the laboratory experiments seeking to confirm the existence of the MRI, near-critical shear is in fact a very important regime.
+The implication of a three-dimensional MRI are twofold: the non-normal nature of the linear operator driving the instability implies that transient growth is always important in the linear phase.
+More importantly, the existance of a three-dimensional unstable mode implies that the MRI may drive dynamo action, amplifying magnetic fields even in the absence of turbulence driven by secondary instabilities.
 
+Our results have important implications for both laboratory studies of the MRI as well as its effects in stellar interiors. 
+The MRI is also an important magnetohydrodynamic process in its own right: it is of interest to plasma physicists and it bears deep connections to viscoelastic Taylor-Couette flow that have been studied in the fluid dynamics community.
+Thus, because our work constitutes an important new direction in MRI research, we believe it is well suited for the diverse readership of \emph{Nature Physics}.
+We look forward to corresponding further with you about our work.
+
+\closing{Sincerely, \\
 \fromsig{\includegraphics[scale=0.5]{jso_sig.pdf}}\\
-\fromname{Jeffrey S. Oishi,\\on behalf of the authors}}
+\fromname{Jeffrey S. Oishi,\\on behalf of the authors:\\
+Geoffrey M.\ Vasil\\
+Morgan Baxter\\
+Andrew Swan\\
+Keaton J. Burns\\
+Daniel Lecoanet\\
+Benjamin P. Brown}}
 

### 2019-09-05 paper/nature_phys_cover_letter.tex
Final natphys cover letter

--- 2019-09-04 paper/nature_phys_cover_letter.tex
@@
 \documentclass{letter}
+\usepackage[hmargin=1in,bottom=0.5in]{geometry}
+\usepackage{fancyhdr}
+\usepackage{graphicx}
+\usepackage{fontspec} 
+\renewcommand{\headrulewidth}{0pt}
+\fancypagestyle{firstpage}{\fancyhf{}\fancyhead[L]{\includegraphics[height=0.4in, keepaspectratio=true]{bates-wordmark-201.eps}\\Department of\\Physics and Astronomy}\cfoot{44 Campus Ave | Lewiston, ME 04240 | (207) 786-6325}}
+\fancypagestyle{plain}{\fancyhf{}\fancyhead[L]{\includegraphics[height=0.4in, keepaspectratio=true]{bates-wordmark-201.eps}\\Department of\\Physics and Astronomy}\cfoot{44 Campus Ave | Lewiston, ME 04240 | (207) 786-6325}}
 
-\usepackage{graphicx}
+\setromanfont[Mapping=tex-text]{Linux Libertine O}
 
 \begin{document}
-\begin{letter}{Nature Physics\\London, UK}
-
+\begin{letter}{}
   \opening{Dear Editor:}
 
-  We are submitting a manuscript titled ``The Magnetorotational Instability Prefers Three Dimensions'' for consideration in \emph{Nature Physics}.
+  We are submitting a manuscript, ``The Magnetorotational Instability Prefers Three Dimensions,'' for consideration in \emph{Nature Physics}.
+The magnetorotational instability (MRI) is an extremely important part of astrophysical fluid dynamics, as it is believed to play a key role in accretion disks around compact objects and may also play an important role in stellar interiors.
+There are several active laboratory campaigns attempting to make an experimental confirmation of the instability.
 
+Our work demonstrates for the first time that the most unstable MRI mode is three-dimensional at onset when the shear that drives it is close to its critical value.
+Previously, based on theoretical and numerical work far from critical shear, the dominant MRI modes were assumed to be axisymmetric.
+The conditions for three-dimensional instability are likely to be very common, as driving the flow to near-critical shear is how the instability saturates in systems like stars and laboratory experiments.
+This process is similar to convection erasing the background entropy gradient or Taylor-Couette instability forcing the angular momentum gradient toward marginal stability.
+%Thus, in the context of stellar interiors and the laboratory experiments seeking to confirm the existence of the MRI, near-critical shear is in fact a very important regime.
+The implication of a three-dimensional MRI are twofold: the non-normal nature of the linear operator driving the instability implies that transient growth is always important in the linear phase.
+More importantly, the existence of a three-dimensional unstable mode implies that the linear MRI may drive direct dynamo action, amplifying magnetic fields even in the absence of turbulence driven by secondary instabilities.
+
+Our results have important implications for both laboratory studies of the MRI as well as its effects in stellar interiors. 
+However, the MRI is also an important magnetohydrodynamic process in its own right.
+In particular, it is of interest to plasma physicists as a potential dynamo mechanism.
+Our work suggests the existence of a new, laminar dynamo solution.
+Such a solution is of fundamental importance in understanding dynamo generated magnetic fields, which are essential to Solar, geo-, and planetary physics.
+It also bears deep connections to viscoelastic Taylor-Couette flow that have been studied in the fluid dynamics community. 
+Thus, because our work constitutes an important new direction in MRI and dynamo research, we believe it is well suited for the diverse readership of \emph{Nature Physics}.
+We look forward to corresponding further with you about our work.
+
+\closing{Sincerely, \\
 \fromsig{\includegraphics[scale=0.5]{jso_sig.pdf}}\\
-\fromname{Jeffrey S. Oishi,\\on behalf of the authors}}
+\fromname{Jeffrey S. Oishi,\\on behalf of the authors:\\
+Geoffrey M.\ Vasil\\
+Morgan Baxter\\
+Andrew Swan\\
+Keaton J. Burns\\
+Daniel Lecoanet\\
+Benjamin P. Brown}}
 
@@
 \end{document}
+
+%  LocalWords:  Magnetorotational magnetorotational axisymmetric geo
+%  LocalWords:  Couette magnetohydrodynamic viscoelastic Oishi Vasil
+%  LocalWords:  Lecoanet
+%%% Local Variables: 
+%%% coding: utf-8
+%%% mode: latex
+%%% TeX-engine: xetex
+%%% End:

### 2019-09-14 paper/paper.tex
Update Andrew's affiliation.
Connected supplemental info and main text; reformatted for PRSA.
* changed to British spellings

@@
-% PRL Length limits
-% Displayed Math 	The word equivalent for displayed math is 16 words per row for single-column equations. Two-column equations count as 32 words per row.
-% Figures 	To estimate the word equivalent for figures use the figure’s aspect ratio (width / height).
-The estimate is [(150 / aspect ratio) + 20 words] for single-column figures, and [300 / (0.5 * aspect ratio)] + 40 words for double-column figures.
-
-% Fig 1: A=w/h=2. Double column, words=340
-% Fig 2: A=4/3. Single column, words=133
-% Fig 3: A=1. Single column, words=170
-% Fig 4: A=0.5 Single column, words=320
-% equations: 8 * 16=128
-
-% current word count: 2057 + 128 + 340 + 133 + 170 + 320
-% limit: 3750
-\documentclass[aps,prl,reprint,superscriptaddress]{revtex4-1}
+\documentclass[openacc]{rsproca_new}%%%%where rsproca is the template name
 \newcommand{\SSC}{S/S_{c}}
@@
 \newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
+\newcommand{\apj}{The Astrophysical Journal}
 \newcommand{\apjs}{The Astrophysical Journal Supplement Series}
 \newcommand{\apjl}{Astrophysical Journal Letters}
+\newcommand{\nat}{Nature}
+\newcommand{\prl}{Physical Review Letters}
+
+\newcommand\Beq{\begin{eqnarray}} 
+\newcommand\Eeq{\end{eqnarray}}
+
+\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
+\newcommand{\Eqs}[2]{Eqs.~(\ref{#1})--(\ref{#2})}
+\newcommand{\eq}[1]{eq.~(\ref{#1})}
+\newcommand{\eqs}[2]{eqs.~(\ref{#1})--(\ref{#2})}
+
+\newcommand{\Prob}[2]{\mathbb{P}[\, #1 \, | \, #2 \, ] }
+\newcommand{\Pro}[1]{\mathbb{P}[\, #1 \, ] }
+
+\newcommand{\inv}[1]{\frac{1}{#1}}
+\newcommand{\pd}[1]{\partial_{#1}}
+
+\newcommand{\Reyn}{\mathrm{Re}}
+\newcommand{\Reym}{\mathrm{Rm}}
+\newcommand{\Ra}{R}
+\newcommand{\Ve}{\mbox{\textit{Ve}}}
+\newcommand{\eps}{\varepsilon}
+\newcommand{\del}{\delta}
+\newcommand{\e}[1]{\eps^{#1}}
+\newcommand{\f}{2\Omega}
+
+\newcommand{\n}{\hat{n}}
+
+\newcommand{\dat}{\cdot}
+\newcommand{\dd}[1]{\,\mathrm{d}{#1}}
+\newcommand{\cc}{\mathrm{c.c.}}
+\renewcommand{\vec}[1]{\boldsymbol{#1}}
+\newcommand{\grad}{\nabla}
+\renewcommand{\div}{\nabla \dat}
+\newcommand{\Tr}{\text{Tr}}
+
+\newcommand{\pren}[1]{ \left(#1\right) }
+\newcommand{\brac}[1]{\left[\,#1\,\right]}
+
+\newcommand{\ave}[1]{\big<#1\big>}
+
+\renewcommand{\l}{\ell}
+
+\newcommand{\txt}[1]{\quad \text{#1} \quad }
+
 \usepackage{graphicx}
 \usepackage{amsmath}
+%\usepackage{natbib}
 
@@
 
-\author{Jeffrey~S.~Oishi}
-\email[]{joishi@bates.edu}
-\affiliation{Bates College Department of Physics and Astronomy, Lewiston, ME 04240}
+\author{%%%% Author details
+Jeffrey~S.~Oishi$^{1}$, Geoffrey~M.~Vasil$^{2}$, Morgan Baxter$^{1}$,
+Andrew Swan$^{3}$, Keaton~J.~Burns$^{4,5}$, Daniel~Lecoanet$^6$, Benjamin~P.~Brown$^7$}
 
-\author{Geoffrey~M.~Vasil}
-\affiliation{University of Sydney School of Mathematics and Statistics, Sydney, NSW, Australia}
-\author{Morgan Baxter}
-\affiliation{Bates College Department of Physics and Astronomy, Lewiston, ME 04240}
-\author{Andrew Swan}
-\affiliation{Faculty of Mathematics, Cambridge University, Cambridge, United Kingdom}
-\author{Keaton~J.~Burns}
-\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010}
-\affiliation{Massachusetts Institute of Technology Department of Physics, Cambridge, MA 02139}
-\author{Daniel~Lecoanet}
-\affiliation{Princeton University Department of Astrophysical Sciences, Princeton, NJ 08544}
-\author{Benjamin~P.~Brown}
-\affiliation{University of Colorado Laboratory for Atmospheric and Space Physics and Department of Astrophysical and Planetary Sciences, Boulder, CO 80309}
-
-\date{\today}
+%%%%%%%%% Insert author address here
+% \address{$^{1}$Bates College Department of Physics and Astronomy, Lewiston, ME 04240, USA\\
+% $^{2}$University of Sydney School of Mathematics and Statistics, Sydney, NSW, Australia\\
+% $^{3}$Statistical Laboratory, DPMMS, University of Cambridge, Cambridge,
+% UK\\
+% $^{4}5$Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA\\
+% $^{5}$Massachusetts Institute of Technology Department of Physics, Cambridge, MA 02139, USA\\
+% $^{6}$Princeton University Department of Astrophysical Sciences, Princeton, NJ 08544, USA\\
+% $^{7}$University of Colorado Laboratory for Atmospheric and Space Physics and Department of Astrophysical and Planetary Sciences, Boulder, CO 80309, USA}
+\address{Loads of places}
+\corres{Jeffrey~S.~Oishi\\
+\email{joishi@bates.edu}}
 
 \begin{abstract}
-The magnetorotational instability (MRI) occurs when a weak magnetic field destabilizes a rotating electrically conducting fluid with an inwardly increasing angular velocity shear.
-The MRI is considered essential to the operation of astrophysical accretion disks, with the central object supplying a Keplerian rotation profile.
-Internal shear layers in stars may also be unstable to the MRI.
-Stellar-interior shears can take a wide range of profiles, including near-critical, but with negligible dissipation. 
-Here, we show that the fastest growing unstable modes of an ideal magnetofluid are three dimensional provided the shear rate, $S$, is close to the two-dimensional onset value, $S_c$.
-For a Keplerian shear, three-dimensional modes are unstable down to $S\approx0.10S_c$, and dominate the two-dimensional modes until $S\approx2.05S_{c}$.
+The magnetorotational instability (MRI) occurs when a weak magnetic field destabilises a rotating, electrically conducting fluid with inwardly increasing angular velocity.
+The MRI is essential to astrophysical accretion disk theory where the shear is typically Keplerian.
+Internal shear layers in stars may also be unstable to the MRI, and they can take a wide range of profiles, including near-critical.
+Here, we show that the fastest growing modes of an ideal magnetofluid are three dimensional provided the shear rate, $S$, is near the two-dimensional onset value, $S_c$.
+For a Keplerian shear, three-dimensional modes are unstable above $S\approx0.10S_c$, and dominate the two-dimensional modes until $S\approx2.05S_{c}$.
+We also demonstrate that three-dimensional modes dominate for shear profiles relevant to stars and at magnetic Prandtl numbers relevant to liquid-metal laboratory experiments.
 Significant numbers of rapidly growing three-dimensional modes remain well past $2.05S_{c}$. 
 These finding are significant in three ways. 
-First, evidence from weakly nonlinear theory suggests that the MRI saturates by pushing the shear rate to its critical value. 
+First, weakly nonlinear theory suggests that the MRI saturates by pushing the shear rate to its critical value. 
 This can happen for systems, like stars and laboratory experiments, that can rearrange their angular velocity profiles.
-Second, the non-normal character and large transient growth of MRI modes should be important anytime strong three-dimensionality exists.
-Finally, three-dimensional growth suggest the possibility of direct dynamo action driven from the linear instability itself.
-We also demonstrate that three-dimensional modes dominate for shear profiles relevant to stars and at the tiny magnetic Prandtl numbers relevant to liquid-metal laboratory experiments.
+Second, the non-normal character and large transient growth of MRI modes should be important whenever three-dimensionality exists.
+Finally, three-dimensional growth suggests direct dynamo action driven from the linear instability itself.
 \end{abstract}
 
-\pacs{}
 \maketitle
 
+\section{Introduction}
 The magnetorotational instability (MRI) is extremely important in astrophysical fluid dynamics.
-A weak magnetic field catalyzes turbulence in a Keplerian shear by changing the stability criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient \citep[e.g.][]{1998RvMP...70....1B,2010RSPTA.368.1607J}.
-This discovery explained the ubiquitous accretion onto compact objects at rates compatible with observations, and may also influence the formation of planets \citep[e.g.][]{2007Natur.448.1022J}.
-In disks, the gravitational field dominates the local plasma dynamics, and thus the MRI cannot significantly affect the background shear; it must saturate by other means \citep{2018MNRAS.474.3451X}.
+A weak magnetic field catalyses turbulence in a Keplerian shear by changing the stability criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient \cite{1998RvMP...70....1B,2010RSPTA.368.1607J}.
+This discovery explained the ubiquitous accretion onto compact objects at rates compatible with observations, and may also influence the formation of planets \cite{2007Natur.448.1022J}.
+In disks, the gravitational field dominates the local plasma dynamics, and thus the MRI cannot significantly affect the background shear; it must saturate by other means \cite{2018MNRAS.474.3451X}.
 However, stars and liquid metal Taylor-Couette experiments have differential rotation profiles driven by much weaker stresses.
-Where the MRI is active in these flows, it saturates by pushing the background shear close to critical \citep{2015RSPSA.47140699V,2017ApJ...841....1C,2017ApJ...841....2C}, analogous to convection mixing entropy.
+Where the MRI is active in these flows, it saturates by pushing the background shear close to critical \cite{2015RSPSA.47140699V,2017ApJ...841....1C,2017ApJ...841....2C}, analogous to convection mixing entropy.
 Stellar interiors have extremely high fluid and magnetic Reynolds numbers, but can operate at or near the critical shear rate for the MRI.
@@
 Here, we investigate the stability of three-dimensional perturbations near the two-dimensional critical shear rate $S_{c}$ for a nearly inviscid, ideal MHD flow.
-We find that the first destabilized modes are three-dimensional, and thus could act as a dynamo even in the absence of secondary instability.
+We find that the first destabilised modes are three-dimensional, and thus could act as a dynamo even in the absence of secondary instability.
 These results also suggest that the non-normal growth of the MRI is always important, even when axisymmetric modes dominate.
 
+\section{Methods}
+\label{sec:methods}
+
 We numerically solve the linearized magnetohydrodynamic equations in rotating plane Couette geometry.
-This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y}=Sx$; \citep[see][]{2015RSPSA.47140699V}.
+This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y}=Sx$ \cite{2015RSPSA.47140699V}; it is also the narrow-gap limit of the Taylor-Couette geometry.
 We cast the Navier-Stokes equation in the form,
@@
 The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $x$ with Dirichlet boundary conditions.
-We pose and solve equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework \citep{2019arXiv190510388B}.
+We pose and solve equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework \cite{2019arXiv190510388B}.
 
@@
 We confirmed our results for unstable solutions are insensitive to small diffusion. 
-For each $(k_{y},k_{z})$ pair, we solve the eigenvalue problem using $n_{x}=128$ modes; all our results are identical at double the resolution \footnote{see supplemental material for details}; see \protect\url{github.com/jsoishi/mri_prefers_3d} for all code used in this paper.
+For each $(k_{y},k_{z})$ pair, we solve the eigenvalue problem using $n_{x}=128$ modes; all our results are identical at double the resolution \footnote{See \protect\url{github.com/jsoishi/mri_prefers_3d} for all code used in this paper.}
+
+For both the ideal and non-ideal MHD equations, we solve the eigenvalue problem in the $x$ direction using the \texttt{EigenValueProblem} solver in \emph{Dedalus} for a grid of $N_y \times N_z$ modes in the $y$ and $z$ directions.
+For most of our runs, we use a targeted, sparse eigenvalue solver to find the 15 modes closest to a guess for the maximum growth rate.
+In run 3, we have confirmed that dense solvers retrieve identical results.
+For nearly ideal runs, we use the ideal 2D growth rate as input for the smallest $k_y > 0$ mode at each $k_z$ and then use the output from each previous $k_y$ as an input guess for the next mode.
+For those runs that have significant $\eta$, we instead use a dense solve for the $k_y = 0$ modes and then step forward in $k_y$ at each $k_z$ as before.
+Our solver is embarrassingly parallelised over the $k_z$ modes.
+For the spectrum in figure 1 of the main text, we used a dense eigenvalue solver at $(k_y^{max},k_z^{max})$ for $\SSC= 1.02$.
+
+In order to ensure our results are converged, we have repeated runs at $n_x=256$ Chebyshev modes as well as doubling the number of modes in $y$ and $z$.
+
+\begin{tabular}{cccccccccccccc}
+\textbf{Run} & \textbf{$\SSC$} & \textbf{$\nu$} & \textbf{$\eta$} & \textbf{$q$} & \textbf{$N_x$} & \textbf{$N_{k_y}$} & \textbf{$N_{k_z}$} & \textbf{Sparse/Dense}& \\
+1  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & resolution study\\
+2  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 256 & 200 & 200 & sparse & \\
+3  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & dense  & \\
+4  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 512 & 512 & sparse & \\
+5  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 100 & 100 & sparse & \\
+6  &   0.2  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & $\SSC$ variation\\
+7  &   0.3  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+8  &   0.4  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+9  &   0.5  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+10 &   0.64 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+11 & 1.020  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+12 & 1.44   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+13 & 1.75   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+14 & 1.891  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+15 & 2.     & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+16 & 2.01   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+17 & 2.015  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+18 & 2.031  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+19 & 2.05   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+20 & 2.1    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+21 & 2.25   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+22 & 2.5    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+23 & 4      & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+24 & 1.02   & $10^{-6}$ & $10^{-6}$ & 0.75 & 128 & 200 & 200 & sparse & Reynolds number study\\
+25 & 1.02   & $10^{-4}$ & $10^{-4}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+26 & 1.02   & $10^{-3}$ & $10^{-3}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+27 & 1.02   & $10^{-2}$ & $10^{-2}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+28 & 1.02   & $10^{-5}$ & $10^{-5}$ & 0.1  & 128 & 200 & 200 & sparse & Low Rossby\\ 
+29 & 1.02   & $10^{-6}$ & $10^{-2}$ & 0.1  & 128 & 200 & 200 & sparse & Liquid metal case\\ 
+\end{tabular}
+%\label{simulations}
+
 \begin{figure*}[ht]
@@
   a) growth rates $\gamma$ over a grid of $k_{y}$ and $k_{z}$ for four values of supercriticality $\SSC$. 
-  The five black contours are equally spaced between zero and the maximum growth rate.
-  The gray contour highlights $\gamma=0$; the red dots indicate the fastest growing mode. 
+  The black contours are equally spaced between zero and the maximum growth rate.
+  The grey contour highlights $\gamma=0$; the red dots indicate the fastest growing mode. 
 Two dimensional modes occur at $k_y = 0$ on the bottom of each figure.
@@
   The asymptotic form is valid as $\SSC\to1$ and $k_{y}\ll1$.) c) The full discrete spectrum of the MRI for $(k_{y},k_{z})\simeq(0.263,0.447)$. 
-  The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray. 
+  The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are cray. 
   d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes ($\phi=0$) at $\SSC\gtrsim2.05$. }
@@
 \end{figure*}
+
+\section{Results}
+\label{sec:results}
+
+\subsection{Growth Rates and Three-dimensionality}
+\label{sec:growth}
 
@@
 
+\subsection{Asymptotic Calculation}
+\label{sec:asymp}
 The MRI's preference for three-dimensional modes can be predicted analytically.
-We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC=1+\epsilon^{2}R$, $k_{z}\sim\mathcal{O}(\epsilon)$, $\sigma\sim{k_{y}}\sim\mathcal{O}(\epsilon^{2})$,
+We analyse equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC=1+\epsilon^{2}R$, $k_{z}\sim\mathcal{O}(\epsilon)$, $\sigma\sim{k_{y}}\sim\mathcal{O}(\epsilon^{2})$.
+Here, we outline the asymptotic calculation for the leading order correction to the 2D growth rates in when accounting for three dimensional effects via finite $k_y$.
+Being in the rotating frame means the shear has no net
+\Beq
+\int_{-d/2}^{d/2} V_{0,y}(x) \text{d} x  \ = \ 0 \quad \implies  \quad V_{0,y}(x) \ = \ S\, x
+\Eeq
+The full linear ideal equations are:
+\Beq\label{u-eq}
+\pd{t} v_{x} + S x \pd{y} v_{x} - f v_{y} + \pd{x} p - B_{0} \pd{z} b_{x} &=& 0\\
+\pd{t} v_{y} + S x \pd{y} v_{y} + (f+S) v_{x} + \pd{y} p - B_{0} \pd{z} b_{y} &=& 0\label{v-eq} \\
+\pd{t} v_{z} + S x \pd{y} v_{z} + \pd{z} p - B_{0} \pd{z} b_{z} &=& 0 \\
+\pd{x} v_{x} + \pd{y} v_{y} + \pd{z} v_{z}  &=& 0\\
+\pd{t} b_{x} + S x \pd{y} b_{x} - B_{0} \pd{z} v_{x} &=& 0\\
+\pd{t} b_{y} + S x \pd{y} b_{y} - B_{0} \pd{z} v_{y} - S b_{x}  &=& 0\\
+\pd{t} b_{z} + S x \pd{y} b_{z} - B_{0} \pd{z} v_{z} &=& 0 \label{bz-eq}
+\Eeq
+The divergence-free condition for the velocity implies the same for the magnetic field.
+The system is 2nd order because $\pd{x}$ only appears in two places.
+All variables take the form
+\Beq
+p = \hat{P}(x) e^{ i ( \omega t + k z + \ell y ) },
+\Eeq
+where we note that for this section, we use the complex frequency $\omega = -i \sigma$ for the time dependence.
+
+To make analytical progress, define the frequency ``parameters'' 
+\Beq
+\omega_{S}(x) \ \equiv \ \omega + S x \ell, \quad \omega_{A}  \ \equiv \ B_{0} k.
+\Eeq
+We can use equations (\ref{v-eq}--\ref{bz-eq}) to find all amplitudes in terms of $\hat{V}_{x}(x)$ and $\hat{V}_{x}'(x)$. 
+\Beq
+\hat{V}_{y}(x) &=& \frac{i \left(\ell 
+   \hat{V}_{x}'(x) + k^{2} \frac{ \left((f+S) \omega
+   _S^2-S \omega _A^2\right)}{\omega _S
+   \left(\omega _S^2-\omega _A^2\right)} \hat{V}_{x}(x)\right)}{k^2+\ell ^2} \\
+\hat{V}_{z}(x) &=& \frac{i k \left(\hat{V}_{x}'(x)- \frac{ \left((f+S) \omega
+   _S^2-S \omega _A^2\right)}{\omega _S
+   \left(\omega _S^2-\omega _A^2\right)}  \ell 
+   \hat{V}_{x}(x)\right)}{k^2+\ell ^2} \\
+   \hat{P}(x) &=& \frac{i \left(\omega _A^2-\omega _S^2\right)
+   \left(\hat{V}_{x}'(x) -  \ell \frac{ \left((f+S) \omega
+   _S^2-S \omega _A^2\right)}{\omega _S
+   \left(\omega _S^2-\omega _A^2\right)}
+   \hat{V}_{x}(x)\right)}{\left(k^2+\ell ^2\right)
+   \omega _S} \\
+   \hat{B}_{x}(x) &=& \frac{\omega _A \hat{V}_{x}(x)}{\omega _S} \\
+   \hat{B}_{y}(x) &=&\frac{\omega _A \hat{V}_{y}(x)}{\omega _S} -\frac{i S \omega _A \hat{V}_{x}(x)}{\omega _S^2} \\
+   \hat{B}_{z}(x) &=& \frac{\omega _A \hat{V}_{z}(x)}{\omega _S}
+   \Eeq
+One can substitute everything into equation (\ref{u-eq}) to get a second-order equation for $\hat{V}_{x}(x)$ of the general form 
+\Beq
+\alpha(x) \hat{V}_{x}''(x) + \beta(x) \hat{V}_{x}'(x) + \gamma(x) \hat{V}_{x}(x) \ = \ 0,
+\Eeq
+where
+\Beq
+\alpha(x) &=& \omega _S^2-\omega _A^2 \\ 
+   \beta(x) &=& \frac{2 S \ell  \omega _A^2}{\omega _S} \\
+   \gamma(x) &=& \omega _A^2 \left(\frac{f^2 k^2}{\omega
+   _S^2-\omega _A^2}-\frac{2 S^2 \ell
+   ^2}{\omega _S^2}+k^2+\ell ^2\right)+f k^2
+   (f+S)-\left(k^2+\ell ^2\right) \omega _S^2.
+\Eeq
+
+We can eliminate the first-order term via
+\Beq
+\hat{V}_{x}(x) \ = \ \chi(x) \psi(x) \quad \text{where} \quad \chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
+\Eeq
+\Beq
+-\psi''(x) + \Phi(x) \psi(x) \ = \ 0 \label{Schr-eq}
+\Eeq
+\Beq
+\Phi(x)  \ = \ \frac{-2 \beta (x) \alpha '(x)+2 \alpha (x)
+   \left(\beta '(x)-2 \gamma (x)\right)+\beta
+   (x)^2}{4 \alpha (x)^2}.
+\Eeq
+Finally, by putting everything together 
+\Beq
+\Phi(x) \ = \ \frac{S  \left(f k^2-S \ell
+   ^2\right) \omega _A^2 -f k^2 (f+S) \omega
+   _S^2}{\left(\omega _A^2-\omega
+   _S^2\right){}^2} + k^{2} + \ell^{2}.
+\Eeq
+If $\omega_{S}=-i \gamma+Sx\ell$, with $Re(\gamma)\neq0$, the denominator of $\Phi(x)$ cannot vanish. 
+We want to solve the Schr\"{o}dinger-type equation~(\ref{Schr-eq}) with the boundary conditions 
+\Beq
+\psi(x=\pm d/2) = 0.
+\Eeq
+To do so, we solve perturbatively near the critical values for the 2D instability. We rescale each parameter in terms of a bookkeeping parameter, $\eps$.
+\Beq
+S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k \ \to \ \eps k, \quad \ell \ \to \ \eps^{2} \ell,
+\Eeq 
+keeping in mind
+\Beq
+q  \ \equiv  \ \frac{\pi ^2 B_{0}^2}{d^2 f^{2}}.
+\Eeq
+Expanding equation~(\ref{Schr-eq}) to $\mathcal{O}(\eps^{2})$, we arrive at
+\Beq
+\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 q \left(k^2 R+q
+   \ell ^2\right) - B_{0}^4 k^4 +f^2 (q+1) (\omega -f q x \ell
+   )^2  }{B_{0}^4 k^2}.
+\Eeq
+The leading order balance is
+\Beq
+\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0  \quad \implies \quad \psi_{0}(x) \ = \ \cos(\pi x / d),
+\Eeq
+and next order is
+\Beq
+\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) + \Phi_{2}(x)\psi_{0}(x)\ = \ 0.
+\Eeq
+The solvability condition for $\psi_{2}(x)$ is 
+\Beq
+\int_{-d/2}^{d/2} \Phi_{2}(x)\psi_{0}(x)^{2} \text{d} x \ = \ 0.
+\Eeq
+This yields our final result
 \begin{equation}\label{eq:asymp}
@@
 \begin{equation}
-\Upsilon=\frac{\left(\pi^{2}+6\right) q+\pi^{2}-6}{12}\approx1.31\quad\text{for}\quad{q}=\frac{3}{4}.
+\Upsilon=\frac{\left(\pi^{2}+6\right) q+\pi^{2}-6}{12}\approx1.31\quad\text{for}\quad{q}=\frac{3}{4},
 \end{equation}
+and we have switched back to $\sigma$ for time dependence.
 The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
@@
 The plot shows a purely growing/decaying complex-conjugate pair (orange/blue dots on the real axis) consistent with equation~(\ref{eq:asymp}).
-The other stable modes (gray dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
+The other stable modes (grey dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
 We reject spurious eigenvalues by using \texttt{eigentools} \footnote{\protect\url{https://bitbucket.org/jsoishi/eigentools}} to solve equations~(\ref{eq:mhd}-\ref{eq:divu}) at two resolutions and retain only pairs equal to within one part in $10^{-6}$.
+
+\subsection{Consistency with previous simulations}
+\label{sec:prior_work}
+
+Numerical simulations of the MRI in accretion disks consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \cite{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
+These ``channel modes'' are exact non-linear solutions for $x$-shearing-periodic domains, and can only saturate via parasitic shear instabilities \cite{1994ApJ...432..213G}.
+Impenetrable and stress-free boundary conditions are more applicable to stellar interiors. 
+But even in disks, any kind of finite radial extent will cutoff the unbounded growth of channel modes. 
+We reconcile our results with earlier simulations by showing that the MRI indeed prefers two dimensions for the larger values of criticality found in disk simulations. 
+Figure~\ref{fig:growth_rate}d shows the phase angle $\phi=\arctan(k_{y}/k_{z})$ of the fastest growing mode as function of $\SSC$.
+The overall critical shear for three-dimensional modes is $S_{c,\text{3D}}\simeq0.102S_c$.
+Above $\SSC\gtrsim2.05$, $\phi$ becomes zero, indicating that axisymmetric modes have the fastest growth rates.
+For the fiducial run in \cite{1996ApJ...464..690H}, $\SSC\simeq4.84$ (most works use similar values).
+Thus, our theory predicts axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
+
+Even for $\SSC> 2.05$ there are significant swaths of unstable three-dimensional modes with growth rates comparable to the maximum (figure~\ref{fig:growth_rate}a).
+Also, non-normality generically accompanies non-axisymmetry in shearing systems \cite{1992MNRAS.255P..25K}.
+In non-normal systems, transient amplification can occur for stable modes, and can even cause turbulence.
+Therefore significant three dimensionality likely implies important non-normal behaviour near onset.
+
+\subsection{Eigenvectors}
+\label{sec:eigenvectors}
 
@@
 
-Numerical simulations of the MRI in accretion disks consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}.
-These ``channel modes'' are exact non-linear solutions for $x$-shearing-periodic domains, and can only saturate via parasitic shear instabilities \citep{1994ApJ...432..213G}.
-Impenetrable and stress-free boundary conditions are more applicable to stellar interiors. 
-But even in disks, any kind of finite radial extent will cutoff the unbounded growth of channel modes. 
-We reconcile our results with earlier simulations by showing that the MRI indeed prefers two dimensions for the larger values of criticality found in disk simulations. 
-Figure~\ref{fig:growth_rate}d shows the phase angle $\phi=\arctan(k_{y}/k_{z})$ of the fastest growing mode as function of $\SSC$.
-The overall critical shear for three-dimensional modes is $S_{c,\text{3D}}\simeq0.102S_c$.
-Above $\SSC\gtrsim2.05$, $\phi$ becomes zero, indicating that axisymmetric modes have the fastest growth rates.
-For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC\simeq4.84$ (most works use similar values).
-Thus, our theory predicts axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
-
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{{mean_field_alpha_run_11_output.h5}.pdf}
-  \caption{Velocity-magnetic field correlations relevant to the mean electromotive force for the eigenvectors of the fastest growing mode at $\SSC = 1.02$.}
-  \label{fig:correlation}
-\end{figure}
-
-Even for $\SSC> 2.05$ there are significant swaths of unstable three-dimensional modes with growth rates comparable to the maximum (figure~\ref{fig:growth_rate}a).
-Also, non-normality generically accompanies non-axisymmetry in shearing systems \citep[see][]{1992MNRAS.255P..25K}.
-In non-normal systems, transient amplification can occur for stable modes, and can even cause turbulence.
-Therefore significant three dimensionality likely implies important non-normal behavior near onset.
-
 Figure~\ref{fig:eigvec} shows the eigenvectors for the most unstable mode at $\SSC=1.02$ using ideal MHD. 
@@
 
+\subsection{Reynolds Number and Departure from Ideal MHD}
+\label{sec:reyn}
+
+In order to ensure that our calculations are probing the ideal MHD regime we are interested in, we have performed as series of runs (24-27) with the $\eta = \nu$ ranging from $10^{-6}$ to $10^{-2}$.
+In our units, this is equivalent to varying the Reynolds and Magnetic Reynolds numbers $\Reyn = \Reym = 10^2 -- 10^6$.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\textwidth]{re_plots.pdf}
+  \caption{Growth rates at $\SSC = 1.02$ for varying diffusivities.
+Note that the top three panels are essentially identical, suggesting that above $\Reyn \gtrsim 10^4$ we are reasonably close to ideal conditions.}
+  \label{fig:reynolds}
+\end{figure}
+Figure~\pageref{fig:reynolds} shows the growth rates at $\SSC=1.05$ for five values of $\Reyn = \Reym$.
+For $\Reyn \gtrsim 10^4$, the maximum growth rate is located at the same position and the same growth rate; the unstable region expands, as expected.
+This suggests that our results at $\Reyn = 10^5$ are reasonably close to ideal MHD for the purposes of understanding the MRI's preference for 3D modes near shear onset.
+
+\subsection{Mean electromotive forces}
+\label{sec:emf}
 We should expect the generation of a mean electromotive force (EMF) in the linear regime, \textit{i.e}
@@
 This is significant because of the possibility of direct laminar dynamo action in regions of weak shear at large Reynolds number.
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{{mean_field_alpha_run_11_output.h5}.pdf}
+  \caption{Velocity-magnetic field correlations relevant to the mean electromotive force for the eigenvectors of the fastest growing mode at $\SSC = 1.02$.}
+  \label{fig:correlation}
+\end{figure}
 Figure~\ref{fig:correlation} shows the correlation (averaged over $y,\,z$) as a function of $x$ over the domain for $\SSC = 1.02$,
@@
 
-MRI dynamos have been studied in a number of different contexts \citep{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one focused on turbulent dynamos. 
-The lone exception, \citet{2016MNRAS.462..818B}, numerically found the exponential growth of mean magnetic fields during the linear growth phase of the MRI far from stability.
+MRI dynamos have been studied in a number of different contexts \cite{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one focused on turbulent dynamos. 
+The lone exception, \cite{2016MNRAS.462..818B}, numerically found the exponential growth of mean magnetic fields during the linear growth phase of the MRI far from stability.
 Our work explains this result in terms of purely linear dynamics:
@@
 
+\subsection{Extensions to stellar and experimental conditions}
+\label{sec:extensions}
+
+Finally, we vary two important parameters:
+$q$ and the magnetic Prandtl number $\Prm=\nu/\eta$.
 \begin{figure}[h!]
@@
 \end{figure}
-Finally, we vary two important parameters:
-$q$ and the magnetic Prandtl number $\Prm=\nu/\eta$.
 Figure~\ref{fig:other_params} shows the growth rates for $\SSC=1.02$.
 The upper panel shows $q=1/10$ (addressing a possible range within stellar interiors) holding all other parameters equal to their fiducial values.
-The growth rates show similar behavior to the $q=3/4$ cases, \textit{i.e.}, dominated by three-dimensional modes with $\gamma_{\max}/|S|\simeq0.1$.
+The growth rates show similar behaviour to the $q=3/4$ cases, \textit{i.e.}, dominated by three-dimensional modes with $\gamma_{\max}/|S|\simeq0.1$.
 The lower panel shows $\nu=10^{-6}$ and $\eta=0.1$ with all other parameters equal to their fiducial values. 
-This case is relevant to liquid metals e.g., the Princeton MRI experiment \citep{2002JFM...462..365G}
+This case is relevant to liquid metals e.g., the Princeton MRI experiment \cite{2002JFM...462..365G}
 at $\Prm=10^{-5}$. 
-Our results are for the rotating plane Couette geometry (Taylor-Couette in the small-gap limit $R_{2}\approx{R}_{1}$) with highly idealized boundary conditions.
+Our results are for the rotating plane Couette geometry (Taylor-Couette in the small-gap limit $R_{2}\approx{R}_{1}$) with highly idealised boundary conditions.
 It is nevertheless quite interesting that, near onset, the low-$\Prm$ MRI is \emph{only} unstable to three-dimensional modes.
+
+\section{Conclusion}
+\label{sec:conclusion}
 
@@
 First, the Sun possesses two internal shear layers with inwardly increasing shear, the high-latitude tachocline, and the near-surface shear layer (NSSL). 
-Past work has already pointed out the possibility of the small-scale MRI in the Sun using local analysis \citep{2007ApJ...667L.207P,2011MNRAS.411L..26M,2014ApJ...787...21K}.
-The NSSL, in particular, may also host slower MRI-driven dynamics; despite containing small-scale convection.
+Past work has already pointed out the possibility of the small-scale MRI in the Sun using local analysis \cite{2007ApJ...667L.207P,2011MNRAS.411L..26M,2014ApJ...787...21K}.
+The NSSL, in particular, may also host slower MRI-driven dynamics, despite containing small-scale convection.
 It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to convection.
@@
 Determining possible non-axisymmetric signatures in Taylor-Couette experiments requires followup work using more realistic boundary conditions and geometry. 
-Our prior work on the axisymmetric MRI \citep{2017ApJ...841....1C,2017ApJ...841....2C} shows only small differences between rotating plane and cylindrical Taylor-Couette geometries. 
+Our prior work on the axisymmetric MRI \cite{2017ApJ...841....1C,2017ApJ...841....2C} shows only small differences between rotating plane and cylindrical Taylor-Couette geometries. 
 We fully expect the general three-dimensional features of the MRI to persist in more complex applications.
 
-\begin{acknowledgments}
-We thank Steve Tobias for helpful conversations on this work.
-Oishi, Baxter, and Brown acknowledge support from NASA LWS grant No. NNX16AC92G.
-Oishi also acknowledges support from Research Corporation Scialog Collaborative Award (TDA) ID\#24231. 
-Computations were performed on the \emph{Leavitt} cluster at the Bates College High Performance Computing Center.
+%\ethics{Insert ethics text here.}
 
-\end{acknowledgments}
+\dataccess{All code used for this project is available at \url{https://github.com/jsoishi/mri\_prefers\_3D}}
 
+\aucontribute{J\@.S\@.O and G\@.M.\@V\@. led the project.
+G\@.M\@.V performed the asymptotic calculation.
+M.B.\ ran the bulk of the eigenvalue calculations and prepared several figures.
+A\@.S\@. performed preliminary eigenvalue calculations.
+J\@.S\@.O, G\@.M.V\@., K\@.J\@.B., D\@.L\@., and B.P\@.B developed \emph{Dedalus}, which is used for all calculations in this paper.
+All authors contributed to the writing of the manuscript.
+}
+
+\competing{We declare we have no competing interests.}
+
+\funding{Oishi, Baxter, and Brown acknowledge support from NASA LWS grant No. NNX16AC92G.
+Oishi also acknowledges support from Research Corporation Scialog Collaborative Award (TDA) ID \#24231.}
+
+\ack{Computations were performed on the \emph{Leavitt} cluster at the Bates College High Performance Computing Centre.}
+
+%\disclaimer{Insert disclaimer text here.}
+
+\bibliographystyle{RS}
 \bibliography{mri}
@@
 %  LocalWords:  hydrodynamical ODEs Dedalus arXiv supercriticality nd
-%  LocalWords:  criticality eigenvectors fiducial axisymmetry NSSL
+%  LocalWords:  criticality eigenvectors fiducial axisymmetry NSSL Sx
 %  LocalWords:  eigenfunctions eigenmodes diffusivities tachocline
-%  LocalWords:  Tobias JSO LWS NNX Scialog TDA Leavitt mri
+%  LocalWords:  Tobias JSO LWS NNX Scialog TDA Leavitt mri DPMMS rms
+%  LocalWords:  eigentools EigenValueProblem parallelised Chebyshev
+%  LocalWords:  cccccccccccccc analytical Schr dinger perturbatively
+%  LocalWords:  solvability

### 2019-09-20 paper/paper.tex
Modifications to author contributions.
two tiny changes. otherwise all good

@@
 \begin{equation}\label{eq:asymp}
-\sigma^{2}=\frac{q\,B_{0}^{2}}{(q+1)}\left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
+\sigma^{2}\ = \ - \omega^{2} \ = \  \frac{q\,B_{0}^{2}}{(q+1)}\left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
 \end{equation}
@@
 G\@.M\@.V performed the asymptotic calculation.
-M.B.\ ran the bulk of the eigenvalue calculations and prepared several figures.
-A\@.S\@. performed preliminary eigenvalue calculations.
+M.B.\ ran the bulk of the eigenvalue calculations and prepared several preliminary figures.
+A\@.S\@. performed preliminary eigenvalue calculations which led to the fiducial parameter choices used here.
 J\@.S\@.O, G\@.M.V\@., K\@.J\@.B., D\@.L\@., and B.P\@.B developed \emph{Dedalus}, which is used for all calculations in this paper.
-All authors contributed to the writing of the manuscript.
+All authors contributed to the writing of the manuscript and physical interpretation of the results.
 }

### 2019-09-21 paper/paper.tex
one 'typo' fix
As submitted to Proceedings of the Royal Society A.
* had to trim a few words off the abstract to be under 200 words
* renamed one of the figures to make it compile on the PRSA's latex install

@@
 %%%%%%%%% Insert author address here
-% \address{$^{1}$Bates College Department of Physics and Astronomy, Lewiston, ME 04240, USA\\
-% $^{2}$University of Sydney School of Mathematics and Statistics, Sydney, NSW, Australia\\
-% $^{3}$Statistical Laboratory, DPMMS, University of Cambridge, Cambridge,
-% UK\\
-% $^{4}5$Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA\\
-% $^{5}$Massachusetts Institute of Technology Department of Physics, Cambridge, MA 02139, USA\\
-% $^{6}$Princeton University Department of Astrophysical Sciences, Princeton, NJ 08544, USA\\
-% $^{7}$University of Colorado Laboratory for Atmospheric and Space Physics and Department of Astrophysical and Planetary Sciences, Boulder, CO 80309, USA}
-\address{Loads of places}
+\address{$^{1}$Bates College Department of Physics and Astronomy, Lewiston, ME 04240, USA\\
+$^{2}$University of Sydney School of Mathematics and Statistics, Sydney, NSW, Australia\\
+$^{3}$Statistical Laboratory, DPMMS, University of Cambridge, Cambridge,
+UK\\
+$^{4}5$Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA\\
+$^{5}$Massachusetts Institute of Technology Department of Physics, Cambridge, MA 02139, USA\\
+$^{6}$Princeton University Department of Astrophysical Sciences, Princeton, NJ 08544, USA\\
+$^{7}$University of Colorado Laboratory for Atmospheric and Space Physics and Department of Astrophysical and Planetary Sciences, Boulder, CO 80309, USA}
 \corres{Jeffrey~S.~Oishi\\
@@
 The magnetorotational instability (MRI) occurs when a weak magnetic field destabilises a rotating, electrically conducting fluid with inwardly increasing angular velocity.
-The MRI is essential to astrophysical accretion disk theory where the shear is typically Keplerian.
-Internal shear layers in stars may also be unstable to the MRI, and they can take a wide range of profiles, including near-critical.
-Here, we show that the fastest growing modes of an ideal magnetofluid are three dimensional provided the shear rate, $S$, is near the two-dimensional onset value, $S_c$.
+The MRI is essential to astrophysical disk theory where the shear is typically Keplerian.
+Internal shear layers in stars may also be MRI unstable, and they take a wide range of profiles, including near-critical.
+We show that the fastest growing modes of an ideal magnetofluid are three dimensional provided the shear rate, $S$, is near the two-dimensional onset value, $S_c$.
 For a Keplerian shear, three-dimensional modes are unstable above $S\approx0.10S_c$, and dominate the two-dimensional modes until $S\approx2.05S_{c}$.
-We also demonstrate that three-dimensional modes dominate for shear profiles relevant to stars and at magnetic Prandtl numbers relevant to liquid-metal laboratory experiments.
+These three-dimensional modes dominate for shear profiles relevant to stars and at magnetic Prandtl numbers relevant to liquid-metal laboratory experiments.
 Significant numbers of rapidly growing three-dimensional modes remain well past $2.05S_{c}$. 
@@
 Second, the non-normal character and large transient growth of MRI modes should be important whenever three-dimensionality exists.
-Finally, three-dimensional growth suggests direct dynamo action driven from the linear instability itself.
+Finally, three-dimensional growth suggests direct dynamo action driven from the linear instability.
 \end{abstract}
@@
 In order to ensure that our calculations are probing the ideal MHD regime we are interested in, we have performed as series of runs (24-27) with the $\eta = \nu$ ranging from $10^{-6}$ to $10^{-2}$.
-In our units, this is equivalent to varying the Reynolds and Magnetic Reynolds numbers $\Reyn = \Reym = 10^2 -- 10^6$.
+In our units, this is equivalent to varying the Reynolds and Magnetic Reynolds numbers $\Reyn = \Reym = 10^2$ -- $10^6$.
 \begin{figure}[h!]
@@
 \begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{{mean_field_alpha_run_11_output.h5}.pdf}
+  \includegraphics[width=\columnwidth]{mean_field_alpha_run_11_output.pdf}
   \caption{Velocity-magnetic field correlations relevant to the mean electromotive force for the eigenvectors of the fastest growing mode at $\SSC = 1.02$.}

### 2019-11-08 paper/paper.tex
Incorporated referee comments

@@
 \newcommand{\Ra}{R}
+\newcommand{\Ro}{\mathrm{Ro}}
 \newcommand{\Ve}{\mbox{\textit{Ve}}}
@@
 \renewcommand{\l}{\ell}
-
+\renewcommand{\ell}{\ensuremath{k_y}}
 \newcommand{\txt}[1]{\quad \text{#1} \quad }
@@
 Internal shear layers in stars may also be MRI unstable, and they take a wide range of profiles, including near-critical.
-We show that the fastest growing modes of an ideal magnetofluid are three dimensional provided the shear rate, $S$, is near the two-dimensional onset value, $S_c$.
+We show that the fastest growing modes of an ideal magnetofluid are three-dimensional provided the shear rate, $S$, is near the two-dimensional onset value, $S_c$.
 For a Keplerian shear, three-dimensional modes are unstable above $S\approx0.10S_c$, and dominate the two-dimensional modes until $S\approx2.05S_{c}$.
@@
 Stellar interiors have extremely high fluid and magnetic Reynolds numbers, but can operate at or near the critical shear rate for the MRI.
-This finite critical shear results from a finite-channel cutoff.
-Despite the extensive literature on accretion disks (strong shear, low dissipation) and liquid-metal experiments (weak shear, large dissipation), even the linear MRI is not well understood in the weak-shear, low-dissipation regime.
+This finite critical shear results from a finite-channel cutoff in the radial direction.
+Despite the extensive literature on accretion disks (strong shear, low dissipation) and liquid-metal experiments (weak shear, large dissipation), even the linear MRI is not well understood in the weak-shear (e.g.\ $S \gtrsim S_c$), low-dissipation regime.
 
 Here, we investigate the stability of three-dimensional perturbations near the two-dimensional critical shear rate $S_{c}$ for a nearly inviscid, ideal MHD flow.
+Throughout this paper, we use ``three-dimensional'' to refer to non-axisymmetric perturbations (and their local Cartesian equivalents), while ``two-dimensional'' refers to axisymmetric perturbations.
+In both cases, we retain all three components of velocity and magnetic fields.
 We find that the first destabilised modes are three-dimensional, and thus could act as a dynamo even in the absence of secondary instability.
@@
 
-We write the induction equation in terms of the $x$-component magnetic field,
+We write the induction equation in terms of the $x$-component of the magnetic field,
 \begin{equation}\label{eq:Bx}
@@
 \end{equation}
-and the $x$-component current density ($j_{x}=\partial_{y}b_{z}-\partial_{z}b_{y}$),
+and the $x$-component of the current density ($j_{x}=\partial_{y}b_{z}-\partial_{z}b_{y}$),
 \begin{equation}\label{eq:Jx}
@@
 
-The main input parameters are the Coriolis parameter, $f=2 \Omega$; the background shear rate, $S=dV_{y}/dx<0$;  the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0}\rho_{0}=1$).
-Accretion-disk modeling usually considers the Rossby number $q=-S/f<1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear
-\footnote{The astrophysical disk community often uses $q' = -S/\Omega$ to represent the same effective quantity; Keplerian profiles have $q' = 3/2$ in this case.}.
-Unless otherwise stated $q=3/4$ (Keplerian).
+The main input parameters are the Coriolis parameter, $f=2 \Omega$; the background shear rate, $S=dV_{y}/dx<0$; and the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0}\rho_{0}=1$).
+Accretion-disk modelling usually considers the Rossby number $\Ro=-S/f<1$; $\Ro\ge1$ corresponds to purely hydrodynamical Rayleigh unstable shear
+\footnote{The astrophysical disk community often uses $q = 2\Ro$ to represent the same effective quantity; Keplerian profiles have $q = 3/2$ in this case.}.
+Unless otherwise stated $\Ro=3/4$ (Keplerian).
+The solution also depends on the viscosity $\nu$ and resistivity $\eta$; in our non-dimensionalization, these are equivalent to the inverse Reynolds and Magnetic Reynolds numbers, respectively.
+That is, $\Reyn = 1/\nu$ and $\Reym = 1/\eta$.
 
-The MRI is a weak-field instability; in the inviscid, ideal case the critical shear rate for instability in \textit{two dimensions} is
+The MRI is a weak-field instability; in the inviscid, ideal case the critical shear rate for axisymmetric instability (i.e.\ in \emph{two dimensions}) is
 \begin{equation}\label{eq:Sc}
-  S_{c}=-\frac{\pi^{2}B_{0}^2}{fd^2}.
+  S_{c}=-\frac{\pi^{2}B_{0}^2}{fd^2}
 \end{equation}
+\cite{2015RSPSA.47140699V}.
+
 We use $\SSC$ as our instability control parameter.
+Because $\SSC$ serves as a ratio of the dimensionless magnetic field strength $B_0^2/f d$ to the length scale $d$, this serves to specify the background field strength.
 
@@
 Our solver is embarrassingly parallelised over the $k_z$ modes.
-For the spectrum in figure 1 of the main text, we used a dense eigenvalue solver at $(k_y^{max},k_z^{max})$ for $\SSC= 1.02$.
+For the spectrum in figure 1 of the main text, we used a dense eigenvalue solver at $(k_y^{max},k_z^{max})$ for $\SSC= 1.002$.
 
@@
 \begin{tabular}{cccccccccccccc}
-\textbf{Run} & \textbf{$\SSC$} & \textbf{$\nu$} & \textbf{$\eta$} & \textbf{$q$} & \textbf{$N_x$} & \textbf{$N_{k_y}$} & \textbf{$N_{k_z}$} & \textbf{Sparse/Dense}& \\
+\textbf{Run} & \textbf{$\SSC$} & \textbf{$\nu$} & \textbf{$\eta$} & \textbf{$\Ro$} & \textbf{$N_x$} & \textbf{$N_{k_y}$} & \textbf{$N_{k_z}$} & \textbf{Sparse/Dense}& \\
 1  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & resolution study\\
@@
 10 &   0.64 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-11 & 1.020  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-12 & 1.44   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-13 & 1.75   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-14 & 1.891  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-15 & 2.     & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-16 & 2.01   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-17 & 2.015  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-18 & 2.031  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-19 & 2.05   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-20 & 2.1    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-21 & 2.25   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-22 & 2.5    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-23 & 4      & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-24 & 1.02   & $10^{-6}$ & $10^{-6}$ & 0.75 & 128 & 200 & 200 & sparse & Reynolds number study\\
-25 & 1.02   & $10^{-4}$ & $10^{-4}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-26 & 1.02   & $10^{-3}$ & $10^{-3}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-27 & 1.02   & $10^{-2}$ & $10^{-2}$ & 0.75 & 128 & 200 & 200 & sparse & \\
-28 & 1.02   & $10^{-5}$ & $10^{-5}$ & 0.1  & 128 & 200 & 200 & sparse & Low Rossby\\ 
-29 & 1.02   & $10^{-6}$ & $10^{-2}$ & 0.1  & 128 & 200 & 200 & sparse & Liquid metal case\\ 
+11 & 1.002  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+12 & 1.002  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & dense & \\
+13 & 1.44   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+14 & 1.75   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+15 & 1.891  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+16 & 2.     & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+17 & 2.01   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+18 & 2.015  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+19 & 2.031  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+20 & 2.05   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+21 & 2.1    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+22 & 2.25   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+23 & 2.5    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+24 & 4      & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+25 & 1.02   & $10^{-6}$ & $10^{-6}$ & 0.75 & 128 & 200 & 200 & sparse & Reynolds number study\\
+26 & 1.02   & $10^{-4}$ & $10^{-4}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+27 & 1.02   & $10^{-3}$ & $10^{-3}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+28 & 1.02   & $10^{-2}$ & $10^{-2}$ & 0.75 & 128 & 200 & 200 & sparse & \\
+29 & 1.02   & $10^{-5}$ & $10^{-5}$ & 0.1  & 128 & 200 & 200 & sparse & Low Rossby\\ 
+30 & 1.02   & $10^{-6}$ & $10^{-2}$ & 0.1  & 128 & 200 & 200 & sparse & Liquid metal case\\ 
 \end{tabular}
@@
   The asymptotic form is valid as $\SSC\to1$ and $k_{y}\ll1$.) c) The full discrete spectrum of the MRI for $(k_{y},k_{z})\simeq(0.263,0.447)$. 
-  The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are cray. 
+  The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are grey. 
   d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes ($\phi=0$) at $\SSC\gtrsim2.05$. }
@@
 These results are pertinent in two ways. 
-First, the MRI contains a substantial ``Goldilocks regime" with possible direct dynamo action, and this regime likely applies to stellar interiors and laboratory experiments. 
+First, the MRI contains a substantial ``Goldilocks regime'' with possible direct dynamo action, and this regime likely applies to stellar interiors and laboratory experiments. 
 Second, our results accord with well-established results for accretion disks that expect two-dimensional primary modes.
 Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
-At $\SSC=1.02$, the maximum growth rate ($\gamma_{\max}/|S|=0.093$) happens at $(k_{y},k_{z})\simeq(0.263,0.447)$.
+At $\SSC=1.002$, the maximum growth rate ($\gamma_{\max}/|S|=0.093$) happens at $(k_{y},k_{z})\simeq(0.263,0.447)$.
 Two-dimensional instability occurs for modes with finite growth rates at $k_y = 0$ along the bottom of the figure.
 In all cases in figure~\ref{fig:growth_rate}a, the maximum growth rate does not occur along this line, indicating the dominance of three-dimensional modes.
-The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability: the zero growth rate contour (highlighted in grey) does not intercept the $x$-axis.
+The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability: the zero growth rate contour (highlighted in grey) does not intercept the $k_z$-axis.
 
@@
 We analyse equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC=1+\epsilon^{2}R$, $k_{z}\sim\mathcal{O}(\epsilon)$, $\sigma\sim{k_{y}}\sim\mathcal{O}(\epsilon^{2})$.
-Here, we outline the asymptotic calculation for the leading order correction to the 2D growth rates in when accounting for three dimensional effects via finite $k_y$.
+Here, we outline the asymptotic calculation for the leading order correction to the 2D growth rates when accounting for three dimensional effects via finite $k_y$.
 Being in the rotating frame means the shear has no net
@@
 The divergence-free condition for the velocity implies the same for the magnetic field.
-The system is 2nd order because $\pd{x}$ only appears in two places.
+The system is 2nd order in $x$ because $\pd{x}$ only appears in two places.
 All variables take the form
 \Beq
-p = \hat{P}(x) e^{ i ( \omega t + k z + \ell y ) },
+p = \hat{P}(x) e^{ i ( \omega t + k_z z + \ell y ) },
 \Eeq
@@
 \hat{V}_{y}(x) &=& \frac{i \left(\ell 
-   \hat{V}_{x}'(x) + k^{2} \frac{ \left((f+S) \omega
+   \hat{V}_{x}'(x) + k_z^2 \frac{ \left((f+S) \omega
    _S^2-S \omega _A^2\right)}{\omega _S
-   \left(\omega _S^2-\omega _A^2\right)} \hat{V}_{x}(x)\right)}{k^2+\ell ^2} \\
-\hat{V}_{z}(x) &=& \frac{i k \left(\hat{V}_{x}'(x)- \frac{ \left((f+S) \omega
+   \left(\omega _S^2-\omega _A^2\right)} \hat{V}_{x}(x)\right)}{k_z^2+\ell ^2} \\
+\hat{V}_{z}(x) &=& \frac{i k_z \left(\hat{V}_{x}'(x)- \frac{ \left((f+S) \omega
    _S^2-S \omega _A^2\right)}{\omega _S
    \left(\omega _S^2-\omega _A^2\right)}  \ell 
-   \hat{V}_{x}(x)\right)}{k^2+\ell ^2} \\
+   \hat{V}_{x}(x)\right)}{k_z^2+\ell ^2} \\
    \hat{P}(x) &=& \frac{i \left(\omega _A^2-\omega _S^2\right)
@@
    \left(\omega _S^2-\omega _A^2\right)}
-   \hat{V}_{x}(x)\right)}{\left(k^2+\ell ^2\right)
+   \hat{V}_{x}(x)\right)}{\left(k_z^2+\ell ^2\right)
    \omega _S} \\
@@
    \beta(x) &=& \frac{2 S \ell  \omega _A^2}{\omega _S} \\
-   \gamma(x) &=& \omega _A^2 \left(\frac{f^2 k^2}{\omega
+   \gamma(x) &=& \omega _A^2 \left(\frac{f^2 k_z^2}{\omega
    _S^2-\omega _A^2}-\frac{2 S^2 \ell
-   ^2}{\omega _S^2}+k^2+\ell ^2\right)+f k^2
-   (f+S)-\left(k^2+\ell ^2\right) \omega _S^2.
+   ^2}{\omega _S^2}+k_z^2+\ell ^2\right)+f k_z^2
+   (f+S)-\left(k_z^2+\ell ^2\right) \omega _S^2.
 \Eeq
@@
 \Beq
-\Phi(x) \ = \ \frac{S  \left(f k^2-S \ell
-   ^2\right) \omega _A^2 -f k^2 (f+S) \omega
+\Phi(x) \ = \ \frac{S  \left(f k_z^2-S \ell
+   ^2\right) \omega _A^2 -f k_z^2 (f+S) \omega
    _S^2}{\left(\omega _A^2-\omega
-   _S^2\right){}^2} + k^{2} + \ell^{2}.
+   _S^2\right){}^2} + k_z^2 + \ell^{2}.
 \Eeq
@@
 \Beq
-S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k \ \to \ \eps k, \quad \ell \ \to \ \eps^{2} \ell,
+S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k_z \ \to \ \eps k_z, \quad \ell \ \to \ \eps^{2} \ell,
 \Eeq 
@@
 \Beq
-q  \ \equiv  \ \frac{\pi ^2 B_{0}^2}{d^2 f^{2}}.
+\Ro  \ \equiv  \ \frac{\pi ^2 B_{0}^2}{d^2 f^{2}}.
 \Eeq
@@
 \Beq
-\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 q \left(k^2 R+q
-   \ell ^2\right) - B_{0}^4 k^4 +f^2 (q+1) (\omega -f q x \ell
-   )^2  }{B_{0}^4 k^2}.
+\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 \Ro \left(k_z^2 R+\Ro
+   \ell ^2\right) - B_{0}^4 k^4 +f^2 (\Ro+1) (\omega -f \Ro x \ell
+   )^2  }{B_{0}^4 k_z^2}.
 \Eeq
@@
 \begin{equation}\label{eq:asymp}
-\sigma^{2}\ = \ - \omega^{2} \ = \  \frac{q\,B_{0}^{2}}{(q+1)}\left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
+\sigma^{2}\ = \ - \omega^{2} \ = \  \frac{\Ro\,B_{0}^{2}}{(\Ro+1)}\left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
 \end{equation}
@@
 \begin{equation}
-\Upsilon=\frac{\left(\pi^{2}+6\right) q+\pi^{2}-6}{12}\approx1.31\quad\text{for}\quad{q}=\frac{3}{4},
+\Upsilon=\frac{\left(\pi^{2}+6\right) \Ro+\pi^{2}-6}{12}\approx1.31\quad\text{for}\quad{\Ro}=\frac{3}{4},
 \end{equation}
@@
 Figure~\ref{fig:growth_rate}b compares the numerical growth rates for $\SSC=1.002$ between $0\le{k_{y}}\le0.2$ to the asymptotic approximation, showing good agreement where the latter is valid.
-Figure~\ref{fig:growth_rate}c shows the full spectrum for $\SSC=1.02$.
+Figure~\ref{fig:growth_rate}c shows the full spectrum for $\SSC=1.002$.
 The plot shows a purely growing/decaying complex-conjugate pair (orange/blue dots on the real axis) consistent with equation~(\ref{eq:asymp}).
@@
 Thus, our theory predicts axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
+However, it may be that those simulations saw a predominance of axisymmetric modes because of the radial boundary conditions and not their large shear rates.
 
@@
 
-In order to ensure that our calculations are probing the ideal MHD regime we are interested in, we have performed as series of runs (24-27) with the $\eta = \nu$ ranging from $10^{-6}$ to $10^{-2}$.
-In our units, this is equivalent to varying the Reynolds and Magnetic Reynolds numbers $\Reyn = \Reym = 10^2$ -- $10^6$.
+In order to ensure that our calculations are probing the ideal MHD regime we are interested in, we have performed a series of runs (24-27) with $\eta = \nu$ ranging from $10^{-6}$ to $10^{-2}$, equivalent to varying the Reynolds and Magnetic Reynolds numbers $\Reyn = \Reym = 10^2$ -- $10^6$.
 \begin{figure}[h!]
@@
 
-MRI dynamos have been studied in a number of different contexts \cite{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one focused on turbulent dynamos. 
-The lone exception, \cite{2016MNRAS.462..818B}, numerically found the exponential growth of mean magnetic fields during the linear growth phase of the MRI far from stability.
+MRI dynamos have been studied in a number of different contexts \cite{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one focused on non-linear, usually turbulent, dynamos.
+The only exception we are aware of, \cite{2016MNRAS.462..818B}, found exponential growth of mean magnetic fields during the linear growth phase of the MRI far from stability in numerical simulations.
 Our work explains this result in terms of purely linear dynamics:
@@
 Finally, we vary two important parameters:
-$q$ and the magnetic Prandtl number $\Prm=\nu/\eta$.
+$\Ro$ and the magnetic Prandtl number $\Prm=\nu/\eta$.
 \begin{figure}[h!]
@@
   \includegraphics[width=\columnwidth]{low_rossby_liquid_metal_growth_rates.pdf}
-  \caption{(a) Growth rates for $q=1/10$ relevant to the stellar interiors. 
+  \caption{(a) Growth rates for $\Ro=1/10$ relevant to the stellar interiors. 
   (b) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta=0.1$.
@@
 Figure~\ref{fig:other_params} shows the growth rates for $\SSC=1.02$.
-The upper panel shows $q=1/10$ (addressing a possible range within stellar interiors) holding all other parameters equal to their fiducial values.
-The growth rates show similar behaviour to the $q=3/4$ cases, \textit{i.e.}, dominated by three-dimensional modes with $\gamma_{\max}/|S|\simeq0.1$.
+The upper panel shows $\Ro=1/10$ (addressing a possible range within stellar interiors) holding all other parameters equal to their fiducial values.
+The growth rates show similar behaviour to the $\Ro=3/4$ cases, \textit{i.e.}, dominated by three-dimensional modes with $\gamma_{\max}/|S|\simeq0.1$.
 The lower panel shows $\nu=10^{-6}$ and $\eta=0.1$ with all other parameters equal to their fiducial values. 
@@
 Our results show that three-dimensional modes grow faster than two-dimensional modes whenever the MRI is near its critical shear values.
-While we mainly focus on the ideal Keplerian case ($q=3/4$), we also demonstrate robustness for low Rossby ($q=1/10$) and Prandtl numbers ($\Prm=10^{-5}$).
+While we mainly focus on the ideal Keplerian case ($\Ro=3/4$), we also demonstrate robustness for low Rossby ($\Ro=1/10$) and magnetic Prandtl numbers ($\Prm=10^{-5}$).
 There are several important future directions this work suggests.
@@
 It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to convection.
+An important uncertainty with regard to stellar applications is the fact that $\SSC$ depends on the square of the magnetic field strength, a quantity that is rather uncertain.
 Second, our low-$\Prm$ results suggest that three-dimensional MRI modes may be the easiest to excite in liquid metal experiments.
@@
 \aucontribute{J\@.S\@.O and G\@.M.\@V\@. led the project.
+J\@.S\@.O prepared all figures.
 G\@.M\@.V performed the asymptotic calculation.
@@
 %  LocalWords:  cccccccccccccc analytical Schr dinger perturbatively
-%  LocalWords:  solvability
+%  LocalWords:  solvability Center

### 2019-11-09 paper/paper.tex
Fixed URL for eigentools

@@
 The other stable modes (grey dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
-We reject spurious eigenvalues by using \texttt{eigentools} \footnote{\protect\url{https://bitbucket.org/jsoishi/eigentools}} to solve equations~(\ref{eq:mhd}-\ref{eq:divu}) at two resolutions and retain only pairs equal to within one part in $10^{-6}$.
+We reject spurious eigenvalues by using \texttt{eigentools} \footnote{\protect\url{https://github.com/dedalusproject/eigentools}} to solve equations~(\ref{eq:mhd}-\ref{eq:divu}) at two resolutions and retain only pairs equal to within one part in $10^{-6}$.
 

### 2019-11-10 paper/paper.tex
a few small tyopgraphical edits. + one sentance in conclusion about Ro=1/2 in NSSL.

@@
 \renewcommand{\l}{\ell}
-\renewcommand{\ell}{\ensuremath{k_y}}
+\renewcommand{\ell}{\ensuremath{k_{y}}}
 \newcommand{\txt}[1]{\quad \text{#1} \quad }
@@
 The main input parameters are the Coriolis parameter, $f=2 \Omega$; the background shear rate, $S=dV_{y}/dx<0$; and the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0}\rho_{0}=1$).
-Accretion-disk modelling usually considers the Rossby number $\Ro=-S/f<1$; $\Ro\ge1$ corresponds to purely hydrodynamical Rayleigh unstable shear
-\footnote{The astrophysical disk community often uses $q = 2\Ro$ to represent the same effective quantity; Keplerian profiles have $q = 3/2$ in this case.}.
+Accretion-disk modelling usually considers the Rossby number $\Ro=-S/f<1$; The regime $\Ro\ge1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
 Unless otherwise stated $\Ro=3/4$ (Keplerian).
@@
 In run 3, we have confirmed that dense solvers retrieve identical results.
-For nearly ideal runs, we use the ideal 2D growth rate as input for the smallest $k_y > 0$ mode at each $k_z$ and then use the output from each previous $k_y$ as an input guess for the next mode.
-For those runs that have significant $\eta$, we instead use a dense solve for the $k_y = 0$ modes and then step forward in $k_y$ at each $k_z$ as before.
-Our solver is embarrassingly parallelised over the $k_z$ modes.
-For the spectrum in figure 1 of the main text, we used a dense eigenvalue solver at $(k_y^{max},k_z^{max})$ for $\SSC= 1.002$.
+For nearly ideal runs, we use the ideal 2D growth rate as input for the smallest $k_{y} > 0$ mode at each $k_{z}$ and then use the output from each previous $k_{y}$ as an input guess for the next mode.
+For those runs that have significant $\eta$, we instead use a dense solve for the $k_{y} = 0$ modes and then step forward in $k_{y}$ at each $k_{z}$ as before.
+Our solver is embarrassingly parallelised over the $k_{z}$ modes.
+For the spectrum in figure 1 of the main text, we used a dense eigenvalue solver at $(k_{y}^{max},k_{z}^{max})$ for $\SSC= 1.002$.
 
@@
 \begin{tabular}{cccccccccccccc}
-\textbf{Run} & \textbf{$\SSC$} & \textbf{$\nu$} & \textbf{$\eta$} & \textbf{$\Ro$} & \textbf{$N_x$} & \textbf{$N_{k_y}$} & \textbf{$N_{k_z}$} & \textbf{Sparse/Dense}& \\
+\textbf{Run} & \textbf{$\SSC$} & \textbf{$\nu$} & \textbf{$\eta$} & \textbf{$\Ro$} & \textbf{$N_x$} & \textbf{$N_{k_{y}}$} & \textbf{$N_{k_{z}}$} & \textbf{Sparse/Dense}& \\
 1  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & resolution study\\
@@
   The grey contour highlights $\gamma=0$; the red dots indicate the fastest growing mode. 
-Two dimensional modes occur at $k_y = 0$ on the bottom of each figure.
+Two dimensional modes occur at $k_{y} = 0$ on the bottom of each figure.
   At $\SSC=0.640$, there are no two dimensional modes. 
@@
 At $\SSC=1.002$, the maximum growth rate ($\gamma_{\max}/|S|=0.093$) happens at $(k_{y},k_{z})\simeq(0.263,0.447)$.
-Two-dimensional instability occurs for modes with finite growth rates at $k_y = 0$ along the bottom of the figure.
+Two-dimensional instability occurs for modes with finite growth rates at $k_{y} = 0$ along the bottom of the figure.
 In all cases in figure~\ref{fig:growth_rate}a, the maximum growth rate does not occur along this line, indicating the dominance of three-dimensional modes.
-The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability: the zero growth rate contour (highlighted in grey) does not intercept the $k_z$-axis.
+The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability: the zero growth rate contour (highlighted in grey) does not intercept the $k_{z}$-axis.
 
@@
 We analyse equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC=1+\epsilon^{2}R$, $k_{z}\sim\mathcal{O}(\epsilon)$, $\sigma\sim{k_{y}}\sim\mathcal{O}(\epsilon^{2})$.
-Here, we outline the asymptotic calculation for the leading order correction to the 2D growth rates when accounting for three dimensional effects via finite $k_y$.
+Here, we outline the asymptotic calculation for the leading order correction to the 2D growth rates when accounting for three dimensional effects via finite $k_{y}$.
 Being in the rotating frame means the shear has no net
@@
 \Beq
-p = \hat{P}(x) e^{ i ( \omega t + k_z z + \ell y ) },
+p = \hat{P}(x) e^{ i ( \omega t + k_{z} z + \ell y ) },
 \Eeq
@@
 \hat{V}_{y}(x) &=& \frac{i \left(\ell 
-   \hat{V}_{x}'(x) + k_z^2 \frac{ \left((f+S) \omega
-   _S^2-S \omega _A^2\right)}{\omega _S
-   \left(\omega _S^2-\omega _A^2\right)} \hat{V}_{x}(x)\right)}{k_z^2+\ell ^2} \\
-\hat{V}_{z}(x) &=& \frac{i k_z \left(\hat{V}_{x}'(x)- \frac{ \left((f+S) \omega
-   _S^2-S \omega _A^2\right)}{\omega _S
-   \left(\omega _S^2-\omega _A^2\right)}  \ell 
-   \hat{V}_{x}(x)\right)}{k_z^2+\ell ^2} \\
-   \hat{P}(x) &=& \frac{i \left(\omega _A^2-\omega _S^2\right)
+   \hat{V}_{x}'(x) + k_{z}^2 \frac{ \left((f+S) \omega
+   _S^2-S \omega _A^2\right)}{\omega_{S}
+   \left(\omega_{S}^2-\omega _A^2\right)} \hat{V}_{x}(x)\right)}{k_{z}^2+\ell ^2} \\
+\hat{V}_{z}(x) &=& \frac{i k_{z} \left(\hat{V}_{x}'(x)- \frac{ \left((f+S) \omega
+   _S^2-S \omega _A^2\right)}{\omega_{S}
+   \left(\omega_{S}^2-\omega _A^2\right)}  \ell 
+   \hat{V}_{x}(x)\right)}{k_{z}^2+\ell ^2} \\
+   \hat{P}(x) &=& \frac{i \left(\omega _A^2-\omega_{S}^2\right)
    \left(\hat{V}_{x}'(x) -  \ell \frac{ \left((f+S) \omega
-   _S^2-S \omega _A^2\right)}{\omega _S
-   \left(\omega _S^2-\omega _A^2\right)}
-   \hat{V}_{x}(x)\right)}{\left(k_z^2+\ell ^2\right)
-   \omega _S} \\
-   \hat{B}_{x}(x) &=& \frac{\omega _A \hat{V}_{x}(x)}{\omega _S} \\
-   \hat{B}_{y}(x) &=&\frac{\omega _A \hat{V}_{y}(x)}{\omega _S} -\frac{i S \omega _A \hat{V}_{x}(x)}{\omega _S^2} \\
-   \hat{B}_{z}(x) &=& \frac{\omega _A \hat{V}_{z}(x)}{\omega _S}
+   _S^2-S \omega _A^2\right)}{\omega_{S}
+   \left(\omega_{S}^2-\omega _A^2\right)}
+   \hat{V}_{x}(x)\right)}{\left(k_{z}^2+\ell ^2\right)
+   \omega_{S}} \\
+   \hat{B}_{x}(x) &=& \frac{\omega _A \hat{V}_{x}(x)}{\omega_{S}} \\
+   \hat{B}_{y}(x) &=&\frac{\omega _A \hat{V}_{y}(x)}{\omega_{S}} -\frac{i S \omega _A \hat{V}_{x}(x)}{\omega_{S}^2} \\
+   \hat{B}_{z}(x) &=& \frac{\omega _A \hat{V}_{z}(x)}{\omega_{S}}
    \Eeq
@@
 \Beq
-\alpha(x) \hat{V}_{x}''(x) + \beta(x) \hat{V}_{x}'(x) + \gamma(x) \hat{V}_{x}(x) \ = \ 0,
+\alpha(x) \hat{V}_{x}''(x) + 2\beta(x) \hat{V}_{x}'(x) + \gamma(x) \hat{V}_{x}(x) \ = \ 0,
 \Eeq
@@
 \Beq
-\alpha(x) &=& \omega _S^2-\omega _A^2 \\ 
-   \beta(x) &=& \frac{2 S \ell  \omega _A^2}{\omega _S} \\
-   \gamma(x) &=& \omega _A^2 \left(\frac{f^2 k_z^2}{\omega
-   _S^2-\omega _A^2}-\frac{2 S^2 \ell
-   ^2}{\omega _S^2}+k_z^2+\ell ^2\right)+f k_z^2
-   (f+S)-\left(k_z^2+\ell ^2\right) \omega _S^2.
+\alpha(x) &=& \omega_{S}^2-\omega _A^2 \\ 
+   \beta(x) &=& \frac{ S \ell  \omega _A^2}{\omega_{S}} \\
+   \gamma(x) &=& f k_{z}^2 \left(\frac{f \omega_{S}^{2}}{\omega_{S}^{2}-\omega
+   _A^2}+S\right)-\frac{2 S^2 \omega _A^2 k_{y}^2}{\omega_{S}^{2}}-\left(k_{z}^2+\ell^2 \right) (\omega_{S}^2 - \omega_{A}^{2}).
 \Eeq
@@
 \Beq
-\hat{V}_{x}(x) \ = \ \chi(x) \psi(x) \quad \text{where} \quad \chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
+\hat{V}_{x}(x) \ = \ \chi(x)\, \psi(x) \quad \text{where} \quad \frac{\chi'(x)}{\chi(x)} \  = \ -\frac{\beta (x)}{\alpha (x)}
 \Eeq
@@
 \Beq
-\Phi(x)  \ = \ \frac{-2 \beta (x) \alpha '(x)+2 \alpha (x)
-   \left(\beta '(x)-2 \gamma (x)\right)+\beta
-   (x)^2}{4 \alpha (x)^2}.
+\Phi(x)  \ = \ \frac{d}{dx}\! \left( \! \frac{\beta(x)}{\alpha(x)} \! \right) + \left( \! \frac{\beta(x)}{\alpha(x)} \! \right)^{\!2} - \frac{\gamma(x)}{\alpha(x)} .
 \Eeq
@@
 \Beq
-\Phi(x) \ = \ \frac{S  \left(f k_z^2-S \ell
-   ^2\right) \omega _A^2 -f k_z^2 (f+S) \omega
+\Phi(x) \ = \ \frac{S  \left(f k_{z}^2-S \ell
+   ^2\right) \omega _A^2 -f k_{z}^2 (f+S) \omega
    _S^2}{\left(\omega _A^2-\omega
-   _S^2\right){}^2} + k_z^2 + \ell^{2}.
+   _S^2\right){}^2} + k_{z}^2 + \ell^{2}.
 \Eeq
@@
 \Beq
-S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k_z \ \to \ \eps k_z, \quad \ell \ \to \ \eps^{2} \ell,
+S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k_{z} \ \to \ \eps k_{z}, \quad \ell \ \to \ \eps^{2} \ell,
 \Eeq 
@@
 \Beq
-\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 \Ro \left(k_z^2 R+\Ro
-   \ell ^2\right) - B_{0}^4 k^4 +f^2 (\Ro+1) (\omega -f \Ro x \ell
-   )^2  }{B_{0}^4 k_z^2}.
+\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 \Ro\, \left(k_{z}^2 R+\Ro
+   \ell ^2\right) - B_{0}^4 k^4 +f^2 (\Ro+1) (\omega -f \Ro\, x \ell
+   )^2  }{B_{0}^4 k_{z}^2}.
 \Eeq
@@
 \begin{equation}
-\Upsilon=\frac{\left(\pi^{2}+6\right) \Ro+\pi^{2}-6}{12}\approx1.31\quad\text{for}\quad{\Ro}=\frac{3}{4},
+\Upsilon \ = \  \frac{\left(\pi^{2}+6\right) \Ro+\pi^{2}-6}{12} \ \approx \ 1.31\quad\text{for}\quad{\Ro} \ = \ 0.75,
 \end{equation}
@@
 The NSSL, in particular, may also host slower MRI-driven dynamics, despite containing small-scale convection.
+The NSSL contains the strongest shear anywhere within the solar interior;  $\Omega(r) \ \propto \ 1/r$, impliying $\Ro = 1/2$.
 It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to convection.

### 2019-11-10 paper/paper.tex
last changes to asymptotic presentation

@@
 The main input parameters are the Coriolis parameter, $f=2 \Omega$; the background shear rate, $S=dV_{y}/dx<0$; and the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0}\rho_{0}=1$).
-Accretion-disk modelling usually considers the Rossby number $\Ro=-S/f<1$; The regime $\Ro\ge1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
+Accretion-disk modelling usually considers the Rossby number $\Ro=-S/f<1$. 
+The regime $\Ro\ge1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
 Unless otherwise stated $\Ro=3/4$ (Keplerian).
@@
 \Beq
-\alpha(x) \hat{V}_{x}''(x) + 2\beta(x) \hat{V}_{x}'(x) + \gamma(x) \hat{V}_{x}(x) \ = \ 0,
+ \hat{V}_{x}''(x) + 2 C_{1}(x) \hat{V}_{x}'(x) + C_{0}(x) \hat{V}_{x}(x) \ = \ 0,
 \Eeq
@@
 \Beq
-\alpha(x) &=& \omega_{S}^2-\omega _A^2 \\ 
-   \beta(x) &=& \frac{ S \ell  \omega _A^2}{\omega_{S}} \\
-   \gamma(x) &=& f k_{z}^2 \left(\frac{f \omega_{S}^{2}}{\omega_{S}^{2}-\omega
-   _A^2}+S\right)-\frac{2 S^2 \omega _A^2 k_{y}^2}{\omega_{S}^{2}}-\left(k_{z}^2+\ell^2 \right) (\omega_{S}^2 - \omega_{A}^{2}).
+   C_{1}(x) &=& \frac{ S \ell  \omega _A^2}{\omega_{S} ( \omega_{S}^2-\omega _A^2 )} \\
+   C_{0}(x) &=&  \frac{\frac{f^{2} k_{z}^2 \omega_{S}^{2}}{\omega_{S}^{2}-\omega
+   _A^2}+f  S  k_{z}^2 -\frac{2 S^2  k_{y}^2 \omega _A^2}{\omega_{S}^{2}}}{\omega_{S}^2-\omega _A^2}  -\left(k_{z}^2+\ell^2 \right) .
 \Eeq
@@
 \Beq
-\hat{V}_{x}(x) \ = \ \chi(x)\, \psi(x) \quad \text{where} \quad \frac{\chi'(x)}{\chi(x)} \  = \ -\frac{\beta (x)}{\alpha (x)}
+\hat{V}_{x}(x) \ = \ \chi(x)\, \psi(x) \quad \text{where} \quad \frac{\chi'(x)}{\chi(x)} \  = \ - C_{1} (x)
 \Eeq
@@
 \Beq
-\Phi(x)  \ = \ \frac{d}{dx}\! \left( \! \frac{\beta(x)}{\alpha(x)} \! \right) + \left( \! \frac{\beta(x)}{\alpha(x)} \! \right)^{\!2} - \frac{\gamma(x)}{\alpha(x)} .
+\Phi(x)  \ = \ C_{1}'(x) + C_{1} (x)^{2} - C_{0}(x).
 \Eeq
@@
    _S^2}{\left(\omega _A^2-\omega
-   _S^2\right){}^2} + k_{z}^2 + \ell^{2}.
+   _S^2\right)^2} + k_{z}^2 + \ell^{2}.
 \Eeq
-If $\omega_{S}=-i \gamma+Sx\ell$, with $Re(\gamma)\neq0$, the denominator of $\Phi(x)$ cannot vanish. 
+The denominator of $\Phi(x)$ cannot vanish if $\omega_{S}=-i \gamma+Sx\ell$, with $Re(\gamma)\neq0$. 
 We want to solve the Schr\"{o}dinger-type equation~(\ref{Schr-eq}) with the boundary conditions 
@@
 \Eeq
-To do so, we solve perturbatively near the critical values for the 2D instability. We rescale each parameter in terms of a bookkeeping parameter, $\eps$.
+The boundary conditions for $\psi(x)$ are the same as $V_{x}(x)$ because 
+\Beq
+\chi(x) \ \propto \ \frac{\omega_{S}(x)}{\sqrt{\omega_{S}(x)^2-\omega _A^2}} \  \ne \ 0 \quad \text{for} \quad \gamma \ \ne \ 0.
+\Eeq
+We solve perturbatively near the critical values for the 2D instability. We rescale each parameter in terms of a bookkeeping parameter, $\eps$.
 \Beq
@@
 \Beq
-\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 \Ro\, \left(k_{z}^2 R+\Ro
-   \ell ^2\right) - B_{0}^4 k^4 +f^2 (\Ro+1) (\omega -f \Ro\, x \ell
+\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 \Ro\, \left(k_{z}^2 R+\Ro\, \ell ^2\right) - B_{0}^4 k^4 +f^2 (\Ro+1) (\omega -f\, \Ro\, x\, \ell
    )^2  }{B_{0}^4 k_{z}^2}.

### 2019-11-10 paper/paper.tex
just kidding. This is the last set of changes

@@
 
-\renewcommand{\l}{\ell}
-\renewcommand{\ell}{\ensuremath{k_{y}}}
 \newcommand{\txt}[1]{\quad \text{#1} \quad }
@@
 \Beq
-p = \hat{P}(x) e^{ i ( \omega t + k_{z} z + \ell y ) },
+p = \hat{P}(x) e^{ i ( \omega t + k_{z} z + k_{y} y ) },
 \Eeq
@@
 \Beq
-\omega_{S}(x) \ \equiv \ \omega + S x \ell, \quad \omega_{A}  \ \equiv \ B_{0} k.
+\omega_{S}(x) \ \equiv \ \omega + S x k_{y}, \quad \omega_{A}  \ \equiv \ B_{0} k.
 \Eeq
@@
 \Beq
-\hat{V}_{y}(x) &=& \frac{i \left(\ell 
+\hat{V}_{y}(x) &=& \frac{i \left(k_{y} 
    \hat{V}_{x}'(x) + k_{z}^2 \frac{ \left((f+S) \omega
    _S^2-S \omega _A^2\right)}{\omega_{S}
-   \left(\omega_{S}^2-\omega _A^2\right)} \hat{V}_{x}(x)\right)}{k_{z}^2+\ell ^2} \\
+   \left(\omega_{S}^2-\omega _A^2\right)} \hat{V}_{x}(x)\right)}{k_{z}^2+k_{y} ^2} \\
 \hat{V}_{z}(x) &=& \frac{i k_{z} \left(\hat{V}_{x}'(x)- \frac{ \left((f+S) \omega
    _S^2-S \omega _A^2\right)}{\omega_{S}
-   \left(\omega_{S}^2-\omega _A^2\right)}  \ell 
-   \hat{V}_{x}(x)\right)}{k_{z}^2+\ell ^2} \\
+   \left(\omega_{S}^2-\omega _A^2\right)}  k_{y} 
+   \hat{V}_{x}(x)\right)}{k_{z}^2+k_{y} ^2} \\
    \hat{P}(x) &=& \frac{i \left(\omega _A^2-\omega_{S}^2\right)
-   \left(\hat{V}_{x}'(x) -  \ell \frac{ \left((f+S) \omega
+   \left(\hat{V}_{x}'(x) -  k_{y} \frac{ \left((f+S) \omega
    _S^2-S \omega _A^2\right)}{\omega_{S}
    \left(\omega_{S}^2-\omega _A^2\right)}
-   \hat{V}_{x}(x)\right)}{\left(k_{z}^2+\ell ^2\right)
+   \hat{V}_{x}(x)\right)}{\left(k_{z}^2+k_{y} ^2\right)
    \omega_{S}} \\
@@
 \Beq
-   C_{1}(x) &=& \frac{ S \ell  \omega _A^2}{\omega_{S} ( \omega_{S}^2-\omega _A^2 )} \\
-   C_{0}(x) &=&  \frac{\frac{f^{2} k_{z}^2 \omega_{S}^{2}}{\omega_{S}^{2}-\omega
-   _A^2}+f  S  k_{z}^2 -\frac{2 S^2  k_{y}^2 \omega _A^2}{\omega_{S}^{2}}}{\omega_{S}^2-\omega _A^2}  -\left(k_{z}^2+\ell^2 \right) .
+   C_{1}(x) \ = \  \frac{ S k_{y}  \omega _A^2}{\omega_{S} ( \omega_{S}^2-\omega _A^2 )} , \quad C_{0}(x) \ = \   \frac{\frac{f^{2} k_{z}^2 \omega_{S}^{2}}{\omega_{S}^{2}-\omega
+   _A^2}+f  S  k_{z}^2 -\frac{2 S^2  k_{y}^2 \omega _A^2}{\omega_{S}^{2}}}{\omega_{S}^2-\omega _A^2}  -\left(k_{y}^{2} + k_{z}^2 \right) .
 \Eeq
-
 We can eliminate the first-order term via
 \Beq
-\hat{V}_{x}(x) \ = \ \chi(x)\, \psi(x) \quad \text{where} \quad \frac{\chi'(x)}{\chi(x)} \  = \ - C_{1} (x)
+\hat{V}_{x}(x) \ = \ \chi(x)\, \psi(x) \quad \text{where} \quad \frac{\chi'(x)}{\chi(x)} \  = \ - C_{1} (x).
 \Eeq
+Therefore 
 \Beq
--\psi''(x) + \Phi(x) \psi(x) \ = \ 0 \label{Schr-eq}
-\Eeq
-\Beq
+-\psi''(x) + \Phi(x) \psi(x) \ = \ 0 \label{Schr-eq}, \quad \text{where}\quad
 \Phi(x)  \ = \ C_{1}'(x) + C_{1} (x)^{2} - C_{0}(x).
@@
 \Beq
-\Phi(x) \ = \ \frac{S  \left(f k_{z}^2-S \ell
+\Phi(x) \ = \ \frac{S  \left(f k_{z}^2-S k_{y}
    ^2\right) \omega _A^2 -f k_{z}^2 (f+S) \omega
    _S^2}{\left(\omega _A^2-\omega
-   _S^2\right)^2} + k_{z}^2 + \ell^{2}.
+   _S^2\right)^2} + k_{y}^{2} + k_{z}^2 .
 \Eeq
-The denominator of $\Phi(x)$ cannot vanish if $\omega_{S}=-i \gamma+Sx\ell$, with $Re(\gamma)\neq0$. 
+The denominator of $\Phi(x)$ cannot vanish if $\omega_{S}=-i \gamma+Sxk_{y}$, with $Re(\gamma)\neq0$. 
 We want to solve the Schr\"{o}dinger-type equation~(\ref{Schr-eq}) with the boundary conditions 
@@
 \Beq
-S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k_{z} \ \to \ \eps k_{z}, \quad \ell \ \to \ \eps^{2} \ell,
+S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k_{z} \ \to \ \eps k_{z}, \quad k_{y} \ \to \ \eps^{2} k_{y},
 \Eeq 
@@
 \Beq
-\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 \Ro\, \left(k_{z}^2 R+\Ro\, \ell ^2\right) - B_{0}^4 k^4 +f^2 (\Ro+1) (\omega -f\, \Ro\, x\, \ell
+\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 \Ro\, \left(k_{z}^2 R+\Ro\, k_{y} ^2\right) - B_{0}^4 k^4 +f^2 (\Ro+1) (\omega -f\, \Ro\, x\, k_{y}
    )^2  }{B_{0}^4 k_{z}^2}.

### 2019-11-11 paper/paper.tex
fixed sign error in one equation. results are still just fine.
i keep finding little things to improve. i'll stop soon.
ok. last tiny edit
New section on analysis on 3D -> 2D transition.

@@
 \newcommand{\n}{\hat{n}}
+
+\newcommand{\s}{\text{s}}
+\newcommand{\x}{\text{x}}
+\newcommand{\y}{\text{y}}
 
@@
 \Beq
-\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 \Ro\, \left(k_{z}^2 R+\Ro\, k_{y} ^2\right) - B_{0}^4 k^4 +f^2 (\Ro+1) (\omega -f\, \Ro\, x\, k_{y}
-   )^2  }{B_{0}^4 k_{z}^2}.
+\Phi(x) \ = \ -\frac{\pi^{2}}{d^{2}} - \frac{\pi^{2} \eps^{2}}{d^{2}}\left[R - \frac{d^2 }{\pi^2}k_z^2 + \frac{\Ro^2 B_0^2  k_y^2+(1+\Ro) (\omega + f \Ro\, x k_y) ^2}{ \Ro\, B_0^2
+   k_z^2} \right].
 \Eeq
@@
 \Beq
-\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) + \Phi_{2}(x)\psi_{0}(x)\ = \ 0.
+\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) \ = \  \Phi_{2}(x)\psi_{0}(x) .
 \Eeq
@@
 \begin{equation}\label{eq:asymp}
-\sigma^{2}\ = \ - \omega^{2} \ = \  \frac{\Ro\,B_{0}^{2}}{(\Ro+1)}\left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
+\sigma^{2}\ = \ - \omega^{2} \ = \  \frac{\Ro\,B_{0}^{2}}{1+\Ro}\left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
 \end{equation}
@@
 \begin{equation}
-\Upsilon \ = \  \frac{\left(\pi^{2}+6\right) \Ro+\pi^{2}-6}{12} \ \approx \ 1.31\quad\text{for}\quad{\Ro} \ = \ 0.75,
+\Upsilon \ = \  \Ro + (1+ \Ro) \frac{\pi^{2}-6}{12}  \ \approx \ 1.31\quad\text{for}\quad{\Ro} \ = \ 0.75,
 \end{equation}
@@
 We reject spurious eigenvalues by using \texttt{eigentools} \footnote{\protect\url{https://github.com/dedalusproject/eigentools}} to solve equations~(\ref{eq:mhd}-\ref{eq:divu}) at two resolutions and retain only pairs equal to within one part in $10^{-6}$.
+
+\subsection{Analysis of the transition to 2D}
+\label{sec:2D_transition}
+
+We can understand the transition from three- to two-dimensional instability semi-analytically.
+We do this by conducting a similar analysis as the above asymptotic calculation, only close to the fastest growing 2D modes.
+We can look for values for $\SSC$ where the analogous value of $\Upsilon(\SSC) = 0$.
+
+The calculation becomes very messy. We therefore only sketch the details, and report the main results.
+We analyse the growth rate as a function of  $k_{y},k_{z}$ and $\SSC$ near $k_{y} \approx0$.  We therefore expand the whole problem in a power-series ansatz
+\Beq
+\gamma(\SSC,k_{y},k_{z}) \ = \   \sum_{n\ge 0} \gamma_{n}(\SSC,k_{z}) k_{y}^{n}. 
+\Eeq
+We use the same asymptotic techniques as before to obtain each progressive correction.
+We solve the leading order by virtue that all non-constant coefficients vanish for 2D modes.
+The following dispersion relations determine the maximum growth rate and the corresponding wavenumber,
+\Beq
+\mathfrak{D}_{0}(k_{z}) &=& \frac{f k_z^2 \left(B_0^2 S k_z^2+\gamma _0^2 (f+S)\right)}{\left(B_0^2 k_z^2+\gamma _0^2\right){}^2}+\frac{\pi ^2}{d^2}+k_z^2 \ = \ 0 \\
+\frac{\mathfrak{D}_{0}'(k_{z})}{2k_{z}} &=& \frac{\gamma _0^2 f \left(B_0^2 k^2 (S-f)+\gamma _0^2 (f+S)\right)}{\left(B_0^2 k^2+\gamma _0^2\right)^3}+1 \ = \ 0.
+\Eeq
+We can transform the these relations into a polynomials in terms of non-dimensional variables,
+\Beq
+\label{dispersion}
+ \Ro^2 \s^2 \y^2+   (2 \Ro + \s - \Ro \, \s + \Ro^2 \s \y)\s  \y \x + (1 - \s + 2 \Ro\, \s \, \y) \x^2 +\x^3 \ = \ 0 \\ 
+(2 \Ro + \s - \Ro \, \s + \Ro^2 \s \y) \s  \y + 2(1 - \s + 2 \Ro\, \s \, \y) \x +3\x^2 \ = \ 0
+\label{critical-dispersion}
+\Eeq
+where
+\Beq
+\x  \ = \ \frac{d^2 k_z^2}{\pi ^2}, \quad  \quad \y \ = \ \frac{\gamma_{0}^{2}}{S^{2}}, \quad \quad \s  \ = \ \frac{S}{S_{c}}.
+\Eeq
+These are directly related to the parameters reported on figure~\ref{fig:growth_rate}.
+The two polynomials are straightforward to solve numerically for $\x,\y$ given $\s, \Ro$.
+However we  still want to find critical values for $\s$, where the fastest growing modes are two dimensional.
+From now on, the calculation becomes sufficiently complicated that we only outline the methodology.
+
+We find the higher-order $\gamma_{n}(\s,k_{z})$ coefficients by applying order-by-order solvability conditions.
+We first find that $\gamma_{1}(\s, k_{z}) = 0$ identically.
+This renders the first-order correction problem solvable, which means solving it and moving on the the 2nd-order corrections.
+The structure of the growth rate is therefore
+\Beq
+\gamma(\s,k_{y},k_{z}) \ = \ \gamma_{0}(\s,\x) + \gamma_{2}(\s,\x) k_{y}^{2} \ + \ \ldots 
+\Eeq
+This is the same structure as the asymptotic calculation in the previous section.
+Now effectively $\Upsilon \propto \gamma_{2}$.
+The case with dominant 2D modes corresponds to $\gamma_{2}(\s,\x)=0$.
+Finding this condition is a formidable calculation, which produces the polynomial expression
+\Beq
+\Ro^6 \s^6 \y^6 + \Ro^4 \s^5 \y^4 (3 \xi  (\Ro-1) \s+\Ro (6 \y-1)) \x + \nonumber \\  
+\frac{\Ro^2 \s^4 \y^3}{12}  \left(12 \Ro^2 (15 \y-4)-3 (4 \xi -3)
+   (\Ro-1)^2 \s^2+24 \xi  (4 \Ro+1) \Ro \, \s\right)\x^{2} + \nonumber \\ 
+\frac{\Ro\, \s^3 \y^2}{6}  \left(-3 (4 \xi -3) \left(\Ro^2-1\right) \s^2+12
+   \Ro^2 (10 \y-3)+36 \xi  (\Ro+2) \Ro\, \s\right)\x^{3} +\nonumber \\ 
+   \left(\Ro^2 \s^2 \y (15 \y-4)-\frac{1}{4} (4 \xi -3) (\Ro+1)^2 \s^4 \y+6 \xi 
+   \Ro\, \s^3 \y\right) \x^{4} +\nonumber \\ 
+   -\s (\xi  \Ro \, \s-6 \Ro \, \y+\Ro+\xi  \s) \x^{5} +  
+   \x^{6} \ = \ 0.
+   \label{2D-condition}
+\Eeq
+The new coefficient is
+\Beq
+\xi \ = \ \frac{\pi^{2}-6}{12} \ = \ \frac{\int_{-d/2}^{d/2} x^2 \cos ^2\left(\frac{\pi  x}{d}\right) \dd{x}}{\frac{d^{2}}{\pi^{2}}\int_{-d/2}^{d/2}  \cos ^2\left(\frac{\pi  x}{d}\right) \dd{x}}.
+\Eeq
+This results from projecting the squared shear onto the leading-order eigenmodes, just as in the previous section.
+
+Given $\Ro$, we solve \eq{dispersion}, \eq{critical-dispersion}, and  \eq{2D-condition} for $\x,\y,\s$ using Newton's method.
+
+\begin{center}
+\begin{tabular}{cccccccccccccc}
+\textbf{$\Ro$} & \textbf{$k_{z} d/\pi$}  & \textbf{$\gamma_{\max}/|S_{2D}|$}  &  \textbf{$S_{2D}/S_{c}$} \\
+                   1.00  & 0.708276 & 0.209120 &   2.13892 \\ 
+                   0.75 & 0.682819 & 0.214895 & 2.05136 \\
+                   0.50 & 0.651764 & 0.220102 & 1.95000 \\
+                   0.25 & 0.615396 & 0.224939 & 1.83881 \\
+                   0.10 & 0.593345 & 0.229047 & 1.77526
+\end{tabular}
+\end{center}
+The numbers for $\Ro = 0.75$ match the numerical results shown in Figure~\ref{fig:growth_rate} excellently.
+
+Finally, we can continue with the calculation conceptually. Because of symmetry $\gamma_{3}(\s,\x) = 0$ identically. This means that to leading order 
+\Beq
+\gamma(\s,k_{y},k_{z}) \ = \ \gamma_{0}(\s,\x) + \gamma_{2}(\s,\x) k_{y}^{2}  + \gamma_{4}(\s,\x) k_{y}^{4} \ + \ \ldots
+\Eeq 
+For $\SSC > S_{2D}/S_{c}$, $\gamma_{2} > 0$. This is sufficient to determine the most unstable 3D wavenumber near the transition to dominant 2D modes, 
+\Beq
+k_{y}^{2} \ \approx \ -\frac{\gamma _2}{2 \gamma_4} \ \propto \  S_{2D}/S_{c} - \SSC . 
+\Eeq 
+Generically $\gamma_{4} \ne 0$.
+We assume that $\gamma_{4} < 0$, but we do not know for sure if this is the case.
+The calculation is considerably more challenging that what we already have done.
+If there is a negative 4th-order feedback, then
+\Beq
+k_{y} \ \sim \ \sqrt{S_{2D}/S_{c} - \SSC}. 
+\Eeq
+This is consistent with Figure~\ref{fig:growth_rate}d near the critical point.
+If $\gamma_{4} > 0$, then the calculation would proceeded to 6th order (and so on), and the critical exponent will be smaller.
+Nevertheless the general picture will stay the same.
 

### 2019-11-11 paper/paper.tex
corrected first draft

@@
 \mathfrak{D}_{0}(k_{z}) &=& \frac{f k_z^2 \left(B_0^2 S k_z^2+\gamma _0^2 (f+S)\right)}{\left(B_0^2 k_z^2+\gamma _0^2\right){}^2}+\frac{\pi ^2}{d^2}+k_z^2 \ = \ 0 \\
-\frac{\mathfrak{D}_{0}'(k_{z})}{2k_{z}} &=& \frac{\gamma _0^2 f \left(B_0^2 k^2 (S-f)+\gamma _0^2 (f+S)\right)}{\left(B_0^2 k^2+\gamma _0^2\right)^3}+1 \ = \ 0.
+\frac{\mathfrak{D}_{0}'(k_{z})}{2k_{z}} &=& \frac{\gamma _0^2 f \left(B_0^2 k_{z}^2 (S-f)+\gamma _0^2 (f+S)\right)}{\left(B_0^2 k_{z}^2+\gamma _0^2\right)^3}+1 \ = \ 0.
 \Eeq
-We can transform the these relations into a polynomials in terms of non-dimensional variables,
+We transform these relations into polynomials in terms of non-dimensional variables,
 \Beq
 \label{dispersion}
- \Ro^2 \s^2 \y^2+   (2 \Ro + \s - \Ro \, \s + \Ro^2 \s \y)\s  \y \x + (1 - \s + 2 \Ro\, \s \, \y) \x^2 +\x^3 \ = \ 0 \\ 
-(2 \Ro + \s - \Ro \, \s + \Ro^2 \s \y) \s  \y + 2(1 - \s + 2 \Ro\, \s \, \y) \x +3\x^2 \ = \ 0
+\y^2 \s^2 \Ro^2 +  \y \, \s  \, ( \y \, \s \, \Ro^2 + (2  - \s)  \, \Ro + \s ) \x + ( 2 \y \,\s \,  \Ro - \s  + 1 ) \x^2 +\x^3 &=& 0 \\ 
+\y \, \s  \, (\y \, \s \, \Ro^2 + (2  - \s)  \, \Ro + \s    )  + 2 ( 2 \y \,\s \,  \Ro - \s  + 1 ) \x +3\x^2 &=& 0,
 \label{critical-dispersion}
@@
 The case with dominant 2D modes corresponds to $\gamma_{2}(\s,\x)=0$.
-Finding this condition is a formidable calculation, which produces the polynomial expression
+Finding this condition is a formidable calculation, which produces the degree-6 polynomial expression
 \Beq
-\Ro^6 \s^6 \y^6 + \Ro^4 \s^5 \y^4 (3 \xi  (\Ro-1) \s+\Ro (6 \y-1)) \x + \nonumber \\  
-\frac{\Ro^2 \s^4 \y^3}{12}  \left(12 \Ro^2 (15 \y-4)-3 (4 \xi -3)
-   (\Ro-1)^2 \s^2+24 \xi  (4 \Ro+1) \Ro \, \s\right)\x^{2} + \nonumber \\ 
-\frac{\Ro\, \s^3 \y^2}{6}  \left(-3 (4 \xi -3) \left(\Ro^2-1\right) \s^2+12
-   \Ro^2 (10 \y-3)+36 \xi  (\Ro+2) \Ro\, \s\right)\x^{3} +\nonumber \\ 
-   \left(\Ro^2 \s^2 \y (15 \y-4)-\frac{1}{4} (4 \xi -3) (\Ro+1)^2 \s^4 \y+6 \xi 
-   \Ro\, \s^3 \y\right) \x^{4} +\nonumber \\ 
-   -\s (\xi  \Ro \, \s-6 \Ro \, \y+\Ro+\xi  \s) \x^{5} +  
+& \y^{6} \s^6\Ro^6  \ + \
+ \y^4 \s^5 \Ro^4   \left( (3 \xi  \s+6 \y-1)\Ro -3 \xi  \s \right) \, \x +
+  \nonumber \\  
+& \frac{ \y^3 \s^4 \Ro^2  }{4}  \left( (60 \y -4 \xi  (\s-8) \s-16 + 3 \s^2)\Ro^{2} + 2 \s (4 \xi  (\s+1)-3 \s) \Ro - (4 \xi -3 ) \s^2 \right) \, \x^{2} + \nonumber \\ 
+& \frac{\y^2  \s^3 \Ro }{2}  \left( (3 \s^2-4 \xi  (\s-3) \s+40 \y-12) \Ro^{2} + 24 \xi \s \Ro + (4 \xi -3) s^2 \right) \, \x^{3} + 
+\nonumber \\ 
+&  \frac{\y \,\s^2}{4} \left( (60 \y -(4 \xi -3) \s^2 - 16)\Ro^{2} + 2\s(3 \s-4 \xi  (\s-3)) \Ro - (4 \xi -3) \s^2 \right)\, \x^{4} +\nonumber \\ 
+ &  \s\, (( 6 \y -\xi  \s-1) \Ro- \xi  \s) \, \x^{5}  \ + \   
    \x^{6} \ = \ 0.
@@
 \textbf{$\Ro$} & \textbf{$k_{z} d/\pi$}  & \textbf{$\gamma_{\max}/|S_{2D}|$}  &  \textbf{$S_{2D}/S_{c}$} \\
-                   1.00  & 0.708276 & 0.209120 &   2.13892 \\ 
+                   1.00  & 0.689610 & 0.208727 &   2.09694 \\ 
                    0.75 & 0.682819 & 0.214895 & 2.05136 \\
-                   0.50 & 0.651764 & 0.220102 & 1.95000 \\
-                   0.25 & 0.615396 & 0.224939 & 1.83881 \\
-                   0.10 & 0.593345 & 0.229047 & 1.77526
+                   0.50 & 0.676062 & 0.221309 & 2.00061 \\
+                   0.25 & 0.671528 & 0.228551 & 1.94963 \\
+                   0.10 & 0.673164 & 0.234376 & 1.92692 \\
+                   0.00 & 0.462324 & 0.240212 & 1.92465
 \end{tabular}

### 2019-11-11 paper/paper.tex
i've checked it as well as i can at this point.

@@
 We assume that $\gamma_{4} < 0$, but we do not know for sure if this is the case.
-The calculation is considerably more challenging that what we already have done.
+The calculation is considerably more challenging than anything so far.
 If there is a negative 4th-order feedback, then
 \Beq
-k_{y} \ \sim \ \sqrt{S_{2D}/S_{c} - \SSC}. 
+k_{y} \ \sim \ \sqrt{S_{2D}/S_{c} - \SSC}, 
 \Eeq
-This is consistent with Figure~\ref{fig:growth_rate}d near the critical point.
-If $\gamma_{4} > 0$, then the calculation would proceeded to 6th order (and so on), and the critical exponent will be smaller.
-Nevertheless the general picture will stay the same.
+which is consistent with Figure~\ref{fig:growth_rate}d near the critical point.
+Furthermore, a scenario where $\gamma_{4} > 0$ would require behaviour that is inconsistent with Figure~\ref{fig:growth_rate}d.
 

### 2019-11-11 paper/paper.tex
edits to section

@@
 
-The calculation becomes very messy. We therefore only sketch the details, and report the main results.
 We analyse the growth rate as a function of  $k_{y},k_{z}$ and $\SSC$ near $k_{y} \approx0$.  We therefore expand the whole problem in a power-series ansatz
 \Beq
-\gamma(\SSC,k_{y},k_{z}) \ = \   \sum_{n\ge 0} \gamma_{n}(\SSC,k_{z}) k_{y}^{n}. 
+\gamma(\SSC,k_{y},k_{z}) \ = \   \sum_{n\ge 0} \gamma_{n}(\SSC,k_{z}) k_{y}^{n}, \quad \psi(x) \ = \ \sum_{n\ge 0} \psi_{n}(x) k_{y}^{n}
 \Eeq
 We use the same asymptotic techniques as before to obtain each progressive correction. 
+\Beq
+\psi_{0}'' - \Phi_{0} \psi_{0} & = & 0 \label{eq:psi0} \\
+\psi_{1}'' - \Phi_{0} \psi_{1} & = & \Phi_{1} \psi_{0} \label{eq:psi1} \\
+\psi_{2}'' - \Phi_{0} \psi_{2} & = & \Phi_{2} \psi_{0} + \Phi_{1} \psi_{1} \label{eq:psi2}  
+\Eeq
 We solve the leading order by virtue that all non-constant coefficients vanish for 2D modes.
+The leading-order eigenfunction is $\psi_{0} = \cos(\pi x / d)$, just as before.
 The following dispersion relations determine the maximum growth rate and the corresponding wavenumber,
@@
 We first find that $\gamma_{1}(\s, k_{z}) = 0$ identically.
-This renders the first-order correction problem solvable, which means solving it and moving on the the 2nd-order corrections.
+This renders the first-order correction \eq{eq:psi1} solvable.
+The 1st-order eigenfunction takes the form
+\Beq
+\psi_{1}(x) \propto F(\pi x/d) \quad \text{where} \quad  F(\vartheta) = \vartheta (\vartheta \sin (\vartheta)+\cos (\vartheta))-\frac{\pi^{2}}{4} \sin (\vartheta).
+\Eeq
 The structure of the growth rate is therefore 
@@
 The case with dominant 2D modes corresponds to $\gamma_{2}(\s,\x)=0$.
-Finding this condition is a formidable calculation, which produces the degree-6 polynomial expression
+The final result derives from the 2nd-order solvability condition
+\Beq
+\int_{-d/2}^{d/2} ( \Phi_{2} \psi_{0} + \Phi_{1} \psi_{1} ) \psi_{0} \dd{x}  \ = \ 0,
+\Eeq
+which produces the degree-6 polynomial expression  
 \Beq

### 2019-11-12 paper/referee_response.tex
Final revisions and response to referees.

\documentclass[letterpaper, 12pt]{article}

% Imports
\usepackage{amsmath, amssymb}
\usepackage{mathtools}
\usepackage{graphicx}

% Layout
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

% Body font
\usepackage{soul, color}
\usepackage[colorlinks, allcolors=blue]{hyperref}
\usepackage{newtxtext}

% Reviews
\usepackage{framed}
\newenvironment{review}{\vspace{1em}\begin{leftbar}}{\end{leftbar}}

\begin{document}

Dear Professor Tobias,

We thank the referees for constructive reports. We have made all of the changes they suggested. In addition, we added section 3c detailing an asymptotic confirmation of the 3D to 2D transition reported numerically in section 3a to at least three significant figures. If this section is too much of an addition at this stage, we can relegate it to an appendix or remove it.

Please see our responses in line below.

\begin{review}
Comments to the Author(s)
This is an interesting manuscript that reveals a new and important result regarding the magnetorotational instability (MRI) in a radially confined geometry such as a Taylor-Couette experiment between differentially rotating cylinders. The authors show that 3D (non-axisymmetric) modes can be preferred to 2D (axisymmetric) ones, especially near the onset of the 2D instability. This might be important for the outcome of MRI experiments such as the one at Princeton, and it also has ramifications for dynamo action. The authors obtain good agreement, where expected between their numerical calculations of unstable modes and their asymptotic analysis.

The manuscript is mostly well written and the figures are well presented (although some are too small in the current format - in particular, Fig. 3 is hard to read).

I recommend publication in Proceedings A provided the following matters can be addressed.

1. It would be helpful to explain that "three-dimensional" means "non-axisymmetric" and "two-dimensional" means "axisymmetric" (but with three-component velocity and magnetic fields). (Of course, "axisymmetric" is to be interpreted as "y-independent" in this local model.)
\end{review}
We have fixed this.

\begin{review}  
2. Section 1: Perhaps clarify that the "finite-channel cutoff" is due to the finite width, not height.
\end{review}
Fixed.

\begin{review}
3. Section 1: What does "weak-shear" mean here: weak compared to what?  
\end{review}

Fixed

\begin{review}
4. Section 2: The notation could be defined more clearly by saying what v, b and j represent. We certainly need to know that f = 2 Omega.
\end{review}

We are unclear what the referee refers to here; these are already in the text.

\begin{review}
5. Section 2: The channel is periodic in z, but it seems that the periodicity length is unspecified, so that any value of the wavenumber $k_z$ is permissible. Please clarify this.
\end{review}

Fixed.

\begin{review}
6. Section 2: Personally, I find it confusing to have the symbol q denoting the Rossby number when (as said in the footnote) in most of the MRI literature q denotes twice the Rossby number. Would it not be better to use a different symbol for the Rossby number?
\end{review}

Fixed

\begin{review}
7. Section 2: I believe that q > 1 (rather than q = 1, as stated) corresponds to Rayleigh-unstable shear. (Are we to assume throughout that q > 0, i.e. S < 0?)
\end{review}
Fixed

\begin{review}
8. Around eq. 2.6: Please clarify the meaning of "in two dimensions" (i.e. axisymmetric, or y-independent, perturbations). Is there a convenient reference for eq. 2.6, as the result is not obvious? I believe it comes from taking the limit $k_z d \ll 1$, so the periodicity length in z has to be very long.
\end{review}

Fixed.

\begin{review}
9. Section 2: An important point is that the system of units has not been clarified and the relevant dimensionless parameters have not all been identified. Everything is expressed in dimensional terms until we are told that nu = eta = 1e-5. In what units are these values expressed?
  \end{review}

We have clarified this point while simultaneously addressing point 16.

\begin{review}
10. Section 2: I thought initially that there was a missing, and unspecified, dimensionless parameter such as $B_0 / f d$, which is a measure of the field strength. I see now that this can be determined from the given values of $q$ and $S/S_c$, but it might be better to explain this in the text.
\end{review}

We have done so.

\begin{review}
11. Figure 1 caption: Presumably "cray" should be "grey".  
\end{review}

Fixed

\begin{review}
12. Before eq. 3.1: Presumably "in when" should be "when".  
\end{review}

Fixed

\begin{review}
13. Before eq. 3.9: "2nd order" should probably be "2nd order in x".  
\end{review}

Fixed

\begin{review}
14. Eq. 3.9: Why are $k_z$ and $k_x$ renamed $k$ and $l$ here?  
\end{review}

We have corrected this.

\begin{review}
15. Section 3(c): It seems to me that the prevalence of axisymmetric modes in previous numerical simulations may have more to do with the radial boundary conditions than with the degree of supercriticality.
\end{review}

A note was added to this effect.

\begin{review}
16. Section 3(e): It would have been better to define Re and Rm back in Section 2. Again, the units in which eta and nu are given are not specified. There are a couple of typos in the first sentence: "as series" should be "a series", and "with the" should be "with".
\end{review}

Typos fixed and Reynolds numbers moved upwards.

\begin{review}
17. I disagree that there is only one previous study of non-turbulent MRI dynamos; indeed, reference [15] and related papers provide examples of this.  
\end{review}

This is completely correct; our original paper was unclear. We meant to express that there are no other *linear* MRI dynamos, not *laminar* ones!

\begin{review}
Referee: 2

Comments to the Author(s)
Thank you very much for sending your article! I find the subject, three-dimensional (3d) modes of the magneto-rotational instability (MRI), a very interesting aspect that so far has received much less attention than the corresponding two-dimensional (2d) modes. Your analysis is very well suited to the task, the results are interesting and presented clearly and in detail and set into the context of conditions in stars and lab experiments . Hence I recommend the article for publication. I have, however, a few minor points to
raise. The primary one is a question regarding the range of applicability of the 3d results, in particular in stars. You vary the ratio between the shear, S, and the critical shear value, $S_c$, given in equation (2.6), from values below unity to a maximum of 4. You note that for $S/S_c >~ 2.05$, the linear dynamics should be dominated by 2d
modes. $S_c$ involves, among other quantities, the square of the magnetic field strength, which can vary by (or is uncertain to) quite large factors. Thus, even moderate variations of the field may change $S_c$ such that $S/S_c$ no longer favours 3d modes. My question is, thus, if this apparent need for what might be considered fine-tuning limits the applicability of the results to, e.g., stars.
\end{review}
Thank you for pointing this out. We have added a cautionary note regarding this point.

\begin{review}
Apart from that, I have only small points:
- Section 3(a): the text mentions $S/S_c = 1.02$, but Fig. 1 has 1.002 and 1.2. Also, at the end of the paragraph: "does not intercept the x-axis"; one might put "$k_z$-axis" to avoid confusion with the axes of the simulation box.
- Section 3(f), after the equation: "rms is denotes...": remove one of the two verbs.
- Also 3(f): does the integral of the EMF over x vanish for reasons of symmetry?
\end{review}

We have fixed all of these.

\vspace{1em}

Sincerely, \\
Jeffrey S. Oishi, Geoffrey M. Vasil, Morgan Baxter, Andrew Swan, Keaton J. Burns, Daniel Lecoanet, \& Benjamin P. Brown
\end{document}

%%% Local Variables:
%%% ispell-local-dictionary: "british"
%%% End:

%  LocalWords:  Tobias magnetorotational Couette axisymmetric Rossby
%  LocalWords:  wavenumber eq nd supercriticality rms Oishi Vasil
%  LocalWords:  Lecoanet

### 2019-11-12 paper/paper.tex
checked all new equations. did calc. from start. everything checks out.
Final revisions and response to referees.

@@
 In order to ensure our results are converged, we have repeated runs at $n_x=256$ Chebyshev modes as well as doubling the number of modes in $y$ and $z$.
-
+\begin{table}
 \begin{tabular}{cccccccccccccc}
@@
 \end{tabular}
+\caption{Eigenvalue calculations performed.}
+\label{tab:runs}
+\end{table}
 %\label{simulations}
@@
 We can understand the transition from three- to two-dimensional instability semi-analytically.
-We do this by conducting a similar analysis as the above asymptotic calculation, only close to the fastest growing 2D modes.
-We can look for values for $\SSC$ where the analogous value of $\Upsilon(\SSC) = 0$.
+We do this by conducting a similar analysis as the asymptotic calculation in section~\ref{sec:asymp} near the fastest growing 2D modes, looking for values for $\SSC$ where the analogous value of $\Upsilon(\SSC)$ equals zero.
 
-We analyse the growth rate as a function of  $k_{y},k_{z}$ and $\SSC$ near $k_{y} \approx0$.  We therefore expand the whole problem in a power-series ansatz
+The growth rate can be written as a function of  $k_{y},k_{z}$ and $\SSC$ near $k_{y} \approx0$.  We therefore expand the whole problem in a power-series 
 \Beq
-\gamma(\SSC,k_{y},k_{z}) \ = \   \sum_{n\ge 0} \gamma_{n}(\SSC,k_{z}) k_{y}^{n}, \quad \psi(x) \ = \ \sum_{n\ge 0} \psi_{n}(x) k_{y}^{n}
+\gamma(\SSC,k_{y},k_{z}) \ = \   \sum_{n\ge 0} \gamma_{n}(\SSC,k_{z}) k_{y}^{n}, \quad \psi(x) \ = \ \sum_{n\ge 0} \psi_{n}(x) k_{y}^{n},
 \Eeq
-We use the same asymptotic techniques as before to obtain each progressive correction. 
+and use the same asymptotic techniques as before to obtain each progressive correction,
 \Beq
@@
 \psi_{1}'' - \Phi_{0} \psi_{1} & = & \Phi_{1} \psi_{0} \label{eq:psi1} \\
-\psi_{2}'' - \Phi_{0} \psi_{2} & = & \Phi_{2} \psi_{0} + \Phi_{1} \psi_{1} \label{eq:psi2}  
+\psi_{2}'' - \Phi_{0} \psi_{2} & = & \Phi_{2} \psi_{0} + \Phi_{1} \psi_{1}.
+\label{eq:psi2}  
 \Eeq
-We solve the leading order by virtue that all non-constant coefficients vanish for 2D modes.
+At leading order, we making use of the fact that all non-constant coefficients vanish for 2D modes.
 The leading-order eigenfunction is $\psi_{0} = \cos(\pi x / d)$, just as before.
@@
 \Eeq
-These are directly related to the parameters reported on figure~\ref{fig:growth_rate}.
-The two polynomials are straightforward to solve numerically for $\x,\y$ given $\s, \Ro$.
-However we  still want to find critical values for $\s$, where the fastest growing modes are two dimensional.
-From now on, the calculation becomes sufficiently complicated that we only outline the methodology.
+These can be computed from the parameters reported in figure~\ref{fig:growth_rate} and table~\ref{tab:runs}.
+The two polynomials are straightforward to solve numerically for $\x,\y$ given $\s, \Ro$. However, we want to find critical values for $\s$, where the fastest growing modes are two dimensional.
 
@@
 \Beq
-\psi_{1}(x) \propto F(\pi x/d) \quad \text{where} \quad  F(\vartheta) = \vartheta (\vartheta \sin (\vartheta)+\cos (\vartheta))-\frac{\pi^{2}}{4} \sin (\vartheta).
+\psi_{1}(x) \ = \ A_{1} F(\pi x/d) \quad \text{where} \quad  F(\vartheta) = \vartheta (\vartheta \sin (\vartheta)+\cos (\vartheta))-\frac{\pi^{2}}{4} \sin (\vartheta).
+\Eeq
+and
+\Beq
+A_{1} \ = \ \frac{i \gamma _0 d^3 f S k_z^2 \left(B_0^2 (f-S) k_z^2-\gamma _0^2
+   (f+S)\right)}{2 \pi ^3 \left(B_0^2 k_z^2+\gamma _0^2\right){}^{\!3}}.
 \Eeq
@@
 \Eeq
-This is the same structure as the asymptotic calculation in the previous section.
-Now effectively $\Upsilon \propto \gamma_{2}$.
+This is the same structure as the asymptotic calculation in section~\ref{sec:asymp}.
+Here, $\Upsilon \propto \gamma_{2}$ effectively.
 The case with dominant 2D modes corresponds to $\gamma_{2}(\s,\x)=0$.
@@
   \nonumber \\  
-& \frac{ \y^3 \s^4 \Ro^2  }{4}  \left( (60 \y -4 \xi  (\s-8) \s-16 + 3 \s^2)\Ro^{2} + 2 \s (4 \xi  (\s+1)-3 \s) \Ro - (4 \xi -3 ) \s^2 \right) \, \x^{2} + \nonumber \\ 
+& \frac{ \y^3 \s^4 \Ro^2  }{4}  \left( (60 \y -4 \xi  (\s-8) \s -16 + 3 \s^2)\Ro^{2} + 2 \s (4 \xi  (\s+1)-3 \s) \Ro - (4 \xi -3 ) \s^2 \right) \, \x^{2} + \nonumber \\ 
 & \frac{\y^2  \s^3 \Ro }{2}  \left( (3 \s^2-4 \xi  (\s-3) \s+40 \y-12) \Ro^{2} + 24 \xi \s \Ro + (4 \xi -3) s^2 \right) \, \x^{3} + 
@@
 \Beq
-\xi \ = \ \frac{\pi^{2}-6}{12} \ = \ \frac{\int_{-d/2}^{d/2} x^2 \cos ^2\left(\frac{\pi  x}{d}\right) \dd{x}}{\frac{d^{2}}{\pi^{2}}\int_{-d/2}^{d/2}  \cos ^2\left(\frac{\pi  x}{d}\right) \dd{x}}.
+\xi \ = \ \frac{\pi^{2}-6}{12} \ = \ \frac{\int_{-d/2}^{d/2} x^2 \cos ^2\left(\frac{\pi  x}{d}\right) \dd{x}}{\frac{d^{2}}{\pi^{2}}\int_{-d/2}^{d/2}  \cos ^2\left(\frac{\pi  x}{d}\right) \dd{x}} \ \approx \ 0.322467.
 \Eeq
@@
 
-Given $\Ro$, we solve \eq{dispersion}, \eq{critical-dispersion}, and  \eq{2D-condition} for $\x,\y,\s$ using Newton's method.
+We solve \eq{dispersion}, \eq{critical-dispersion}, and  \eq{2D-condition} for $\x,\y,\s$ using Newton's method for a variety of $\Ro$, yielding
 
@@
 \end{center}
-The numbers for $\Ro = 0.75$ match the numerical results shown in Figure~\ref{fig:growth_rate} excellently.
+The value of $\SSC = 2.05136$ at the 2D to 3D transition when $\Ro = 0.75$ matches the numerical results shown in Figure~\ref{fig:growth_rate} two all reported significant figures in the numerical calculation.
 
-Finally, we can continue with the calculation conceptually. Because of symmetry $\gamma_{3}(\s,\x) = 0$ identically. This means that to leading order 
+We conclude this section with some speculations concerning the higher order corrections. Because of symmetry, $\gamma_{3}(\s,\x) = 0$ identically. This means that to leading order 
 \Beq
@@
 \Eeq 
-Generically $\gamma_{4} \ne 0$.
-We assume that $\gamma_{4} < 0$, but we do not know for sure if this is the case.
-The calculation is considerably more challenging than anything so far.
-If there is a negative 4th-order feedback, then
+In general, $\gamma_{4} \ne 0$. We speculate that $\gamma_{4} < 0$, though the calculation is outside the scope of this work. If there is a negative 4th-order feedback, then 
 \Beq
@@
 \Eeq
-which is consistent with Figure~\ref{fig:growth_rate}d near the critical point.
-Furthermore, a scenario where $\gamma_{4} > 0$ would require behaviour that is inconsistent with Figure~\ref{fig:growth_rate}d.
+which is consistent with Figure~\ref{fig:growth_rate}d near the critical point. However, a scenario where $\gamma_{4} > 0$ would require behaviour that is inconsistent with our numerical calculations.
 
@@
 The NSSL, in particular, may also host slower MRI-driven dynamics, despite containing small-scale convection.
-The NSSL contains the strongest shear anywhere within the solar interior;  $\Omega(r) \ \propto \ 1/r$, impliying $\Ro = 1/2$.
+The NSSL contains the strongest shear anywhere within the solar interior;  $\Omega(r) \ \propto \ 1/r$, implying $\Ro = 1/2$.
 It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to convection.
@@
 
+%%% Local Variables:
+%%% ispell-local-dictionary: "british"
+%%% End:
+
 %  LocalWords:  Magnetorotational Oishi Lewiston Vasil NSW Lecoanet
@@
 %  LocalWords:  criticality eigenvectors fiducial axisymmetry NSSL Sx
-%  LocalWords:  eigenfunctions eigenmodes diffusivities tachocline
+%  LocalWords:  eigenfunctions eigenmodes diffusivities tachocline eq
 %  LocalWords:  Tobias JSO LWS NNX Scialog TDA Leavitt mri DPMMS rms
@@
 %  LocalWords:  cccccccccccccc analytical Schr dinger perturbatively
-%  LocalWords:  solvability Center
+%  LocalWords:  solvability Center resistivity dimensionalization Sxk
+%  LocalWords:  eigenfunction wavenumber th

### 2019-11-12 paper/paper.tex
Shortened abstract to comply with PRSA rules.

@@
 $^{6}$Princeton University Department of Astrophysical Sciences, Princeton, NJ 08544, USA\\
-$^{7}$University of Colorado Laboratory for Atmospheric and Space Physics and Department of Astrophysical and Planetary Sciences, Boulder, CO 80309, USA}
+$^{7}$University of Colorado Department of Astrophysical and Planetary Sciences, Boulder, CO 80309, USA}
 \corres{Jeffrey~S.~Oishi\\
@@
 The MRI is essential to astrophysical disk theory where the shear is typically Keplerian.
-Internal shear layers in stars may also be MRI unstable, and they take a wide range of profiles, including near-critical.
-We show that the fastest growing modes of an ideal magnetofluid are three-dimensional provided the shear rate, $S$, is near the two-dimensional onset value, $S_c$.
-For a Keplerian shear, three-dimensional modes are unstable above $S\approx0.10S_c$, and dominate the two-dimensional modes until $S\approx2.05S_{c}$.
-These three-dimensional modes dominate for shear profiles relevant to stars and at magnetic Prandtl numbers relevant to liquid-metal laboratory experiments.
-Significant numbers of rapidly growing three-dimensional modes remain well past $2.05S_{c}$. 
-These finding are significant in three ways. 
-First, weakly nonlinear theory suggests that the MRI saturates by pushing the shear rate to its critical value. 
-This can happen for systems, like stars and laboratory experiments, that can rearrange their angular velocity profiles.
-Second, the non-normal character and large transient growth of MRI modes should be important whenever three-dimensionality exists.
-Finally, three-dimensional growth suggests direct dynamo action driven from the linear instability.
+Shear layers in stars take a wide range of profiles that may be MRI unstable.
+We show that the fastest growing modes are 3D when the shear rate, $S$, is near the 2D onset value, $S_c$.
+For Keplerian shear, 3D modes are unstable above $S\approx0.10S_c$, and dominate the 2D modes until $S\approx2.05S_{c}$.
+Three-dimensional modes dominate for shear profiles relevant to stars and at magnetic Prandtl numbers found in liquid-metal laboratory experiments.
+Theory suggests that the MRI saturates by pushing the shear rate to critical in stars and laboratory experiments, where the shear profiles can relax.
+whenever three-dimensionality exists.
+Finally, 3D growth suggests a laminar dynamo driven by the linear instability.
 \end{abstract}

### 2019-11-13 paper/paper.tex
Fixed longer abstract to break across pages

--- 2019-11-12 paper/paper.tex
@@
 For a Keplerian shear, three-dimensional modes are unstable above $S\approx0.10S_c$, and dominate the two-dimensional modes until $S\approx2.05S_{c}$.
+(continued)
+\end{abstract}
+
+\maketitle
+
+\section{Abstract continued}
 These three-dimensional modes dominate for shear profiles relevant to stars and at magnetic Prandtl numbers relevant to liquid-metal laboratory experiments.
@@
 Finally, three-dimensional growth suggests direct dynamo action driven from the linear instability.
-\end{abstract}
-
-\maketitle
 
